### 2009-12-09 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Graph Algoriths in MapReduce}

  RMAT generation, pagerank, CC, triangle-finding, Luby, SSSP, other?
  how many variants of something like CC do we want to include?
    I would vote for just one
  discuss possible inefficiency due to a giant CC in CC algorithm

### 2009-12-09 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Introduction}

  what is MR
  why implement in MPI
  our motivation = graph algorithms, large data sets
    but lib can also be used for any MR algorithm, data agnostic
  novel features of our MR implementation
    C++ lib, open-source, MPI, out-of-core
    no fault tolerance

### 2009-12-09 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{MapReduce in MPI}

  keys and values, KV and KMV data structures
  map and reduce are naturally parallel on distributed KV, KMV
    user owns data, just set of bytes passed to library
    flavors of map()
  collate is equivalant to all2all() on distributed hash table
    how we do irregular communication
  big difference is we are doing the comm in syncronous manner
    data is reorganized locally, once it is all on-proc
  other related operations: compress, gather, collapse, sort

### 2009-12-09 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Out-of-core Issues}

  what part needs to be out-of-core in MPI context
  each proc does it's own disk IO
  basic idea: modified data structures for KV and KMV
  map() operation
  aggreate() operation
  convert() operation
  reduce() operation

### 2009-12-09 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.25in}

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

\title{Map{R}educe in {MPI} for Large-scale Graph Algorithms}

\author{
Steven J.~Plimpton, Karen D.~Devine, Jon?, Cohen?, Others? \\
Sandia National Laboratories \\
Albuquerque, NM \\
sjplimp@sandia.gov
}

\date{}

\maketitle

\centerline{Keywords: MapReduce, message-passing, MPI, graph
algorithms, RMAT matrices}

\vspace*{0.4in}

\begin{abstract}

We describe a library that allows parallel algorithms to be expressed
in the MapReduce paradigm, where the user does not need to write
explicit parallel code, but instead provides a collection of ``map''
and ``reduce'' functions which operate on portions of a data set.  The
library itself uses MPI message passing calls to perform the needed
parallel operations across distributed memory.  We highlight how
MapReduce operations work in an MPI context, and also how they can be
coded in an out-of-core manner for data sets that do not fit within
the aggregate memory of a parallel machine.  Our motivation for
creating this library was to enable graph algorithms to be coded as
MapReduce operations, allowing processing of Terabyte-scale data sets.
We outline a MapReduce version of several such algorithms: vertex
ranking via PageRank, connected component identification, triangle
finding, Luby's algorithm for maximally independent sets, and
single-source shortest-path.  To input arbitrarily large artificial
graphs we generate RMAT matrices in parallel; we also provide a
MapReduce version of this operation.  Performance and scalability
results for the various algorithms are presented for graphs of varying
sizes on a distributed-memory cluster.  For some cases, we compare our
results with timings for different algorithms, different machines, and
different MapReduce software, namely Hadoop.  Our open-source library
is written in C++, is callable from C++, C, Fortran, or scripting
languages such as Python, and can run on any parallel platform that
supports MPI.

\end{abstract}

\pagebreak

\input{intro}
\input{mr}
\input{outcore}
\input{graph}
\input{results}
\input{thanks}

\bibliographystyle{aip}
\bibliography{paper}

\end{document}

### 2009-12-09 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Performance Results}

  ideally, for each problem:
    define chosen characteristics of graph, e.g. edges/vertex
    choose 3 sizes:
      small = fits on 1 proc)
      medium = out-of-core on 1 proc, but not on P procs
      huge = impressive size
  show performance plots for each size running on 1 to P procs
    would like to see kink in plots as go out-of-core
    be able to judge and explain basic scaling results
  what machine: Odin or Tbird or RStorm?
    something that maps to a vanilla cluster
  where we have data:
    show multiple machines
    alternate non-MR algorithms
      discuss alternate CC algorithms via Trilinos and hybrid
    alternate MR implementations (Hadoop)

### 2009-12-09 paper/thanks.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Acknowledgements}

Sandia is a multiprogram laboratory operated by Sandia Corporation, a
Lockheed Martin Company, for the United States Department of
Energy's National Nuclear Security Administration under contract
DE-AC04-94AL85000.

Hadoop folks.

### 2010-03-26 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@478 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-  what is MR
-  why implement in MPI
-  our motivation = graph algorithms, large data sets
-    but lib can also be used for any MR algorithm, data agnostic
-  novel features of our MR implementation
-    C++ lib, open-source, MPI, out-of-core
-    no fault tolerance
+MapReduce is the programming paradigm popularized by Google
+researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
+enable analysis programs to be rapidly developed and deployed within
+Google to operate on massive data sets residing on their large
+distributed clusters.  They introduced a novel way of thinking about
+certain kinds of large-scale computations as a "map" operation
+followed by a "reduce" operation, possibly repeated.  The power of the
+paradigm is that when cast in this way, a traditionally serial
+algorithm now becomes two highly parallel operations which work on
+data local to each processor, sandwiched around an intermediate
+data-shuffling operation that requires inter-processor communication.
+The user need only write serial code for the application-specific map
+and reduce functions; the parallel data shuffle can be encapsulated in
+a library since its operation is independent of the application.
+
+The Google implementation of MapReduce is a C++ library with
+communication between networked machines via remote procedure calls.
+It allows for fault tolerance when large numbers of machines are used,
+and can use disks as out-of-core memory to process petabyte-scale data
+sets.  Tens of thousands of MapReduce programs have since been written
+by Google researchers and are a significant part of the daily compute
+tasks run by the company \cite{Dean2}.
+
+Similarly, the Hadoop implementation of MapReduce \cite{Hadoop}, has
+become widely popular in the past few years for parallel processing of
+large-scale data sets at Yahoo and other data-centric companies, as
+well as in university and laboratory research groups, due to its
+open-source availability.  MapReduce programs in Hadoop are typically
+written in Java, though it also supports use of stand-alone map and
+reduce kernels, which can be written as shell scripts or in other
+languages.
+
+More recently, MapReduce formulations of more traditional
+number-crunching kinds of scientific computational tasks have been
+described, such as post-processing analysis of simulation data
+\cite{Tu}, graph algorithmics \cite{Cohen}, and linear algebra
+operations \cite{Fox}.  The paper by Tu et al \cite{Tu} was
+particularly insightful to us, because it described how a MapReduce
+formalism could be implemented on top of the ubiquitous
+distributed-memory message-passing interface (MPI), and how the
+intermediate data-shuffle operation is conceptually identical to the
+familiar MPI_Alltoall() operation.  Their implementation of MapReduce
+was within a Python wrapper to simplify the writing of user programs.
+The paper motivated us to develop our own library built on top of MPI
+for use in graph analytics, which we initially released as open-source
+software in mid-2009 \cite{MRMPI}.  We have since worked to optimize
+several of the library's underlying algorithms and to enable its
+operation in out-of-core mode on larger data sets.  These algorithmic
+improvements are described in this paper and are part of the current
+downloadable version \cite{MRMPI}.
+
+The MapReduce-MPI (MR-MPI) library described in this paper is a simple
+and lightweight implementation of basic MapReduce functionality, with
+the following features and limitations:
+
+\begin{itemize}
+
+\item {\it C++ library using MPI for inter-processor communication:}
+The user writes a (typically) simple main program which runs on each
+processor in a parallel machine, and makes calls to the MR-MPI
+library.  For map and reduce operations, the library calls back to
+user provided map() and reduce() functions.  The use of C++ allows
+precise control over the memory and format of data allocated by each
+processor during a MapReduce.  Operations like the overall map,
+reduce, and data shuffle, are syncrhonous, meaning all the processors
+participate and finish the operation before proceeding.  Similarly the
+use of MPI within the library is the traditional synchronous mode of
+MPI Send and Recv calls on large aggregated data buffers, preferred
+for large-scale data movement.  A recent paper by \cite{Dongarra} also
+describes the MapReduce formalism from an MPI perspective, though they
+advocate a more asyncrhonous approach, using one-way communication of
+small messages.
+
+\item {Small, portable:} The entire library is a few thousand lines of
+standard C++ code.  For parallel operation, the program is linked with
+MPI, a standard message passing library available on all distributed
+memory machines.  For serial operation, a dummy MPI library (provided)
+can be substituted.
+
+\item {\it In-core or out-of-core operation:} Each MapReduce object
+allocates per-processor "pages" of memory, where the page size is
+determined by the user.  Typical MapReduce operations can be performed
+using a few such pages.  If the data set fits in a single page (per
+processor), then the library performs its operations in-core.  If the
+data set exceeds the page size, then each processor writes to
+temporary disk files (local or to a parallel disk system) as needed
+and subsequently reads from them.  This allows processing of data sets
+larger than the aggregate memory of all the processors, i.e. up to the
+aggregate available disk space.  No HDFS.
+
+\item store state, other MPI calls:} An advantage of writing a
+MapReduce program in C++ on top of MPI, is that the user program can
+easitly store state.  And it can make calls to MPI, e.g an AllReduce()
+that are not obvious how to do in MR formalism, such as Hadoop.  These
+were often quite nice to exploit in the graph algorihtms of Section 3.
+
+\item {\it C++, C, and Python interfaces provided:} A C++ interface
+means a user program instantiates and then invokes methods in one or
+more MapReduce objects.  A C interface means the library can also be
+called from C or other hi-level languages such as Fortran.  A Python
+interface means the library can be called from a Python script,
+allowing the user to write serial map() and reduce() functions in
+Python.  If a parallel machine can run Python in parallel, a parallel
+MapReduce can also be run in this mode.  The Python wrapper can be
+installed on any machine with a version of Python that includes the
+ctypes module, typically Python 2.5 or later.
+
+\item {No fault tolerance:} Current MPI implementations do not enable
+easy detection of a dead processor, or retrieval of the data it was
+working on.  So like most MPI programs, a MapReduce program calling
+the MR-MPI library will hang or crash if a processor goes away.
+
+The remainder of the paper is organized as follows.  The next two
+sections \ref{} and \ref[} describe how in-core and out-of-core
+MapReduce primitives are formulated as MPI-based operations in the
+MR-MPI library.  Section \ref{} briefly describes how several common
+graph-based algorithms can be formulated as MapReduce operations.
+Section \ref{} gives performance results for these algorithms running
+on a parallel cluster on a range of graph sizes, ranging from 1 Mb to
+1 Tb in vertex/edge counts.  We highlight the performance and
+complexity trade-offs of a MapReduce approach versus other more
+special-purpose algorithms.  The latter generally perform better but
+are harder to implement efficiently on distributed memory machines,
+due to the explicit management of parallelism required, particularly
+for out-of-core data sets.  We conclude with some lessons
+learned in Section \ref{}.
+
+Lessons learned at end:
+
+MapReduce algorithms are hard to design, but easy to write and debug.
+Thinking about a computational task in a MR mode is different than
+used to.  Often seems heretical to intentinally avoid data loca;lity
+in an MPI persepctive.  However writing small kernel map and reduce
+functions is eays. And avoiding parallel code is amazing.  And if you
+get the algorithmn right, sot that it to run on 1 processors, it often
+works without need for any additional debugging on P processors.
+
+MapReduce achieves parallelism thru randomized distribution of data
+across processors, i.e. intentionally ignore locality.  Requires
+maximal communication, but achieves good balance, even on
+hard-to-balance irregular data.  More tranditional parallel
+algorithsm, for things like matrix operations, grid or particle codes,
+tend to work hard to localize data, and minimize communication.  This
+means they have a lot of specialized, application-specific logic.
+
+Importance of state:
+algs could be improved by allowing state to be paged in at same
+time MR KV data is paged in, or just kept in core if possible
+  this requires ability to hash to desired processor
+  more processor centric view of data than Hadooop
+  possibly at odds with HDFS and data-centric view of Hadoop or MR
+
+Finally, though this paper focuses on graph operations in MapReduce,
+there is nothing about the MR-MPI library itself that is graph
+centric.  We hope the library can be generally useful for a variety of
+data-intense or compute-intense problems that are amenable to solution
+using a MapReduce framework on traditional large-scale parallel
+machines which support MPI.

### 2010-03-26 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@478 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-We describe a library that allows parallel algorithms to be expressed
-in the MapReduce paradigm, where the user does not need to write
-explicit parallel code, but instead provides a collection of ``map''
-and ``reduce'' functions which operate on portions of a data set.  The
-library itself uses MPI message passing calls to perform the needed
-parallel operations across distributed memory.  We highlight how
-MapReduce operations work in an MPI context, and also how they can be
-coded in an out-of-core manner for data sets that do not fit within
+We describe a parallel library that allows algorithms to be expressed
+in the MapReduce paradigm, where the calling program does not need to
+include explicit parallel code, but instead provides ``map'' and
+``reduce'' functions which operate in parallel on elements of a data
+set distributed across processors.  The library performs needed
+parallel operations using message-passing (MPI) calls.  We describe
+how typical MapReduce operations can be implemented in an MPI context,
+and also in an out-of-core manner for data sets that do not fit within
 the aggregate memory of a parallel machine.  Our motivation for
-creating this library was to enable graph algorithms to be coded as
+creating this library was to enable graph algorithms to be written as
 MapReduce operations, allowing processing of Terabyte-scale data sets.
-We outline a MapReduce version of several such algorithms: vertex
-ranking via PageRank, connected component identification, triangle
-finding, Luby's algorithm for maximally independent sets, and
-single-source shortest-path.  To input arbitrarily large artificial
-graphs we generate RMAT matrices in parallel; we also provide a
-MapReduce version of this operation.  Performance and scalability
-results for the various algorithms are presented for graphs of varying
-sizes on a distributed-memory cluster.  For some cases, we compare our
-results with timings for different algorithms, different machines, and
-different MapReduce software, namely Hadoop.  Our open-source library
-is written in C++, is callable from C++, C, Fortran, or scripting
-languages such as Python, and can run on any parallel platform that
-supports MPI.
+We outline MapReduce versions of several such algorithms: vertex
+ranking via PageRank, triangle finding, connected component
+identification, Luby's algorithm for maximally independent sets, and
+single-source shortest-path.  To test the algorithms on arbitrarily
+large artificial graphs we generate randomized RMAT matrices in
+parallel; we also describe a MapReduce version of this operation.
+Performance and scalability results for the various algorithms are
+presented for graphs of varying sizes on a distributed-memory cluster.
+For some cases, we compare the results with non-MapReduce algorithms,
+different machines, and different MapReduce software, namely Hadoop.
+Our open-source library is written in C++, is callable from C++, C,
+Fortran, or scripting languages such as Python, and can run on any
+parallel platform that supports MPI.
 

### 2010-03-26 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/intro.tex
@@
 \section{Introduction}
+\label{sec:intro}
 
-  what is MR
-  why implement in MPI
-  our motivation = graph algorithms, large data sets
-    but lib can also be used for any MR algorithm, data agnostic
-  novel features of our MR implementation
-    C++ lib, open-source, MPI, out-of-core
-    no fault tolerance
+MapReduce is the programming paradigm popularized by Google
+researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
+enable rapid development and deployment of analysis programs to
+operate on massive data sets residing on Google's large distributed
+clusters.  They introduced a novel way of thinking about certain kinds
+of large-scale computations as a "map" operation followed by a
+"reduce" operation.  The power of the paradigm is that when cast in
+this way, a nominally serial algorithm now becomes two highly parallel
+operations working on data local to each processor, sandwiched around
+an intermediate data-shuffling operation that requires inter-processor
+communication.  The user need only write serial code for the
+application-specific map and reduce functions; the parallel data
+shuffle can be encapsulated in a library since its operation is
+independent of the application.
+
+The Google implementation of MapReduce is a C++ library with
+communication between networked machines via remote procedure calls.
+It allows for fault tolerance when large numbers of machines are used,
+and can use disks as out-of-core memory to process petabyte-scale data
+sets.  Tens of thousands of MapReduce programs have since been written
+by Google researchers and are a significant part of the daily compute
+tasks run by the company \cite{Dean2}.
+
+Similarly, the open-source Hadoop implementation of MapReduce
+\cite{Hadoop}, has become widely popular in the past few years for
+parallel analysis of large-scale data sets at Yahoo and other
+data-centric companies, as well as in university and laboratory
+research groups, due to its free availability.  MapReduce programs in
+Hadoop are typically written in Java, though it also supports use of
+stand-alone map and reduce kernels, which can be written as shell
+scripts or in other languages.
+
+More recently, MapReduce formulations of traditional number-crunching
+kinds of scientific computational tasks have been described, such as
+post-processing analysis of simulation data \cite{Tu}, graph
+algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
+The paper by Tu et al \cite{Tu} was particularly insightful to us,
+because it described how MapReduce could be implemented on top of the
+ubiquitous distributed-memory message-passing interface (MPI), and how
+the intermediate data-shuffle operation is conceptually identical to
+the familiar $MPI_Alltoall$ operation.  Their implementation of
+MapReduce was within a Python wrapper to simplify the writing of user
+programs.  The paper motivated us to develop our own C++ library built
+on top of MPI for use in graph analytics, which we initially released
+as open-source software in mid-2009 \cite{MRMPI}.  We have since
+worked to optimize several of the library's underlying algorithms and
+to enable its operation in out-of-core mode on larger data sets.
+These algorithmic improvements are described in this paper and are
+part of the current downloadable version \cite{MRMPI}.
+
+The MapReduce-MPI (MR-MPI) library described in this paper is a
+simple, lightweight implementation of basic MapReduce functionality,
+with the following features and limitations:
+
+\begin{itemize}
+
+\item {\it C++ library using MPI for inter-processor communication:}
+The user writes a (typically) simple main program which runs on each
+processor of a parallel machine, making calls to the MR-MPI library.
+For map and reduce operations, the library calls back to user-provided
+map() and reduce() functions.  The use of C++ allows precise control
+over the memory and format of data allocated by each processor during
+a MapReduce.  Library calls for performing a map, reduce, or data
+shuffle, are synncrhonous, meaning all the processors participate and
+finish the operation before proceeding.  Similarly, the use of MPI
+within the library is the traditional mode of $MPI_Send()$ and
+$MPI_Recv$ calls between processor pairs using large aggregated
+messages to improve bandwidth performance and reduce latency costs.  A
+recent paper by \cite{Dongarra} also outlines the MapReduce formalism
+from an MPI perspective, though they advocate a more asyncrhonous
+approach, using one-way communication of small messages.
+
+\item {Small, portable:} The entire MR-MPI library is a few thousand
+lines of standard C++ code.  For parallel operation, the program is
+linked with MPI, a standard message passing library available on all
+distributed memory machines.  For serial operation, a dummy MPI
+library (provided) can be substituted.
+
+\item {\it In-core or out-of-core operation:} Each MapReduce object a
+processor defines, allocates per-processor "pages" of memory, where
+the page size is determined by the user.  Typical MapReduce operations
+can be performed using a few such pages.  If the data set fits in a
+single page (per processor), then the library performs its operations
+in-core.  If the data set exceeds the page size, then processors each
+write to temporary disk files (to local disk or a parallel file
+system) as needed and subsequently read from them.  This allows
+processing of data sets larger than the aggregate memory of all the
+processors, i.e. up to the available aggregate disk space.
+
+\item {\it Flexible programmability:} An advantage of writing a
+MapReduce program on top of MPI, is that the user program can invoke
+MPI calls directly, if desired.  For example, one-line calls to
+$MPI_Allreduce$ are often useful in determining the status of an
+iterative graph algorithm, as described in Section \ref{sec:graph}.
+The library interface also provides a user data pointer as an argument
+passed back to all callback functions, so it is easy for the user
+program to store ``state'' on each processor, accessible during the
+map and reduce operaions.  For example, various flags can be stored
+that alter the operation of a map or reduce function, as can richer
+data structures, that accumulate the results.
+
+\item {\it C++, C, and Python interfaces:} A C++ interface to the
+MR-MPI library means a user program instantiates and then invokes
+methods in one or more MapReduce objects.  A C interface means the
+library can also be called from C or other hi-level languages such as
+Fortran.  A C interface also means the library can be easily wrapped
+by Python via the Python ``ctypes'' module.  The library can then be
+called from a Python script, allowing the user to write map() and
+reduce() callback functions in Python.  If a machine supports running
+Python in parallel, a parallel MapReduce can also be run in this mode.
+
+\item {No fault tolerance:} Current MPI implementations do not enable
+easy detection of a dead processor, or retrieval of the data it was
+working on.  So like most MPI programs, a parallel program calling the
+MR-MPI library will hang or crash if a processor goes away.  Unlike
+Hadoop, and its HDFS file system which provides for data redundancy,
+the MR-MPI library simply reads and writes simple, flat files.  It can
+use local per-processor disks, or a parallel file system, if
+available, but these typically provide no data redundancy.
+
+\end{itemize}
+
+The remainder of the paper is organized as follows.  The next two
+sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
+out-of-core MapReduce primitives are formulated as MPI-based
+operations in the MR-MPI library.  Section \ref{sec:graph} briefly
+describes the formulation of several common graph algorithms as
+MapReduce operations.  Section \ref{sec:results} gives performance
+results for these algorithms running on a parallel cluster for graphs
+ranging in size from 1 million to 1 trillion vertices or edges.  In
+this section, we highlight the performance and complexity trade-offs
+of a MapReduce approach versus other more special-purpose algorithms.
+The latter generally perform better but are harder to implement
+efficiently on distributed memory machines, due to the required
+explicit management of parallelism, particularly for large out-of-core
+data sets.  Section \ref{sec:lessons} summarizes some lessons learned
+from the implementation and use of our library.

### 2010-03-26 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{MapReduce in MPI}
+\label{sec:mr}
 
@@
   other related operations: compress, gather, collapse, sort
+
+MR-MPI~\cite{MRMPIManual} is a C++ library using MPI for
+inter-processor communication.  The basic data units in MapReduce are
+key/value pairs.  The data in keys and values can have any type or
+length; they are stored in simple character arrays.  A related data
+type is the key/multivalue pair, where all values associated with the
+same key are collected and stored together as a multivalue.
+
+The three basic operations in a typical MR-MPI program are {\it
+map()}, {\it collate()}, and {\it reduce()}.  The {\it map()}
+operation generates key/value pairs on each processor.  Typically,
+this is done from data read in from files, but a {\it map()} may
+generate data itself or process existing key/value pairs to generate
+new ones.  The key/value pairs produced are stored locally on each
+processor; a {\it map()} requires no inter-processor communication.
+Users provide a pointer to their own map function; the MR-MPI {\it
+map()} operation invokes the user function as a callback.
+
+The {\it collate()} operation identifies unique keys and collects all
+the values associated with those keys to create key/multivalue pairs.
+This is done in two stages, the first of which requires communication,
+since key/value pairs with the same key may be owned by any processor.
+Each processor hashes its keys to an ``owning'' processor.  It then
+sends each of its key/value pairs to the processor which owns the key.
+After receiving key/value pairs, the second stage is an on-processor
+computation, whereby each processor constructs key/multivalue pairs
+for each unique key it now owns, using a local hash table.  Thus the
+overall {\it collate()} is equivalent to an all-to-all communication
+using a distributed hash table followed by a local reordering of the
+data.  We use synchronous point-to-point irregular communication
+instead of the MPI\_Alltoall() library call, since this allows control
+of message sizes and sequencing.
+
+The {\it reduce()} operation processes key/multivalue pairs and can
+produce new key/value pairs for continued computation.  Each processor
+operates on the key/multivalue pairs it owns; no communication is
+required.  As with the {\it map()}, users provide pointers to their
+own reduce functions.  The MR-MPI {\it reduce()} invokes the user's
+function once for each key/multivalue pair.  Output from a {\it
+reduce()} operation can serve as input to a subsequent {\it map()} or
+{\it collate()}.
+
+Several related operations are provided by MR-MPI.  For example, a
+{\it compress()} combines on-processor key/value pairs to eliminate
+duplicate keys, forming single key/value pairs with an aggregated
+value.  The {\it gather()} operation collects key/values pairs from
+all processors to a subset of processors; it is useful for doing
+output from a single or subset of processors.  Library calls for
+sorting data by key or value or within multivalues are also provided.
+
+Since the MapReduce paradigm was designed to enable processing of very
+large data sets, we provide some details of how MR-MPI manages memory
+and large disk files.  The user (optionally) specifies a {\it memsize}
+parameter when instantiating a MapReduce object, e.g. 1 Gbyte.  This
+amount of memory is allocated by each processor for storing key/value
+and key/multivalue pairs, as well as communication buffers and a hash
+table for the {\it collate{}} operation.
+
+If the data generated by various MR-MPI operations can be stored by
+each processor within its allocated chunk of memory, we refer to this
+as ``in-core'' processing and no disk files are written or read (aside
+from initial input data if it exists or final output data if
+generated).  Three additional points are worth noting.
+
+At some stages in a MapReduce algorithm, the MapReduce object will
+typically need to store both key/value and key/multivalue pairs at the
+same time.  For example, during a {\it reduce()} operation,
+key/multivalue pairs serve as input to the user's reduce function,
+while the user's function may be generating new key/value pairs.
+
+The memory available for storing key/value and key/multivalue pairs is
+{\it memsize} summed across all processors.  So long as the key/value
+pairs remain reasonably evenly distributed across processors then this
+can be a significant amount of memory, e.g. 1 Tbyte on a
+1024-processor machine.  The hash function used by the {\it collate()}
+operation attempts to randomize key assignment to processors so
+effective load-balance occurs.
+
+For some MapReduce algorithms, it is convenient to create multiple
+MapReduce objects and pass data back and forth between them.
+Obviously, the user must then partition physical memory judiciously
+between these various objects if all the data is to stay in-core.  The
+{\it memsize} parameter can be specified for each MapReduce object
+created.

### 2010-03-26 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{Out-of-core Issues}
+\label{sec:outcore}
 
@@
   reduce() operation
+
+If the set of key/value or key/multivalue pairs generated or owned by
+a processor exceeds the {\it memsize} of its allocated memory, then
+MR-MPI switches to ``out-of-core'' processing.  Each processor uses
+temporary files to store data that does not fit in its memory.
+Depending on the parallel machine's hardware configuration, these
+files can reside on disks local to each processor, on the front end
+(typically a NSF-mounted file system for a parallel machine), or on a
+parallel file system/disk array.  The files are read/written as large
+contiguous ``pages'', which are 1/4 the size of the {\it memsize}
+setting discussed in the previous section.
+
+We now explain how the MapReduce primitives described above operate in
+out-of-core mode.  The {\it map()} and {\it reduce()} operations are
+relatively simple.  As a {\it map()} generates key/value pairs, it
+fills up a page, one key/value pair at a time.  When the page is full,
+it is written to disk.  If the source of data for the {\it map()}
+operation is an existing set of key/value pairs, then those datums are
+read in, one page at a time, and each key/value pair is given to the
+user map() function.  Similarly, a {\it reduce()} reads one page of
+key/multivalue pairs from disk, passes them one at a time to the user
+reduce() function, which typically generates new key/value pairs.  The
+generated pairs fill up a new page, which is written to disk when
+full, just as with the {\it map()} operation.  Thus for both a {\it
+map()} and {\it reduce()}, out-of-core disk files are read and written
+sequentially, one page at a time.
+
+A special case is when a single key/multivalue pair is larger than a
+single page.  This can happen, for example, in a connected component
+finding algorithm if the graph collapses into one or a few giant
+components.  In this case, the entire set of values (graph vertices
+and edges) associated with the unique key (the component ID), may not
+fit in one page of memory.  The individual values in the multivalue
+will then be spread across as many pages as needed.  The user reduce()
+function is passed a flag indicating it is receiving only a portion of
+the values.  Once it has processed them, it can request a new set of
+values (from a new page) from the MR-MPI library.
+
+Performing a {\it collate()} operation out-of-core is more complex.
+Recall that a collate() operation occurs in two stages.  First, keys
+are hashed to ``owning'' processors and the key/value data is
+communicated in an all-to-all fashion.  This is done one page of
+key/value data at a time.  Each processor reads in a page of key/value
+data, the communication pattern is determined (who needs what datums),
+and the communication is performed.  So long as each processor
+receives no more than 2x its share of the datums (one page), it has
+room to receive it within the {\it memsize} restriction.  If the
+key/value data is load-imbalanced so that this limit is exceeded, then
+the all-to-all communication can be performed in multiple passes.
+
+When the first stage is done, each processor now has an out-of-core
+set of key/value datums that need to be reorganized into an
+out-of-core set of key/multivalue pairs.  In principle, to create one
+page of key/multivalue data, the entire set of key/value pairs needs
+to be scanned to find all keys which contribute values to that page.
+Reading all the key/value pages from disk for each page of
+key/multivalue output could be prohibitively expensive.  A related
+issue is that a hash table is needed to match new keys with previously
+encountered keys.  But there is no guarantee that the hash table
+itself will fit within the memory specified by the user via the {\it
+memsize} parameter.  Our solution to these issues is to partition the
+{\it memsize} memory into 3 pieces: 2 pieces each of size {\it
+memsize}/4 for reading/writing pages of key/value and key/multivalue
+pairs, and the remaining {\it memsize}/2 piece for a hash table.  We
+can then scan the key/value pairs at most four times to produce all
+the pages of key/multivalue output, as diagrammed in Figure
+\ref{f:collate}.
+
+\begin{figure}
+\includegraphics[width=\textwidth,angle=-90]{fig_collate.pdf}
+\caption{Four-pass algorithm for converting key/value data to
+key/multivalue data.  The vertical lines represent out-of-core data
+sets.  KV is key/value pairs; HT is an in-memory hash table; KMV is
+key/multivalue pairs.}
+\label{f:collate}
+\end{figure}
+
+\begin{itemize}
+
+\item (Pass 1) The key/values are read and split into two files.  The
+first file contains all the key/values pairs whose keys fit into the
+finite-size hash table.  The second contains those which do not.
+
+\item (Pass 2) The size of the first file (generally smaller) can be
+used to estimate how many chunks the second file (generally larger)
+needs to be partitioned into so that each chunk contains a set of
+unique keys than can fit into the hash table.  This estimate is made
+conservatively so that additional passes through the data are almost
+never required.  The second pass thus reads the larger file and writes
+out smaller files, similar in size to the first file created by the
+first pass.
+
+\item (Pass 3) The unique keys in each small file of key/value pairs
+can now be hashed, but the file may contain more value data than will
+fit in one key/multivalue page.  As each small file is read, and its
+keys hashed, the accumulating hash table statistics can be used to
+determine which key/values correspond to single pages of
+key/multivalue output.  The key/values can thus be split into even
+smaller files, one for each key/multivalue page.
+
+\item (Pass 4) Each smaller file of key/value pairs is read, its keys
+hashed, and the key/value data reorganized into sequential
+key/multivalue pairs which fill a key/multivalue page, which is then
+written out to disk.
+
+\end{itemize}
+
+Note that depending on the data set size and its key distribution all
+four passes through the data may not be necessary, but the upper limit
+is four, assuming the chunk-size estimations are accurate.  By
+comparison, a merge sort of key/value pairs would require $log_2(N)$
+passes (read/write) through the out-of-core data set, where $N$ is its
+number of pages.  The MR-MPI {\it collate()} operation does not
+actually sort the key/multivalue pairs by key; the final ordering is
+indeterminate as it is created by the hash function.
+
+The other MR-MPI library calls ({\it compress()}, {\it gather()},
+sorting by key or value, etc) can also operate in out-of-core mode,
+typically with one pass through the key/value or key/multivalue data.
+Sorting is an exception, as an out-of-core merge sort requires
+$log_2(N)$ passes, as just mentioned, where $N$ is the number of
+pages.

### 2010-03-26 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \author{
-Steven J.~Plimpton, Karen D.~Devine, Jon?, Cohen?, Others? \\
+Steven J.~Plimpton, Karen D.~Devine, Others? \\
 Sandia National Laboratories \\
@@
 
-We describe a parallel library that allows algorithms to be expressed
-in the MapReduce paradigm, where the calling program does not need to
-include explicit parallel code, but instead provides ``map'' and
-``reduce'' functions which operate in parallel on elements of a data
-set distributed across processors.  The library performs needed
-parallel operations using message-passing (MPI) calls.  We describe
-how typical MapReduce operations can be implemented in an MPI context,
-and also in an out-of-core manner for data sets that do not fit within
-the aggregate memory of a parallel machine.  Our motivation for
-creating this library was to enable graph algorithms to be written as
-MapReduce operations, allowing processing of Terabyte-scale data sets.
-We outline MapReduce versions of several such algorithms: vertex
-ranking via PageRank, triangle finding, connected component
-identification, Luby's algorithm for maximally independent sets, and
-single-source shortest-path.  To test the algorithms on arbitrarily
+We describe a parallel library written with message-passing (MPI)
+calls that allows algorithms to be expressed in the MapReduce
+paradigm.  This means the calling program does not need to include
+explicit parallel code, but instead provides ``map'' and ``reduce''
+functions which operate independently on elements of a data set
+distributed across processors.  The library performs needed data
+movement between processors.  We describe how typical MapReduce
+funtionality can be implemented in an MPI context, and also in an
+out-of-core manner for data sets that do not fit within the aggregate
+memory of a parallel machine.  Our motivation for creating this
+library was to enable graph algorithms to be written as MapReduce
+operations, allowing processing of Terabyte-scale data sets.  We
+outline MapReduce versions of several such algorithms: vertex ranking
+via PageRank, triangle finding, connected component identification,
+Luby's algorithm for maximally independent sets, and single-source
+shortest-path calculation.  To test the algorithms on arbitrarily
 large artificial graphs we generate randomized RMAT matrices in
-parallel; we also describe a MapReduce version of this operation.
+parallel; a MapReduce version of this operation is also described.
 Performance and scalability results for the various algorithms are
-presented for graphs of varying sizes on a distributed-memory cluster.
+presented for varying size graphs on a distributed-memory cluster.
 For some cases, we compare the results with non-MapReduce algorithms,
@@
 \input{results}
+\input{lessons}
 \input{thanks}

### 2010-03-26 paper/thanks.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{Acknowledgements}
+\label{sec:thanks}
 
-Sandia is a multiprogram laboratory operated by Sandia Corporation, a
-Lockheed Martin Company, for the United States Department of
-Energy's National Nuclear Security Administration under contract
-DE-AC04-94AL85000.
+We thank the following individuals for their contributions to this
+paper: Greg Bayer and Todd Plantenga (Sandia) for explaining Hadoop
+concepts to us, and for the Hadoop implementations and timings of
+Section \ref{sec:results}; Jon Cohen (NSA) for fruitful discussions
+about his MapReduce graph algorithms \cite{Cohen}; Brian Barrett
+(Sandia) for the PBGL results of Section \ref{sec:results}; Jon Berry
+(Sandia) for the MTGL results of Section \ref{sec:results}, and for
+his overall support of this work and many useful discussions.
 
-Hadoop folks.
+Sandia National Laboratories is a multi-program laboratory operated by
+Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
+company, for the U.S. Department of Energy's National Nuclear Security
+Administration under contract DE-AC04-94AL85000.

### 2010-03-26 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@480 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/mr.tex
@@
 \section{MapReduce in MPI}
+\label{sec:mr}
 
-  keys and values, KV and KMV data structures
-  map and reduce are naturally parallel on distributed KV, KMV
-    user owns data, just set of bytes passed to library
-    flavors of map()
-  collate is equivalant to all2all() on distributed hash table
-    how we do irregular communication
-  big difference is we are doing the comm in syncronous manner
-    data is reorganized locally, once it is all on-proc
-  other related operations: compress, gather, collapse, sort
+The basic datums stored and operated on by any MapReduce framweork are
+key/value (KV) pairs.  In the MR-MPI library, individual keys or
+values can be of any data type or length, or combinations of multiple
+types (one integer, a string of characters, two integers and a double,
+etc); they are simply treated as byte strings by the library.  A KV
+pair always has a key; its value may be NULL.  A related data type is
+the key/multivalue (KMV) pair, where all values associated with the
+same key are collected and stored contiguously as a multivalue, which
+is just a longer byte string with an associated vector of integer
+lengths, one per value.
+
+A typical MR-MPI program makes at least 3 calls to the MR-MPI library,
+to perform a {\it map()}, {\it collate()}, and {\it reduce()}
+operation.  In a {\it map()}, zero or more key/value pairs are
+generated by each processor.  Typically, this is done using data read
+in from files, but a {\it map()} may generate data itself or process
+existing KV pairs to create new ones.  The KV pairs produced are
+stored locally by each processor; a {\it map()} thus requires no
+inter-processor communication.  Users call the library with a count of
+tasks to perform and a pointer to a user function; the MR-MPI {\it
+map()} operation invokes the user function multiple times as a
+callback.  Depending on which variant of {\it map()} is called, the
+user function may be passed a file name, a chunk of bytes from a large
+file, or a task ID.  Options for assigning map tasks to processors.
+
+The {\it collate()} operation (or data shuffle in Hadoop) identifies
+unique keys and collects all the values associated with those keys to
+create KMV pairs.  This is done in two stages, the first of which
+requires communication, since KV pairs with the same key may be owned
+by any processor.  Each processor hashes each of its keys to determine
+the processor who ``owns'' it.  The $N$-byte length key is hashed into
+a 32-bit value whose remainder modulo $P$ generates the owning
+processor ID.  Alternatively, the user can provide a hash function
+which converts the key into a processor ID.  Each processor then sends
+each of its KV pairs to the owning processor.
+
+After receiving new KV pairs, the second stage is an on-processor
+computation, requiring no further communication.  Each processor
+reorganizes its KV pairs into KMV pairs, one for each unique key it
+owns.  This is done using a hash table, rather than a sort.  A sort
+requires $\log(N)$ passes through the $N$ KV pairs.  the list of KMV
+pairs can be created in 2 passes through the KV pairs, one to populate
+the hash table with needed count and length information, and the
+second pass to copy the key and value datums into the appropriate
+location in a new KMV data structure.  Since the lookup of a key in a
+well-formed hash table is a constant-time $O(1)$ operation, the cost
+of the data reorganization is $O(N)$, rather than $N\log(N)$ for full
+sort.
+
+Note that the first portion of the {\it collate()} operation involves
+all-to-all communication (each processor sends and receives data from
+every other processor) driven by a distributed hash table.  The
+communication can either be done via a MPI\_Alltoall() library call,
+or by a custom routine that aggregates messages and invokes pointwise
+MPI\_Send() and MPI\_IRecv() calls.
+
+The {\it reduce()} operation processes KMV pairs and can produce new
+KV pairs for continued computation.  Each processor operates only on
+the KMV pairs it owns; no communication is required.  As with the {\it
+map()}, users call the library with a pointer to a user function.  The
+MR-MPI {\it reduce()} operation invokes the user function, once for
+each KMV pair.
+
+Several related MapReduce operations are provided by the library.  For
+example, the {\it compress()} operation combines on-processor KV pairs
+to eliminate duplicate keys, forming new KV pairs with an aggregated
+value.  The { \it convert()} operation turns a list of KV pairs into
+KMV pairs, with one value per key.  The {collapse()} operation turns
+$N$ KV pairs into one KMV pair, with the keys and values of the KV
+pairs becoming $2N$ values of a single multivalue assigned to a new
+key on the KMV.  The {\it gather()} operation collects KV pairs from
+all processors to a subset of processors; it is useful for doing
+output from a single or subset of processors.  Library calls for
+sorting datums by key or value or within multivalues are also
+provided.  These routines invoke the C-library quicksort() function to
+compute the sorted ordering of the KV pairs (or values within a
+multivalue), using a user-provided comparison function.  The KV pairs
+(or values in a multivalue) are then copied one-by-one into a new data
+structure in sorted order.
+
+The purpose of providing an interface to various lo-level operations,
+is that a user program can string them together in various ways to
+produce interesting MapReduce algorithms.  For example, output from a
+{\it reduce()} can serve as input to a subsequent {\it map()} or {\it
+collate()}.  KV pairs from mutiple MapReduce objects can be combined
+together to perform new sequences of {\it map()}, {\it collate()}, and
+{\it reduce()} operations.
+
+The above discussion assumed that the KV or KMV pairs stored by a
+processor fit in its physical memory.  If this is the case, then we
+refer to this as ``in-core'' processing and no disk files are written
+or read by any of the processors, aside from initial input data if it
+exists or final output data if it is generated.  Note that the
+aggregate physical memory of large parallel machines can be multiple
+Tbytes, which allows for large data sets to be processed in-core,
+assuming the KV and KMV pairs remain evenly distributed across
+processors throughout the sequence of MapReduce operations.  The use
+of randomized hashing to assign keys to processors is designed for
+such load-balancing.  What happens when data sets do not
+fit in available memory, is the subject of the next section.

### 2010-03-26 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@480 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
   reduce() operation
+
+Since the MapReduce paradigm was designed to enable processing of
+extremely large data sets, the MR-MPI library also allows for
+``out-of-core'' processing, which is triggered when the KV or KMV
+pairs owned by a processor do not fit in its memory.
+
+When the user program creates a MapReduce object, the user specifies
+a 
+The memory available for storing key/value and key/multivalue pairs is
+{\it memsize} summed across all processors.  So long as the key/value
+pairs remain reasonably evenly distributed across processors then this
+can be a significant amount of memory, e.g. 1 Tbyte on a
+1024-processor machine.  The hash function used by the {\it collate()}
+operation attempts to randomize key assignment to processors so
+effective load-balance occurs.
 

### 2010-03-26 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@481 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 of the data reorganization is $O(N)$, rather than $N\log(N)$ for full
-sort.
+sort.  It also has the added benefit that the number of values in each
+KMV pair is known, which is passed to the user function during a
+reduce.  For some reduce operations, this is all the information a
+reduce requires; thus the values need not be looped over.  This is not
+the case for a Hadoop-style data reorg which sorts the values.  It
+does not know a priori how many values are associated with the same
+key without iterating over them.
 

### 2010-03-26 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@481 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/outcore.tex
@@
 \section{Out-of-core Issues}
+\label{sec:outcore}
 
-  what part needs to be out-of-core in MPI context
-  each proc does it's own disk IO
-  basic idea: modified data structures for KV and KMV
-  map() operation
-  aggreate() operation
-  convert() operation
-  reduce() operation
+Since the MapReduce paradigm was designed to enable processing of
+extremely large data sets, the MR-MPI library also allows for
+``out-of-core'' processing, which is triggered when the KV or KMV
+pairs owned by a processor do not fit in its memory.  When this
+occurs, a processor writes one or more temporary files to disk,
+containing KV or KMV pairs, and reads them when required.  Depending
+on the parallel machine's hardware configuration, these files can
+reside on disks local to each processor, on the front end (typically a
+NSF-mounted file system for a parallel machine), or on a parallel file
+system/disk array.
+
+When a user program creates a MapReduce object, a ``pagesize'' can be
+specified, which defaults to 64 Mbytes.  As described below, each
+MR-MPI operation, is constrained to use no more than a handful of
+these pages.  The {\it pagesize} setting can be as small as 1 Mbyte or
+as large as desired, though the user should insure the allocated pages
+fit in physical memory, else the MR-MPI library may allocate slow
+virtual memory.  The {\it pagesize} is also the typical size of
+individual reads and writes to the temporary disk files; hence a
+reasonable pagesize insures good I/O performance.
+
+We now explain how the MapReduce operations described in the
+preceeding section work in out-of-core mode.  The {\it map()} and {\it
+reduce()} operations are relatively simple.  As a {\it map()}
+generates KV pairs via the user function, a page of memory is filled
+up, one KV pair at a time.  When the page is full, it is written to
+disk.  If the source of data for the {\it map()} operation is an
+existing set of KV pairs, then those datums are read in, one page at a
+time, and a pointer to each KV pair is given to the user function.
+Similarly, a {\it reduce()} reads one page of KMV pairs from disk,
+passes a pointer to each pair, one at a time to the user function,
+which typically generates new KV pairs.  The generated pairs fill up a
+new page, which is written to disk when full, just as with the {\it
+map()} operation.  Thus for both a {\it map()} and {\it reduce()},
+out-of-core disk files are read and written sequentially, one page at
+a time, requiring at most two pages of memory.
+
+A special case is when a single KMV pair is larger than a single page.
+This can happen, for example, in a connected component finding
+algorithm if the graph collapses into one or a few giant components.
+In this case, the set of values (graph vertices and edges) associated
+with a unique key (the component ID), may not fit in one page of
+memory.  The individual values in the multivalue will then be spread
+across as many pages as needed.  The user function that processes the
+KMV pair is passed a flag indicating it is receiving only a portion of
+the values.  Once it has processed them, it can request a new set of
+values from the MR-MPI library, which reads in a new page.
+
+Performing a {\it collate()} operation out-of-core is more complex.
+Recall that a collate() operation occurs in two stages.  First, keys
+are hashed to ``owning'' processors and the KV datums are communicated
+in an all-to-all fashion.  In out-of-core mode, this is done one page
+of KV data at a time.  Each processor reads in a page of KV pairs, the
+communication pattern for the datums in that page is determined (who
+needs what datums), and all-to-all communication is performed.  Each
+processor allocates a two-page chunk of memory to receive incoming KV
+pairs.  On average each processor should receive KV pairs that fill
+only one page; the two-page allocation allows for some load-imbalance.
+If the KV pairs are distributed unevenly so that this limit is
+exceeded, then the all-to-all communication is performed in smaller,
+multiple passes until the full page (contributed by each processor) is
+completed.  Performing an MPI\_Alltoall(), or using the custom version
+provided by the MR-MPI library, requires auxiliary arrays that store
+individual datum lengths and pointers.  In the worst-case scenario
+(many small KV pairs), the total number of pages required to perform
+the all-to-all communication, including the source and target KV pages
+is 7.
+
+When the communication stage of the {\it collate()} operation is
+complete, each processor now has a set of KV pages, stored in an
+out-of-core KV file, that need to be reorganized into a set of KMV
+pages, stored in new out-of-core KMV file.  In principle, to create
+one page of KMV pairs, requires all KV pages to be scanned to find all
+the keys which contribute values to that page.  Doing this for each
+output page of KMV pairs could be prohibitively expensive.  A related
+issue is that a hash table is needed to match new keys with previously
+encountered keys, as described in the previous section.  But there is
+no guarantee that the hash table itself will not grow arbitrarily
+large for some data sets with many unique keys.  We need an algorithm
+for generating KMV pairs which operates within a small number of
+memory pages and performs a minimal number of passes through the KV
+disk file.  An algorithm that meets these goals is diagrammed in
+Figure \ref{f:collate}; it requires 4 read passes of the KV pages and
+3 write passes.  It also uses a finite-size in-memory hash table,
+which stores unique keys as they are encountered in the loops over KV,
+as well as auxiliary information needed to construct the output pages
+of KMV pairs.
+
+\begin{figure}
+\includegraphics[width=\textwidth,angle=-90]{fig_collate.pdf}
+\caption{Multi-pass algorithm for converting KV data to KMV data.  The
+vertical lines represent out-of-core data sets.  KV is key/value
+pairs; HT is an in-memory hash table; KMV is key/multivalue pairs.}
+\label{f:collate}
+\end{figure}
+
+\begin{itemize}
+
+\item (Pass 1) The KV pairs are read and split into ``partition''
+files.  Each partition file contains KV pairs whose unique keys will
+(likely) fit in the hash table (HT).  Intially all KV pairs are
+assigned to the first partition file as the HT is populated.  When the
+HT becomes full, e.g. at the point represented by the horizontal line
+in a downward scan of the leftmost vertical line, then the fraction of
+KV pairs read thus far is used to estimate the number of additional
+partition files needed.  As subsequent KV pairs are read in, they are
+written into either the original partition file or one of the new
+ones.  The former if its key exists in the current HT; the latter if
+not, and a subset of bits in the hash value of its key is used to
+assign it to one of the partition files.  The number of needed
+partition files is estimated conservatively, rounding up to the next
+power-of-two, so that extra passes through the KV pairs are almost
+never needed, and so the bit masking can be done quickly.  This pass
+requires both a read and write of all the KV pairs.
+
+\item (Pass 2) The KV pairs in a partition file are read, and their
+unique keys are hashed into the HT, accumulating data about the number
+and size of values associated with each key.  Before the next
+partition file is read, passes 3 and 4 are completed for this
+partition.  But eventually, this pass reads all partition files,
+requiring a read of all KV pairs.
+
+\item (Pass 3) The unique keys in the HT are scanned.  The associated
+values may create KMV pairs which span many pages.  If this is the
+case, the associated partition file of KV pairs is read, and assigned
+to smaller ``set'' files, using the accumulated HT statistics.  Each
+set file corresponds to the KV pairs which will contribute to the set
+of KMV pairs that will populate one page of KMV output.  If a KMV pair
+spans multiple pages, because it has a very large number of values,
+then the corresponding KV pairs will still be part of a single set
+file.  Eventually, this pass reads all partition files, requiring both
+a read and write of all KV pairs.
+
+\item {Pass 4) A set file is read and the key/value data for each of
+its KV pairs is copied into the appropriate place in the page of KMV
+pairs, using information stored in the HT.  When complete, the KMV
+page is written to disk.  This pass thus eventually reads all set
+files, requiring a read of all KV pairs.  It also writes all KMV
+pairs.  If each KV pair has a unique key, the size of KMV output is
+roughly the same as that of the KV input.
+
+\end{itemize}
+
+In summary, this data reorganization for the {collate()} operation is
+somewhat complex, but requires a small, constant number of passes
+through the data.  Namely, at most 4 read passes through the pages of
+KV pairs, and 3 write passes.  The first two write passes reorganize
+KV pairs into partition and set files; the final write pass creates
+the KMV data set.  By contrast, a full out-of-core sort() of the KV
+pairs, performed via a merge sort as discussed below, requires
+$O(\log{M})$ read/write passes through the KV data, where $M$ is the
+number of pages of KV pairs, which can be a large number for big data
+sets.
+
+Intermediate passes done via spool files.  Can't bound precisely,
+since depends on data characteristics (see tech discussion in manual),
+but generally fit wihtin 7 pages of aggregate.
+
+Note that depending on the data set size and its key distribution all
+four passes through the data may not be necessary, but the upper limit
+is four, assuming the chunk-size estimations are accurate.  By
+comparison, a merge sort of KV pairs would require $log_2(N)$
+passes (read/write) through the out-of-core data set, where $N$ is its
+number of pages.  The MR-MPI {\it collate()} operation does not
+actually sort the KMV pairs by key; the final ordering is
+indeterminate as it is created by the hash function.
+
+Avoid extra passes if not needed.  E.g. one partition file, then never
+even written.  First partition file not re-read to popluate HT.
+
+Uneven distribution of unique keys in KV pair could induce extra
+passes due to not enough partitions, but quite rare.
+
+The other various MR-MPI library calls ({\it compress()}, {\it
+gather(), etc}, can also operate in out-of-core mode, typically with
+one pass through the KV or KMV data.  Sorting, by
+key or value, is an exception.  as an out-of-core merge sort requires
+$log_2(N)$ passes, as just mentioned, where $N$ is the number of
+pages.

### 2010-03-26 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@482 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{Graph Algoriths in MapReduce}
+\label{sec:graph}
 
-  RMAT generation, pagerank, CC, triangle-finding, Luby, SSSP, other?
-  how many variants of something like CC do we want to include?
-    I would vote for just one
-  discuss possible inefficiency due to a giant CC in CC algorithm
+RMAT generation
+
+pagerank
+
+triangle finding
+
+connected component identification
+
+do wee need differnt variants of CC?
+
+could just do one, discuss possible inefficiency due to giant CC
+
+Luby algorithm for maximally independent sets
+
+SSSP
+
+discuss new optimization in SSSP?

### 2010-03-29 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@483 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 ``out-of-core'' processing, which is triggered when the KV or KMV
-pairs owned by a processor do not fit in its memory.  When this
-occurs, a processor writes one or more temporary files to disk,
-containing KV or KMV pairs, and reads them when required.  Depending
-on the parallel machine's hardware configuration, these files can
-reside on disks local to each processor, on the front end (typically a
-NSF-mounted file system for a parallel machine), or on a parallel file
-system/disk array.
+pairs owned by oen or more processors do not fit in its local memory.
+When this occurs, a processor writes one or more temporary files to
+disk, containing KV or KMV pairs, and reads them back in when
+required.  Depending on the parallel machine's hardware configuration,
+these files can reside on disks local to each processor, on the front
+end (typically a NSF-mounted file system for a parallel machine), or
+on a parallel file system/disk array.
 
@@
 individual reads and writes to the temporary disk files; hence a
-reasonable pagesize insures good I/O performance.
+reasonable {\it pagesize} insures good I/O performance.
 
-We now explain how the MapReduce operations described in the
-preceeding section work in out-of-core mode.  The {\it map()} and {\it
-reduce()} operations are relatively simple.  As a {\it map()}
-generates KV pairs via the user function, a page of memory is filled
-up, one KV pair at a time.  When the page is full, it is written to
-disk.  If the source of data for the {\it map()} operation is an
-existing set of KV pairs, then those datums are read in, one page at a
-time, and a pointer to each KV pair is given to the user function.
-Similarly, a {\it reduce()} reads one page of KMV pairs from disk,
-passes a pointer to each pair, one at a time to the user function,
-which typically generates new KV pairs.  The generated pairs fill up a
-new page, which is written to disk when full, just as with the {\it
-map()} operation.  Thus for both a {\it map()} and {\it reduce()},
-out-of-core disk files are read and written sequentially, one page at
-a time, requiring at most two pages of memory.
+We now explain how the MapReduce operations described in the previous
+section work in out-of-core mode.  The {\it map()} and {\it reduce()}
+operations are relatively simple.  As a {\it map()} generates KV pairs
+via the user function, a page of memory fills up, one KV pair at a
+time.  When the page is full, it is written to disk.  If the source of
+data for the {\it map()} operation is an existing set of KV pairs,
+then those datums are read in, one page at a time, and a pointer to
+each KV pair is given to the user function.  Similarly, a {\it
+reduce()} reads one page of KMV pairs from disk, passes a pointer, one
+at a time, to each pair to the user function which typically generates
+new KV pairs.  The generated pairs fill up a new page, which is
+written to disk when full, just as with the {\it map()} operation.
+Thus for both a {\it map()} and {\it reduce()}, out-of-core disk files
+are read and written sequentially, one page at a time, which requires
+at most two pages of memory.
 
@@
 across as many pages as needed.  The user function that processes the
-KMV pair is passed a flag indicating it is receiving only a portion of
-the values.  Once it has processed them, it can request a new set of
-values from the MR-MPI library, which reads in a new page.
+KMV pair is passed a flag indicating it received only a portion of the
+values.  Once it has processed them, it can request a new set of
+values from the MR-MPI library, which reads in a new page from disk.
 
-Performing a {\it collate()} operation out-of-core is more complex.
-Recall that a collate() operation occurs in two stages.  First, keys
-are hashed to ``owning'' processors and the KV datums are communicated
-in an all-to-all fashion.  In out-of-core mode, this is done one page
-of KV data at a time.  Each processor reads in a page of KV pairs, the
-communication pattern for the datums in that page is determined (who
-needs what datums), and all-to-all communication is performed.  Each
-processor allocates a two-page chunk of memory to receive incoming KV
-pairs.  On average each processor should receive KV pairs that fill
-only one page; the two-page allocation allows for some load-imbalance.
-If the KV pairs are distributed unevenly so that this limit is
-exceeded, then the all-to-all communication is performed in smaller,
-multiple passes until the full page (contributed by each processor) is
-completed.  Performing an MPI\_Alltoall(), or using the custom version
-provided by the MR-MPI library, requires auxiliary arrays that store
-individual datum lengths and pointers.  In the worst-case scenario
-(many small KV pairs), the total number of pages required to perform
-the all-to-all communication, including the source and target KV pages
-is 7.
+Performing an our-of-core {\it collate()} operation is more complex.
+Recall that the operation occurs in two stages.  First, keys are
+hashed to ``owning'' processors and the KV datums are communicated to
+new processors in an all-to-all fashion.  In out-of-core mode, this is
+done one page of KV data at a time.  Each processor reads in a page of
+KV pairs, the communication pattern for the datums in that page is
+determined (who needs what datums), and all-to-all communication is
+performed.  Each processor allocates a two-page chunk of memory to
+receive incoming KV pairs.  On average each processor should receive
+one page of KV pairs; the two-page allocation allows for some
+load-imbalance.  If the KV pairs are distributed unevenly so that this
+limit is exceeded, then the all-to-all communication is performed in
+smaller, multiple passes until the full page contributed by every
+processor has been communicated.  Performing an MPI\_Alltoall(), or
+using the custom all-to-all routines provided by the MR-MPI library,
+requires allocation of auxiliary arrays that store processor indices,
+datum pointers, and datum lengths.  For the case of many tiny KV
+datums, the total number of memory pages required to perform the
+all-to-all communication, including the source and target KV pages is
+at most 7.
 
@@
 out-of-core KV file, that need to be reorganized into a set of KMV
-pages, stored in new out-of-core KMV file.  In principle, to create
-one page of KMV pairs, requires all KV pages to be scanned to find all
-the keys which contribute values to that page.  Doing this for each
-output page of KMV pairs could be prohibitively expensive.  A related
-issue is that a hash table is needed to match new keys with previously
-encountered keys, as described in the previous section.  But there is
-no guarantee that the hash table itself will not grow arbitrarily
-large for some data sets with many unique keys.  We need an algorithm
-for generating KMV pairs which operates within a small number of
-memory pages and performs a minimal number of passes through the KV
-disk file.  An algorithm that meets these goals is diagrammed in
-Figure \ref{f:collate}; it requires 4 read passes of the KV pages and
-3 write passes.  It also uses a finite-size in-memory hash table,
-which stores unique keys as they are encountered in the loops over KV,
-as well as auxiliary information needed to construct the output pages
-of KMV pairs.
+pages, also stored in new out-of-core KMV file.  In principle, to
+create one page of KMV pairs, would require all KV pages be scanned,
+to find all the keys which contribute values to that page.  Doing this
+for each output page of KMV pairs could be prohibitively expensive.  A
+related issue is that, as described in the previous section, a hash
+table is needed to match new keys with previously encountered keys.
+But there is no guarantee that the hash table itself will not grow
+arbitrarily large for data sets with many unique keys.  We need an
+algorithm for generating KMV pairs which operates within a small
+number of memory pages and performs a minimal number of passes through
+the KV disk file.  An algorithm that meets these goals is diagrammed
+in Figure \ref{fig:collate}; it reads the KV pages at most 4 times, and
+writes out new KV or KMV pages at most 3 times.  It also uses a
+finite-size in-memory hash table, which stores unique keys as they are
+encountered while looping over the KV pages, as well as auxiliary
+information needed to construct the output pages of KMV pairs.
 
 \begin{figure}
-\includegraphics[width=\textwidth,angle=-90]{fig_collate.pdf}
+\includegraphics[width=\textwidth,angle=-90]{fig_collate2.pdf}
 \caption{Multi-pass algorithm for converting KV data to KMV data.  The
@@
 pairs; HT is an in-memory hash table; KMV is key/multivalue pairs.}
-\label{f:collate}
+\label{fig:collate}
 \end{figure}
@@
 
-\item (Pass 1) The KV pairs are read and split into ``partition''
-files.  Each partition file contains KV pairs whose unique keys will
-(likely) fit in the hash table (HT).  Intially all KV pairs are
-assigned to the first partition file as the HT is populated.  When the
-HT becomes full, e.g. at the point represented by the horizontal line
-in a downward scan of the leftmost vertical line, then the fraction of
-KV pairs read thus far is used to estimate the number of additional
-partition files needed.  As subsequent KV pairs are read in, they are
-written into either the original partition file or one of the new
-ones.  The former if its key exists in the current HT; the latter if
-not, and a subset of bits in the hash value of its key is used to
-assign it to one of the partition files.  The number of needed
-partition files is estimated conservatively, rounding up to the next
-power-of-two, so that extra passes through the KV pairs are almost
-never needed, and so the bit masking can be done quickly.  This pass
-requires both a read and write of all the KV pairs.
+\item (Pass 1) The KV pairs are read, one page at a time, and split
+into ``partition'' files, represented by KVp in the figure.  Each
+partition file contains KV pairs whose unique keys will (likely) fit
+in the hash table (HT).  Initially, all KV pairs are assigned to the
+first partition file as the HT is populated.  When the HT becomes
+full, e.g. at the point represented by the horizontal line shown on
+the leftmost vertical line as a downward scan is performed, then the
+fraction of KV pairs read thus far is used to estimate the number of
+additional partition files needed.  As subsequent KV pairs are read
+in, they are assigned to the original partition file if the KV pair's
+key is already in the current HT.  If not, then a subset of bits in
+the hash value of the key is used to assign the KV pair to one of the
+new partition files.  The number of needed partition files is
+estimated conservatively, rounding up to the next power-of-two, so
+that extra passes through the KV pairs are almost never needed, and so
+that the bit masking can be done quickly.  This first pass entails
+both a read and write of all the KV pairs.
 
-\item (Pass 2) The KV pairs in a partition file are read, and their
-unique keys are hashed into the HT, accumulating data about the number
-and size of values associated with each key.  Before the next
-partition file is read, passes 3 and 4 are completed for this
-partition.  But eventually, this pass reads all partition files,
-requiring a read of all KV pairs.
+\item (Pass 2) The partition files are read, one at a time.  The key
+for each KV pair is hashed into the HT, accumulating data about the
+number and size of values associated with each unique key.  Before the
+next partition file is read, passes 3 and 4 are completed for the
+current partition.  Eventually this pass entails a read of all KV
+pairs.
 
-\item (Pass 3) The unique keys in the HT are scanned.  The associated
-values may create KMV pairs which span many pages.  If this is the
-case, the associated partition file of KV pairs is read, and assigned
-to smaller ``set'' files, using the accumulated HT statistics.  Each
-set file corresponds to the KV pairs which will contribute to the set
-of KMV pairs that will populate one page of KMV output.  If a KMV pair
-spans multiple pages, because it has a very large number of values,
-then the corresponding KV pairs will still be part of a single set
-file.  Eventually, this pass reads all partition files, requiring both
+\item (Pass 3) The unique keys in the HT for one partition are
+scanned.  The associated values may create KMV pairs which will span
+$M$ memory pages.  If $M > 1$ the associated partition file of KV
+pairs is read again, and each KV pair is assigned to one of $M$
+smaller ``set'' files, represented by KVs in the figure.  Each set
+file contains the KV pairs which will contribute to the KMV pairs that
+will populate one page of KMV output.  If a single KMV pair spans
+multiple pages, because it contains a very large number of values,
+then the corresponding KV pairs will still be written to a single set
+file.  Eventually, this pass reads all partition files, entailing both
 a read and write of all KV pairs.
@@
 \item {Pass 4) A set file is read and the key/value data for each of
-its KV pairs is copied into the appropriate place in the page of KMV
-pairs, using information stored in the HT.  When complete, the KMV
-page is written to disk.  This pass thus eventually reads all set
-files, requiring a read of all KV pairs.  It also writes all KMV
-pairs.  If each KV pair has a unique key, the size of KMV output is
-roughly the same as that of the KV input.
+its KV pairs is copied into the appropriate location in the page of
+KMV pairs, using information stored in the HT.  When complete, the KMV
+page is written to disk.  This pass eventually reads all set files,
+entailing a final read of all KV pairs.  It also writes all KMV pairs
+to disk.  If each KV pair has a unique key, the volume of KMV output
+is roughly the same as that of the KV input.
 
@@
 
-In summary, this data reorganization for the {collate()} operation is
-somewhat complex, but requires a small, constant number of passes
-through the data.  Namely, at most 4 read passes through the pages of
-KV pairs, and 3 write passes.  The first two write passes reorganize
-KV pairs into partition and set files; the final write pass creates
-the KMV data set.  By contrast, a full out-of-core sort() of the KV
-pairs, performed via a merge sort as discussed below, requires
-$O(\log{M})$ read/write passes through the KV data, where $M$ is the
-number of pages of KV pairs, which can be a large number for big data
-sets.
+In summary, this data reorganization for the {\it collate()} operation
+is somewhat complex, but requires only a small, constant number of
+passes through the data.  The KV datums are read from disk at most 4
+times, and written to disk 3 times.  The first two write passes
+reorganize KV pairs into partition and set files; the final write pass
+creates the KMV data set.  Depending on the size and characteristics
+of the KV datums (e.g. how many unique keys it contains), some of
+these passes may not be needed.  By contrast, a full out-of-core sort
+of the KV pairs, performed via a merge sort as discussed below,
+requires $O(\log{M})$ read/write passes through the KV data, where $M$
+is the number of pages of KV pairs, which can be a large number for
+big data sets.
 
-Intermediate passes done via spool files.  Can't bound precisely,
-since depends on data characteristics (see tech discussion in manual),
-but generally fit wihtin 7 pages of aggregate.
+The memory page requirements for the above algorithm are as follows.
+Two contiguous pages of memory are used for the hash table.
+Intermediate passes 2 and 3 can require numerous partition or set
+files to be opened simultaneously and written to.  To avoid small
+writes of individual KV datums, a buffer of at least 16K bytes is
+allocated for each file.  The precise number of simultaneously open
+files is difficult to bound, but typically one or two additional pages
+memory suffice for all needed buffers.  Thus the total pages required
+for the data reorganization stage of the {\it collate()} operation is
+less than the 7 used by the all-to-all communication stage.
 
-Note that depending on the data set size and its key distribution all
-four passes through the data may not be necessary, but the upper limit
-is four, assuming the chunk-size estimations are accurate.  By
-comparison, a merge sort of KV pairs would require $log_2(N)$
-passes (read/write) through the out-of-core data set, where $N$ is its
-number of pages.  The MR-MPI {\it collate()} operation does not
-actually sort the KMV pairs by key; the final ordering is
-indeterminate as it is created by the hash function.
-
-Avoid extra passes if not needed.  E.g. one partition file, then never
-even written.  First partition file not re-read to popluate HT.
-
-Uneven distribution of unique keys in KV pair could induce extra
-passes due to not enough partitions, but quite rare.
-
-The other various MR-MPI library calls ({\it compress()}, {\it
-gather(), etc}, can also operate in out-of-core mode, typically with
-one pass through the KV or KMV data.  Sorting, by
-key or value, is an exception.  as an out-of-core merge sort requires
-$log_2(N)$ passes, as just mentioned, where $N$ is the number of
-pages.
+Other MR-MPI library calls discussed in the previous section, such as
+{\it compress()}, {\it convert()}, {\it collapse()}, {\it gather()},
+etc, can also operate in out-of-core mode, typically with one pass
+through the KV or KMV data and the use of one or two memory pages.
+Sorting KV pairs, by key or value, is an exception.  An out-of-core
+merge sort is performed as follows.  Two pages of KV datums are read
+from disk.  Each is sorted using the C-library quicksort() function,
+as discussed in the previous section.  The two pages are then scanned
+and merged into a new file which is written to disk as its associated
+memory page fills up.  This process requires 5 memory pages, which
+includes vectors of KV datum pointers and lengths needed by the
+in-memory quicksort() operation.  Once all pairs of pages have been
+merged, the sort continues in a recursive fashion, merging pairs of
+files into a new third file, without the need to perform additional
+in-memory sorts.  Thus the overall memory requirement is 5 pages.  The
+number of read/write passes through the KV data set for the merge sort
+is $O(\log{M})$, where $M$ is the number of pages of KV pairs.

### 2010-03-29 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/graph.tex
@@
 \section{Graph Algoriths in MapReduce}
+\label{sec:graph}
 
-  RMAT generation, pagerank, CC, triangle-finding, Luby, SSSP, other?
-  how many variants of something like CC do we want to include?
-    I would vote for just one
-  discuss possible inefficiency due to a giant CC in CC algorithm
+We begin with a MapReduce procedure for creating large, sparse,
+randomized graphs, since they are the input for the algorithms
+discussed below.  R-MAT graphs ~\cite{RMAT} are recursively generated
+graphs with power-law degree distributions.  They are commonly used to
+represent web and social networks.  The user specifices 6 parameters
+which define the graph: the number of vertices $N$ and edges $M$, and
+4 parameters $a$, $b$, $c$, $d$ which sum to 1.0 and are discussed
+below.  The algorithm in Figure \ref{fig:rmat} generates $N_z = MN$
+unique non-zero entries in a sparse $NxN$ matrix $A$, where each entry
+$A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for R-MAT graph generation.}
+
+ \label{fig:rmat}
+\end{figure}
+
+In the map() operation, each of $P$ processors is tasked with
+generating a $1/P$ fraction of the desired edges.  A single random
+$i,j$ edge is computed recursively as follows.  Pick a random quadrant
+of the $A$ matrix with relative probabilities $a$, $b$, $c$, and $d$.
+Treat the chosen quadrant as a sub-matrix and select a random quadrant
+within it, in the same manner.  Repeat this process $n$ times where $N
+= 2^n$.  At the end of the recursion, the final ``quadrant'' is
+non-zero matrix element $A_{ij}$.
+
+The map() will often generate some small number of duplicate edges.
+The collate() and reduce() operations remove the duplicates.  The
+entire map-collate-reduce sequence is repeated until the number of
+resulting key/value pairs $N_{kv}$ equals $N_z$.  For reasonably
+sparse graphs this typically takes only a few iterations.
+
+Note that the degree distribution of vertices in the graph depends on
+the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
+values is larger than the other 3, a highly skewed distribution
+results.  Variants of the above algorithm can be used when $N$ is not
+a power-of-two, to generate graphs with weighted edges (assign a
+random numeric value to the $A_{ij}$ edge), graphs without self edges
+(require $i != j$), or graphs with undirected edges (require $i < $j).
+
+Discussion of PageRank and its matvec kernel.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for PageRank vertex ranking.}
+
+ \label{fig:rmat}
+\end{figure}
+
+Discussion of triangle finding algorithm.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for triangle finding.}
+
+ \label{fig:rmat}
+\end{figure}
+
+Discuss connected component identification algorithm.
+
+Do we need differnt variants of CC?
+
+Could just do one, discuss possible inefficiency due to giant CC.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for connected component identification.}
+
+ \label{fig:rmat}
+\end{figure}
+
+Discussion of Luby algorithm for maximally independent sets.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for R-MAT graog generation.}
+
+ \label{fig:rmat}
+\end{figure}
+
+Discussion of SSSP.
+
+Discuss new optimization in SSSP?
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$N_{\rm remain} = N_z$ \\
+while $N_{\rm remain} > 0$: \\
+\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate() \\
+\> Reduce(): \> {\bf Remove} duplicate edges \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $N_{\rm remain} = N_z - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for single-source shortest path (SSSP)
+ determination.}
+
+ \label{fig:rmat}
+\end{figure}

### 2010-03-29 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-\item {Pass 4) A set file is read and the key/value data for each of
+\item (Pass 4) A set file is read and the key/value data for each of
 its KV pairs is copied into the appropriate location in the page of

### 2010-03-29 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \centerline{Keywords: MapReduce, message-passing, MPI, graph
-algorithms, RMAT matrices}
+algorithms, R-MAT matrices}
 
@@
 shortest-path calculation.  To test the algorithms on arbitrarily
-large artificial graphs we generate randomized RMAT matrices in
+large artificial graphs we generate randomized R-MAT matrices in
 parallel; a MapReduce version of this operation is also described.

### 2010-03-29 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{Performance Results}
+\label{sec:results}
 
-  ideally, for each problem:
-    define chosen characteristics of graph, e.g. edges/vertex
-    choose 3 sizes:
-      small = fits on 1 proc)
-      medium = out-of-core on 1 proc, but not on P procs
-      huge = impressive size
-  show performance plots for each size running on 1 to P procs
-    would like to see kink in plots as go out-of-core
-    be able to judge and explain basic scaling results
-  what machine: Odin or Tbird or RStorm?
-    something that maps to a vanilla cluster
-  where we have data:
-    show multiple machines
-    alternate non-MR algorithms
-      discuss alternate CC algorithms via Trilinos and hybrid
-    alternate MR implementations (Hadoop)
+In this section, we present performance results for the MapReduce
+graph algorithms of the preceeding section, implemented as small C++
+programs calling our MR-MPI library.
+
+The benchmarks were run on a medium-sized Linux cluster.  Describe
+Odin.
+
+We ran each of the algorithms on 3 artificial R-MAT graphs of
+different sizes, each on a varying number of processors.  The {\it
+small} problem size (around 1M edges) was a problem that could be run
+on a single processor.  The {\it medium} problem size (around 1B
+edges) could be run on a few processors.  The {\it large} problem size
+(around 1T edges) required most of the machine to run.
+
+The resulting timings give a sense of the inherent scalability of the
+MapReduce algorithms as graph size grows on a fixed number of
+processors, and of the parallel scalability for computing on a graph
+of fixed size on a growing number of processors.
+
+Discuss R-MAT generation times.
+
+Discuss PageRank performance.  Include Trilinos and other
+algorithms and machines?
+
+Discuss triangle finding times.
+
+Discuss connected component identification times.
+
+Discuss Luby maximally independent sets times.
+
+Discuss SSSP times.

### 2010-03-29 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@486 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-Discussion of PageRank and its matvec kernel.
+The PageRank algorithm assings a relative numeric rank to each
+vertex in a graph.
+
+It models the web as a directed graph $G(V,E)$, with each vertex $v
+\in V$ representing a web page and each edge $e_{ij} \in E$
+representing a hyperlink from $v_i$ to $v_j$.  The probability of
+moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
+(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
+0.8-0.9), $d_{out}(v)$ is the outdegree of vertex $v$, and $|V|$ is
+the cardinality of $V$.  The first term represents the probability of
+following a given link on page $v_i$; the second represents the
+probability of moving to a random page.  For pages with no outlinks,
+the first term is $\alpha/|V|$, indicating equal likelihood to move to
+any other page.  Equivalently, the graph can be represented by a
+matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
+\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
+algorithm, then, is simply a power-method iteration in which the
+dominating computation is matrix-vector multiplication $A^T x=y$,
+where $x$ is the PageRank vector from the previous iteration.
+
+The algorithm in Figure \ref{fig:pr} performs these
+iterations.
 
@@
 
- \label{fig:rmat}
+ \label{fig:pr}
 \end{figure}
 
-Discussion of triangle finding algorithm.
+The MapReduce implementation performs two {\it map()} operations to
+initialize the graph matrix $A$ and PageRank vector $x$.  A {\it
+collate()} operation gathers all row entries $a_{ij}$ with their
+associated $x_i$ entry, and a {\it reduce()} computes $a_{ij} x_i$.  A
+second {\it collate()} gathers, for each $j$, all contributions to the
+column sum $\sum a_{ij} x_i$, which is computed by a second {\it
+reduce()}.  MPI\_Allreduce calls are used to compute global norms and
+residuals.
+
+A triangle in a graph is a clique of size three.  The algorithm in
+Figure \ref{fig:tri} enumerates all such triplets of vertices (Vi, Vj,
+Vk) where the three edges (Vi,Vj), (Vj,Vk), (Vi,Vk) exist in the
+graph.  The input to the first map() is an undirected graph, a lower
+triangular matrix from R-MAT generation, so that $V_i$ < $V_j$ for
+every edge.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxxxxxx\=xxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+Copy(): store $G_0$ = copy of KV pairs for input graph \\
+Map(): \> {\bf Vertices} Convert edges to vertices \\
+           \> \> input: Key = $(V_i,V_j)$, Value = NULL \\
+           \> \> output: Key = $V_i$, Value = $V_j$ \\
+           \> \> output: Key = $V_j$, Value = $V_i$ \\
+Collate() \\
+Reduce(): \> {\bf First Degree} add degree to one vertex in edge \\
+              \> \> input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> for each value in MultiValue: \\
+              \> \> output: Key = $(V_i,V_j)$, Value = $(D_i,0)$ if $Vi < Vj$ \\
+              \> \> output: Key = $(V_j,V_i)$, Value = $(0,D_i)$ if $Vj < Vi$ \\
+Collate() \\
+Reduce(): \> {\bf Second Degree} add degree to second vertex in edge \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
+              \> \> output: Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with Vi < Vj \\
+Map(): \> {\bf Low Degree} vertex emits edge \\
+           \> \> if $Di < Dj$, output: Key = $V_i$, Value = $V_j$ \\
+           \> \> else if $Dj < Di$, output: Key = $V_j$, Value = $V_i$ \\
+           \> \> else output: Key = $V_j$, Value = $V_i$ \\
+Collate() \\
+Reduce(): \> {\bf Angles} find angles for each central vertex $Vi$ \\
+              \> \> input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> Loop $J$ over values in MultiValue: \\
+	        \> \> Loop $K > J$ over values in MultiValue: \\
+		  \> \> if $Vj < Vk$, output: Key = $(V_j,V_k)$, Value = $Vi$ \\
+		  \> \> else output: Key = $(V_k,V_j)$, Value = Vi \\
+Merge() \\ {\bf Add} KV pairs from original graph $G_0$ to Angles reduce() KV pairs \\
+Collate() \\
+Reduce(): \> {\bf Emit Triangles} \\
+              \> \> input: Key = $(V_i,V_j)$, MultiValue = $(Vk, Vl, Vm, NULL, ...)$ \\
+              \> \> if NULL exists in MultiValue, \\
+	        \> \> output: Key = $(V_i,V_j,V_k)$, Value = NULL for each non-NULL value in MultiValue
 
@@
 
- \label{fig:rmat}
+ \label{fig:tri}
 \end{figure}
+
+Further interations could find rectangles from open triangles, etc.
 

### 2010-03-30 src/mapreduce.cpp
Minor change to remove compiler warnings.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@487 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -2342,7 +2342,7 @@ uint64_t MapReduce::roundup(uint64_t n, int nalign)
    write a histogram of value to screen with title
 ------------------------------------------------------------------------- */
 
-void MapReduce::write_histo(double value, char *title)
+void MapReduce::write_histo(double value, const char *title)
 {
   int histo[10],histotmp[10];
   double ave,max,min;

### 2010-03-30 src/mapreduce.h

@@ -211,7 +211,7 @@ class MapReduce {
   void file_stats(int);
   uint64_t roundup(uint64_t, int);
   void start_timer();
-  void write_histo(double, char *);
+  void write_histo(double, const char *);
   void histogram(int, double *, double &, double &, double &,
 		 int, int *, int *);
 

### 2010-03-30 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@488 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2010-03-29 paper/graph.tex
@@
 while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+\> 1 Map: \> Generate $N_{\rm remain}/P$ random edges on each processor \\
+           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
+\> 1 Collate \\
+\> 1 Reduce: \> Remove duplicate edges \\
+              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
 \> $N_{\rm remain} = N_z - N_{kv}$
@@
 \> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
 \> Collate() \\
 \> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
+              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
 \> $N_{\rm remain} = N_z - N_{kv}$
@@
 
- \label{fig:rmat}
+ \label{fig:pr}
 \end{figure}
 
-Discussion of triangle finding algorithm.
+The MapReduce implementation performs two {\it map()} operations to
+initialize the graph matrix $A$ and PageRank vector $x$.  A {\it
+collate()} operation gathers all row entries $a_{ij}$ with their
+associated $x_i$ entry, and a {\it reduce()} computes $a_{ij} x_i$.  A
+second {\it collate()} gathers, for each $j$, all contributions to the
+column sum $\sum a_{ij} x_i$, which is computed by a second {\it
+reduce()}.  MPI\_Allreduce calls are used to compute global norms and
+residuals.
+
+A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
+the three edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
+\ref{fig:tri} outlines a MapReduce algorithm that enumerates all such
+triplets, assuming an input graph of undirected edge keys $(V_i,V_j)$
+where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
+generated matrix.  This exposition follows the triangle-finding
+algorithm presented in \cite{Cohen}.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxxxxxxxxxx\=xxx\=xxx\=xxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
+1 Map: \> Convert edges to vertices \\
+           \> \> input Key = $(V_i,V_j)$, Value = NULL \\
+           \> \> output Key = $V_i$, Value = $V_j$ \\
+           \> \> output Key = $V_j$, Value = $V_i$ \\
+1 Collate \\
+1 Reduce: \> Add first degree to one vertex in edge \\
+              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> for each V in MultiValue: \\
+              \> \> \> if $V_i < V$: output Key = $(V_i,V)$, Value = $(D_i,0)$ \\
+              \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
+2 Collate \\
+2 Reduce: \> Add second degree to other vertex in edge \\
+              \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
+              \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
+3 Map: \> Low degree vertex emits edges \\
+           \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
+           \> \> else if $D_j < D_i$: output Key = $V_j$, Value = $V_i$ \\
+           \> \> else: output Key = $V_i$, Value = $V_j$ \\
+3 Collate \\
+3 Reduce: \> Emit angles of each vertex \\
+              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> for each $V_1$ in MultiValue: \\
+	      \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
+	      \> \> \> \> if $V_1 < V_2$: output Key = $(V_1,V_2)$, Value = $V_i$ \\
+	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
+4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs \\
+4 Collate \\
+4 Reduce: \> Emit triangles \\
+              \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
+              \> \> if NULL exists in MultiValue: \\
+	      \> \> \> for each non-NULL V in MultiValue: \\
+	      \> \> \> \> output Key = $(V_i,V_j,V)$, Value = NULL
 
@@
 
- \caption{MapReduce algorithm for triangle finding.}
+ \caption{MapReduce algorithm for triangle enumeration.}
 
- \label{fig:rmat}
+ \label{fig:tri}
 \end{figure}
+
+The initial step is to store a copy of the graph edges as key/value
+(KV) pairs in a auxiliary MapReduce object $G_0$, for use later in the
+algorithm.  The first {\it map()} operation converts edge keys to
+vertex keys with edge values.  After the {\it collate()}, each vertex
+has a list of vertices it is connected to; the first {\it reduce()}
+can thus flag one vertex $V_i$ in each edge with a degree count $D_i$.
+The second {\it collate()} and {\it reduce()} assign a degree count
+$D_j$ to the other vertex in each edge.  In the third {\it map()},
+only the lower-degree vertex in each edge emits its edges as key/value
+(KV) pairs.  The task of the third {\it reduce()} is to emit
+``angles'' for each of these low-degree vertices.  An ``angle'' is a
+root vertex $V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a
+triangle without the third edge $(V_1,V_2)$.  The {\it reduce()} emits
+a list of all angles of vertex $V_i$, by a double loop over the edges
+of $V_i$.  Note that the aggregate volume of KV pairs emitted at this
+stage is minimized by having only the low-degree vertex in each edge
+generate angles.
+
+In stage 4, the KV pairs in the original graph $G_0$ are added to the
+current working set of KV pairs.  The KV pairs in $G_0$ are edges that
+complete triangles for the angle KV pairs just generated.  After the
+fourth {\it collate()}, a pair of vertices $(V_i,V_j)$ is the key, and
+the multivalue is the list of all root vertices in angles that contain
+$V_i$ and $V_j$.  If the multivalue also contains a NULL, contributed
+by $G_0$, then there is a $(V_i,V_j)$ edge in the graph.  Thus all
+vertices in the multivalue are roots of angles which are complete
+triangles and can be emitted as a triplet key.
 

### 2010-04-01 programs/zerokv.cpp
Test program to exhibit a bug for Steve.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@496 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -1,0 +1,60 @@
+// Program to test case where a processor has zero KVs to convert to a KMV.
+// Karen Devine, April 2010
+//
+// Generate ten KVs on each processor, all with the same key so they all map to
+// one processor during aggregate().  Then call aggregate() and convert() 
+// and see what happens.
+//
+// To build:  mpic++ -m64 -I../src zerokv.cpp -L../src -lmrmpi
+// To run:    mpirun -np 2 a.out
+
+#include <mpi.h>
+#include <iostream>
+#include <string.h>
+#include "mapreduce.h"
+#include "keyvalue.h"
+
+using namespace MAPREDUCE_NS;
+
+//////////////////////////////////////////////////////////////////////////////
+// Generate KVs that all have the same key.
+void all_to_one(int itask, KeyValue *kv, void *ptr)
+{
+  const int cnt = 10;
+  const char *key = "kdd";
+  int len = strlen(key) * sizeof(char);
+  for (int i = 0; i < cnt; i++)
+    kv->add((char *) key, len, NULL, 0);
+}
+
+//////////////////////////////////////////////////////////////////////////////
+// Print the number of occurrences of key.
+void count(char *key, int keysize, char *multivalue, int nvalues, 
+           int *valuebytes, KeyValue *kv, void *ptr)
+{
+  std::cout << " AFTER CONVERT " << key << " has " << nvalues 
+            << " occurrences" << std::endl;
+}
+
+//////////////////////////////////////////////////////////////////////////////
+int main(int narg, char *arg[])
+{
+  MPI_Init(&narg, &arg);
+  int me, np;
+  MPI_Comm_rank(MPI_COMM_WORLD, &me);
+  MPI_Comm_size(MPI_COMM_WORLD, &np);
+
+  MapReduce *mr = new MapReduce;
+  mr->map(np, &all_to_one, NULL);
+  std::cout << me << " AFTER MAP, MY nkv = " << mr->kv->nkv << std::endl;
+
+  mr->aggregate(NULL);
+  std::cout << me << " AFTER AGGREGATE, MY nkv = " << mr->kv->nkv << std::endl;
+
+  mr->convert();
+
+  mr->reduce(&count, NULL);
+
+  delete mr;
+  MPI_Finalize();
+}

### 2010-04-01 programs/rmat.hpp
RMAT-generate as a function call now appears to work (at least for small cases that don't ever have zero KVs going into convert -- that bug is reported to Steve and exercised with zerokv.cpp).
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@497 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -12,13 +12,17 @@
 ------------------------------------------------------------------------- */
 
 // MapReduce random RMAT matrix generation example in C++
-// Syntax: rmat N Nz a b c d frac seed {outfile}
+// Parameters:  N Nz a b c d frac seed {outfile} printstats
 //   2^N = # of rows in RMAT matrix
 //   Nz = non-zeroes per row
 //   a,b,c,d = RMAT params (must sum to 1.0)
 //   frac = RMAT randomization param (frac < 1, 0 = no randomization)
 //   seed = RNG seed (positive int)
 //   outfile = output RMAT matrix to this filename (optional)
+//
+// Resulting vertices are numbered 1, 2, ... 2^N, to be consistent with
+// Matrix-Market (which is one-based) and 
+// Karl's data (where 0 is a non-valid vertex ID).
 
 #include "mpi.h"
 #include "math.h"
@@ -38,7 +42,8 @@ void generate_edge(int, KeyValue *, void *);
 void final_edge(char *, int, char *, int, int *, KeyValue *, void *);
 void cull(char *, int, char *, int, int *, KeyValue *, void *);
 void output(char *, int, char *, int, int *, KeyValue *, void *);
-void nonzero(char *, int, char *, int, int *, KeyValue *, void *);
+void nonzero_in_row(char *, int, char *, int, int *, KeyValue *, void *);
+void nonzero_in_col(char *, int, char *, int, int *, KeyValue *, void *);
 void degree(char *, int, char *, int, int *, KeyValue *, void *);
 void histo(char *, int, char *, int, int *, KeyValue *, void *);
 int ncompare(char *, int, char *, int);
@@ -61,7 +66,7 @@ public:
   FILE *fp;
   int me;
   int nprocs;
-  bool printstats;
+  int printstats;
 
   GenerateRMAT(int, char**);
   ~GenerateRMAT() {delete [] outfile;};
@@ -71,13 +76,12 @@ public:
 ////////////////////////////////////////////////////////////////////////////
 GenerateRMAT::GenerateRMAT(int narg, char **args)
 {
-  int me;
   MPI_Comm_rank(MPI_COMM_WORLD,&me);
   MPI_Comm_size(MPI_COMM_WORLD,&nprocs);
 
   // parse command-line args
 
-  if (me == 0) printf("Syntax for rmat: -rn N -rz z -ra a -rb b -rc c -rd d -rf frac -rs seed {-ro outfile} -rp\n");
[…]

### 2010-04-01 programs/sssp.cpp

@@ -262,8 +262,8 @@ void output_distances(char *key, int keybytes, char *multivalue,
 //  else
 //    *fp << *((VERTEX *)key) << "   " << -1. << endl;
 
-//  *fp << *((VERTEX *)key) << "   " << *e << endl;
-  *fp << *((VERTEX *)key) << "   " << e->wt << endl;
+  *fp << *((VERTEX *)key) << "   " << *e << endl;
+//  *fp << *((VERTEX *)key) << "   " << e->wt << endl;
 
 //  if (keybytes == 16)
 //    fprintf(fp, "%lld %lld    %lld %lld  %ld\n",
@@ -448,8 +448,12 @@ bool SSSP<VERTEX, EDGE>::get_next_source(
       }
     }
     else if (filetype == RMAT) {
+      static uint64_t vv = 1;
+      if (vv <= nverts) source->v[0] = vv;
+      else source->reset();
+      vv++;
       // Randomly select a vertex in range [0..nverts-1].
-      source->v[0] = ((uint64_t) (drand48() * nverts)) + 1;
+      // source->v[0] = ((uint64_t) (drand48() * nverts)) + 1;
     }
     else {
       static bool firsttime = true;

### 2010-04-01 programs/sssp_v2.cpp
Made sssp_v2 work with new MR-MPI library. Need to test on odin.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@498 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -20,11 +20,23 @@
 #include "read_fb_data.hpp"
 #include "read_mm_data.hpp"
 #include "shared.hpp"
+#include "rmat.hpp"
 
 using namespace std;
 using namespace MAPREDUCE_NS;
 
 #define MAX_NUM_EXPERIMENTS 50
+
+#define FBFILE 0
+#define MMFILE 1
+#define RMAT 2
+
+uint64_t NVtxLabeled = 0;  // Global variable (yuk!) counting number of 
+                           // vertices labeled on this processor.
+                           // Used as a diagnostic.
+uint64_t GNVtxLabeled;     // Global variable (yuk!) counting number of
+                           // vertices labeled across all processors.
+                           // Used as a diagnostic.
 
 /////////////////////////////////////////////////////////////////////////////
 // Class used to pass distance information through the MapReduce system.
@@ -64,14 +76,14 @@ void add_source(int nmap, KeyValue *kv, void *ptr)
   d.current = false;
   kv->add((char *) v, sizeof(VERTEX),
           (char *) &d, sizeof(DISTANCE<VERTEX, EDGE>));
-  cout << "KDDKDD add_source " << *v << endl;
+//  cout << "KDDKDD add_source " << *v << endl;
 }
 
 /////////////////////////////////////////////////////////////////////////////
 // move_to_new_mr:  Move KV from existing MR to new MR provided in ptr.
 // Map:    Input:   KVs in exisitng MR object, new MR object in ptr.
 //         Output:  No KVs in existing MR object; they are all added to new MR.
-void move_to_new_mr(int itask, char *key, int keybytes, 
+void move_to_new_mr(uint64_t itask, char *key, int keybytes, 
                     char *value, int valuebytes, 
                     KeyValue *kv, void *ptr)
 {
@@ -85,7 +97,7 @@ void move_to_new_mr(int itask, char *key, int keybytes,
 // Map:    Input:   KV  key = Vtx ID; value = NULL
 //         Output:  KV  key = Vtx ID; value = initial distance
 template <typename VERTEX, typename EDGE>
-void initialize_vertex_distances(int itask, char *key, int keybytes, 
+void initialize_vertex_distances(uint64_t itask, char *key, int keybytes, 
[…]

### 2010-04-01 src/irregular.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@500 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -205,7 +205,7 @@ int Irregular::setup(int n, int *proclist, int *sizes, int *reorder,
    sizes = byte count of each datum
    reorder (for custom) = contiguous send indices for each send
    copy = buffer to pack send datums into
-   recv = buffer to recv all aatums into
+   recv = buffer to recv all datums into
 ------------------------------------------------------------------------- */
 
 void Irregular::exchange(int n, int *proclist, char **ptrs, int *sizes, 
@@ -234,6 +234,8 @@ void Irregular::exchange_all2all(int n, int *proclist, char **ptrs,
     memcpy(cptrs[iproc],ptrs[i],sizes[i]);
     cptrs[iproc] += sizes[i];
   }
+
+  delete [] cptrs;
 
   MPI_Alltoallv(copy,sendbytes,sdispls,MPI_BYTE,
 		recv,recvbytes,rdispls,MPI_BYTE,comm);

### 2010-04-01 src/keymultivalue.cpp

@@ -765,6 +765,11 @@ int KeyMultiValue::unique2kmv_all()
   ptr = page;
   newflag = 1;
 
+  if (nunique == 0) {
+    sets[nset-1].nunique = 0;
+    sets[nset-1].extended = 0;
+  }
+    
   for (int i = 0; i < nunique; i++) {
     if (newflag) {
       newflag = 0;

### 2010-04-01 programs/test_pagerank.cpp
Reducing default memsizes.  Having memsize too big is deadly.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@501 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -387,9 +387,6 @@ int main(int narg, char **args)
   double alpha = 0.8;
   double tolerance = 0.00001;
   int NumberOfPageranks = 1;
-#ifdef NEW_OUT_OF_CORE
-  int memsize = 1024;
-#endif
 
   // Parse the command line.
   int ch;

### 2010-04-14 programs/mrmatrix2.cpp
Compiles now.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@541 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -23,7 +23,7 @@ using namespace std;
 // initialize_matrix map function
 // Given MapReduce object containing edges, create matrix non-zeros.
 template <typename IDTYPE>
-static void initialize_matrix(uint64_t itask, char *key, int keybytes,
+static void mrm_initialize_matrix(uint64_t itask, char *key, int keybytes,
                               char *value, int valuebytes, 
                               KeyValue *kv, void *ptr)
 {
@@ -35,7 +35,7 @@ static void initialize_matrix(uint64_t itask, char *key, int keybytes,
 
 //--------------------------------------------------------------
 template <typename IDTYPE>
-static void initialize_transpose_matrix(uint64_t itask, char *key, int keybytes,
+static void mrm_initialize_transpose_matrix(uint64_t itask, char *key, int keybytes,
                                         char *value, int valuebytes,
                                         KeyValue *kv, void *ptr)
 {
@@ -46,7 +46,7 @@ static void initialize_transpose_matrix(uint64_t itask, char *key, int keybytes,
 }
 
 //--------------------------------------------------------------
-static void save_empty_rows(char *key, int keybytes, 
+static void mrm_save_empty_rows(char *key, int keybytes, 
                             char *multivalue, int nvalues, int *valuebytes,
                             KeyValue *kv, void *ptr)
 {
@@ -79,24 +79,24 @@ MRMatrix<IDTYPE>::MRMatrix(
   transposeFlag = transpose;
   if (transpose) {
     N = m; M = n;
-    mr->map(mredge, &initialize_transpose_matrix<IDTYPE>, NULL);
+    mr->map(mredge, mrm_initialize_transpose_matrix<IDTYPE>, NULL);
   }
   else {
     N = n; M = m;
-    mr->map(mredge, &initialize_matrix<IDTYPE>, NULL);
+    mr->map(mredge, mrm_initialize_matrix<IDTYPE>, NULL);
   }
 
   // Identify empty rows of matrix A (leaf nodes in graph).
   emptyRows = new MRVector<IDTYPE>(N, pagesize, fpath);
   MapReduce *emr = emptyRows->mr;
   emr->add(mredge);
-  nEmptyRows = emr->compress(&save_empty_rows, NULL);
+  nEmptyRows = emr->compress(mrm_save_empty_rows, NULL);
 }
[…]

### 2010-04-14 programs/mrvector2.cpp

@@ -19,7 +19,7 @@ using namespace std;
 // Vector entries are initialized to zero.
 // Output:  Key = global index;  Value = 0.
 template <typename IDTYPE>
-static void initialize_vec(int itask, KeyValue *kv, void *ptr)
+static void mrv_initialize_vec(int itask, KeyValue *kv, void *ptr)
 {
   double zero = 0.;
   int nprocs;
@@ -57,7 +57,7 @@ MRVector<IDTYPE>::MRVector(
   mr->set_fpath(fpath);
 
   // Emit vector values
-  mr->map(mr->num_procs(), &initialize_vec, &n);
+  mr->map(mr->num_procs(), mrv_initialize_vec<IDTYPE>, &n);
 
   // Gather values to processors based on vector index.
   mr->aggregate(NULL);
@@ -75,7 +75,7 @@ void MRVector<IDTYPE>::MakeEmpty()
 // MRVector::Print
 // Print the vector entries; nothing fancy for parallel here.
 template <typename IDTYPE>
-static void printvector(uint64_t itask, char *key, int keybytes,
+static void mrv_printvector(uint64_t itask, char *key, int keybytes,
                         char *value, int valuebytes, KeyValue *kv, void *ptr)
 {
   cout << *((IDTYPE *) key) << "\t" << *((double *) value) << endl;
@@ -86,13 +86,13 @@ void MRVector<IDTYPE>::Print()
 {
   int me = mr->my_proc();
   cout << "Vector on processor " << me << endl;
-  mr->map(mr, &printvector<IDTYPE>, NULL, 1);  // set addflag to keep mr as is.
+  mr->map(mr, mrv_printvector<IDTYPE>, NULL, 1);  // set addflag to keep mr as is.
 }
 
 /////////////////////////////////////////////////////////////////////////////
 // MRVector::PutScalar
 // Assign a scalar value to each entry of a vector.
-static void setvector(uint64_t itask, char *key, int keybytes, 
+static void mrv_setvector(uint64_t itask, char *key, int keybytes, 
                       char *value, int valuebytes, KeyValue *kv, void *ptr)
 {
   double d = *((double *) ptr);
@@ -102,13 +102,13 @@ static void setvector(uint64_t itask, char *key, int keybytes,
 template <typename IDTYPE>
 void MRVector<IDTYPE>::PutScalar(double d)
 {
-  mr->map(mr, &setvector, &d);
+  mr->map(mr, mrv_setvector, &d);
 }
 
[…]

### 2010-04-14 programs/test_pagerank2.cpp

@@ -30,6 +30,8 @@
 #include "keyvalue.h"
 #include "mrmatrix2.h"
 #include "mrvector2.h"
+#include "mrmatrix2.cpp"
+#include "mrvector2.cpp"
 #include "blockmacros.hpp"
 #include "localdisks.hpp"
 #include "read_fb_data.hpp"
@@ -41,8 +43,8 @@ using namespace std;
 
 ////////////////////////////////////////////////////////////////////////////
 // Global ordinal type for matrix row/column numbers.
-#define IDTYPE int64_t
-#define MPI_IDTYPE MPI_LONG
+#define IDXTYPE int64_t
+#define MPI_IDXTYPE MPI_LONG
 
 // Input file types
 #define FBFILE 0
@@ -75,8 +77,8 @@ void allzero_contribution(char *key, int keybytes, char *multivalue,
 }
 
 double compute_local_allzero_adj(
-  MRMatrix<IDTYPE> *A,
-  MRVector<IDTYPE> *x,
+  MRMatrix<IDXTYPE> *A,
+  MRVector<IDXTYPE> *x,
   double alpha
 )
 {
@@ -110,17 +112,17 @@ void compute_lmax_residual(char *key, int keybytes,
 }
 
 ////////////////////////////////////////////////////////////////////////////
-MRVector<IDTYPE> *pagerank(
-  MRMatrix<IDTYPE> *A,
+MRVector<IDXTYPE> *pagerank(
+  MRMatrix<IDXTYPE> *A,
   double alpha,
   double tolerance
 )
 {
   int me = A->mr->my_proc();
   if (me == 0) {cout << "Initializing vectors..." << endl; flush(cout);}
-  MRVector<IDTYPE> *x = new MRVector<IDTYPE>(A->NumRows(),
+  MRVector<IDXTYPE> *x = new MRVector<IDXTYPE>(A->NumRows(),
                                              A->mr->memsize, A->mr->fpath);
-  MRVector<IDTYPE> *y = new MRVector<IDTYPE>(x->GlobalLen(), 
+  MRVector<IDXTYPE> *y = new MRVector<IDXTYPE>(x->GlobalLen(), 
                                              A->mr->memsize, A->mr->fpath);
 
[…]

### 2010-04-14 programs/test_pagerank2.cpp
Moved cpp to hpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@542 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -28,10 +28,8 @@
 #include <mpi.h>
 #include "mapreduce.h"
 #include "keyvalue.h"
-#include "mrmatrix2.h"
-#include "mrvector2.h"
-#include "mrmatrix2.cpp"
-#include "mrvector2.cpp"
+#include "mrmatrix2.hpp"
+#include "mrvector2.hpp"
 #include "blockmacros.hpp"
 #include "localdisks.hpp"
 #include "read_fb_data.hpp"

### 2010-04-14 programs/mrmatrix2.h
Time to start testing
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@543 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -1,61 +1,0 @@
-// MapReduce Matrix class
-// Karen Devine, 1416
-// April 2010
-//
-// Store matrix as a MapReduce object, with different storage depending on
-// whether matrix A is stored or transpose A^T is stored.
-// For A:  (Key, Value) = (row index i, (col index j, a_ij))
-// For A^T:  (Key, Value) = (col index j, (row index i, a_ij))
-// 
-// Provides method for matrix-vector multiplication  A*x=y.
-//
-#include <iostream>
-#include <list>
-#include "mpi.h"
-#include "mapreduce.h"
-#include "keyvalue.h"
-
-using namespace MAPREDUCE_NS;
-using namespace std;
-
-#include "mrall2.h"
-#include "mrvector2.h"
-
-#ifndef __MRMATRIX
-#define __MRMATRIX
-
-////////////////////////////////////////////////////////////////////////////
-
-template <typename IDTYPE>
-class MRNonZero {
-public:
-  IDTYPE ij;        // column (if A) or row (if A^T) index
-  double nzv;       // non-zero value.
-  MRNonZero() {};
-  ~MRNonZero() {};
-};
-
-template <typename IDTYPE>
-class MRMatrix {
-  public:  
-    MRMatrix(IDTYPE, IDTYPE, MapReduce *, bool transpose=0, 
-             int pagesize=64, const char *fpath=".");
-    ~MRMatrix() {delete mr;};
-
-    IDTYPE NumRows() { return N; };
-    IDTYPE NumCols() { return M; };
-    void MatVec(MRVector<IDTYPE> *, MRVector<IDTYPE> *);
-    void Scale(double);
-    void MakeEmpty() {delete mr;};
-    MapReduce *mr;  // Actual storage; perhaps should be private.
-    uint64_t nEmptyRows;  // Number of rows of A with no nonzeros.
-    MRVector<IDTYPE> *emptyRows; // Indices of rows of A with no nonzeros (leaf nodes).
-  private:
-    IDTYPE N;  // Number of rows
-    IDTYPE M;  // Number of cols
-    bool transposeFlag;  // State variable; indicates whether matrix is stored
-                         // as A or as A^T.
-};
-
-
-#endif

### 2010-04-14 programs/mrmatrix2.hpp

@@ -1,6 +1,13 @@
-// MapReduce Matrix Class
+// MapReduce Matrix class
 // Karen Devine, 1416
-// June 2008
+// April 2010
+//
+// Store matrix as a MapReduce object, with different storage depending on
+// whether matrix A is stored or transpose A^T is stored.
+// For A:  (Key, Value) = (row index i, (col index j, a_ij))
+// For A^T:  (Key, Value) = (col index j, (row index i, a_ij))
+// 
+// Provides method for matrix-vector multiplication  A*x=y.
 //
 // The dimensions of the matrix A are given by N and M (N rows, M columns).
 
@@ -9,12 +16,53 @@
 #include "mpi.h"
 #include "mapreduce.h"
 #include "keyvalue.h"
-#include "mrmatrix2.h"
-#include "mrvector2.h"
+#include "mrvector2.hpp"
 #include "blockmacros.hpp"
+#include "mrall2.h"
 
 using namespace MAPREDUCE_NS;
 using namespace std;
+
+#ifndef __MRMATRIX
+#define __MRMATRIX
+
+/////////////////////////////////////////////////////////////////////////////
+/////////////////////////////////////////////////////////////////////////////
+
+template <typename IDTYPE>
+class MRNonZero {
+public:
+  IDTYPE ij;        // column (if A) or row (if A^T) index
+  double nzv;       // non-zero value.
+  MRNonZero() {};
+  ~MRNonZero() {};
+};
+
+template <typename IDTYPE>
+class MRMatrix {
+  public:  
+    MRMatrix(IDTYPE, IDTYPE, MapReduce *, bool transpose=0, 
+             int pagesize=64, const char *fpath=".");
+    ~MRMatrix() {delete mr;};
+
+    IDTYPE NumRows() { return N; };
+    IDTYPE NumCols() { return M; };
+    void MatVec(MRVector<IDTYPE> *, MRVector<IDTYPE> *);
+    void Scale(double);
+    void MakeEmpty() {delete mr;};
+    MapReduce *mr;  // Actual storage; perhaps should be private.
+    uint64_t nEmptyRows;  // Number of rows of A with no nonzeros.
+    MRVector<IDTYPE> *emptyRows; // Indices of rows of A with no nonzeros (leaf nodes).
+  private:
+    IDTYPE N;  // Number of rows
+    IDTYPE M;  // Number of cols
+    bool transposeFlag;  // State variable; indicates whether matrix is stored
+                         // as A or as A^T.
+};
+
+
[…]

### 2010-04-14 programs/mrvector2.h

@@ -1,46 +1,0 @@
-// MapReduce Vector class
-// Karen Devine, 1416
-// April 2010
-//
-// Stores a vector as a MapReduce object with 
-// (Key, Value) = (Global Index i, vector value i)
-
-#include <iostream>
-#include <list>
-#include "mpi.h"
-#include "mapreduce.h"
-#include "keyvalue.h"
-
-#ifndef __MRVECTOR
-#define __MRVECTOR
-
-using namespace MAPREDUCE_NS;
-using namespace std;
-
-#include "mrall2.h"
-
-/////////////////////////////////////////////////////////////////////////////
-template <typename IDTYPE>
-class MRVector {
-  public:
-    MRVector(IDTYPE, int pagesize=64, const char *fpath = ".");
-    ~MRVector() {MakeEmpty();}
-    void MakeEmpty();
-    void PutScalar(double);
-    void AddScalar(double);
-    void Scale(double);
-    double GlobalSum();
-    double GlobalMax();
-    double GlobalMin();
-    double LocalSum();
-    double LocalMax();
-    double LocalMin();
-    IDTYPE GlobalLen() {return global_len;};
-    void Print();
-    MapReduce *mr;   // Where the data is actually stored; probably should
-                     // be private.
-  private:
-    IDTYPE global_len;
-};
-
-#endif

### 2010-04-14 programs/mrvector2.hpp

@@ -1,4 +1,8 @@
-// MapReduce Vectorr Class
+// MapReduce Vector Class compatible with out-of-core MR-MPI.
+//
+// Stores a vector as a MapReduce object with 
+// (Key, Value) = (Global Index i, vector value i)
+//
 // Karen Devine, 1416
 // April 2010
 
@@ -8,10 +12,43 @@
 #include "mpi.h"
 #include "mapreduce.h"
 #include "keyvalue.h"
-#include "mrvector2.h"
+#include "mrall2.h"
 
 using namespace MAPREDUCE_NS;
 using namespace std;
+
+#ifndef __MRVECTOR
+#define __MRVECTOR
+
+/////////////////////////////////////////////////////////////////////////////
+/////////////////////////////////////////////////////////////////////////////
+
+template <typename IDTYPE>
+class MRVector {
+  public:
+    MRVector(IDTYPE, int pagesize=64, const char *fpath = ".");
+    ~MRVector() {MakeEmpty();}
+    void MakeEmpty();
+    void PutScalar(double);
+    void AddScalar(double);
+    void Scale(double);
+    double GlobalSum();
+    double GlobalMax();
+    double GlobalMin();
+    double LocalSum();
+    double LocalMax();
+    double LocalMin();
+    IDTYPE GlobalLen() {return global_len;};
+    void Print();
+    MapReduce *mr;   // Where the data is actually stored; probably should
+                     // be private.
+  private:
+    IDTYPE global_len;
+};
+
+/////////////////////////////////////////////////////////////////////////////
+/////////////////////////////////////////////////////////////////////////////
+
 
 /////////////////////////////////////////////////////////////////////////////
 // initialize_vec map() function
@@ -237,3 +274,4 @@ double MRVector<IDTYPE>::LocalMax()
   return max;
 }
 
+#endif

### 2010-04-14 programs/test_pagerank2.cpp

@@ -28,8 +28,8 @@
 #include <mpi.h>
 #include "mapreduce.h"
 #include "keyvalue.h"
-#include "mrmatrix2.hpp"
 #include "mrvector2.hpp"
+#include "mrmatrix2.hpp"
 #include "blockmacros.hpp"
 #include "localdisks.hpp"
 #include "read_fb_data.hpp"

### 2010-04-14 programs/mrmatrix2.hpp
Added matrix initialization
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@544 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -41,18 +41,19 @@ public:
 template <typename IDTYPE>
 class MRMatrix {
   public:  
-    MRMatrix(IDTYPE, IDTYPE, MapReduce *, bool transpose=0, 
+    MRMatrix(IDTYPE, IDTYPE, MapReduce *, MapReduce *, 
              int pagesize=64, const char *fpath=".");
     ~MRMatrix() {delete mr;};
 
     IDTYPE NumRows() { return N; };
     IDTYPE NumCols() { return M; };
     void MatVec(MRVector<IDTYPE> *, MRVector<IDTYPE> *);
+    void Transpose();
     void Scale(double);
     void MakeEmpty() {delete mr;};
     MapReduce *mr;  // Actual storage; perhaps should be private.
     uint64_t nEmptyRows;  // Number of rows of A with no nonzeros.
-    MRVector<IDTYPE> *emptyRows; // Indices of rows of A with no nonzeros (leaf nodes).
+    MapReduce *emptyRows; // Indices of rows of A with no nonzeros (leaf nodes).
   private:
     IDTYPE N;  // Number of rows
     IDTYPE M;  // Number of cols
@@ -69,78 +70,118 @@ class MRMatrix {
 // Takes a graph as input and constructs an appropriate matrix.
 //
 // initialize_matrix map function
-// Given MapReduce object containing edges, create matrix non-zeros.
+// Given MapReduce object containing edges and vertices, 
+// create matrix non-zeros.  Give initial value of 1/(outdegree of v_i) to
+// each nonzero in row i.  
+// For rows with outdegree = 0, add a KV to the MapReduce object provided
+// in ptr.
 template <typename IDTYPE>
-static void mrm_initialize_matrix(uint64_t itask, char *key, int keybytes,
-                              char *value, int valuebytes, 
-                              KeyValue *kv, void *ptr)
+static void mrm_initialize_matrix(char *key, int keybytes, 
+                  char *multivalue, int nvalues, int *valuebytes,
+                  KeyValue *kv, void *ptr)
 {
-  MRNonZero<IDTYPE> v;
-  v.ij = *((IDTYPE *) value);
-  v.nzv = 0.;
-  kv->add(key, keybytes, (char *) &v, sizeof(MRNonZero<IDTYPE>));
-}
+  // To receive entries for all-zero rows...
+  KeyValue *emptyRowsKV = (KeyValue *) ptr;
 
[…]

### 2010-04-14 programs/read_mm_data.hpp

@@ -126,7 +126,7 @@ void ReadMMData::run(
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();
 
-  if (me == 0) printf("Reading input files...\n");
+  if (me == 0) printf("Reading input files %s...\n", filename);
   *nrawedges = mrraw->map(np,1,&filename,'\n',80,&readMM_fileread,this);
   int tmp = N;
   MPI_Allreduce(&tmp, &N, 1, MPI_INT, MPI_MAX, MPI_COMM_WORLD);
@@ -211,7 +211,8 @@ static void readMM_vertex_emit(int itask, KeyValue *kv, void *ptr)
 #endif
 
 
-static void readMM_fileread(int itask, char *bytes, int nbytes, KeyValue *kv, void *ptr)
+static void readMM_fileread(int itask, char *bytes, int nbytes, 
+                            KeyValue *kv, void *ptr)
 {
   ReadMMData *rfp = (ReadMMData *) ptr;
   uint64_t edge[4] = {0, 0, 0, 0};

### 2010-04-14 programs/test_pagerank2.cpp

@@ -82,7 +82,7 @@ double compute_local_allzero_adj(
 {
   double sum = 0.;
 
-  MapReduce *emr = A->emptyRows->mr;
+  MapReduce *emr = A->emptyRows;
 
   emr->add(x->mr);
   emr->compress(allzero_contribution, &sum);
@@ -131,6 +131,9 @@ MRVector<IDXTYPE> *pagerank(
   A->Scale(alpha);
   x->PutScalar(1./x->GlobalLen());
 
+cout << "KDDKDD After PutScalar" << endl;
+x->Print();
+
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();
 
@@ -145,7 +148,8 @@ MRVector<IDXTYPE> *pagerank(
 
     // Compute local adjustment for all-zero rows.
     double allzeroadj = 0.;
-    allzeroadj = compute_local_allzero_adj(A, x, alpha);
+    if (A->nEmptyRows)
+      allzeroadj = compute_local_allzero_adj(A, x, alpha);
     ladj += allzeroadj;
 
     // Compute global adjustment via all-reduce-like operation.
@@ -155,14 +159,23 @@ MRVector<IDXTYPE> *pagerank(
     // Compute global adjustment.
     A->MatVec(x, y);
 
+cout << "KDDKDD After MatVec" << endl;
+y->Print();
+
     // Add adjustment to product vector in mr.
     y->AddScalar(gadj);
+
+cout << "KDDKDD After AddScalar" << endl;
+y->Print();
 
     // Compute max-norm of vector in mr.
     double gmax = y->GlobalMax();
 
     // Scale vector in mr by 1/maxnorm.
     y->Scale(1./gmax);
+
+cout << "KDDKDD After Scale" << endl;
+y->Print();
 
     // Compute local residual; this also empties x->mr.
     double lresid = 0.;
@@ -177,6 +190,8 @@ MRVector<IDXTYPE> *pagerank(
     MRVector<IDXTYPE> *tmp = x;
     x = y;
     y = tmp;
+cout << "KDDKDD After Swap" << endl;
+x->Print();
 
     if (me == 0) {
       cout << "iteration " << iter+1 << " resid " << gresid << endl; 
@@ -277,6 +292,12 @@ int main(int narg, char **args)
   int pagesize = 64;
   int filetype = RMAT;
 
+  if (me == 0) 
+    cout << "Syntax: pagerank "
+         << "[-a alpha] [-t tolerance] "
[…]

### 2010-04-14 app/cc2.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@545 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -65,11 +65,6 @@ void CC2::command(int narg, char **arg)
 
   mrv->map(nprocs,vs->appmap,vs->appptr);
 
-  //if (me == 0) printf("VERTEX ZONE pre-list:\n");
-  //mrv->print(-1,1,2,2);
-  //if (me == 0) printf("EDGE pre-list:\n");
-  //mre->print(-1,1,7,2);
-
   // loop until zones do not change
 
   MPI_Barrier(world);
@@ -85,9 +80,6 @@ void CC2::command(int narg, char **arg)
     mr->collate(NULL);
     mr->reduce(ezone->appreduce,ezone->appptr);
 
-    //if (me == 0) printf("EDGE ZONE list:\n");
-    //mr->print(-1,1,7,2);
-
     mr->collate(NULL);
     zwin->flag = 0;
     uint64_t foo = mr->reduce(zwin->appreduce,zwin->appptr);
@@ -95,24 +87,15 @@ void CC2::command(int narg, char **arg)
     MPI_Allreduce(&zwin->flag,&flagall,1,MPI_INT,MPI_SUM,world);
     if (!flagall) break;
 
-    //if (me == 0) printf("ZONE ZONE list:\n");
-    //mr->print(-1,1,2,2);
-
     mrv->map(mrv,im->appmap_mr,im->appptr);
     mrv->map(mr,zm->appmap_mr,zm->appptr,1);
     mrv->collate(NULL);
     mrv->reduce(zr2->appreduce,zr2->appptr);
-    
-    //if (me == 0) printf("VERTEX ZONE list:\n");
-    //mrv->print(-1,1,2,2);
   }
 
   // strip any hi-bits from final (Vi,Zi) key/values
 
   mrv->map(mrv,strip->appmap_mr,strip->appptr);
-
-  //if (me == 0) printf("VERTEX ZONE final list:\n");
-  //mrv->print(-1,1,2,2);
 
   MPI_Barrier(world);
   double tstop = MPI_Wtime();

### 2010-04-14 app/in.cc

@@ -1,7 +1,7 @@
 # RMAT matrix generation & statistics
 
 variable order equal 15
-variable nz equal 1
+variable nz equal 2
 variable nrows equal 2^v_order
 variable n equal v_nrows*v_nz
 
@@ -22,8 +22,8 @@ mrz.memsize(1)
 mrz.verbosity(1)
 mrz.timer(1)
 
-#cc(${nrows},mre,mrv,mrz)
-cc2(${nrows},0,mre,mrv,mrz)
+cc(${nrows},mre,mrv,mrz)
+#cc2(${nrows},0,mre,mrv,mrz)
 
 # print CC stats, destroys mrv
 

### 2010-04-14 app/reduce_zone_reassign2.cpp

@@ -49,7 +49,7 @@ void ReduceZoneReassign2::reduce(char *key, int keybytes,
   int thresh = data->thresh;
   uint64_t lmask = data->lmask;
   
-  int i,hinew;
+  int i,hnew;
   char *value;
   uint64_t znew;
 
@@ -58,8 +58,9 @@ void ReduceZoneReassign2::reduce(char *key, int keybytes,
 
   uint64_t zcount = 0;
   uint64_t zone = *(uint64_t *) key;
-  int hibit = zone >> 63;
+  int hkey = zone >> 63;
   zone &= lmask;
+  int hwinner = 0;
 
   uint64_t nvalues_total;
   CHECK_FOR_BLOCKS(multivalue,valuebytes,nvalues,nvalues_total)
@@ -69,11 +70,11 @@ void ReduceZoneReassign2::reduce(char *key, int keybytes,
   for (i = 0; i < nvalues; i++) {
     if (valuebytes[i] != sizeof(uint64_t)) {
       znew = *(uint64_t *) value;
-      hinew = znew >> 63;
+      hnew = znew >> 63;
       znew &= INT64MAX;
       if (znew < zone) {
 	zone = znew;
-	hibit = hinew;
+	hwinner = hnew;
       }
       zcount++;
     }
@@ -85,7 +86,8 @@ void ReduceZoneReassign2::reduce(char *key, int keybytes,
   // emit one KV per vertex with zone ID as value
   // add hi-bit to zone if necessary
 
-  if (hibit || nvalues_total-zcount > thresh) zone |= HIBIT;
+  if (hkey || hwinner) zone |= HIBIT;
+  else if (nvalues_total-zcount > thresh) zone |= HIBIT;
 
   BEGIN_BLOCK_LOOP(multivalue,valuebytes,nvalues)
 

### 2010-04-14 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@546 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-In the map() operation, each of $P$ processors is tasked with
-generating a $1/P$ fraction of the desired edges.  A single random
-$i,j$ edge is computed recursively as follows.  Pick a random quadrant
-of the $A$ matrix with relative probabilities $a$, $b$, $c$, and $d$.
-Treat the chosen quadrant as a sub-matrix and select a random quadrant
-within it, in the same manner.  Repeat this process $n$ times where $N
-= 2^n$.  At the end of the recursion, the final ``quadrant'' is
-non-zero matrix element $A_{ij}$.
+In the map() operation, each of $P$ processors generates a $1/P$
+fraction of the desired edges.  A single random $i,j$ edge is computed
+recursively as follows.  Pick a random quadrant of the $A$ matrix with
+relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
+quadrant as a sub-matrix and select a random quadrant within it, in
+the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
+the end of the recursion, the final ``quadrant'' is non-zero matrix
+element $A_{ij}$.
 
@@
 a power-of-two, to generate graphs with weighted edges (assign a
-random numeric value to the $A_{ij}$ edge), graphs without self edges
+numeric value to the $A_{ij}$ edge), graphs without self edges
 (require $i != j$), or graphs with undirected edges (require $i < $j).
+Or the general R-MAT matrix can be further processed by MapReduce
+operations to meet these requirements.
 
@@
 A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
-the three edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
-\ref{fig:tri} outlines a MapReduce algorithm that enumerates all such
-triplets, assuming an input graph of undirected edge keys $(V_i,V_j)$
+the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
+\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
+triangles, assuming an input graph of undirected edges $(V_i,V_j)$
 where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
-generated matrix.  This exposition follows the triangle-finding
-algorithm presented in \cite{Cohen}.
+matrix.  This exposition follows the triangle-finding algorithm
+presented in \cite{Cohen}.
 
@@
 
-Discuss connected component identification algorithm.
-
-Do we need differnt variants of CC?
-
-Could just do one, discuss possible inefficiency due to giant CC.
+A connected component of a graph is a set of vertices where all pairs
+of vertices in the set are connected by a path of edges.  A sparse
+graph may contain many such components.  Figure \ref{fig:cc} outlines
+a MapReduce algorithm that labels each vertex in a graph with a
+component ID.  All vertices in the same component are labelled with
+the same ID, which is the ID of a vertex in the component.  We assume
+an input graph of undirected edges $(V_i,V_j)$.  This exposition also
+follows the connected-component algorithm presented in \cite{Cohen},
+with the addition of logic that load-balances data across processors
+when one or a few giant components exist in the graph.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+Iterate: \\
+\> 1 Map: \> Convert edges to vertices \\
+    \> \> \> input Key = $E_{ij}$, Value = NULL \\
+    \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
+    \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
+\> 1 Add: \> Zone assignment of each vertex \\
+    \> \> \> output Key = $V_i$, Value = $Z_i$ \\
+\> 1 Collate: \> Vertex as key \\
+\> 1 Reduce: \> Emit edges of each vertex with zone of vertex \\
+       \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
+       \> \> \> for each $E$ in MultiValue: \\
+      \> \> \> \> output Key = $E_{ij}$, Value = $Z_i$ \\
+
+\> 2 Collate: \> Edge as key \\
+\> 2 Reduce: \> Emit zone re-assignments \\
+       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
+       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
+       \> \> \> if $Z_i$ and $Z_j$ are different: \\
+      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
+\> 2 Exit: \> if no output by Reduce 2 \\
+\> 3 Map: \> Invert vertex/zone pairs \\
+    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
+    \> \> \> if $Z_i$ is not partitioned: \\
+   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
+    \> \> \> else: \\
+   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
+\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
+    \> \> \> if $Z_i$ is not partitioned: \\
+   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
+    \> \> \> else: \\
+   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
+   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
+\> 3 Collate: \> Zone ID as key \\
+\> 3 Reduce: \> Emit new zone assignment of each vertex \\
+       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
+       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
+       \> \> \> partition $Z_{new}$ if number of $V$ > threshhold \\
+       \> \> \> for each $V$ in MultiValue: \\
+      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$
 
@@
 
- \caption{MapReduce algorithm for connected component identification.}
+ \caption{MapReduce algorithm for connected component labelling.}
 
- \label{fig:rmat}
+ \label{fig:cc}
 \end{figure}
 
+The algorithm begins (before the iteration loop) by assigning each
+vertex to its own component or ``zone'', so that $Z_i$ = $V_i$.  Each
+iteration will grow the zones, one layer of neighbors at a time.  As
+zones collide due to shared edges, a winner is chosen (the smaller
+zone ID), and vertices in the losing zone are reassigned to the
+winning zone.  When the iterations complete, each zone will have
+become a fully connected component.  The algorithm thus finds all
+connected components in the graph simultaneously.  The number of
+iterations required depends on the largest diameter of any component
+in the graph.
+
+The first {\it map()} operation emits the vertices in each edge as
+keys, with the edge as a value.  The current zone assignment of each
+vertex is added to the set of key/value pairs.  The first collate()
+operation collects all the edges of a vertex and its zone assignment
+together in one multi-value.  The first reduce operation then re-emits
+each edge, tagged by the zone assignment of one of its vertices.
+
+Since each edge was emitted twice, the second collate operation
+collects the zone assignments for its two vertices together.  If the
+two zone IDs are different, the second reduce operation chooses a
+winner (the min of the 2 IDs), and emits the loser ID as a key, with
+the winning ID as a value.  If no zone ID changes are emitted, the
+algorithm is finished, and the iteration exits.
+
+The third map() operation inverts the vertex/zone key/value pairs to
+become zone/vertex pairs.  The third add operation adds the changing
+zone assignments to the set of key/value pairs.  The third collate()
+can then collects all the vertices of a zone and zero or more
+reassignments for the zone ID.  Since a zone could collide with
+multiple other zones on the same iteration due to shared edges, the
+new zone ID becomes the minimum ID of any of the neighboring zones.
+If no zone reassignment value appears in the multi-value, the zone ID
+is unchanged.  The final reduce() a key/value pair for each vertex in
+the zone, with the vertex as a key and the new zone ID as a value.
+
+Note that if a graph has only a few components, then the third collate
+operation, which keys on the zone ID, may generate a few very large
+key/multi-value (KMV) pairs.  For example, if the graph is fully
+connected, then on the last iteration, a single KMV pair will contain
+all vertices in the graph and be assigned to one processor.  This
+imbalance in memory and computational work can lead to poor parallel
+performance of the overall algorithm.  To counter this effect, the
+various operations of stage 3 include extra logic.  The idea is to
+partition zones whose vertex count exceeds a user-defined threshhold
+into $P$ sub-zones, where $P$ is the number of processors.  The 64-bit
+integer that stores the zone ID also stores a bit flag indicating the
+zone has been partitioned and a set of bits that encode the processor
+ID.
+
+During the third map() operation, if the zone has been partitioned,
+then the vertex is assigned to a random processor and the processor ID
+bits are added to the zone ID, as indicated by the $Z_i^+$ notation in
+Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
+third add() operation emits the zone ID change key/value pair
+($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In this case
+($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once for
+each processor, and once as if $Z_i$ had not been partitioned (for a
+reason discussed below).
+
+This additional partitioning logic means that the third collate()
+operation, which keys on zone IDs, some of which now include processor
+bits, will collect only a $1/P$ subset of the vertices in large zones
+onto each processor.  But the multivalue on each processor contains
+all the zone-reassignments relevant to the unpartitioned zone.  This
+allows the third reduce operation to change the zone ID (if necessary)
+in a consistent manner across all $P$ multi-values that contain the
+zone's vertices.  When the reduce() operation emits new zone
+assignments for each vertex, the zone retains its partitioned status,
+and the partition bit is also explicitly set if the vertex count
+exceeds the threshhold for the first time.
+
+Note that this logic does not guarantee that the partition bits of the
+zone IDs for all the vertices in a single zone will be set
+consistently on a given iteration.  For example, an unpartitioned zone
+with a small ID may consume a partitioned zone.  The vertices from the
+partitioned zone will retain their partitioned status, but the
+original vertices in the small zone may not set the partition bit of
+their zone IDs.  On subsequent iterations, the third add() operation
+emits $P+1$ copies of new zone reassignments for both partitioned and
+unpartioned zone IDs, to insure all vertices in the zone will know the
+reassignment information.
+
 Discussion of Luby algorithm for maximally independent sets.
+
+footnote on NP-complete for maximal; MapReduce is unlikely to
+help with that issue.
 

### 2010-04-14 src/mapreduce.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@547 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -1398,6 +1398,36 @@ uint64_t MapReduce::map(MapReduce *mr,
 }
 
 /* ----------------------------------------------------------------------
+   debug print of KV or KMV pairs
+   if all procs are printing, pass print token from proc to proc
+------------------------------------------------------------------------- */
+
+void MapReduce::print(int proc, int nstride, int kflag, int vflag)
+{
+  MPI_Status status;
+
+  if (kv == NULL && kmv == NULL)
+    error->all("Cannot print without KeyValue or KeyMultiValue");
+  if (kflag < 0 || vflag < 0) error->all("Invalid print args");
+  if (kflag > 7 || vflag > 7) error->all("Invalid print args");
+
+  if (proc == me) {
+    if (kv) kv->print(nstride,kflag,vflag);
+    if (kmv) kmv->print(nstride,kflag,vflag);
+  }
+
+  if (proc >= 0) return;
+
+  int token;
+  MPI_Barrier(comm);
+  if (me > 0) MPI_Recv(&token,0,MPI_INT,me-1,0,comm,&status);
+  if (kv) kv->print(nstride,kflag,vflag);
+  if (kmv) kmv->print(nstride,kflag,vflag);
+  if (me < nprocs-1) MPI_Send(&token,0,MPI_INT,me+1,0,comm);
+  MPI_Barrier(comm);
+}
+
+/* ----------------------------------------------------------------------
    create a KV from a KMV via a parallel reduce operation for nmap tasks
    make one call to appreduce() for each KMV pair
    each proc processes its owned KMV pairs
@@ -2161,36 +2191,6 @@ void MapReduce::cummulative_stats(int level, int reset)
     rsize = wsize = 0;
     cssize = crsize = 0;
   }
-}
-
-/* ----------------------------------------------------------------------
-   debug print of KV or KMV pairs
-   if all procs are printing, pass print token from proc to proc
-------------------------------------------------------------------------- */
-
-void MapReduce::print(int proc, int nstride, int kflag, int vflag)
-{
-  MPI_Status status;
-
-  if (kv == NULL && kmv == NULL)
-    error->all("Cannot print without KeyValue or KeyMultiValue");
-  if (kflag < 0 || vflag < 0) error->all("Invalid print args");
-  if (kflag > 7 || vflag > 7) error->all("Invalid print args");
-
[…]

### 2010-04-14 src/mapreduce.h

@@ -85,6 +85,7 @@ class MapReduce {
 				     class KeyValue *, void *),
 	       void *, int addflag = 0);
 
+  void print(int, int, int, int);
   uint64_t reduce(void (*)(char *, int, char *,
 			   int, int *, class KeyValue *, void *),
 		  void *);
@@ -100,7 +101,6 @@ class MapReduce {
   void kv_stats(int);
   void kmv_stats(int);
   void cummulative_stats(int, int);
-  void print(int, int, int, int);
 
   void set_fpath(const char *);
 

### 2010-04-15 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@549 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
+\subsection{PageRank}
+
 The PageRank algorithm assings a relative numeric rank to each
@@
 residuals.
+
+\subsection{Triangle enumeration}
 
@@
 
+\subsection{Connected component labeling}
+
 A connected component of a graph is a set of vertices where all pairs
@@
 
- \caption{MapReduce algorithm for connected component labelling.}
+ \caption{MapReduce algorithm for connected component labeling.}
 
@@
 keys, with the edge as a value.  The current zone assignment of each
-vertex is added to the set of key/value pairs.  The first collate()
-operation collects all the edges of a vertex and its zone assignment
-together in one multi-value.  The first reduce operation then re-emits
-each edge, tagged by the zone assignment of one of its vertices.
+vertex is added to the set of key/value pairs.  The first {\it
+collate()} operation collects all the edges of a vertex and its zone
+assignment together in one multi-value.  The first {\it reduce()}
+operation then re-emits each edge, tagged by the zone assignment of
+one of its vertices.
 
-Since each edge was emitted twice, the second collate operation
-collects the zone assignments for its two vertices together.  If the
-two zone IDs are different, the second reduce operation chooses a
-winner (the min of the 2 IDs), and emits the loser ID as a key, with
-the winning ID as a value.  If no zone ID changes are emitted, the
-algorithm is finished, and the iteration exits.
+Since each edge was emitted twice, the second {\it collate()}
+operation collects the zone assignments for its two vertices together.
+If the two zone IDs are different, the second {\it reduce()} operation
+chooses a winner (the min of the 2 IDs), and emits the loser ID as a
+key, with the winning ID as a value.  If no zone ID changes are
+emitted, the algorithm is finished, and the iteration exits.
 
-The third map() operation inverts the vertex/zone key/value pairs to
-become zone/vertex pairs.  The third add operation adds the changing
-zone assignments to the set of key/value pairs.  The third collate()
-can then collects all the vertices of a zone and zero or more
-reassignments for the zone ID.  Since a zone could collide with
-multiple other zones on the same iteration due to shared edges, the
-new zone ID becomes the minimum ID of any of the neighboring zones.
-If no zone reassignment value appears in the multi-value, the zone ID
-is unchanged.  The final reduce() a key/value pair for each vertex in
-the zone, with the vertex as a key and the new zone ID as a value.
+The third {\it map()} operation inverts the vertex/zone key/value
+pairs to become zone/vertex pairs.  The third {\it add()} operation
+adds the changing zone assignments to the set of key/value pairs.  The
+third collate() can then collects all the vertices of a zone and zero
+or more reassignments for the zone ID.  Since a zone could collide
+with multiple other zones on the same iteration due to shared edges,
+the new zone ID becomes the minimum ID of any of the neighboring
+zones.  If no zone reassignment value appears in the multi-value, the
+zone ID is unchanged.  The final reduce() a key/value pair for each
+vertex in the zone, with the vertex as a key and the new zone ID as a
+value.
 
-Note that if a graph has only a few components, then the third collate
-operation, which keys on the zone ID, may generate a few very large
-key/multi-value (KMV) pairs.  For example, if the graph is fully
-connected, then on the last iteration, a single KMV pair will contain
-all vertices in the graph and be assigned to one processor.  This
-imbalance in memory and computational work can lead to poor parallel
-performance of the overall algorithm.  To counter this effect, the
-various operations of stage 3 include extra logic.  The idea is to
-partition zones whose vertex count exceeds a user-defined threshhold
-into $P$ sub-zones, where $P$ is the number of processors.  The 64-bit
-integer that stores the zone ID also stores a bit flag indicating the
-zone has been partitioned and a set of bits that encode the processor
-ID.
+Note that if a graph has only a few components, then the third {\it
+collate()} operation, which keys on the zone ID, may generate a few
+very large key/multi-value (KMV) pairs.  For example, if the graph is
+fully connected, then on the last iteration, a single KMV pair will
+contain all vertices in the graph and be assigned to one processor.
+This imbalance in memory and computational work can lead to poor
+parallel performance of the overall algorithm.  To counter this
+effect, the various operations of stage 3 include extra logic.  The
+idea is to partition zones whose vertex count exceeds a user-defined
+threshhold into $P$ sub-zones, where $P$ is the number of processors.
+The 64-bit integer that stores the zone ID also stores a bit flag
+indicating the zone has been partitioned and a set of bits that encode
+the processor ID.
 
-During the third map() operation, if the zone has been partitioned,
-then the vertex is assigned to a random processor and the processor ID
-bits are added to the zone ID, as indicated by the $Z_i^+$ notation in
-Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
-third add() operation emits the zone ID change key/value pair
-($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In this case
-($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once for
-each processor, and once as if $Z_i$ had not been partitioned (for a
-reason discussed below).
+During the third {\it map()} operation, if the zone has been
+partitioned, then the vertex is assigned to a random processor and the
+processor ID bits are added to the zone ID, as indicated by the
+$Z_i^+$ notation in Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been
+partitioned, the third {\it add()} operation emits the zone ID change
+key/value pair ($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In
+this case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times,
+once for each processor, and once as if $Z_i$ had not been partitioned
+(for a reason discussed below).
 
-This additional partitioning logic means that the third collate()
-operation, which keys on zone IDs, some of which now include processor
-bits, will collect only a $1/P$ subset of the vertices in large zones
-onto each processor.  But the multivalue on each processor contains
-all the zone-reassignments relevant to the unpartitioned zone.  This
-allows the third reduce operation to change the zone ID (if necessary)
-in a consistent manner across all $P$ multi-values that contain the
-zone's vertices.  When the reduce() operation emits new zone
-assignments for each vertex, the zone retains its partitioned status,
-and the partition bit is also explicitly set if the vertex count
-exceeds the threshhold for the first time.
+This additional partitioning logic means that the third {\it
+collate()} operation, which keys on zone IDs, some of which now
+include processor bits, will collect only a $1/P$ subset of the
+vertices in large zones onto each processor.  But the multivalue on
+each processor contains all the zone-reassignments relevant to the
+unpartitioned zone.  This allows the third {\it reduce()} operation to
+change the zone ID (if necessary) in a consistent manner across all
+$P$ multi-values that contain the zone's vertices.  When the {\it
+reduce()} operation emits new zone assignments for each vertex, the
+zone retains its partitioned status, and the partition bit is also
+explicitly set if the vertex count exceeds the threshhold for the
+first time.
 
@@
 original vertices in the small zone may not set the partition bit of
-their zone IDs.  On subsequent iterations, the third add() operation
-emits $P+1$ copies of new zone reassignments for both partitioned and
-unpartioned zone IDs, to insure all vertices in the zone will know the
-reassignment information.
+their zone IDs.  On subsequent iterations, the third {\it add()}
+operation emits $P+1$ copies of new zone reassignments for both
+partitioned and unpartioned zone IDs, to insure all vertices in the
+zone will know the reassignment information.
 
-Discussion of Luby algorithm for maximally independent sets.
+\subsection{Maximal independent set identification}
 
-footnote on NP-complete for maximal; MapReduce is unlikely to
-help with that issue.
+An ``independent'' set of vertices in a graph is one where no pair of
+vertices in the set share an edge.  The set is ``maximal'' if no
+additional vertex can be added to it \footnote{A maximal set is
+``maximum'' if it is the largest maximal set.  Finding a graph's
+maximum independent set is an NP-hard problem, which MapReduce is
+unlikely to help with.}.  Finding a maximal independent set is useful
+in several contexts, such as identifying large numbers of indepenent
+starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
+is a well-known parallel method for finding a minimal set.  Figure
+\ref{fig:luby} outlines a MapReduce version of Luby's algorithm.
 
@@
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+Iterate: \\
+\> 1 Map: \> Convert edges to vertices \\
+    \> \> \> input Key = $E_{ij}$, Value = NULL \\
+    \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
+    \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
+\> 1 Add: \> Zone assignment of each vertex \\
+    \> \> \> output Key = $V_i$, Value = $Z_i$ \\
+\> 1 Collate: \> Vertex as key \\
+\> 1 Reduce: \> Emit edges of each vertex with zone of vertex \\
+       \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
+       \> \> \> for each $E$ in MultiValue: \\
+      \> \> \> \> output Key = $E_{ij}$, Value = $Z_i$ \\
+
+\> 2 Collate: \> Edge as key \\
+\> 2 Reduce: \> Emit zone re-assignments \\
+       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
+       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
+       \> \> \> if $Z_i$ and $Z_j$ are different: \\
+      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
+\> 2 Exit: \> if no output by Reduce 2 \\
+\> 3 Map: \> Invert vertex/zone pairs \\
+    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
+    \> \> \> if $Z_i$ is not partitioned: \\
+   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
+    \> \> \> else: \\
+   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
+\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
+    \> \> \> if $Z_i$ is not partitioned: \\
+   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
+    \> \> \> else: \\
+   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
+   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
+\> 3 Collate: \> Zone ID as key \\
+\> 3 Reduce: \> Emit new zone assignment of each vertex \\
+       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
+       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
+       \> \> \> partition $Z_{new}$ if number of $V$ > threshhold \\
+       \> \> \> for each $V$ in MultiValue: \\
+      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$
 
@@
 
- \caption{MapReduce algorithm for R-MAT graog generation.}
+ \caption{MapReduce algorithm for finding amaximal independent set of
+ graph vertices.}
 
- \label{fig:rmat}
+ \label{fig:luby}
 \end{figure}
@@
 
- \label{fig:rmat}
+ \label{fig:sssp}
 \end{figure}

### 2010-04-15 python/mrmpi.py

@@ -107,6 +107,10 @@ class mrmpi:
     n = self.lib.MR_clone(self.mr)
     return n
 
+  def close(self):
+    n = self.lib.MR_close(self.mr)
+    return n
+
   def collapse(self,key):
     ckey = dumps(key,1)
     n = self.lib.MR_collapse(self.mr,ckey,len(ckey))
@@ -238,6 +242,10 @@ class mrmpi:
     value = loads(cvalue[:valuebytes])
     if self.map_argcount == 4: self.map_caller(itask,key,value,self)
     else: self.map_caller(itask,key,value,self,self.map_ptr)
+
+  def open(self,addflag=0):
+    if not addflag: self.lib.MR_open(self.mr)
+    else: self.lib.MR_open_add(self.mr,addflag)
 
   def print(self,proc,nstride,kflag,vflag):
     self.lib.print(self.mr,proc,nstride,kflag,vflag)

### 2010-04-15 src/cmapreduce.cpp

@@ -70,6 +70,12 @@ uint64_t MR_clone(void *MRptr)
   return mr->clone();
 }
 
+uint64_t MR_close(void *MRptr)
+{
+  MapReduce *mr = (MapReduce *) MRptr;
+  return mr->close();
+}
+
 uint64_t MR_collapse(void *MRptr, char *key, int keybytes)
 {
   MapReduce *mr = (MapReduce *) MRptr;
@@ -214,6 +220,18 @@ uint64_t MR_map_mr_add(void *MRptr, void *MRptr2,
   MapReduce *mr2 = (MapReduce *) MRptr2;
   MapFunc *appmap = (MapFunc *) mymap;
   return mr->map(mr2,appmap,APPptr,addflag);
+}
+
+void MR_open(void *MRptr)
+{
+  MapReduce *mr = (MapReduce *) MRptr;
+  return mr->open();
+}
+
+void MR_open_add(void *MRptr, int addflag)
+{
+  MapReduce *mr = (MapReduce *) MRptr;
+  return mr->open(addflag);
 }
 
 void MR_print(void *MRptr, int proc, int nstride, int kflag, int vflag)

### 2010-04-15 src/cmapreduce.h

@@ -31,6 +31,7 @@ void *MR_copy(void *MRptr);
 uint64_t MR_add(void *MRptr, void *MRptr2);
 uint64_t MR_aggregate(void *MRptr, int (*myhash)(char *, int));
 uint64_t MR_clone(void *MRptr);
+uint64_t MR_close(void *MRptr);
 uint64_t MR_collapse(void *MRptr, char *key, int keybytes);
 uint64_t MR_collate(void *MRptr, int (*myhash)(char *, int));
 uint64_t MR_compress(void *MRptr, 
@@ -82,6 +83,8 @@ uint64_t MR_map_mr_add(void *MRptr, void *MRptr2,
 		       void (*mymap)(uint64_t, char *, int, char *, int, 
 				     void *KVptr, void *APPptr),
 		       void *APPptr, int addflag);
+void MR_open(void *MRptr);
+void MR_open_add(void *MRptr, int addflag);
 void MR_print(void *MRptr, int proc, int nstride, int kflag, int vflag);
 uint64_t MR_reduce(void *MRptr,
 		   void (*myreduce)(char *, int, char *,

### 2010-04-15 src/mapreduce.cpp

@@ -497,6 +497,25 @@ uint64_t MapReduce::clone()
 }
 
 /* ----------------------------------------------------------------------
+   close a KV that KV pairs were added to by another MR's map()
+------------------------------------------------------------------------- */
+
+uint64_t MapReduce::close()
+{
+  if (kv == NULL) error->all("Cannot close without KeyValue");
+  if (timer) start_timer();
+  if (verbosity) file_stats(0);
+
+  kv->complete();
+
+  stats("Complete",0);
+
+  uint64_t nkeyall;
+  MPI_Allreduce(&kv->nkv,&nkeyall,1,MPI_UNSIGNED_LONG,MPI_SUM,comm);
+  return nkeyall;
+}
+
+/* ----------------------------------------------------------------------
    collapse KV into a KMV with a single key/value
    each proc collapses only its data
    new key = provided key name (same on every proc)
@@ -1395,6 +1414,30 @@ uint64_t MapReduce::map(MapReduce *mr,
   uint64_t nkeyall;
   MPI_Allreduce(&kv->nkv,&nkeyall,1,MPI_UNSIGNED_LONG,MPI_SUM,comm);
   return nkeyall;
+}
+
+/* ----------------------------------------------------------------------
+   open a KV so KV pairs can be added to it by another MR's map
+------------------------------------------------------------------------- */
+
+void MapReduce::open(int addflag)
+{
+  if (!allocated) allocate();
+  if (kmv) myfree(kmv->memtag);
+  delete kmv;
+  kmv = NULL;
+
+  if (addflag == 0) {
+    if (kv) myfree(kv->memtag);
+    delete kv;
+    kv = new KeyValue(this,kalign,valign,memory,error,comm);
+    kv->set_page();
+  } else if (kv == NULL) {
+    kv = new KeyValue(this,kalign,valign,memory,error,comm);
+    kv->set_page();
+  } else {
+    kv->append();
+  }
 }
 
 /* ----------------------------------------------------------------------

### 2010-04-15 src/mapreduce.h

@@ -63,6 +63,7 @@ class MapReduce {
   uint64_t add(MapReduce *);
   uint64_t aggregate(int (*)(char *, int));
   uint64_t clone();
+  uint64_t close();
   uint64_t collapse(char *, int);
   uint64_t collate(int (*)(char *, int));
   uint64_t compress(void (*)(char *, int, char *,
@@ -85,6 +86,7 @@ class MapReduce {
 				     class KeyValue *, void *),
 	       void *, int addflag = 0);
 
+  void open(int addflag = 0);
   void print(int, int, int, int);
   uint64_t reduce(void (*)(char *, int, char *,
 			   int, int *, class KeyValue *, void *),

### 2010-04-15 programs/mrmatrix2.hpp
Working but slow
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@554 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -41,7 +41,7 @@ public:
 template <typename IDTYPE>
 class MRMatrix {
   public:  
-    MRMatrix(IDTYPE, IDTYPE, MapReduce *, MapReduce *, 
+    MRMatrix(IDTYPE, IDTYPE, MapReduce *, MapReduce *, bool transpose = 0,
              int pagesize=64, const char *fpath=".");
     ~MRMatrix() {delete mr;};
 
@@ -54,11 +54,12 @@ class MRMatrix {
     MapReduce *mr;  // Actual storage; perhaps should be private.
     uint64_t nEmptyRows;  // Number of rows of A with no nonzeros.
     MapReduce *emptyRows; // Indices of rows of A with no nonzeros (leaf nodes).
+    void Print();
+    bool transposeFlag;  // State variable; indicates whether matrix is stored
+                         // as A or as A^T.
   private:
     IDTYPE N;  // Number of rows
     IDTYPE M;  // Number of cols
-    bool transposeFlag;  // State variable; indicates whether matrix is stored
-                         // as A or as A^T.
 };
 
 
@@ -73,6 +74,10 @@ class MRMatrix {
 // Given MapReduce object containing edges and vertices, 
 // create matrix non-zeros.  Give initial value of 1/(outdegree of v_i) to
 // each nonzero in row i.  
+// Store KVs by column index to make MatVec more efficient: 
+// key = j, value = (i, a_ij).
+// If transposeFlag is set, store A^T (i.e., store key=i, value=(j, a_ij)).
+// Aggregate before returning from constructor.
 // For rows with outdegree = 0, add a KV to the MapReduce object provided
 // in ptr.
 template <typename IDTYPE>
@@ -81,7 +86,9 @@ static void mrm_initialize_matrix(char *key, int keybytes,
                   KeyValue *kv, void *ptr)
 {
   // To receive entries for all-zero rows...
-  KeyValue *emptyRowsKV = (KeyValue *) ptr;
+  MRMatrix<IDTYPE> *A = (MRMatrix<IDTYPE> *) ptr;
+  KeyValue *emptyRowsKV = A->emptyRows->kv;
+  bool transposeFlag = A->transposeFlag;
 
   uint64_t totalnvalues;
   CHECK_FOR_BLOCKS(multivalue, valuebytes, nvalues, totalnvalues)
[…]

### 2010-04-15 programs/mrvector2.hpp

@@ -122,7 +122,8 @@ template <typename IDTYPE>
 void MRVector<IDTYPE>::Print()
 {
   int me = mr->my_proc();
-  cout << "Vector on processor " << me << endl;
+  cout << "Vector on processor " << me 
+       << " Local Length = " << mr->kv->nkv << endl;
   mr->map(mr, mrv_printvector<IDTYPE>, NULL, 1);  // set addflag to keep mr as is.
 }
 

### 2010-04-15 programs/test_pagerank.cpp

@@ -198,6 +198,8 @@ MPI_Barrier(MPI_COMM_WORLD);
 KDDtmp = MPI_Wtime();
 #endif //KDDTIME
 
+double tt = MPI_Wtime();
+
     // Compute adjustment for irreducibility (1-alpha)/n
     double ladj = 0.;
     double gadj = 0.;
@@ -209,10 +211,13 @@ KDDtmp = MPI_Wtime();
                                            storage_aware);
     ladj += allzeroadj;
 
+cout << "allzero rows " << MPI_Wtime() - tt << endl;
+
     // Compute global adjustment via all-reduce-like operation.
     // Cheating here!  Should be done through MapReduce.
     MPI_Allreduce(&ladj, &gadj, 1, MPI_DOUBLE, MPI_SUM, MPI_COMM_WORLD);
 
+tt = MPI_Wtime();
 #ifdef KDDTIME
 KDDallzero += (MPI_Wtime() - KDDtmp);
 MPI_Barrier(MPI_COMM_WORLD);
@@ -222,6 +227,8 @@ KDDtmp = MPI_Wtime();
     // Compute global adjustment.
     A->MatVec(mr, x, y, 1, storage_aware);
 
+cout << "matvec " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
 #ifdef KDDTIME
 KDDmatvec += (MPI_Wtime() - KDDtmp);
 MPI_Barrier(MPI_COMM_WORLD);
@@ -231,12 +238,18 @@ KDDtmp = MPI_Wtime();
     // Add adjustment to product vector in mr.
     y->AddScalar(gadj);
 
+cout << "addscalar " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
     // Compute max-norm of vector in mr.
     double gmax = y->GlobalMax(mr);
 
+cout << "globalmax " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
     // Scale vector in mr by 1/maxnorm.
     y->Scale(1./gmax);
 
+cout << "scale " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
 #ifdef KDDTIME
 KDDadjust += (MPI_Wtime() - KDDtmp);
 MPI_Barrier(MPI_COMM_WORLD);
@@ -261,6 +274,7 @@ KDDtmp = MPI_Wtime();
       mr->collate(NULL);
       mr->reduce(compute_lmax_residual, &lresid);
     }
[…]

### 2010-04-15 programs/test_pagerank2.cpp

@@ -131,9 +131,6 @@ MRVector<IDXTYPE> *pagerank(
   A->Scale(alpha);
   x->PutScalar(1./x->GlobalLen());
 
-cout << "KDDKDD After PutScalar" << endl;
-x->Print();
-
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();
 
@@ -141,6 +138,8 @@ x->Print();
   // PageRank iteration
   for (iter = 0; iter < maxniter; iter++) {
 
+MPI_Barrier(MPI_COMM_WORLD);
+double tt = MPI_Wtime();
     // Compute adjustment for irreducibility (1-alpha)/n
     double ladj = 0.;
     double gadj = 0.;
@@ -152,35 +151,44 @@ x->Print();
       allzeroadj = compute_local_allzero_adj(A, x, alpha);
     ladj += allzeroadj;
 
+cout << "allzero rows " << MPI_Wtime() - tt << endl;
+
     // Compute global adjustment via all-reduce-like operation.
     // Cheating here!  Should be done through MapReduce.
     MPI_Allreduce(&ladj, &gadj, 1, MPI_DOUBLE, MPI_SUM, MPI_COMM_WORLD);
 
+tt = MPI_Wtime();
     // Compute global adjustment.
     A->MatVec(x, y);
 
-cout << "KDDKDD After MatVec" << endl;
-y->Print();
+cout << "matvec " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
 
     // Add adjustment to product vector in mr.
     y->AddScalar(gadj);
 
-cout << "KDDKDD After AddScalar" << endl;
-y->Print();
+cout << "addscalar " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
 
     // Compute max-norm of vector in mr.
     double gmax = y->GlobalMax();
 
+cout << "globalmax " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
+
     // Scale vector in mr by 1/maxnorm.
     y->Scale(1./gmax);
 
-cout << "KDDKDD After Scale" << endl;
-y->Print();
+cout << "scale " << MPI_Wtime() - tt << endl;
+tt = MPI_Wtime();
 
     // Compute local residual; this also empties x->mr.
     double lresid = 0.;
     x->mr->add(y->mr);
     x->mr->compress(compute_lmax_residual, &lresid);
+
[…]

### 2010-04-15 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@555 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 vertices in the set share an edge.  The set is ``maximal'' if no
-additional vertex can be added to it \footnote{A maximal set is
-``maximum'' if it is the largest maximal set.  Finding a graph's
-maximum independent set is an NP-hard problem, which MapReduce is
-unlikely to help with.}.  Finding a maximal independent set is useful
-in several contexts, such as identifying large numbers of indepenent
+additional vertex can be added \footnote{A maximal set is ``maximum''
+if it is the largest maximal set.  Finding a graph's maximum
+independent set is an NP-hard problem, which MapReduce is unlikely to
+help with.}.  Finding a maximal independent set (MIS) is useful in
+several contexts, such as identifying large numbers of indepenent
 starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
-is a well-known parallel method for finding a minimal set.  Figure
-\ref{fig:luby} outlines a MapReduce version of Luby's algorithm.
+is a well-known parallel method for finding a MIS.  Figure
+\ref{fig:luby} outlines a MapReduce version of Luby's algorithm,
+assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
+V_j$ for every edge, i.e. an upper-triangular R-MAT matrix.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-Iterate: \\
-\> 1 Map: \> Convert edges to vertices \\
-    \> \> \> input Key = $E_{ij}$, Value = NULL \\
-    \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
-    \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
-\> 1 Add: \> Zone assignment of each vertex \\
-    \> \> \> output Key = $V_i$, Value = $Z_i$ \\
+Assign random values to each vertex \\
+Clone: convert edge key/values directly to key/multivalues \\
+Create empty MapReduce MRv for maximal independent set vertices \\
+\\
+while $N_{edges} > 0$: \\
+\> 1 Reduce: \> Determine winner/loser vertex of each edge \\
+       \> \> \> input Key = $E_{ij}$, Multivalue = NULLS or LOSER \\
+       \> \> \> if any value is LOSER, emit nothing \\
+       \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
+       \> \> \> output Key = $V_w$, Value = $V_l$ WINNER \\
+       \> \> \> output Key = $V_l$, Value = $V_w$ LOSER \\
 \> 1 Collate: \> Vertex as key \\
-\> 1 Reduce: \> Emit edges of each vertex with zone of vertex \\
-       \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
-       \> \> \> for each $E$ in MultiValue: \\
-      \> \> \> \> output Key = $E_{ij}$, Value = $Z_i$ \\
-
-\> 2 Collate: \> Edge as key \\
-\> 2 Reduce: \> Emit zone re-assignments \\
-       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
-       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
-       \> \> \> if $Z_i$ and $Z_j$ are different: \\
-      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
-\> 2 Exit: \> if no output by Reduce 2 \\
-\> 3 Map: \> Invert vertex/zone pairs \\
-    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
-    \> \> \> if $Z_i$ is not partitioned: \\
-   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
+\> 2 Reduce: \> Find WINNER vertices \\
+       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> \> if all $V$ are WINNERs: \\
+      \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
+      \> \> \> \> for each $V$ in Multivalue: \\
+     \> \> \> \> \> output Key = $V_j$, Value = $Vi$ LOSER \\
        \> \> \> else: \\
-   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
-\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
-    \> \> \> if $Z_i$ is not partitioned: \\
-   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
+      \> \> \> \> for each $V$ in Multivalue: \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
+\> 2 Collate: \> Vertex as key \\
+\> 3 Reduce: \> Find LOSER vertices, delete their edges \\
+       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> \> if any $V$ is LOSER: \\
+      \> \> \> \> for each $V$ in Multivalue: \\
+     \> \> \> \> \> output Key = $E_{ij}$, Value = LOSER \\
 	\> \> \> else: \\
-   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
-   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
-\> 3 Collate: \> Zone ID as key \\
-\> 3 Reduce: \> Emit new zone assignment of each vertex \\
-       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
-       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
-       \> \> \> partition $Z_{new}$ if number of $V$ > threshhold \\
-       \> \> \> for each $V$ in MultiValue: \\
-      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$
+      \> \> \> \> for each $V$ in Multivalue: \\
+     \> \> \> \> \> output Key = $E_{ij}$, Value = NULL \\
+\> 3 Collate: \> Edge as key \\
 
@@
 
- \caption{MapReduce algorithm for finding amaximal independent set of
+ \caption{MapReduce algorithm for finding a maximal independent set of
  graph vertices.}
@@
 \end{figure}
+
+Before the iterations begin, a random value is assigned to each
+vertex, which will be used to compare pairs of vertices.  The vertex
+ID itself could be used as the random value, but since the vertices
+eventually selected for the MIS depend on the randomization, this may
+introduce an unwanted bias.  Instead a random number generator can be
+used to assign a random value $R_i$ to each vertex (during a map()
+operation), which is then carried along with the vertex ID through
+each stage of the algorithm.  In the notation of Figure
+\ref{fig:luby}, each $V_i$ is then really two quantities, a vertex ID
+(1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used to relabel
+the vertices in a random manner, assigning each a new, unique vertex
+ID from 1 to $N$.  In this case, $V_i$ is simply the new vertex ID,
+and the vertices in the final MIS could be remapped to recover their
+original IDs.  These operations can be performed with a few extra
+MapReduce stages, outside the iteration loop.
+
+The pre-iteration clone() operation converts the initial list of edge
+key/values one by one to key/multivalue pairs, with the edge as the
+key, and NULL as the value.  This is so that the iteration can begin
+with a reduce() operation, without need for a collate() operation.  A
+second empty MapReduce object is also initialized, which will
+accumulate MIS vertices as the iterations proceed.
+
+At each iteration, the current set of edges are examined to identify
+winning vertices.  There are removed and added to the MIS.  Vertices
+adjacent to winners are flagged as losers, and all edges adjacent to
+losers are flagged for removal at the start of the next iteration.
+Thus each iteration removes edges from the initially full graph, and
+adds vertices to the MIS.  When the graph is empty, the MIS is
+complete.
+
+The first reduce() operation flags the two vertices in each edge as a
+potential winner or loser.  This is done by comparing the two random
+values for the vertices (or the randomized vertex IDs as explained
+above).  The first collate() operation thus produces a multivalue for
+each vertex which contains all the vertices it shares an edge with and
+associated winner/loser flags for $V_i$.  In the second reduce()
+operation, if $V_i$ won the comparison with all its edge neighbors,
+then it is a ``winner'' vertex and is emitted to the second MapReduce
+object.  All vertices connected to the winner are emitted with a loser
+flag.  If $V_i$ is not an overall winner, its edge vertices are
+emitted without the loser flag.  The second collate() operation
+collects this new information for each vertex.
+
+In the third reduce() operation, if $V_i$ was flagged as a loser by
+any edge neighbor who was a winner, then it becomes a loser.  All of
+its edges are emitted with a loser flag.  If $V_i$ was not flagged as
+a loser, its edges are emitted with no flag.  The final collate()
+operation gathers this information for each edge; each edge will have
+two values (from each vertex), which are either a loser flag or NULL.
+At the first stage of the next iteration, any edge flagged as a loser
+does not compare its two vertices; it is effectively deleted from the
+graph.
+
+The Luby iterations end when all edges have been removed from the
+graph.  At this point, the second MapReduce object contains a complete
+MIS of vertices.
+
+\subsection{Single source shortest path}
 

### 2010-04-22 programs/inferlinks.cpp
Changes to exploit new default compile-time flags. Removed many calls to set_fpath and set memsize, since we can compile the defaults into the library now. Changed my flag names to Steve's new names to be consistent with library.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@576 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -91,7 +91,7 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD, &me);
 
   if (np < 100) greetings();
-#ifdef LOCALDISK
+#ifdef MRMPI_FPATH
   // Test the file system for writing; some nodes seem to have 
   // trouble writing to local disk.
   test_local_disks();

### 2010-04-22 programs/link2graph_weighted.cpp

@@ -94,7 +94,7 @@ int main(int narg, char **args)
 
   if (nprocs < 100) greetings();
 
-#ifdef LOCALDISK
+#ifdef MRMPI_FPATH
   // On odin, test the file system for writing; some nodes seem to have 
   // trouble writing to local disk.
   test_local_disks();
@@ -234,7 +234,7 @@ int main(int narg, char **args)
     if (me == 0) {
       printf("Printing hash-key files ...\n");
       char fname[254];
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
       sprintf(fname,"%s/%s.header",MYLOCALDISK,hfile);
 #else
       sprintf(fname,"%s.header",hfile);
@@ -245,7 +245,7 @@ int main(int narg, char **args)
     }
     
     char fname[254];
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
     sprintf(fname,"%s/%s.%d",MYLOCALDISK,hfile,me);
 #else
     sprintf(fname,"%s.%d",hfile,me);
@@ -287,21 +287,21 @@ int main(int narg, char **args)
     
     char fname[254];
     FILE *fp[3];
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
     sprintf(fname,"%s/%s.srcs.%d",MYLOCALDISK,tsfile,me);
 #else
     sprintf(fname,"%s.srcs.%d",tsfile,me);
 #endif
     fp[0] = fopen(fname,"w");
 
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
     sprintf(fname,"%s/%s.dests.%d",MYLOCALDISK,tsfile,me);
 #else
     sprintf(fname,"%s.dests.%d",tsfile,me);
 #endif
     fp[1] = fopen(fname,"w");
 
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
     sprintf(fname,"%s/%s.freq.%d",MYLOCALDISK,tsfile,me);
 #else
     sprintf(fname,"%s.freq.%d",tsfile,me);
@@ -323,7 +323,7 @@ int main(int narg, char **args)
     // sort; output to files.
     // Files then must be cat'ed in order of processor number.
     // Processor ID in filename should allow that.
-#ifdef LOCALDISK
+#ifndef KEEP_OUTPUT
     sprintf(fname, "%s/%s.vmap.%04d", MYLOCALDISK, tsfile, me);
[…]

### 2010-04-22 programs/localdisks.hpp

@@ -6,23 +6,27 @@
 #define _LOCALDISKS_HPP
 
 /////////////////////////////////////////////////////////////////////////////
-// LOCALDISK can be defined in the build environment, allowing it to be
+// MRMPI_FPATH can be defined in the build environment, allowing it to be
 // specified easily on different architectures.  If it is not specified,
 // the current working directory is used. 
 // Applications can then specify MYLOCALDISK as a macro when setting 
 // the local disk for a MapReduce object.
 
-#ifdef LOCALDISK
-// This magic converts a LOCALDISK specified with a -D compiler flag
+#ifdef MRMPI_FPATH
+// This magic converts a MRMPI_FPATH specified with a -D compiler flag
 // to a string that can be passed as an argument to MapReduce methods.
 #define _QUOTEME(x) #x
 #define QUOTEME(x) _QUOTEME(x)
-#define MYLOCALDISK QUOTEME(LOCALDISK)
+#define MYLOCALDISK QUOTEME(MRMPI_FPATH)
 
 #else
 
 #define MYLOCALDISK "."
 
+#endif
+
+#ifndef MRMPI_MEMSIZE
+#define MRMPI_MEMSIZE 64
 #endif
 
 /////////////////////////////////////////////////////////////////////////////

### 2010-04-22 programs/mrmatrix2.hpp

@@ -19,6 +19,7 @@
 #include "mrvector2.hpp"
 #include "blockmacros.hpp"
 #include "mrall2.h"
+#include "localdisks.hpp"
 
 using namespace MAPREDUCE_NS;
 using namespace std;
@@ -42,7 +43,7 @@ template <typename IDTYPE>
 class MRMatrix {
   public:  
     MRMatrix(IDTYPE, IDTYPE, MapReduce *, MapReduce *, bool transpose = 0,
-             int pagesize=64, const char *fpath=".");
+             int pagesize=MRMPI_MEMSIZE, const char *fpath=MYLOCALDISK);
     ~MRMatrix() {delete mr; delete emptyRows;};
 
     IDTYPE NumRows() { return N; };
@@ -171,6 +172,7 @@ MRMatrix<IDTYPE>::MRMatrix(
   // Create MapReduce object for all-zero rows.
   MapReduce *emr = new MapReduce(MPI_COMM_WORLD);
   emr->set_fpath(fpath);
+  emr->memsize = pagesize;
   emr->map(1, mrm_do_nothing_map, NULL);
   
   // The initial call to convert incurs overhead for a mapreduce object.

### 2010-04-22 programs/mrvector2.hpp

@@ -13,6 +13,7 @@
 #include "mapreduce.h"
 #include "keyvalue.h"
 #include "mrall2.h"
+#include "localdisks.hpp"
 
 using namespace MAPREDUCE_NS;
 using namespace std;
@@ -26,7 +27,7 @@ using namespace std;
 template <typename IDTYPE>
 class MRVector {
   public:
-    MRVector(IDTYPE, int pagesize=64, const char *fpath = ".");
+    MRVector(IDTYPE, int pagesize=MRMPI_MEMSIZE, const char *fpath=MYLOCALDISK);
     ~MRVector() {delete mr;}
     void MakeEmpty();
     void PutScalar(double);

### 2010-04-22 programs/read_fb_data.hpp

@@ -160,10 +160,6 @@ void ReadFBData::run(
   MapReduce *mrraw = new MapReduce(MPI_COMM_WORLD);
   mrraw->verbosity = 0;
 
-#ifdef NEW_OUT_OF_CORE
-  mrraw->set_fpath((char *) MYLOCALDISK);
-  mrraw->memsize = MRMEMSIZE;
-#endif
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();
 

### 2010-04-22 programs/read_mm_data.hpp

@@ -121,10 +121,6 @@ void ReadMMData::run(
   // mrraw = all edges in file data
   MapReduce *mrraw = new MapReduce(MPI_COMM_WORLD);
   mrraw->verbosity = 0;
-#ifdef NEW_OUT_OF_CORE
-  mrraw->set_fpath((char *) MYLOCALDISK);
-  mrraw->memsize = MRMEMSIZE;
-#endif
 
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();

### 2010-04-22 programs/rmat.hpp

@@ -181,7 +181,6 @@ void GenerateRMAT::run(
   // Each processor generates a range of the vertices.
   if (me == 0) cout << "Generating vertices..." << endl;
   MapReduce *mrvert = new MapReduce(MPI_COMM_WORLD);
-  mrvert->set_fpath(MYLOCALDISK);
   mrvert->map(nprocs, rmat_generate_vertex, this);
   if (me == 0) cout << "Vertex aggregate..." << endl;
   mrvert->aggregate(NULL); 
@@ -192,7 +191,6 @@ void GenerateRMAT::run(
   // Now generate mredge; this is harder, as it requires the RMAT algorithm.
   if (me == 0) cout << "Generating edges..." << endl;
   MapReduce *mredge = new MapReduce(MPI_COMM_WORLD);
-  mredge->set_fpath(MYLOCALDISK);
   // Put mredge in state where it has a KeyValue (empty).
   mredge->map(1,&rmat_do_nothing,NULL);
 //  mredge->verbosity = 2;
@@ -209,7 +207,6 @@ void GenerateRMAT::run(
     ngenerate = nremain/nprocs;
     if ((unsigned) me < (nremain % nprocs)) ngenerate++;
     MapReduce mrnew(MPI_COMM_WORLD);
-    mrnew.set_fpath(MYLOCALDISK);
     mrnew.map(nprocs,&rmat_generate_edge,this);
     mrnew.aggregate(NULL);
     mredge->add(&mrnew);

### 2010-04-22 programs/shared.hpp

@@ -8,7 +8,6 @@
 #ifndef _SHARED_HPP
 #define _SHARED_HPP
 
-#define MRMEMSIZE 64
 #include <stdint.h>
 #include <iostream>
 

### 2010-04-22 programs/sssp.cpp

@@ -497,8 +497,6 @@ bool SSSP<VERTEX, EDGE>::run()
     return false;  // no unique source remains; quit execution and return.
 
   MapReduce *mrpath = new MapReduce(MPI_COMM_WORLD);
-  mrpath->set_fpath(MYLOCALDISK); 
-  mrpath->memsize = MRMEMSIZE;
   mrpath->verbosity = 0;
 
   if (me == 0) cout << counter << ": BEGINNING SOURCE " << source << endl;
@@ -587,11 +585,7 @@ bool SSSP<VERTEX, EDGE>::run()
 #else
     //  Single filename per processor; will be rewritten for each source, 
     //  so it is useful only for timings.
-#ifdef LOCALDISK
     sprintf(filename, "%s/distance.%03d", MYLOCALDISK, me);
-#else
-    sprintf(filename, "distance.%03d", me);
-#endif
 #endif
 
     ofstream fp;
@@ -621,7 +615,7 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD, &me);
 
   if (np < 100) greetings();
-#ifdef LOCALDISK
+#ifdef MRMPI_FPATH
   // Test the file system for writing; some nodes seem to have 
   // trouble writing to local disk.
   test_local_disks();

### 2010-04-22 programs/sssp_v2.cpp

@@ -552,8 +552,6 @@ bool SSSP<VERTEX, EDGE>::run()
   mrvert->map(mrvert, initialize_vertex_distances<VERTEX, EDGE>, (void *) NULL, 0);
 
   MapReduce *mrpath = new MapReduce(MPI_COMM_WORLD);
-  mrpath->set_fpath(MYLOCALDISK); 
-  mrpath->memsize = MRMEMSIZE;
 
   if (me == 0) cout << counter << ": BEGINNING SOURCE " << source << endl;
 
@@ -651,11 +649,7 @@ bool SSSP<VERTEX, EDGE>::run()
 #else
     //  Single filename per processor; will be rewritten for each source, 
     //  so it is useful only for timings.
-#ifdef LOCALDISK
     sprintf(filename, "%s/distance.%03d", MYLOCALDISK, me);
-#else
-    sprintf(filename, "distance.%03d", me);
-#endif
 #endif
 
     ofstream fp;
@@ -686,7 +680,7 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD, &me);
 
   if (np < 100) greetings();
-#ifdef LOCALDISK
+#ifdef MRMPI_FPATH
   // Test the file system for writing; some nodes seem to have 
   // trouble writing to local disk.
   test_local_disks();

### 2010-04-22 programs/stress_test1.cpp

@@ -77,9 +77,6 @@ int main(int narg, char **args)
   }
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK); 
-#endif
   mr->verbosity = 2;
   mr->timer = 1;
   mr->all2all = 1;
@@ -87,9 +84,7 @@ int main(int narg, char **args)
   int N = atoi(args[2]);
   if (N > 3) FREQ = (1 << (N-3));
   else FREQ = 1;
-#ifdef NEW_OUT_OF_CORE
   mr->memsize=atoi(args[4]);
-#endif
 
   if (me == 0) {printf("KDD: genwords...map\n"); fflush(stdout);}
   MPI_Barrier(MPI_COMM_WORLD);

### 2010-04-22 programs/stress_test2.cpp

@@ -74,18 +74,13 @@ int main(int narg, char **args)
   }
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK); 
-#endif
   mr->verbosity = 1;
 // mr->timer = 1;
 
   int N = atoi(args[2]);
   if (N > 3) FREQ = (1 << (N-3))/nprocs;
   else FREQ = 1;
-#ifdef NEW_OUT_OF_CORE
   mr->memsize=atoi(args[4]);
-#endif
 
   if (me == 0) {printf("KDD: genwords...map\n"); fflush(stdout);}
   MPI_Barrier(MPI_COMM_WORLD);
@@ -139,9 +134,7 @@ int main(int narg, char **args)
 
   if (me == 0) printf("nwords = %llu : nunique = %llu : Time = %f\n", nwords, nunique, tstop - tstart);
 
-#ifdef NEW_OUT_OF_CORE
   mr->cummulative_stats(2, 0);
-#endif
   delete mr;
 
   MPI_Finalize();

### 2010-04-22 programs/test_cc.cpp

@@ -181,9 +181,6 @@ int main(int narg, char **args)
   // find connected components via MapReduce
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK);
-#endif
   mr->verbosity = 0;
 
   if (cc.input == FILES) {

### 2010-04-22 programs/test_ccnd.cpp

@@ -67,7 +67,7 @@ int main(int narg, char **args)
   // parse command-line args
 
   cc.root = -1;
-  cc.maxmem = 100;
+  cc.maxmem = MRMPI_MEMSIZE;
   cc.input = NOINPUT;
   cc.nfiles = 0;
   cc.permute = 0;
@@ -162,10 +162,7 @@ int main(int narg, char **args)
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
   mr->verbosity = 0;
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK);
   mr->memsize = cc.maxmem;
-#endif
 
 
   if (cc.input == FILES) {

### 2010-04-22 programs/test_gs.cpp

@@ -37,9 +37,6 @@ int main(int narg, char **args)
   MPI_Barrier(MPI_COMM_WORLD);
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK);
-#endif
   mr->verbosity = 0;
 
   int gsum = MRGlobalSum(mr, me);

### 2010-04-22 programs/test_matvec.cpp

@@ -72,9 +72,6 @@ int main(int narg, char **args)
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
   mr->verbosity = 0;
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK);
-#endif
 
   // Persistent storage of the matrix. Will be loaded from files initially.
   MRMatrix A(mr, N, M, args[1]);

### 2010-04-22 programs/test_pagerank.cpp

@@ -393,7 +393,7 @@ int main(int narg, char **args)
   opterr = 0;
   const char *optstring = "a:t:m:n:g:s:";
 
-  int pagesize = 64;
+  int pagesize = MRMPI_MEMSIZE;
 
   while ((ch = getopt(narg, args, optstring)) != -1) {
     switch (ch) {
@@ -463,7 +463,6 @@ int main(int narg, char **args)
   mr->verbosity = 0;
   mr->timer = 0;
   mr->mapstyle = 1;  // mapstyle == 0 does not work for this code.
-  mr->set_fpath(MYLOCALDISK);
   mr->memsize = pagesize;
 
   // Persistent storage of the matrix. Will be loaded from files initially.

### 2010-04-22 programs/test_pagerank2.cpp

@@ -278,7 +278,7 @@ int main(int narg, char **args)
   double alpha = 0.8;
   double tolerance = 0.00001;
   int NumberOfPageranks = 1;
-  int pagesize = 64;
+  int pagesize = MRMPI_MEMSIZE;
   int filetype = RMAT;
 
   if (me == 0) 

### 2010-04-22 programs/tri_new.cpp

@@ -66,11 +66,8 @@ int main(int narg, char **args)
   strcpy(outfile,args[2]);
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-  mr->set_fpath(MYLOCALDISK);
   mr->verbosity = 1;
   mr->timer = 1;
-  //mr->memsize = 1;
-  //mr->memsize = 128;
 
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();

### 2010-04-22 programs/wordfreq.cpp

@@ -78,7 +78,7 @@ int main(int narg, char **args)
   bool postprocess = false;
   enum GenPowerLawEnum gen_flag = GEN_DEFAULT;
   int N = -1;
-  int memsize = 64;
+  int memsize = MRMPI_MEMSIZE;
   const char *optstring = "bcf:m:n:p";
 
   char ch;
@@ -133,10 +133,7 @@ int main(int narg, char **args)
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
 // mr->verbosity = 1;
 // mr->timer = 1;
-#ifdef NEW_OUT_OF_CORE
-  mr->set_fpath(MYLOCALDISK);
   mr->memsize=memsize;
-#endif
 
   MAPFN *genmap;
   int nmaptasks;
@@ -194,9 +191,7 @@ int main(int narg, char **args)
     mr->reduce(&genPowerLaw_RedistributeReduce, NULL);
   }
 
-#ifdef NEW_OUT_OF_CORE
   mr->cummulative_stats(2, 1);
-#endif
 
   MPI_Barrier(MPI_COMM_WORLD);
   double tread = MPI_Wtime();
@@ -229,17 +224,11 @@ int main(int narg, char **args)
   MPI_Barrier(MPI_COMM_WORLD);
   double tstop = MPI_Wtime();
 
-#ifdef NEW_OUT_OF_CORE
   mr->cummulative_stats(2, 0);
-#endif
 
   if (postprocess) {
     if (me == 0) {printf("Begin post-process...copy\n"); fflush(stdout);}
-#ifdef NEW_OUT_OF_CORE
     MapReduce *mrout = mr->copy();
-#else
-    MapReduce *mrout = new MapReduce(*mr);
-#endif
     if (me == 0) {printf("                  ...sort_values\n");fflush(stdout);}
     mrout->sort_values(&ncompare);
 
@@ -248,11 +237,7 @@ int main(int narg, char **args)
     count.limit = 10;
     count.flag = 0;
     if (me == 0) {printf("                  ...map\n");fflush(stdout);}
-#ifdef NEW_OUT_OF_CORE
     mrout->map(mrout, &output, &count);
-#else
-    mrout->map(mrout->kv, &output, &count);
-#endif
   
     if (me == 0) {printf("                  ...gather\n");fflush(stdout);}
     mrout->gather(1);
@@ -263,11 +248,7 @@ int main(int narg, char **args)
     count.limit = 10;
     count.flag = 1;
[…]

### 2010-04-22 paper/lessons.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@577 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Lessons Learned}

We conclude with several observations about performing MapReduce
operations on distributed-memory parallel machines via MPI.

MapReduce achieves parallelism through randomizing the distribution of
data across processors, which often intentionally ignores data
locality.  This translates into maximal data movement (during a
shuffle) with communication between all pairs of processors.  But the
benefit is often good load-balance, even for hard-to-balance irregular
data sets.  By contrast, more traditional distributed-memory parallel
algorithms, e.g. for matrix operations, or grid-based or
particle-based simulation codes, tend to work hard to localize data
and minimize communication.  To do this they typically require a lot
of application-specific logic and parallel communication coding, to
create and maintain a data decomposition, generate ghost versions of
nearby spatially-decomposed data, etc.

MapReduce algorithms can be hard to design, but are often relatively
easy to write and debug.  Thinking about a computational task from a
MapReduce perspective is different than traditional distributed-memory
parallel algorithm design.  For example, with an MPI mindset, it often
seems heretical to intentionally ignore data locality.  However,
writing small map() and reduce() functions is typically easy.  And
writing an algorithm that involves complex parallel operations,
without actually needing to write application-specific parallel code
to move data via MPI calls, is often a pleasant surprise.  Moreover,
if the MapReduce algorithm is initially coded so that it runs
correctly on one processor, it often works out-of-the-box on hundreds
or thousands of processors, without the need for additional debugging.

Performing MapReduce operations on a fixed allocation of processors on
a traditional MPI-based parallel machine is a somewhat different
conceptual model than that of cloud-computing MapReduces using (for
example) Hadoop.  In the former case, one can potentially control
which processor owns which data at various stages of an algorithm.
This is somewhat hidden from the user in a typical cloud-computing
model, where data simply exists somehere in the cloud, and Hadoop
ensures data moves where it needs to and is operated on by some
processor.  The cloud model is a nice data-centric abstraction which
allows for fault tolerance both to data loss (via redundant storage)
and to processor failure (via reassignment of work), neither of which
is typically possible on current MPI-based parallel machines.

However, since the MPI implementation of MapReduce, at least as
described in this paper, is processor-centric, one can sometimes
fruitfully exploit the possibility for processors to maintain
``state'' over the course of multiple map and reduce operations.  By
controlling where data resides for maps and reduces (e.g. via a
user-specified hash function), and by assuming that processor will
always be available, more efficient operations with less data movement
are sometimes possible.  The discussion of enhanced graph algorithms
in Section \ref{sec:graph} illustrated this.  To fully exploit this
idea for large data sets, mechanisms are needed for processors to
store, retrieve, and efficiently find needed datums on local disk
(e.g. static edges of a graph), so that archived state can be used
efficiently during subsequent map and reduce operations.

Finally, though this paper focuses on graph algorithms expressed as
MapReduce operations, there is nothing about the MR-MPI library itself
that is graph centric.  We hope the library can be generally useful on
large-scale monolithic or cloud-style parallel machines which support
MPI, for a variety of data-intense or compute-intense problems that
are amenable to solution using a MapReduce paradigm.

### 2010-04-23 programs/mrvector2.hpp
Modified vector initialization
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@579 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -67,8 +67,10 @@ static void mrv_initialize_vec(int itask, KeyValue *kv, void *ptr)
   IDTYPE share = n / nprocs;
   IDTYPE rem = n % nprocs;
   IDTYPE first = itask * share + 1;  // Matrix-market is one-based.
-  first += (itask < rem);
-  IDTYPE last = first + share + ((itask + 1) < rem);
+
+  first += MIN(itask, rem);
+
+  IDTYPE last = first + share + (itask < rem);
 
   for (IDTYPE i = first; i < last; i++)
     kv->add((char *) &i, sizeof(i), (char *)&zero, sizeof(zero));

### 2010-04-23 app/in.cc
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@580 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -1,4 +1,4 @@
-# RMAT matrix generation & statistics
+# connected-component identification
 
 variable order equal 15
 variable nz equal 2

### 2010-04-23 app/style_command.h

@@ -1,4 +1,5 @@
 #include "cc.h"
 #include "cc2.h"
+#include "luby.h"
 #include "rmat.h"
 #include "shell.h"

### 2010-04-23 programs/test_pagerank2.cpp
Removed two passes over vectors.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@581 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -52,42 +52,95 @@ using namespace std;
 #define ABS(a) ((a) >= 0. ? (a) : -1*(a));
 
 ////////////////////////////////////////////////////////////////////////////
-// Compute local contribution to adjustment for allzero rows.
+// Compute local contribution to adjustment for randomlinks and allzero rows.
 //
-// allzero_contribution reduce() function
-// Performs a dot product of emptyRows and vector x.
+struct LADJUSTMENT {
+  double randomlink;
+  double alpha_over_len;
+  double sum;
+};
+
+// adjustment_contribution reduce() function
+// Computes alpha / N * dot_product(emptyRows, x) + randomlink * x->LocalSum().
 // Re-emits emptyRows for next iteration.
 // Input:  vector entries (i, v_i) + indices of allzero rows (i, 0)
-// Output:  sum of v_j for each allzero row j; stored in ptr.
-void allzero_contribution(char *key, int keybytes, char *multivalue, 
-                          int nvalues, int *valuebytes,
-                          KeyValue *kv, void *ptr)
+// Output:  adjustment computed, stored in ptr; re-emit emptyRows.
+void adjustment_contribution(char *key, int keybytes, char *multivalue, 
+                             int nvalues, int *valuebytes,
+                             KeyValue *kv, void *ptr)
 {
-  if (nvalues == 2) {
+  struct LADJUSTMENT *adj = (struct LADJUSTMENT *) ptr;
+  assert (nvalues == 1 || nvalues == 2);
+  if (nvalues == 1) {
+    // This is not an allzero row; 
+    // multivalue has only v_j (from x).
+    // Compute adjustment due to randomlink only.
+    adj->sum += adj->randomlink * *((double*)multivalue);
+  }
+  else {
     // This is an allzero row; 
     // multivalue has 0 (from emptyRows) and v_j (from x).
+    // Compute adjustment, then re-emit the emptyRow.
     const double zero = 0.;
-    double *sum = (double *) ptr;
     double *values = (double *) multivalue;
-    *sum += (values[0] + values[1]);
+    adj->sum += (adj->alpha_over_len + adj->randomlink) * (values[0]+values[1]);
     kv->add(key, keybytes, (char *) &zero, sizeof(double));
   }
 }
 
[…]

### 2010-04-23 paper/graph.tex
Starting PageRank algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@583 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
-\section{Graph Algoriths in MapReduce}
+\section{Graph Algorithms in MapReduce}
 \label{sec:graph}
@@
 
-It models the web as a directed graph $G(V,E)$, with each vertex $v
-\in V$ representing a web page and each edge $e_{ij} \in E$
-representing a hyperlink from $v_i$ to $v_j$.  The probability of
+It models the web as a directed graph $G(V,E)$, with each vertex $v_i
+\in V$ representing a web page and each edge $E_{ij} \in E$
+representing a hyperlink from $v_i$ TO $v_j$.  The probability of
 moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
 (1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
-0.8-0.9), $d_{out}(v)$ is the outdegree of vertex $v$, and $|V|$ is
+0.8-0.9), $d_{out}(v_i)$ is the outdegree of vertex $v_i$, and $|V|$ is
 the cardinality of $V$.  The first term represents the probability of
@@
 \alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
-algorithm, then, is simply a power-method iteration in which the
+algorithm, then, is essentially a power-method iteration in which the
 dominating computation is matrix-vector multiplication $A^T x=y$,
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+Input from R-MAT generator: \\
+\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
+\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
+Initialize MapReduce object for matrix $A$  and \\
+MapReduce object $MT$ containing all vertices with no outlinks.\\
+\> Copy: $A = E$ \\
+\> Add: $V$ to $A$ \\
+\> Collate $A$: Vertex is Key \\
+\> Reduce $A$:  \\
+\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
+\> \> If (${nvalues} > 1$) \\
+\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
+\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
+\> \> else \\
+\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
+Initialize MapReduce object for pagerank vector $x$. \\
+\> Map:   \\
+\> \> Input:  Graph vertices $V$. \\
+\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
+While (${residual} > {tolerance}$) \\
+\> Compute contribution for random jumps and vertices with no outlinks. \\
+\> \> Add: $V$ to $MT$. \\
+\> \> Collate $MT$:  Vertex is Key. \\
+\> \> Reduce $MT$:   \\
+\> \> \> Input:  Key = $v_i$; MultiValue = $x_i, 0$ if $v_i \in {MT}$; MultiValue = $x_i$ if $v_i \notin {MT}$. \\
+\> \> \> Compute:  Contributions; store in data pointer. \\
+\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
+\> \> MPI\_Allreduce with MPI\_SUM to compute global contributions. \\
+\> Compute $y = A^T x$.
 

### 2010-04-28 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@587 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 4 parameters $a$, $b$, $c$, $d$ which sum to 1.0 and are discussed
-below.  The algorithm in Figure \ref{fig:rmat} generates $N_z = MN$
-unique non-zero entries in a sparse $NxN$ matrix $A$, where each entry
+below.  The algorithm in Figure \ref{fig:rmat} generates $M$ unique
+non-zero entries in a sparse $NxN$ matrix $A$, where each entry
 $A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.
@@
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> 1 Map: \> Generate $N_{\rm remain}/P$ random edges on each processor \\
+$M_{\rm remain} = M$ \\
+while $M_{\rm remain} > 0$: \\
+\> 1 Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
            \> \> output Key = $(V_i,V_j)$, Value = NULL \\
@@
               \> \> output Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+\> $M_{\rm remain} = M - N_{kv}$
 
@@
 entire map-collate-reduce sequence is repeated until the number of
-resulting key/value pairs $N_{kv}$ equals $N_z$.  For reasonably
-sparse graphs this typically takes only a few iterations.
+resulting key/value pairs $N_{kv} = M$.  For reasonably sparse graphs
+this typically takes only a few iterations.
 
@@
 
-Assign random values to each vertex \\
+Map: assign random values to each vertex \\
 Clone: convert edge key/values directly to key/multivalues \\
-Create empty MapReduce MRv for maximal independent set vertices \\
+Create: empty MapReduce MRv for maximal independent set vertices \\
 \\
 while $N_{edges} > 0$: \\
-\> 1 Reduce: \> Determine winner/loser vertex of each edge \\
-       \> \> \> input Key = $E_{ij}$, Multivalue = NULLS or LOSER \\
-       \> \> \> if any value is LOSER, emit nothing \\
+\> 1 Reduce: \> Determine WINNER/LOSER vertex of each edge \\
+       \> \> \> input Key = $E_{ij}$ with $V_i < V_j$, Multivalue = NULL or LOSER \\
+       \> \> \> if either value is LOSER, emit nothing \\
        \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
@@
        \> \> \> if all $V$ are WINNERs: \\
-      \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
       \> \> \> \> for each $V$ in Multivalue: \\
@@
 \> 2 Collate: \> Vertex as key \\
-\> 3 Reduce: \> Find LOSER vertices, delete their edges \\
+\> 3 Reduce: \> Find LOSER vertices \\
        \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
@@
       \> \> \> \> for each $V$ in Multivalue: \\
-     \> \> \> \> \> output Key = $E_{ij}$, Value = LOSER \\
+     \> \> \> \> \> output Key = $V_j$, Value = $Vi$ LOSER \\
        \> \> \> else: \\
       \> \> \> \> for each $V$ in Multivalue: \\
-     \> \> \> \> \> output Key = $E_{ij}$, Value = NULL \\
-\> 3 Collate: \> Edge as key \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
+\> 3 Collate: \> Vertex as key \\
+\> 4 Reduce: \> Emit WINNER vertices and LOSER edges \\
+       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> \> if all $V$ are LOSERs: \\
+      \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
+       \> \> \> for each $V$ in Multivalue: \\
+      \> \> \> \> if $V$ is LOSER: \\
+     \> \> \> \> \> output Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
+      \> \> \> \> else: \\
+     \> \> \> \> \> output Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
+\> 4 Collate: \> Edge as key \\
 
@@
  \caption{MapReduce algorithm for finding a maximal independent set of
- graph vertices.}
+ graph vertices via Luby's algorithm.}
 
@@
 introduce an unwanted bias.  Instead a random number generator can be
-used to assign a random value $R_i$ to each vertex (during a map()
-operation), which is then carried along with the vertex ID through
-each stage of the algorithm.  In the notation of Figure
-\ref{fig:luby}, each $V_i$ is then really two quantities, a vertex ID
-(1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used to relabel
-the vertices in a random manner, assigning each a new, unique vertex
-ID from 1 to $N$.  In this case, $V_i$ is simply the new vertex ID,
-and the vertices in the final MIS could be remapped to recover their
-original IDs.  These operations can be performed with a few extra
-MapReduce stages, outside the iteration loop.
+used to assign a consistent random value $R_i$ to each vertex in each
+edge (via a map() operation), which is then carried along with the
+vertex ID through each stage of the algorithm.  In the notation of
+Figure \ref{fig:luby}, each $V_i$ is then really two quantities, a
+vertex ID (1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used
+to relabel the vertices in a random manner, assigning each a new,
+unique vertex ID from 1 to $N$.  In this case, $V_i$ is simply the new
+vertex ID, and the vertices in the final MIS could be remapped to
+recover their original IDs.  These operations could be performed with
+a few extra MapReduce steps, outside the iteration loop.
 
@@
 key/values one by one to key/multivalue pairs, with the edge as the
-key, and NULL as the value.  This is so that the iteration can begin
-with a reduce() operation, without need for a collate() operation.  A
+key, and NULL as the value.  This is so the iterations can begin with
+a reduce() operation, without need for a collate() operation.  A
 second empty MapReduce object is also initialized, which will
@@
 At each iteration, the current set of edges are examined to identify
-winning vertices.  There are removed and added to the MIS.  Vertices
-adjacent to winners are flagged as losers, and all edges adjacent to
-losers are flagged for removal at the start of the next iteration.
-Thus each iteration removes edges from the initially full graph, and
-adds vertices to the MIS.  When the graph is empty, the MIS is
-complete.
+winning vertices.  Vertices adjacent to winners are flagged as losers,
+and all edges adjacent to losers are flagged for removal at the start
+of the next iteration.  Vertices with all losing edges are flagged as
+winners.  At each iteration edges are removed from the graph, and
+winning vertices are added to the MIS.  When the graph is empty, the
+MIS is complete.
 
@@
 above).  The first collate() operation thus produces a multivalue for
-each vertex which contains all the vertices it shares an edge with and
-associated winner/loser flags for $V_i$.  In the second reduce()
+each vertex $V_i$ which contains all the vertices it shares an edge
+with, and associated winner/loser flags.  In the second reduce()
 operation, if $V_i$ won the comparison with all its edge neighbors,
-then it is a ``winner'' vertex and is emitted to the second MapReduce
-object.  All vertices connected to the winner are emitted with a loser
-flag.  If $V_i$ is not an overall winner, its edge vertices are
-emitted without the loser flag.  The second collate() operation
-collects this new information for each vertex.
+then it is a ``winner'' vertex.  All vertices connected to the winner
+are emitted with a loser flag.  If $V_i$ is not an overall winner, its
+edge vertices are emitted without the loser flag.  The second
+collate() operation collects this new information for each vertex.
 
 In the third reduce() operation, if $V_i$ was flagged as a loser by
-any edge neighbor who was a winner, then it becomes a loser.  All of
-its edges are emitted with a loser flag.  If $V_i$ was not flagged as
-a loser, its edges are emitted with no flag.  The final collate()
-operation gathers this information for each edge; each edge will have
-two values (from each vertex), which are either a loser flag or NULL.
-At the first stage of the next iteration, any edge flagged as a loser
-does not compare its two vertices; it is effectively deleted from the
-graph.
+any edge neighbor who was a winner, then it becomes a losing vertex
+and emits all its edge vertices with a loser flag.  Otherwise it emits
+all its edge vertices without the flag.
 
-The Luby iterations end when all edges have been removed from the
-graph.  At this point, the second MapReduce object contains a complete
-MIS of vertices.
+In the fourth reduce() operation, some vertices will have all their
+edge vertices flagged as losers.  These vertices are all winners,
+either the ones identified in the second reduce() operation, or
+additional vertices who would become singletons (having no edges) when
+edges containing a loser vertex are removed (on the next iteration).
+All of these winning vertices are emitted to the MapReduce object that
+stores the accumulating MIS.
+
+All edges of each vertex are also emitted, retaining the loser flag if
+they have it.  They are emitted with $E_ij$ as their key, with $V_i <
+V_j$.  This is to insure both copies of the edge come together via the
+fourth collate() operation.  The multivalue for each edge will now
+have two values, either of which may be NULL or a loser flag.
+
+At the first stage of the next iteration, any edge flagged by either
+value as a loser does not compare its two vertices; it is effectively
+deleted from the graph.  The Luby iterations end when all edges have
+been removed from the graph.  At this point, the second MapReduce
+object contains a complete MIS of vertices.
 

### 2010-04-30 paper/graph.tex
Added improved R-MAT generating algorithm and finished PageRank algorithms.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@594 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
+By using more than one MapReduce object, we can improve the performance
+of the R-MAT generation algorithm.  In the improved algorithm
+(Figure~\ref{fig:rmat2}, we emit newly created edges into MapReduce object
+$E_{new}$ and aggregate the edges to processors.  We then add those edges
+to the MapReduce object $E_{old}$, containing all previously generated
+unique edges.  We convert the total set of key/values to key/multivalues
+and use a reduce operation to remove duplicates.  Performance improvements
+in this algorithm result from the reduced number of key/values in the aggregate
+operation, compared to the collate operation in the original algorithm, 
+particularly in later iterations of the generation where few
+edges are created.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+$M_{\rm remain} = M$ \\
+while $M_{\rm remain} > 0$: \\
+\> 1 Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
+           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
+\> 1 Aggregate $E_{new}$. \\
+\> 1 Add $E_{new}$ to $E_{old}$. \\
+\> 1 Convert $E_{old}$ from KV to KMV. \\
+\> 1 Reduce $E_{old}$: \> Remove duplicate edges \\
+              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
+\> $M_{\rm remain} = M - N_{kv}$
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{Improved MapReduce algorithm for R-MAT graph generation.}
+
+ \label{fig:rmat2}
+\end{figure}
+
 \subsection{PageRank}
@@
 
-Input from R-MAT generator: \\
-\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
-\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
-Initialize MapReduce object for matrix $A$  and \\
-MapReduce object $MT$ containing all vertices with no outlinks.\\
-\> Copy: $A = E$ \\
-\> Add: $V$ to $A$ \\
-\> Collate $A$: Vertex is Key \\
-\> Reduce $A$:  \\
-\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
-\> \> If (${nvalues} > 1$) \\
-\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
-\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
-\> \> else \\
-\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
-Initialize MapReduce object for pagerank vector $x$. \\
-\> Map:   \\
-\> \> Input:  Graph vertices $V$. \\
-\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
+%Input from R-MAT generator: \\
+%\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
+%\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
+%Initialize MapReduce object for matrix $A$  and \\
+%MapReduce object $MT$ containing all vertices with no outlinks.\\
+%\> Copy: $A = E$ \\
+%\> Add: $V$ to $A$ \\
+%\> Collate $A$: Vertex is Key \\
+%\> Reduce $A$:  \\
+%\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
+%\> \> If (${nvalues} > 1$) \\
+%\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
+%\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
+%\> \> else \\
+%\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
+%Initialize MapReduce object for pagerank vector $x$. \\
+%\> Map:   \\
+%\> \> Input:  Graph vertices $V$. \\
+%\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
+Input: \\
+\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
+\> MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
+\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
 While (${residual} > {tolerance}$) \\
-\> Compute contribution for random jumps and vertices with no outlinks. \\
-\> \> Add: $V$ to $MT$. \\
+\> Compute contribution for random jumps and zero-outdegree vertices. \\
+\> \> Add: $x$ to $MT$. \\
 \> \> Collate $MT$:  Vertex is Key. \\
 \> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $v_i$; MultiValue = $x_i, 0$ if $v_i \in {MT}$; MultiValue = $x_i$ if $v_i \notin {MT}$. \\
-\> \> \> Compute:  Contributions; store in data pointer. \\
+\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
+\> \> \> Compute:  Local contributions $cont$. \\
 \> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
-\> \> MPI\_Allreduce with MPI\_SUM to compute global contributions. \\
-\> Compute $y = A^T x$.
-
+\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
+\> Compute $y = A^T x$.  \\
+\> \> Copy $x$ to $y$. \\
+\> \> Add $A$ to $y$. \\
+\> \> Collate $y$:  Key is row index of $A$. \\
+\> \> Reduce $y$: \\
+\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
+\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
+\> \> Collate $y$:  Key is column index of $A$. \\
+\> \> Reduce $y$:  \\
+\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
+\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
+\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
+\> \> Map $y$:  \\
+\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
+\> \> \> Output:  Key = $i$, Value = $y_i$. \\
+\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
+\> Scale $y$ by $gnorm$. \\
+\> \> Map $y$: \\
+\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
+\> Compute residual. \\
+\> \> Add $y$ to $x$. \\
+\> \> Collate $x$:  Key is index $i$. \\
+\> \> Reduce $x$:   \\
+\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
+\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
+\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
+\> $x = y$; \\
+Output:  PageRank vector $x$.
   \end{tabbing}
@@
 
-The MapReduce implementation performs two {\it map()} operations to
-initialize the graph matrix $A$ and PageRank vector $x$.  A {\it
+%The MapReduce implementation performs two {\it map()} operations to
+%initialize the graph matrix $A$ and PageRank vector $x$.  
+The key kernel of PageRank is the matrix-vector multiplication $y = A^T x$.
+In it, a {\it
 collate()} operation gathers all row entries $a_{ij}$ with their
@@
 second {\it collate()} gathers, for each $j$, all contributions to the
-column sum $\sum a_{ij} x_i$, which is computed by a second {\it
+column sum $\sum_i a_{ij} x_i$, which is computed by a second {\it
 reduce()}.  MPI\_Allreduce calls are used to compute global norms and
 residuals.
+
+The entire PageRank algorithm can be made more efficient by calling
+{\it aggregate()} on the MapReduce objects representing the matrix $A$, 
+the empty matrix rows $MT$, and the vector $x$ once before the PageRank 
+iterations begin.  Since these three MapReduce
+objects have the same key types, pre-aggregating the objects moves all
+Key-Value pairs with a given key to a single processor.  Thus, many of
+the PageRank operations become local operations, replacing 
+the {\it collate()} operation (which is equivalent to {\it aggregate()} 
+followed by {\it convert()}) with only a local {\it convert()} call. 
+Figure~\ref{fig:pr2} shows the
+improved algorithm; modified operations are marked with an asterisk.
+This strategy of pre-aggregating to improve data locality is useful in
+many MapReduce algorithms, but assumes that keys are always mapped to
+the same processor.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+Input: \\
+$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
+$*$\> Pre-aggregated MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
+$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
+While (${residual} > {tolerance}$) \\
+\> Compute contribution for random jumps and zero-outdegree vertices. \\
+\> \> Add: $x$ to $MT$. \\
+$*$\> \> Convert $MT$:  Vertex is Key. \\
+\> \> Reduce $MT$:   \\
+\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
+\> \> \> Compute:  Local contributions $cont$. \\
+\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
+\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
+\> Compute $y = A^T x$.  \\
+\> \> Copy $x$ to $y$. \\
+\> \> Add $A$ to $y$. \\
+$*$\> \> Convert $y$:  Key is row index of $A$. \\
+\> \> Reduce $y$: \\
+\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
+\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
+\> \> Collate $y$:  Key is column index of $A$. \\
+\> \> Reduce $y$:  \\
+\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
+\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
+\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
+\> \> Map $y$:  \\
+\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
+\> \> \> Output:  Key = $i$, Value = $y_i$. \\
+\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
+\> Scale $y$ by $gnorm$. \\
+\> \> Map $y$: \\
+\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
+\> Compute residual. \\
+\> \> Add $y$ to $x$. \\
+$*$\> \> Convert $x$:  Key is index $i$. \\
+\> \> Reduce $x$:   \\
+\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
+\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
+\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
+\> $x = y$; \\
+Output:  PageRank vector $x$.
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{Improved MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
+operations to be replaced by with strictly local {\it convert()} operations (as marked by the asterisks).}
+
+ \label{fig:pr2}
+\end{figure}
 

### 2010-04-30 paper/graph.tex
Minor typo
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@595 3fc257bb-8550-0410-81b3-a5c103da83f7
Added SSSP algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@596 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
        \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
-       \> \> \> partition $Z_{new}$ if number of $V$ > threshhold \\
+       \> \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
        \> \> \> for each $V$ in MultiValue: \\
@@
 
-Discussion of SSSP.
+The single-source shortest path (SSSP) algorithm computes, for a directed graph
+with weighted edges,
+the shortest weighted distance from a given source vertex to
+all other vertices in the graph.  This operation is straightforward
+when global graph information is available.  The source vertex is labeled with
+distance zero.  Then in each iteration, edge adjacencies of labeled vertices
+are followed, and adjacent vertices are labeled with updated distances.  When
+no distances change, the iterations are complete.  Most importantly, only
+edges from modified vertices are visited in each iteration.
 
-Discuss new optimization in SSSP?
+In MapReduce, global information about edge adjacencies and labeled vertices
+is not available.  Instead, following the Bellman-Ford-style 
+algorithm of~\cite{SSSPMapReduce, BellmanFord},
+every edge of the graph is visited in each iteration,
+although labels for an edge's vertices are updated only if the breadth-first
+traversal has reached those vertices.  The algorithm begins with a {\it map()}
+function loading edges $(v_i, v_j)$ and source vertex $v_s$ with distance 0.
+In each iteration, the edges $(v_i, v_j)$ are collated with respect to
+$v_i$, and a {\it reduce()} operation updates each $v_j$'s distance if
+$v_i$'s distance is non-zero.  The algorithm is fully described in
+Figure~\ref{fig:sssp}.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxx\=xxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-$N_{\rm remain} = N_z$ \\
-while $N_{\rm remain} > 0$: \\
-\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
-           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> Collate() \\
-\> Reduce(): \> {\bf Remove} duplicate edges \\
-              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
-\> $N_{\rm remain} = N_z - N_{kv}$
+Input:  \\
+\> Source vertex $v_s$.\\
+\> MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$ (initially $\infty$):  \\
+\> \> Key = $v_i$; Value = $\infty$.\\
+\> MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
+\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
+Map $V$:  Set distance for source vertex $v_s$ to zero. \\
+\> Input:  Key = $v_i$; Value = $\infty$ \\
+\> Output:  Key = $v_i$; Value = 0 if $v_i = v_s$, otherwise Value = $\infty$. \\
+while not done: \\
+\> done = 1; \\
+\> Add $E$ to $V$. \\
+\> Collate $V$:  Key is vertex $v_i$. \\
+\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
+\> \> Input:  Key = $v_i$; Multivalue = edges $[v_j, w_{ij}]$, candidate distances $d_i$. \\
+\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
+\> \> \> If $d_{min} \ne d_i$, done = 0; \\
+\> \> Output:  Key = $v_i$; Value = $d_{min}$ (updated distance for $v_i$) \\
+\> \> \> and if $d_{min} \ne d_i$, Key = $v_j$; Value = $d_{min} + w_{ij}$ for each edge ($v_i, v_j$). \\
+\> MPI\_Allreduce on done with MPI\_MIN. \\
+Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.
 
@@
 \end{figure}
+
+As with PageRank and R-MAT generation, 
+significant savings in communication and execution time
+can be achieved by pre-aggregating certain MapReduce objects and 
+storing data in more than one MapReduce object.  The improved algorithm
+is shown in Figure~\ref{fig:sssp2}.

### 2010-04-30 paper/graph.tex
Added discussion of improved SSSP algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@597 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 storing data in more than one MapReduce object.  The improved algorithm
-is shown in Figure~\ref{fig:sssp2}.
+is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects storing
+vertices and edges are aggregated to processors only once at the beginning
+of the computation.  Then, in each iteration of 
+SSSP, only a small number of new candidate distances are generated.
+The improved algorithm uses MapReduce object $MR$ to store these updates.
+All communication is performed using $MR$, so the total amount of communication
+needed is small compared to the original algorithm (which aggregates all
+graph edges in each iteration).  Updated vertex distances are stored back
+in the the MapReduce object for vertices.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+Input:  \\
+\> Source vertex $v_s$.\\
+\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
+\> \> Key = $v_i$; Value = $\infty$.\\
+\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
+\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
+$U$ = new MapReduce object to store updates to distances. \\
+Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
+\> Input:  $v_s$ \\
+\> Output:  Key = $v_s$; Value = 0. \\
+while not done: \\
+\> done = 1; \\
+\> Aggregate $U$. \\
+\> Add $U$ to $V$. Empty $U$.\\
+\> Convert $V$:  Key is vertex $v_i$. \\
+\> Reduce $V$:  Pick best candidate distance for each vertex. \\
+\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
+\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
+\> \> \> If $d_{min} \ne d_i$, done = 0; \\
+\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
+\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
+\> MPI\_Allreduce on done with MPI\_MIN. \\
+\> if not done \\
+\> \> Add $U$ to $E$.  Empty $U$. \\
+\> \> Convert $E$. \\
+\> \> Reduce $E$:  Generate candidate distances for neighbors of changed vertices. \\
+\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
+\> \> \> \> and distance $d_i$ if updated in previous step. \\
+\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
+\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
+Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{MapReduce algorithm for single-source shortest path (SSSP)
+ determination.}
+
+ \label{fig:sssp2}
+\end{figure}

### 2010-04-30 paper/graph.tex
minor fix
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@598 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 SSSP, only a small number of new candidate distances are generated.
-The improved algorithm uses MapReduce object $MR$ to store these updates.
-All communication is performed using $MR$, so the total amount of communication
+The improved algorithm uses MapReduce object $U$ to store these updates.
+Only the updates are communicated; each update is aggregated to the 
+same processor that stores the corresponding vertex and its edges.
+Communication is performed using $U$, so the total amount of communication
 needed is small compared to the original algorithm (which aggregates all

### 2010-04-30 paper/results.tex
More edits
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@599 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \label{sec:results}
+
+\subsection{Comparisons to Other Implementations}
 
@@
 programs calling our MR-MPI library.
+The benchmarks were run on a medium-sized Linux cluster made up of 2 GHz
+dual-core AMD Opteron processors with a Myrinet network.  Most importantly
+for MR-MPI, each node of the cluster has one local disk, which is used
+for out-of-core operations in MR-MPI.  To avoid contention for disk I/O, we 
+ran all experiments with one MPI process per node.
 
-The benchmarks were run on a medium-sized Linux cluster.  Describe
-Odin.
+We ran each of the algorithms on three R-MAT graphs of
+different sizes, each on a varying number of processors.  Details of the
+input data are shown in Table~\ref{t:rmats}.
+The {\it small} problem size (around 134M edges) was a 
+problem that could be run
+on a single processor.  The {\it medium} problem size (around 2B
+edges) could be run on a few processors.  The {\it large} problem size
+(around 34B edges) required most of the machine to run.
+All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.
 
-We ran each of the algorithms on 3 artificial R-MAT graphs of
-different sizes, each on a varying number of processors.  The {\it
-small} problem size (around 1M edges) was a problem that could be run
-on a single processor.  The {\it medium} problem size (around 1B
-edges) could be run on a few processors.  The {\it large} problem size
-(around 1T edges) required most of the machine to run.
+\begin{table}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+Data & \# of    & \# of & Maximum \\
+Set  & vertices & edges & vertex degree\\
+\hline
+small  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
+medium &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
+large  &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
+\hline
+\end{tabular}
+\caption{Characteristics of R-MAT input data for graph algorithm
+experiments.}
+\label{t:rmats}
+\end{table}
 
@@
 Discuss SSSP times.
+
+Finally, we compare our MapReduce implementations with a multi-threaded
+implementations
+in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}, and
+distributed-memory matrix-based implementations using the linear
+algebra library Trilinos~\cite{Trilinos-Overview}.
+In the MTGL~\cite{MTGL} implementation,
+rank propagates via adjacency list traversal
+in a compressed sparse-row data structure.
+To maintain scalability, code must
+be written so that a single thread spawns the loop that processes all
+in-neighbors of a given vertex; this detail enables the compiler to generate
+hotspot-free code.
+The matrix-based distributed-memory implementation of PageRank
+uses Trilinos' Epetra matrix/vector classes to represent the graph and
+PageRank vector.
+Rows of matrix $A$ and the associated entries of the PageRank vector $x$
+are uniquely assigned to processors; a random permutation of the input
+matrix achieves processor load balance.
+Interprocessor communication gathers $x$ values for matrix-vector
+multiplication and sums partial products into the $y$ vector.
+Most communication is point-to-point communication,
+but some global communication is needed for computing
+residuals and norms of $x$ and $y$.
+
+\begin{table}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+Data & R-MAT  & R-MAT  & R-MAT  & R-MAT  & \# of    & \# of & Maximum \\
+Set  & a      & b      & c      & d      & vertices & edges & vertex degree\\
+\hline
+nice  & 0.45 & 0.15 & 0.15 & 0.25 & $2^{25}$ & $2^{28}$ & 1108 \\
+nasty & 0.57 & 0.19 & 0.19 & 0.05 & $2^{25}$ & $2^{28}$ & 230,207\\
+\hline
+\end{tabular}
+\caption{Characteristics of R-MAT input data for PageRank and Connected
+Components experiments.}
+\label{t:rmat}
+\end{table}
+
+We performed our experiments using the R-MAT data sets described in 
+Table~\ref{t:rmat}.  
+In Figure~\ref{f:prbig}, we show the performance 
+of the various PageRank
+implementations on distributed memory and multi-threaded architectures.
+The MR-MPI implementation was run on the Redstorm parallel computer;
+for these experiments,
+the data set fit in the memory of Redstorm, so out-of-core operations
+were not used.
+The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
+however, it required an order-of-magnitude
+more execution time than the matrix-based implementations on Redstorm.  
+This result is likely due to higher volume of
+communication in the MapReduce implementation (where all edges are communicated
+in each iteration), compared to the Trilinos implementation (where only
+vertex-based data are communicated).  The distributed memory matrix-based
+implementations are competitive with the multi-threaded implementation
+in MTGL on the Cray XMT.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_pagerank.pdf}
+\caption{Comparison of PageRank implementations using MapReduce,
+Trilinos, and MTGL on R-MAT data sets.}
+\label{f:prbig}
+\end{figure}

### 2010-05-03 paper/graph.tex
More edits
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@599 3fc257bb-8550-0410-81b3-a5c103da83f7
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \> 1 Add $E_{new}$ to $E_{old}$. \\
-\> 1 Convert $E_{old}$ from KV to KMV. \\
-\> 1 Reduce $E_{old}$: \> Remove duplicate edges \\
+\> 1 Compress $E_{old}$: \> Remove duplicate edges \\
               \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
@@
 \> Add $U$ to $V$. Empty $U$.\\
-\> Convert $V$:  Key is vertex $v_i$. \\
-\> Reduce $V$:  Pick best candidate distance for each vertex. \\
+\> Compress $V$:  Pick best candidate distance for each vertex. \\
 \> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
@@
 \> \> Add $U$ to $E$.  Empty $U$. \\
-\> \> Convert $E$. \\
-\> \> Reduce $E$:  Generate candidate distances for neighbors of changed vertices. \\
+\> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
 \> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
@@
 
- \caption{MapReduce algorithm for single-source shortest path (SSSP)
- determination.}
+ \caption{Improved MapReduce algorithm for single-source shortest path (SSSP)
+ determination.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}
 

### 2010-05-03 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@499 3fc257bb-8550-0410-81b3-a5c103da83f7
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 clusters.  They introduced a novel way of thinking about certain kinds
-of large-scale computations as a "map" operation followed by a
-"reduce" operation.  The power of the paradigm is that when cast in
+of large-scale computations as a ``map'' operation followed by a
+``reduce'' operation.  The power of the paradigm is that when cast in
 this way, a nominally serial algorithm now becomes two highly parallel
@@
 the intermediate data-shuffle operation is conceptually identical to
-the familiar $MPI_Alltoall$ operation.  Their implementation of
+the familiar MPI\_Alltoall operation.  Their implementation of
 MapReduce was within a Python wrapper to simplify the writing of user
@@
 a MapReduce.  Library calls for performing a map, reduce, or data
-shuffle, are synncrhonous, meaning all the processors participate and
+shuffle, are synchronous, meaning all the processors participate and
 finish the operation before proceeding.  Similarly, the use of MPI
-within the library is the traditional mode of $MPI_Send()$ and
-$MPI_Recv$ calls between processor pairs using large aggregated
+within the library is the traditional mode of MPI\_Send and
+MPI\_Recv calls between processor pairs using large aggregated
 messages to improve bandwidth performance and reduce latency costs.  A
@@
 distributed memory machines.  For serial operation, a dummy MPI
-library (provided) can be substituted.
+library (provided) can be substituted.  As a library, it can be
+embedded in other codes \cite{Titan,TitanURL} to enable them to
+perform MapReduces.
 
-\item {\it In-core or out-of-core operation:} Each MapReduce object a
-processor defines, allocates per-processor "pages" of memory, where
+\item {\it In-core or out-of-core operation:} Each MapReduce object that a
+processor defines allocates per-processor ``pages'' of memory, where
 the page size is determined by the user.  Typical MapReduce operations
@@
 MPI calls directly, if desired.  For example, one-line calls to
-$MPI_Allreduce$ are often useful in determining the status of an
+MPI\_Allreduce are often useful in determining the status of an
 iterative graph algorithm, as described in Section \ref{sec:graph}.
@@
 program to store ``state'' on each processor, accessible during the
-map and reduce operaions.  For example, various flags can be stored
+map and reduce operations.  For example, various flags can be stored
 that alter the operation of a map or reduce function, as can richer
@@
 methods in one or more MapReduce objects.  A C interface means the
-library can also be called from C or other hi-level languages such as
+library can also be called from C or other high-level languages such as
 Fortran.  A C interface also means the library can be easily wrapped
@@
 results for these algorithms running on a parallel cluster for graphs
+%KDDKDD CHECK SIZES
 ranging in size from 1 million to 1 trillion vertices or edges.  In
@@
 explicit management of parallelism, particularly for large out-of-core
-data sets.  Section \ref{sec:lessons} summarizes some lessons learned
+data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
 from the implementation and use of our library.

### 2010-05-03 paper/mr.tex
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2010-03-26 paper/mr.tex
@@
 
-A typical MR-MPI program makes at least 3 calls to the MR-MPI library,
+A typical MR-MPI program makes at least three calls to the MR-MPI library,
 to perform a {\it map()}, {\it collate()}, and {\it reduce()}
@@
 user function may be passed a file name, a chunk of bytes from a large
-file, or a task ID.  Options for assigning map tasks to processors.
+file, a task ID, or a KV pair.  Options for assigning map tasks to processors
+are specified by the user and include assigning consecutive 
+chunks of tasks to processors, striding the tasks across processors,
+or using a master-slave model that is useful when tasks have widely
+varying workloads.
 
@@
 by any processor.  Each processor hashes each of its keys to determine
-the processor who ``owns'' it.  The $N$-byte length key is hashed into
-a 32-bit value whose remainder modulo $P$ generates the owning
+which processor will ``own'' it.  The $k$-byte length key is hashed into
+a 32-bit value whose remainder modulo $P$ processors generates the owning
 processor ID.  Alternatively, the user can provide a hash function
@@
 owns.  This is done using a hash table, rather than a sort.  A sort
-requires $\log(N)$ passes through the $N$ KV pairs.  the list of KMV
-pairs can be created in 2 passes through the KV pairs, one to populate
+requires $\log(N)$ passes through the $N$ KV pairs.  With hashing,
+the list of KMV
+pairs can be created in two passes through the KV pairs, one to populate
 the hash table with needed count and length information, and the
@@
 well-formed hash table is a constant-time $O(1)$ operation, the cost
-of the data reorganization is $O(N)$, rather than $N\log(N)$ for full
-sort.
+of the data reorganization is $O(N)$, rather than $O(N\log(N))$ for full
+sorting.  It also has the added benefit that the number of values in each
+KMV pair is known and is passed to the user function during a
+reduce.  For some reduce operations, this count is all the information a
+reduce requires; thus the values need not be looped over.  This feature is not
+available in Hadoop-style data reorganizations, which sort the values; 
+with sorting, the number of values associated with the same key is not known 
+without interating over the values.
 
@@
 
-Several related MapReduce operations are provided by the library.  For
-example, the {\it compress()} operation combines on-processor KV pairs
-to eliminate duplicate keys, forming new KV pairs with an aggregated
-value.  The { \it convert()} operation turns a list of KV pairs into
-KMV pairs, with one value per key.  The {collapse()} operation turns
+Several related MapReduce operations are provided by the library and
+exposed in the library interface.  For example, the {\it collate()} function
+described above
+calls two other functions:  {\it aggregate()}, which performs the all-to-all
+communication, and {\it convert()}, which turns a list of KV pairs into
+KMV pairs.  
+Both {\it aggregate()} and {\it convert()} are exposed in the MR-MPI
+interface for direct access by the user.
+The {\it compress()} function allows on-processor operations to be 
+performed; it is equivalent to a {\it convert()} with on-processor KVs as
+input, followed by a {\it reduce()}.  
+The {\it clone()} operation turns a list of KV pairs into
+KMV pairs, with one value per key.  The {\it collapse()} operation turns
 $N$ KV pairs into one KMV pair, with the keys and values of the KV
@@
 sorting datums by key or value or within multivalues are also
-provided.  These routines invoke the C-library quicksort() function to
+provided.  These routines invoke the C-library {\it quicksort()} function to
 compute the sorted ordering of the KV pairs (or values within a
@@
 
-The purpose of providing an interface to various lo-level operations,
-is that a user program can string them together in various ways to
+The interface to various low-level operations is provided so
+that a user's program can string them together in various ways to
 produce interesting MapReduce algorithms.  For example, output from a
@@
 aggregate physical memory of large parallel machines can be multiple
-Tbytes, which allows for large data sets to be processed in-core,
+terabytes, which allows for large data sets to be processed in-core,
 assuming the KV and KMV pairs remain evenly distributed across
@@
 of randomized hashing to assign keys to processors is designed for
-such load-balancing.  What happens when data sets do not
-fit in available memory, is the subject of the next section.
+such load-balancing.  In the next section, we discuss what happens 
+when data sets do not fit in available memory.

### 2010-05-03 programs/rmat.hpp
Better management of mapreduce objects for faster rmat generation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@601 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -189,39 +189,58 @@ void GenerateRMAT::run(
   *return_mrvert = mrvert;
 
   // Now generate mredge; this is harder, as it requires the RMAT algorithm.
-  if (me == 0) cout << "Generating edges..." << endl;
-  MapReduce *mredge = new MapReduce(MPI_COMM_WORLD);
-  // Put mredge in state where it has a KeyValue (empty).
-  mredge->map(1,&rmat_do_nothing,NULL);
-//  mredge->verbosity = 2;
-//  mredge->timer = 1;
+  // Loop until desired number of unique nonzero entries
 
-  // loop until desired number of unique nonzero entries
+  if (me == 0) cout << "Generating edges..." << endl;
+
+  MapReduce *mrnew = new MapReduce(MPI_COMM_WORLD);
+  mrnew->timer = 1;
+  mrnew->verbosity = 1;
+  MapReduce *mredge = NULL;
 
   int niterate = 0;
   uint64_t one = 1;
   uint64_t ntotal = (one << nlevels) * avgdeg;
   uint64_t nremain = ntotal;
+
   while (nremain) {
     niterate++;
     ngenerate = nremain/nprocs;
     if ((unsigned) me < (nremain % nprocs)) ngenerate++;
-    MapReduce mrnew(MPI_COMM_WORLD);
-    mrnew.map(nprocs,&rmat_generate_edge,this);
-    mrnew.aggregate(NULL);
-    mredge->add(&mrnew);
+    if (me == 0) {cout << "  map rmat_generate_edge..." << endl; flush(cout);}
+    mrnew->map(nprocs,&rmat_generate_edge,this);
+    if (me == 0) {cout << "  aggregate..." << endl; flush(cout);}
+    mrnew->aggregate(NULL);
+    if (me == 0) {cout << "  add..." << endl; flush(cout);}
+    if (niterate > 1) {
+      mredge->add(mrnew);
+      mrnew->map(1, rmat_do_nothing, NULL);  // Empty mrnew.
+    }
+    else {  // First iteration -- no need to add; just set the pointer.
+      mredge = mrnew;
+      mrnew = new MapReduce(MPI_COMM_WORLD);
+      mrnew->timer = 1;
+      mrnew->verbosity = 1;
+    }
+    if (me == 0) {cout << "  convert..." << endl; flush(cout);}
     uint64_t nunique = mredge->convert();
     if (nunique == ntotal) break;
[…]

### 2010-05-03 programs/test_pagerank2.cpp

@@ -327,6 +327,12 @@ int main(int narg, char **args)
   MPI_Comm_size(MPI_COMM_WORLD,&np);
   if (me == 0) {cout << "Here we go..." << endl; flush(cout);}
 
+#ifdef MRMPI_FPATH
+  // Test the file system for writing; some nodes seem to have 
+  // trouble writing to local disk.
+  test_local_disks();
+#endif
+
   // Default parameters
   double alpha = 0.8;
   double tolerance = 0.00001;

### 2010-05-03 paper/graph.tex
More edits; consistent notation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@602 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 graphs with power-law degree distributions.  They are commonly used to
-represent web and social networks.  The user specifices 6 parameters
-which define the graph: the number of vertices $N$ and edges $M$, and
-4 parameters $a$, $b$, $c$, $d$ which sum to 1.0 and are discussed
+represent web and social networks.  The user specifices six parameters
+that define the graph: the number of vertices $N$ and edges $M$, and
+four parameters $a$, $b$, $c$, $d$ that sum to 1.0 and are discussed
 below.  The algorithm in Figure \ref{fig:rmat} generates $M$ unique
-non-zero entries in a sparse $NxN$ matrix $A$, where each entry
+non-zero entries in a sparse $N \times N$ matrix $A$, where each entry
 $A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.
@@
 while $M_{\rm remain} > 0$: \\
-\> 1 Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
-           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
-\> 1 Collate \\
-\> 1 Reduce: \> Remove duplicate edges \\
-              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
+\> Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
+           \> \> output Key = $(i,j)$, Value = NULL \\
+\> Collate \\
+\> Reduce: \> Remove duplicate edges \\
+           \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
+           \> \> output Key = $(i,j)$, Value = NULL \\
 \> $M_{\rm remain} = M - N_{kv}$
@@
 
-In the map() operation, each of $P$ processors generates a $1/P$
-fraction of the desired edges.  A single random $i,j$ edge is computed
+In the {\it map()} operation, each of $P$ processors generates a $1/P$
+fraction of the desired edges.  A single random edge ($i,j$) is computed
 recursively as follows.  Pick a random quadrant of the $A$ matrix with
@@
 
-The map() will often generate some small number of duplicate edges.
-The collate() and reduce() operations remove the duplicates.  The
+The {\it map()} will often generate some small number of duplicate edges.
+The {\it collate()} and {\it reduce()} operations remove the duplicates.  The
 entire map-collate-reduce sequence is repeated until the number of
@@
 the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
-values is larger than the other 3, a highly skewed distribution
+values is larger than the other three, a highly skewed distribution
 results.  Variants of the above algorithm can be used when $N$ is not
@@
 numeric value to the $A_{ij}$ edge), graphs without self edges
-(require $i != j$), or graphs with undirected edges (require $i < $j).
+(require $i \ne j$), or graphs with undirected edges (require $i < j$).
 Or the general R-MAT matrix can be further processed by MapReduce
@@
 of the R-MAT generation algorithm.  In the improved algorithm
-(Figure~\ref{fig:rmat2}, we emit newly created edges into MapReduce object
-$E_{new}$ and aggregate the edges to processors.  We then add those edges
+(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce object
+$E_{new}$ and {\it aggregate()} the edges to processors.  
+We then add those edges
 to the MapReduce object $E_{old}$, containing all previously generated
-unique edges.  We convert the total set of key/values to key/multivalues
-and use a reduce operation to remove duplicates.  Performance improvements
-in this algorithm result from the reduced number of key/values in the aggregate
-operation, compared to the collate operation in the original algorithm, 
+unique edges; the edges of $E_{old}$ are already aggregated to processors
+using the same mapping as $E_{new}$.  We can then use the {\it compress()}
+function to reduce duplicates locally on each processor.
+Performance improvements
+in this algorithm result from the reduced number of KV pairs communicated in
+{\it aggregate()},
+compared to the collate operation in the original algorithm, 
 particularly in later iterations of the generation where few
@@
 while $M_{\rm remain} > 0$: \\
-\> 1 Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
-           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
-\> 1 Aggregate $E_{new}$. \\
-\> 1 Add $E_{new}$ to $E_{old}$. \\
-\> 1 Compress $E_{old}$: \> Remove duplicate edges \\
-              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
-              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
+\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
+           \> \> output Key = $(i,j)$, Value = NULL \\
+\> Aggregate $E_{new}$. \\
+\> Add $E_{new}$ to $E_{old}$. \\
+\> Compress $E_{old}$: Remove duplicate edges \\
+            \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
+            \> \> output Key = $(i,j)$, Value = NULL \\
 \> $M_{\rm remain} = M - N_{kv}$
@@
 vertex in a graph.
-
-It models the web as a directed graph $G(V,E)$, with each vertex $v_i
-\in V$ representing a web page and each edge $E_{ij} \in E$
-representing a hyperlink from $v_i$ TO $v_j$.  The probability of
-moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
+It models the web as a directed graph $G(V,E)$, with each vertex $V_i
+\in V$ representing a web page and each edge $(V_i, V_j) \in E$
+representing a hyperlink from $V_i$ to $V_j$.  The probability of
+moving from $V_i$ to another vertex $V_j$ is $\alpha/d_{out}(V_i) +
 (1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
-0.8-0.9), $d_{out}(v_i)$ is the outdegree of vertex $v_i$, and $|V|$ is
+0.8-0.9), $d_{out}(V_i)$ is the outdegree of vertex $V_i$, and $|V|$ is
 the cardinality of $V$.  The first term represents the probability of
-following a given link on page $v_i$; the second represents the
+following a given link on page $V_i$; the second represents the
 probability of moving to a random page.  For pages with no outlinks,
@@
 matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
-\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
+\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  The PageRank
 algorithm, then, is essentially a power-method iteration in which the
@@
 where $x$ is the PageRank vector from the previous iteration.
-
 The algorithm in Figure \ref{fig:pr} performs these
@@
 %Input from R-MAT generator: \\
-%\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
-%\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
+%\> MapReduce object $V$ containing all graph vertices:  Key=$V_i$, Value = NULL.  \\
+%\> MapReduce object $E$ containing all graph edges: Key=$V_i$, Value = $V_j$. \\
 %Initialize MapReduce object for matrix $A$  and \\
@@
 %\> Reduce $A$:  \\
-%\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
+%\> \> Input:  Key = $V_i$; MultiValue = NULL,$V_j$,$V_k$,... \\
 %\> \> If (${nvalues} > 1$) \\
-%\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
-%\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
+%\> \> \> Compute: $d_{out}(V_i) = {nvalues} - 1$. \\
+%\> \> \> Output:  For each $V_j$ in MultiValue, Key = $V_i$; Value = ($V_j$, $\alpha/d_{out}(V_i)$). \\
 %\> \> else \\
-%\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
+%\> \> \> Output to $MT$: Key = $V_i$; Value = 0. \\
 %Initialize MapReduce object for pagerank vector $x$. \\
@@
 %\> \> Input:  Graph vertices $V$. \\
-%\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
+%\> \> Output:  For each vertex, Key = $V_i$; Value = $1/|V|$. \\
 Input: \\
@@
 \> \> \> Compute:  Local contributions $cont$. \\
-\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
+\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
 \> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
@@
 traversal has reached those vertices.  The algorithm begins with a {\it map()}
-function loading edges $(v_i, v_j)$ and source vertex $v_s$ with distance 0.
-In each iteration, the edges $(v_i, v_j)$ are collated with respect to
-$v_i$, and a {\it reduce()} operation updates each $v_j$'s distance if
-$v_i$'s distance is non-zero.  The algorithm is fully described in
+function loading edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.
+In each iteration, the edges $(V_i, V_j)$ are collated with respect to
+$V_i$, and a {\it reduce()} operation updates each $V_j$'s distance if
+$V_i$'s distance is non-zero.  The algorithm is fully described in
 Figure~\ref{fig:sssp}.
@@
 Input:  \\
-\> Source vertex $v_s$.\\
-\> MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$ (initially $\infty$):  \\
-\> \> Key = $v_i$; Value = $\infty$.\\
-\> MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
-\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
-Map $V$:  Set distance for source vertex $v_s$ to zero. \\
-\> Input:  Key = $v_i$; Value = $\infty$ \\
-\> Output:  Key = $v_i$; Value = 0 if $v_i = v_s$, otherwise Value = $\infty$. \\
+\> Source vertex $V_s$.\\
+\> MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$ (initially $\infty$):  \\
+\> \> Key = $V_i$; Value = $\infty$.\\
+\> MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
+\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
+Map $V$:  Set distance for source vertex $V_s$ to zero. \\
+\> Input:  Key = $V_i$; Value = $\infty$ \\
+\> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$. \\
 while not done: \\
@@
 \> Add $E$ to $V$. \\
-\> Collate $V$:  Key is vertex $v_i$. \\
+\> Collate $V$:  Key is vertex $V_i$. \\
 \> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
-\> \> Input:  Key = $v_i$; Multivalue = edges $[v_j, w_{ij}]$, candidate distances $d_i$. \\
-\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
+\> \> Input:  Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$, candidate distances $d_i$. \\
+\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
 \> \> \> If $d_{min} \ne d_i$, done = 0; \\
-\> \> Output:  Key = $v_i$; Value = $d_{min}$ (updated distance for $v_i$) \\
-\> \> \> and if $d_{min} \ne d_i$, Key = $v_j$; Value = $d_{min} + w_{ij}$ for each edge ($v_i, v_j$). \\
+\> \> Output:  Key = $V_i$; Value = $d_{min}$ (updated distance for $V_i$) \\
+\> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$). \\
 \> MPI\_Allreduce on done with MPI\_MIN. \\
-Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.
+Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
 
@@
 Input:  \\
-\> Source vertex $v_s$.\\
-\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
-\> \> Key = $v_i$; Value = $\infty$.\\
-\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
-\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
+\> Source vertex $V_s$.\\
+\> Pre-aggregated MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$:  \\
+\> \> Key = $V_i$; Value = $\infty$.\\
+\> Pre-aggregated MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
+\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
 $U$ = new MapReduce object to store updates to distances. \\
-Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
-\> Input:  $v_s$ \\
-\> Output:  Key = $v_s$; Value = 0. \\
+Map $U$:  Add $V_s$ with distance $d_s = 0$ to $U$. \\
+\> Input:  $V_s$ \\
+\> Output:  Key = $V_s$; Value = 0. \\
 while not done: \\
@@
 \> Compress $V$:  Pick best candidate distance for each vertex. \\
-\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
-\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
+\> \> Input:  Key = $V_i$; Multivalue = candidate distances $d_i$. \\
+\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
 \> \> \> If $d_{min} \ne d_i$, done = 0; \\
-\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
-\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
+\> \> Output to $V$:  Key = $V_i$; Value = $d_{min}$. \\
+\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$. \\
 \> MPI\_Allreduce on done with MPI\_MIN. \\
@@
 \> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
-\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
+\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ for edges ($V_i, V_j$) \\
 \> \> \> \> and distance $d_i$ if updated in previous step. \\
 \> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
-\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
-Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.
+\> \> \> \> Key = $V_j$; Value = $d_i+w_{ij}$ for each edge ($V_i, V_j$). \\
+Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
 

### 2010-05-03 paper/outcore.tex
More edits; consistent notation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@602 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 ``out-of-core'' processing, which is triggered when the KV or KMV
-pairs owned by oen or more processors do not fit in its local memory.
+pairs owned by one or more processors do not fit in local memory.
 When this occurs, a processor writes one or more temporary files to
@@
 specified, which defaults to 64 Mbytes.  As described below, each
-MR-MPI operation, is constrained to use no more than a handful of
+MR-MPI operation is constrained to use no more than a handful of
 these pages.  The {\it pagesize} setting can be as small as 1 Mbyte or
-as large as desired, though the user should insure the allocated pages
-fit in physical memory, else the MR-MPI library may allocate slow
+as large as desired, though the user should ensure the allocated pages
+fit in physical memory; otherwise, the MR-MPI library may allocate slow
 virtual memory.  The {\it pagesize} is also the typical size of
 individual reads and writes to the temporary disk files; hence a
-reasonable {\it pagesize} insures good I/O performance.
+reasonable {\it pagesize} ensures good I/O performance.
 
@@
 hashed to ``owning'' processors and the KV datums are communicated to
-new processors in an all-to-all fashion.  In out-of-core mode, this is
-done one page of KV data at a time.  Each processor reads in a page of
-KV pairs, the communication pattern for the datums in that page is
-determined (who needs what datums), and all-to-all communication is
+new processors in an all-to-all fashion.  In out-of-core mode, this 
+operation is
+performed on one page of KV data at a time.  Each processor reads in a page of
+KV pairs, the communication pattern (which processors will receive 
+which datums in that page) is
+determined, and all-to-all communication is
 performed.  Each processor allocates a two-page chunk of memory to
@@
 all-to-all communication, including the source and target KV pages is
-at most 7.
+at most seven.
 
 When the communication stage of the {\it collate()} operation is
-complete, each processor now has a set of KV pages, stored in an
+complete, each processor has a set of KV pages, stored in an
 out-of-core KV file, that need to be reorganized into a set of KMV
-pages, also stored in new out-of-core KMV file.  In principle, to
-create one page of KMV pairs, would require all KV pages be scanned,
-to find all the keys which contribute values to that page.  Doing this
+pages, stored in a new out-of-core KMV file.  In principle, 
+creating one page of KMV pairs would require all KV pages be scanned
+to find all the keys that contribute values to that page.  Doing this
 for each output page of KMV pairs could be prohibitively expensive.  A
@@
 arbitrarily large for data sets with many unique keys.  We need an
-algorithm for generating KMV pairs which operates within a small
+algorithm for generating KMV pairs that operates within a small
 number of memory pages and performs a minimal number of passes through
 the KV disk file.  An algorithm that meets these goals is diagrammed
-in Figure \ref{fig:collate}; it reads the KV pages at most 4 times, and
-writes out new KV or KMV pages at most 3 times.  It also uses a
+in Figure \ref{fig:collate}; it reads the KV pages at most four times, and
+writes out new KV or KMV pages at most three times.  It also uses a
 finite-size in-memory hash table, which stores unique keys as they are
@@
 \item (Pass 3) The unique keys in the HT for one partition are
-scanned.  The associated values may create KMV pairs which will span
+scanned to determine the total size of KMV for keys in the HT.  
+The associated values may create KMV pairs that will span
 $M$ memory pages.  If $M > 1$ the associated partition file of KV
@@
 smaller ``set'' files, represented by KVs in the figure.  Each set
-file contains the KV pairs which will contribute to the KMV pairs that
+file contains the KV pairs that will contribute to the KMV pairs that
 will populate one page of KMV output.  If a single KMV pair spans
@@
 is somewhat complex, but requires only a small, constant number of
-passes through the data.  The KV datums are read from disk at most 4
-times, and written to disk 3 times.  The first two write passes
+passes through the data.  The KV datums are read from disk at most four
+times, and written to disk three times.  The first two write passes
 reorganize KV pairs into partition and set files; the final write pass
 creates the KMV data set.  Depending on the size and characteristics
-of the KV datums (e.g. how many unique keys it contains), some of
+of the KV datums (e.g., the number of unique keys), some of
 these passes may not be needed.  By contrast, a full out-of-core sort
@@
 allocated for each file.  The precise number of simultaneously open
-files is difficult to bound, but typically one or two additional pages
-memory suffice for all needed buffers.  Thus the total pages required
+files is difficult to bound, but typically one or two additional pages'
+memory suffice for all needed buffers.  Thus the total number of pages required
 for the data reorganization stage of the {\it collate()} operation is
-less than the 7 used by the all-to-all communication stage.
+less than the seven used by the all-to-all communication stage.
 
 Other MR-MPI library calls discussed in the previous section, such as
-{\it compress()}, {\it convert()}, {\it collapse()}, {\it gather()},
-etc, can also operate in out-of-core mode, typically with one pass
+{\it compress()}, {\it convert()}, {\it collapse()}, and {\it gather()},
+can also operate in out-of-core mode, typically with one pass
 through the KV or KMV data and the use of one or two memory pages.
@@
 and merged into a new file which is written to disk as its associated
-memory page fills up.  This process requires 5 memory pages, which
+memory page fills up.  This process requires five memory pages, which
 includes vectors of KV datum pointers and lengths needed by the
@@
 files into a new third file, without the need to perform additional
-in-memory sorts.  Thus the overall memory requirement is 5 pages.  The
+in-memory sorts.  Thus the overall memory requirement is five pages.  The
 number of read/write passes through the KV data set for the merge sort

### 2010-05-04 paper/graph.tex
Minor
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@603 3fc257bb-8550-0410-81b3-a5c103da83f7
Modified pagerank description to make it more like other algorithms.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@604 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
-\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  The PageRank
-algorithm, then, is essentially a power-method iteration in which the
-dominating computation is matrix-vector multiplication $A^T x=y$,
-where $x$ is the PageRank vector from the previous iteration.
+\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  To maintain the 
+sparsity of $A$, terms for random jumps and zero out-degree vertices are
+not explicitly stored in $A$.  Rather, they are computed separately
+as adjustments to the PageRank vector.  The kernel of the PageRank
+algorithm, then, is a power-method iteration consisting of 
+matrix-vector multiplication $A^T x=y$,
+where $x$ is the PageRank vector from the previous iteration, and
+a few dot product and norm computations.
+
 The algorithm in Figure \ref{fig:pr} performs these
-iterations.
+iterations.  The nonzeros of matrix $A$ are initialized as described above;
+the PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
+contains the indices of all-zero rows of $A$, corresponding to vertices
+with no outlinks.  In step 1 of a PageRank interation, $MT$ is used 
+in dot products to compute adjustments to the
+PageRank vector to represent uniform probability of going from a leaf page
+to any other page.  Adjustments for random jumps from any page are also
+computed in this step.  Step 2, the matrix-vector product $A^T x$, is the most
+expensive part of the computation, requiring two passes over the nonzero
+structure of $A$.  In the first pass, {\it collate()} 
+gathers all row entries $a_{ij}$ with their
+associated $x_i$ entry, and {\it reduce()} computes $a_{ij} x_i$.  
+In the second pass, 
+{\it collate()} gathers, for each $j$, all contributions to the
+column sum $\sum_i a_{ij} x_i$, which is computed by a second {\it reduce()}. 
+Steps 3 and 4 adjust and scale the 
+product vector; the work is proportional to $|V|$.  Step 6 computes the
+residual to determine whether the PageRank vector has converged; work is
+again proportional to $|V|$.  Throughout the algorithm, we exploit the
+ability to call MPI\_Allreduce to compute global norms and residuals.
 
@@
 \> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
-\> MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
+\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
 \> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
 While (${residual} > {tolerance}$) \\
-\> Compute contribution for random jumps and zero-outdegree vertices. \\
+\> 1. Compute contribution for random jumps and zero-outdegree vertices. \\
 \> \> Add: $x$ to $MT$. \\
@@
 \> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
-\> \> \> Compute:  Local contributions $cont$. \\
+\> \> \> Compute:  Local contributions $c_{local}$. \\
 \> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
-\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
-\> Compute $y = A^T x$.  \\
+\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
+\> 2. Compute $y = A^T x$.  \\
 \> \> Copy $x$ to $y$. \\
@@
 \> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
-\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
+\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
 \> \> Map $y$:  \\
 \> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
+\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$. \\
 \> \> \> Output:  Key = $i$, Value = $y_i$. \\
-\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
-\> Scale $y$ by $gnorm$. \\
+\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
+\> 4. Scale $y$ by $n_{global}$. \\
 \> \> Map $y$: \\
 \> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
-\> Compute residual. \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
+\> 5. Compute residual. \\
 \> \> Add $y$ to $x$. \\
@@
 \> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
-\> $x = y$; \\
+\> 6. $x = y$; \\
 Output:  PageRank vector $x$.
@@
 \end{figure}
-
-%The MapReduce implementation performs two {\it map()} operations to
-%initialize the graph matrix $A$ and PageRank vector $x$.  
-The key kernel of PageRank is the matrix-vector multiplication $y = A^T x$.
-In it, a {\it
-collate()} operation gathers all row entries $a_{ij}$ with their
-associated $x_i$ entry, and a {\it reduce()} computes $a_{ij} x_i$.  A
-second {\it collate()} gathers, for each $j$, all contributions to the
-column sum $\sum_i a_{ij} x_i$, which is computed by a second {\it
-reduce()}.  MPI\_Allreduce calls are used to compute global norms and
-residuals.
 
@@
 \> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
-\> \> \> Compute:  Local contributions $cont$. \\
+\> \> \> Compute:  Local contributions $c_{local}$. \\
 \> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
-\> \> MPI\_Allreduce with MPI\_SUM to compute global contribution $gcont$. \\
+\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
 \> Compute $y = A^T x$.  \\
@@
 \> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
-\> Add contribution $gcont$ to $y$; compute local max norm $norm$ of $y$. \\
+\> Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
 \> \> Map $y$:  \\
 \> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Compute: $y_i = y_i + gcont$; if $y_i > {norm}$, $norm = y_i$. \\
+\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$. \\
 \> \> \> Output:  Key = $i$, Value = $y_i$. \\
-\> \> MPI\_Allreduce $norm$ with MPI\_MAX to compute global max norm $gnorm$. \\
-\> Scale $y$ by $gnorm$. \\
+\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
+\> Scale $y$ by $n_{global}$. \\
 \> \> Map $y$: \\
 \> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {gnorm}$. \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
 \> Compute residual. \\
@@
 The initial step is to store a copy of the graph edges as key/value
-(KV) pairs in a auxiliary MapReduce object $G_0$, for use later in the
+(KV) pairs in an auxiliary MapReduce object $G_0$, for use later in the
 algorithm.  The first {\it map()} operation converts edge keys to

### 2010-05-04 paper/graph.tex
More edits.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@605 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 four parameters $a$, $b$, $c$, $d$ that sum to 1.0 and are discussed
-below.  The algorithm in Figure \ref{fig:rmat} generates $M$ unique
+below.  The algorithm in Figure~\ref{fig:rmat} generates $M$ unique
 non-zero entries in a sparse $N \times N$ matrix $A$, where each entry
@@
 By using more than one MapReduce object, we can improve the performance
-of the R-MAT generation algorithm.  In the improved algorithm
+of the R-MAT generation algorithm.  In the enhanced algorithm
 (Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce object
@@
 
- \caption{Improved MapReduce algorithm for R-MAT graph generation.}
+ \caption{Enhanced MapReduce algorithm for R-MAT graph generation.}
 
@@
 
-The algorithm in Figure \ref{fig:pr} performs these
+The algorithm in Figure~\ref{fig:pr} performs these
 iterations.  The nonzeros of matrix $A$ are initialized as described above;
@@
 {\it aggregate()} on the MapReduce objects representing the matrix $A$, 
-the empty matrix rows $MT$, and the vector $x$ once before the PageRank 
+the all-zero matrix rows $MT$, and the vector $x$ once before the PageRank 
 iterations begin.  Since these three MapReduce
@@
 Figure~\ref{fig:pr2} shows the
-improved algorithm; modified operations are marked with an asterisk.
+enhanced algorithm; modified operations are marked with an asterisk.
 This strategy of pre-aggregating to improve data locality is useful in
@@
 $*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
-$*$\> Pre-aggregated MapReduce object $MT$ containing indices of empty rows of $A$:  Key = $i$, Value = 0. \\
+$*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
 $*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
@@
 
- \caption{Improved MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
+ \caption{Enhanced MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
 operations to be replaced by with strictly local {\it convert()} operations (as marked by the asterisks).}
@@
 of vertices in the set are connected by a path of edges.  A sparse
-graph may contain many such components.  Figure \ref{fig:cc} outlines
+graph may contain many such components.  Figure~\ref{fig:cc} outlines
 a MapReduce algorithm that labels each vertex in a graph with a
@@
 The algorithm begins (before the iteration loop) by assigning each
-vertex to its own component or ``zone'', so that $Z_i$ = $V_i$.  Each
+vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$.  Each
 iteration will grow the zones, one layer of neighbors at a time.  As
@@
 If the two zone IDs are different, the second {\it reduce()} operation
-chooses a winner (the min of the 2 IDs), and emits the loser ID as a
+chooses a winner (the smaller of the two IDs), and emits the loser ID as a
 key, with the winning ID as a value.  If no zone ID changes are
@@
 adds the changing zone assignments to the set of key/value pairs.  The
-third collate() can then collects all the vertices of a zone and zero
+third collate() can then collect all the vertices of a zone and zero
 or more reassignments for the zone ID.  Since a zone could collide
@@
 
-Note that if a graph has only a few components, then the third {\it
-collate()} operation, which keys on the zone ID, may generate a few
+Note that if a graph has only a few components, the third {\it
+collate()} operation, which collates by the zone ID, may generate a few
 very large key/multi-value (KMV) pairs.  For example, if the graph is
@@
 processor ID bits are added to the zone ID, as indicated by the
-$Z_i^+$ notation in Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been
+$Z_i^+$ notation in Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been
 partitioned, the third {\it add()} operation emits the zone ID change
@@
 operation emits $P+1$ copies of new zone reassignments for both
-partitioned and unpartioned zone IDs, to insure all vertices in the
+partitioned and unpartioned zone IDs, to ensure all vertices in the
 zone will know the reassignment information.
@@
 Clone: convert edge key/values directly to key/multivalues \\
-Create: empty MapReduce MRv for maximal independent set vertices \\
+Create: empty MapReduce object MRv for maximal independent set vertices \\
 \\
@@
        \> \> \> if all $V$ are WINNERs: \\
-      \> \> \> \> for each $V$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $Vi$ LOSER \\
+      \> \> \> \> for each $V_j$ in Multivalue: \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ LOSER \\
        \> \> \> else: \\
-      \> \> \> \> for each $V$ in Multivalue: \\
+      \> \> \> \> for each $V_j$ in Multivalue: \\
      \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
@@
        \> \> \> if any $V$ is LOSER: \\
-      \> \> \> \> for each $V$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $Vi$ LOSER \\
+      \> \> \> \> for each $V_j$ in Multivalue: \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ LOSER \\
        \> \> \> else: \\
-      \> \> \> \> for each $V$ in Multivalue: \\
+      \> \> \> \> for each $V_j$ in Multivalue: \\
      \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
@@
       \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
-       \> \> \> for each $V$ in Multivalue: \\
-      \> \> \> \> if $V$ is LOSER: \\
+       \> \> \> for each $V_j$ in Multivalue: \\
+      \> \> \> \> if $V_j$ is LOSER: \\
      \> \> \> \> \> output Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
@@
 used to assign a consistent random value $R_i$ to each vertex in each
-edge (via a map() operation), which is then carried along with the
+edge (via a {\it map()} operation), which is then carried along with the
 vertex ID through each stage of the algorithm.  In the notation of
-Figure \ref{fig:luby}, each $V_i$ is then really two quantities, a
+Figure~\ref{fig:luby}, each $V_i$ is then really two quantities, a
 vertex ID (1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used
@@
 
-The pre-iteration clone() operation converts the initial list of edge
+The pre-iteration {\it clone()} operation converts the initial list of edge
 key/values one by one to key/multivalue pairs, with the edge as the
-key, and NULL as the value.  This is so the iterations can begin with
-a reduce() operation, without need for a collate() operation.  A
+key, and NULL as the value.  Thus, the iterations can begin with
+a {\it reduce()} operation, without need for a {\it collate()} operation.  A
 second empty MapReduce object is also initialized, which will
@@
 
-At each iteration, the current set of edges are examined to identify
+At each iteration, the current set of edges is examined to identify
 winning vertices.  Vertices adjacent to winners are flagged as losers,
@@
 
-The first reduce() operation flags the two vertices in each edge as a
+The first {\it reduce()} operation flags the two vertices in each edge as a
 potential winner or loser.  This is done by comparing the two random
 values for the vertices (or the randomized vertex IDs as explained
-above).  The first collate() operation thus produces a multivalue for
+above).  The first {\it collate()} operation thus produces a multivalue for
 each vertex $V_i$ which contains all the vertices it shares an edge
-with, and associated winner/loser flags.  In the second reduce()
+with, and associated winner/loser flags.  In the second {\it reduce()}
 operation, if $V_i$ won the comparison with all its edge neighbors,
@@
 edge vertices are emitted without the loser flag.  The second
-collate() operation collects this new information for each vertex.
+{\it collate()} operation collects this new information for each vertex.
 
-In the third reduce() operation, if $V_i$ was flagged as a loser by
+In the third {\it reduce()} operation, if $V_i$ was flagged as a loser by
 any edge neighbor who was a winner, then it becomes a losing vertex
@@
 
-In the fourth reduce() operation, some vertices will have all their
+In the fourth {\it reduce()} operation, some vertices will have all their
 edge vertices flagged as losers.  These vertices are all winners,
-either the ones identified in the second reduce() operation, or
+either the ones identified in the second {\it reduce()} operation, or
 additional vertices who would become singletons (having no edges) when
@@
 All edges of each vertex are also emitted, retaining the loser flag if
-they have it.  They are emitted with $E_ij$ as their key, with $V_i <
-V_j$.  This is to insure both copies of the edge come together via the
-fourth collate() operation.  The multivalue for each edge will now
+they have it.  They are emitted with $E_{ij}$ as their key, with $V_i <
+V_j$.  This is to ensure both copies of the edge come together via the
+fourth {\it collate()} operation.  The multivalue for each edge will now
 have two values, either of which may be NULL or a loser flag.
@@
 \> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$. \\
-while not done: \\
-\> done = 1; \\
+while not $done$: \\
+\> $done$ = 1; \\
 \> Add $E$ to $V$. \\
@@
 \> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
-\> \> \> If $d_{min} \ne d_i$, done = 0; \\
+\> \> \> If $d_{min} \ne d_i$, $done$ = 0; \\
 \> \> Output:  Key = $V_i$; Value = $d_{min}$ (updated distance for $V_i$) \\
 \> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$). \\
-\> MPI\_Allreduce on done with MPI\_MIN. \\
+\> MPI\_Allreduce on $done$ with MPI\_MIN. \\
 Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
@@
 can be achieved by pre-aggregating certain MapReduce objects and 
-storing data in more than one MapReduce object.  The improved algorithm
+storing data in more than one MapReduce object.  The enhanced algorithm
 is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects storing
@@
 SSSP, only a small number of new candidate distances are generated.
-The improved algorithm uses MapReduce object $U$ to store these updates.
+The enhanced algorithm uses MapReduce object $U$ to store these updates.
 Only the updates are communicated; each update is aggregated to the 
@@
 \> Output:  Key = $V_s$; Value = 0. \\
-while not done: \\
-\> done = 1; \\
+while not $done$: \\
+\> $done$ = 1; \\
 \> Aggregate $U$. \\
@@
 \> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
-\> \> \> If $d_{min} \ne d_i$, done = 0; \\
+\> \> \> If $d_{min} \ne d_i$, $done$ = 0; \\
 \> \> Output to $V$:  Key = $V_i$; Value = $d_{min}$. \\
 \> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$. \\
-\> MPI\_Allreduce on done with MPI\_MIN. \\
-\> if not done \\
+\> MPI\_Allreduce on $done$ with MPI\_MIN. \\
+\> if not $done$ \\
 \> \> Add $U$ to $E$.  Empty $U$. \\
@@
 
- \caption{Improved MapReduce algorithm for single-source shortest path (SSSP)
+ \caption{Enhanced MapReduce algorithm for single-source shortest path (SSSP)
  determination.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}

### 2010-05-04 paper/lessons.tex
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7
More edits.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@605 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \section{Lessons Learned}
+\label{sec:lessons}
 
@@
 easy to write and debug.  Thinking about a computational task from a
-MapReduce perspective is different than traditional distributed-memory
+MapReduce perspective is different from traditional distributed-memory
 parallel algorithm design.  For example, with an MPI mindset, it often
 seems heretical to intentionally ignore data locality.  However,
-writing small map() and reduce() functions is typically easy.  And
+writing small {\it map()} and {\it reduce()} functions is typically easy.  And
 writing an algorithm that involves complex parallel operations,
@@
 This is somewhat hidden from the user in a typical cloud-computing
-model, where data simply exists somehere in the cloud, and Hadoop
+model, where data simply exists somewhere in the cloud and Hadoop
 ensures data moves where it needs to and is operated on by some
@@
 that is graph centric.  We hope the library can be generally useful on
-large-scale monolithic or cloud-style parallel machines which support
+large-scale monolithic or cloud-style parallel machines that support
 MPI, for a variety of data-intense or compute-intense problems that

### 2010-05-04 paper/results.tex
Adding results
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@606 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-We ran each of the algorithms on three R-MAT graphs of
+We ran each of the algorithms on four R-MAT graphs of
 different sizes, each on a varying number of processors.  Details of the
 input data are shown in Table~\ref{t:rmats}.
-The {\it small} problem size (around 134M edges) was a 
-problem that could be run
-on a single processor.  The {\it medium} problem size (around 2B
-edges) could be run on a few processors.  The {\it large} problem size
-(around 34B edges) required most of the machine to run.
+The {\it tiny} problem size (around 8M edges) can typically be used
+on a single processor without significant out-of-core operations.
+The {\it small} problem size (around 134M edges) is a 
+problem that can be run
+on a single processor with out-of-core operations; larger processor 
+configurations, however, do not necessary need out-of-core operations.
+The {\it medium} problem size (around 2B
+edges) can be used on a few processors.  The {\it large} problem size
+(around 34B edges) requires most of the machine to run.
 All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.
@@
 \hline
+tiny   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
 small  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
@@
 processors, and of the parallel scalability for computing on a graph
-of fixed size on a growing number of processors.
+of fixed size on a growing number of processors.  Where available, we
+compare the MapReduce algorithm with other parallel implementations.
 

### 2010-05-06 paper/graph.tex
Minor changes to paper.  Some new results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@608 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 Or the general R-MAT matrix can be further processed by MapReduce
-operations to meet these requirements.
+operations to meet these requirements.  For some algorithms below
+(e.g., triangle finding and maximal independent set generation), we
+post-process R-MAT matrices to create upper-triangular R-MAT
+matrices representing undirected graphs.  For each edge $(i,j)$ 
+with $i>j$, we add edge $(j,i)$, essentially symmetrizing the matrix; 
+we then remove duplicate edges and self edges,
+leaving only edges with $i<j$.
 
 By using more than one MapReduce object, we can improve the performance
-of the R-MAT generation algorithm.  In the enhanced algorithm
+of the R-MAT generation algorithm, particularly when out-of-core operations
+are needed.  In the enhanced algorithm
 (Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce object
@@
 function to reduce duplicates locally on each processor.
-Performance improvements
+Time reductions
 in this algorithm result from the reduced number of KV pairs communicated in
@@
 
-Map: assign random values to each vertex \\
+Map: assign random values to each vertex $V_i$ and $V_j$ of edge $E_{ij}$. \\
 Clone: convert edge key/values directly to key/multivalues \\

### 2010-05-26 paper/graph.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 matrix.  This exposition follows the triangle-finding algorithm
-presented in \cite{Cohen}.
+presented in~\cite{Cohen09}.
 
@@
 an input graph of undirected edges $(V_i,V_j)$.  This exposition also
-follows the connected-component algorithm presented in \cite{Cohen},
+follows the connected-component algorithm presented in~\cite{Cohen09},
 with the addition of logic that load-balances data across processors
@@
 several contexts, such as identifying large numbers of indepenent
-starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
-is a well-known parallel method for finding a MIS.  Figure
-\ref{fig:luby} outlines a MapReduce version of Luby's algorithm,
+starting vertices for graph traversals.  Luby's algorithm~\cite{Luby86}
+is a well-known parallel method for finding a MIS.  Figure~\ref{fig:luby} 
+outlines a MapReduce version of Luby's algorithm,
 assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
@@
 is not available.  Instead, following the Bellman-Ford-style 
-algorithm of~\cite{SSSPMapReduce, BellmanFord},
+algorithm of~\cite{SSSPMapReduce, Bellman58, Ford62},
 every edge of the graph is visited in each iteration,

### 2010-05-26 paper/intro.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 MapReduce is the programming paradigm popularized by Google
-researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
+researchers Dean and Ghemawat~\cite{MapReduce2004}.  Their motivation was to
 enable rapid development and deployment of analysis programs to
@@
 by Google researchers and are a significant part of the daily compute
-tasks run by the company \cite{Dean2}.
+tasks run by the company~\cite{MapReduce2008}.
 
-Similarly, the open-source Hadoop implementation of MapReduce
-\cite{Hadoop}, has become widely popular in the past few years for
+Similarly, the open-source Hadoop implementation of MapReduce~\cite{Hadoop}, 
+has become widely popular in the past few years for
 parallel analysis of large-scale data sets at Yahoo and other
@@
 kinds of scientific computational tasks have been described, such as
-post-processing analysis of simulation data \cite{Tu}, graph
-algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
-The paper by Tu et al \cite{Tu} was particularly insightful to us,
+post-processing analysis of simulation data~\cite{Tu08}, graph
+algorithmics~\cite{Cohen09}, and linear algebra operations~\cite{Ekanayake09}.
+The paper by Tu et al.~\cite{Tu08} was particularly insightful to us,
 because it described how MapReduce could be implemented on top of the
@@
 on top of MPI for use in graph analytics, which we initially released
-as open-source software in mid-2009 \cite{MRMPI}.  We have since
+as open-source software in mid-2009~\cite{MRMPI}.  We have since
 worked to optimize several of the library's underlying algorithms and
@@
 These algorithmic improvements are described in this paper and are
-part of the current downloadable version \cite{MRMPI}.
+part of the current downloadable version~\cite{MRMPI}.
 
@@
 messages to improve bandwidth performance and reduce latency costs.  A
-recent paper by \cite{Dongarra} also outlines the MapReduce formalism
+recent paper~\cite{Hoefler09} also outlines the MapReduce formalism
 from an MPI perspective, though they advocate a more asyncrhonous
@@
 library (provided) can be substituted.  As a library, it can be
-embedded in other codes \cite{Titan,TitanURL} to enable them to
+embedded in other codes~\cite{Titan} to enable them to
 perform MapReduces.
@@
 MPI\_Allreduce are often useful in determining the status of an
-iterative graph algorithm, as described in Section \ref{sec:graph}.
+iterative graph algorithm, as described in Section~\ref{sec:graph}.
 The library interface also provides a user data pointer as an argument
@@
 The remainder of the paper is organized as follows.  The next two
-sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
+sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
 out-of-core MapReduce primitives are formulated as MPI-based
-operations in the MR-MPI library.  Section \ref{sec:graph} briefly
+operations in the MR-MPI library.  Section~\ref{sec:graph} briefly
 describes the formulation of several common graph algorithms as
-MapReduce operations.  Section \ref{sec:results} gives performance
+MapReduce operations.  Section~\ref{sec:results} gives performance
 results for these algorithms running on a parallel cluster for graphs
 %KDDKDD CHECK SIZES
-ranging in size from 1 million to 1 trillion vertices or edges.  In
+ranging in size from 1 million to 2 billion vertices or edges.  In
 this section, we highlight the performance and complexity trade-offs

### 2010-05-26 paper/thanks.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/thanks.tex
@@
 \section{Acknowledgements}
+\label{sec:thanks}
 
-Sandia is a multiprogram laboratory operated by Sandia Corporation, a
-Lockheed Martin Company, for the United States Department of
-Energy's National Nuclear Security Administration under contract
-DE-AC04-94AL85000.
+We thank the following individuals for their contributions to this
+paper: Greg Bayer and Todd Plantenga (Sandia) for explaining Hadoop
+concepts to us, and for the Hadoop implementations and timings of
+Section~\ref{sec:results}; Jon Cohen (NSA) for fruitful discussions
+about his MapReduce graph algorithms~\cite{Cohen09}; Brian Barrett
+(Sandia) for the PBGL results of Section~\ref{sec:results}; Jon Berry
+(Sandia) for the MTGL results of Section~\ref{sec:results}, and for
+his overall support of this work and many useful discussions.
 
-Hadoop folks.
+Sandia National Laboratories is a multi-program laboratory operated by
+Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
+company, for the U.S. Department of Energy's National Nuclear Security
+Administration under contract DE-AC04-94AL85000.

### 2010-05-26 paper/graph.tex
Added scalability discussion
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@620 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-The algorithm in Figure~\ref{fig:pr} performs these
+The algorithm in Figure~\ref{fig:pr2} performs these
 iterations.  The nonzeros of matrix $A$ are initialized as described above;
@@
 contains the indices of all-zero rows of $A$, corresponding to vertices
-with no outlinks.  In step 1 of a PageRank interation, $MT$ is used 
+with no outlinks.  
+The PageRank algorithm is made more efficient by calling
+{\it aggregate()} on the MapReduce objects representing the matrix $A$, 
+the all-zero matrix rows $MT$, and the vector $x$ once before the PageRank 
+iterations begin.  Since these three MapReduce
+objects have the same key types, pre-aggregating the objects moves all
+Key-Value pairs with a given key to a single processor.  Thus, many of
+the PageRank operations become local operations, replacing 
+{\it collate()} operations (which is equivalent to {\it aggregate()} 
+followed by {\it convert()}) with only local {\it convert()} calls. 
+In Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
+substituted for {\it convert()} operations are marked with an asterisk.
+This strategy of pre-aggregating to improve data locality is useful in
+many MapReduce algorithms, but assumes that keys are always mapped to
+the same processor.
+
+In step 1 of a PageRank interation, $MT$ is used 
 in dot products to compute adjustments to the
@@
 expensive part of the computation, requiring two passes over the nonzero
-structure of $A$.  In the first pass, {\it collate()} 
-gathers all row entries $a_{ij}$ with their
+structure of $A$.  In the first pass, {\it convert()} 
+gathers all local row entries $a_{ij}$ with their
 associated $x_i$ entry, and {\it reduce()} computes $a_{ij} x_i$.  
@@
 
-\begin{figure}[htb]
- \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
-
-%Input from R-MAT generator: \\
-%\> MapReduce object $V$ containing all graph vertices:  Key=$V_i$, Value = NULL.  \\
-%\> MapReduce object $E$ containing all graph edges: Key=$V_i$, Value = $V_j$. \\
-%Initialize MapReduce object for matrix $A$  and \\
-%MapReduce object $MT$ containing all vertices with no outlinks.\\
-%\> Copy: $A = E$ \\
-%\> Add: $V$ to $A$ \\
-%\> Collate $A$: Vertex is Key \\
-%\> Reduce $A$:  \\
-%\> \> Input:  Key = $V_i$; MultiValue = NULL,$V_j$,$V_k$,... \\
-%\> \> If (${nvalues} > 1$) \\
-%\> \> \> Compute: $d_{out}(V_i) = {nvalues} - 1$. \\
-%\> \> \> Output:  For each $V_j$ in MultiValue, Key = $V_i$; Value = ($V_j$, $\alpha/d_{out}(V_i)$). \\
-%\> \> else \\
-%\> \> \> Output to $MT$: Key = $V_i$; Value = 0. \\
-%Initialize MapReduce object for pagerank vector $x$. \\
-%\> Map:   \\
-%\> \> Input:  Graph vertices $V$. \\
-%\> \> Output:  For each vertex, Key = $V_i$; Value = $1/|V|$. \\
-Input: \\
-\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
-\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
-\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
-While (${residual} > {tolerance}$) \\
-\> 1. Compute contribution for random jumps and zero-outdegree vertices. \\
-\> \> Add: $x$ to $MT$. \\
-\> \> Collate $MT$:  Vertex is Key. \\
-\> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
-\> \> \> Compute:  Local contributions $c_{local}$. \\
-\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
-\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
-\> 2. Compute $y = A^T x$.  \\
-\> \> Copy $x$ to $y$. \\
-\> \> Add $A$ to $y$. \\
-\> \> Collate $y$:  Key is row index of $A$. \\
-\> \> Reduce $y$: \\
-\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
-\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
-\> \> Collate $y$:  Key is column index of $A$. \\
-\> \> Reduce $y$:  \\
-\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
-\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
-\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
-\> \> Map $y$:  \\
-\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i$. \\
-\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
-\> 4. Scale $y$ by $n_{global}$. \\
-\> \> Map $y$: \\
-\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
-\> 5. Compute residual. \\
-\> \> Add $y$ to $x$. \\
-\> \> Collate $x$:  Key is index $i$. \\
-\> \> Reduce $x$:   \\
-\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
-\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
-\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
-\> 6. $x = y$; \\
-Output:  PageRank vector $x$.
-  \end{tabbing}
- \end{minipage}}\end{center}
-
- \caption{MapReduce algorithm for PageRank vertex ranking.}
-
- \label{fig:pr}
-\end{figure}
-
-The entire PageRank algorithm can be made more efficient by calling
-{\it aggregate()} on the MapReduce objects representing the matrix $A$, 
-the all-zero matrix rows $MT$, and the vector $x$ once before the PageRank 
-iterations begin.  Since these three MapReduce
-objects have the same key types, pre-aggregating the objects moves all
-Key-Value pairs with a given key to a single processor.  Thus, many of
-the PageRank operations become local operations, replacing 
-the {\it collate()} operation (which is equivalent to {\it aggregate()} 
-followed by {\it convert()}) with only a local {\it convert()} call. 
-Figure~\ref{fig:pr2} shows the
-enhanced algorithm; modified operations are marked with an asterisk.
-This strategy of pre-aggregating to improve data locality is useful in
-many MapReduce algorithms, but assumes that keys are always mapped to
-the same processor.
+%\begin{figure}[htb]
+% \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+%  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+%Input: \\
+%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
+%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
+%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
+%While (${residual} > {tolerance}$) \\
+%\> 1. Compute contribution for random jumps and zero-outdegree vertices. \\
+%\> \> Add: $x$ to $MT$. \\
+%\> \> Collate $MT$:  Vertex is Key. \\
+%\> \> Reduce $MT$:   \\
+%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
+%\> \> \> Compute:  Local contributions $c_{local}$. \\
+%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
+%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
+%\> 2. Compute $y = A^T x$.  \\
+%\> \> Copy $x$ to $y$. \\
+%\> \> Add $A$ to $y$. \\
+%\> \> Collate $y$:  Key is row index of $A$. \\
+%\> \> Reduce $y$: \\
+%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
+%\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
+%\> \> Collate $y$:  Key is column index of $A$. \\
+%\> \> Reduce $y$:  \\
+%\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
+%\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
+%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
+%\> \> Map $y$:  \\
+%\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$. \\
+%\> \> \> Output:  Key = $i$, Value = $y_i$. \\
+%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
+%\> 4. Scale $y$ by $n_{global}$. \\
+%\> \> Map $y$: \\
+%\> \> \> Input:  Key = $i$, Value = $y_i$. \\
+%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
+%\> 5. Compute residual. \\
+%\> \> Add $y$ to $x$. \\
+%\> \> Collate $x$:  Key is index $i$. \\
+%\> \> Reduce $x$:   \\
+%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
+%\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
+%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
+%\> 6. $x = y$; \\
+%Output:  PageRank vector $x$.
+  %\end{tabbing}
+ %\end{minipage}}\end{center}
+%
+ %\caption{MapReduce algorithm for PageRank vertex ranking.}
+%
+ %\label{fig:pr}
+%\end{figure}
 
@@
 
- \caption{Enhanced MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
+ \caption{MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
 operations to be replaced by with strictly local {\it convert()} operations (as marked by the asterisks).}

### 2010-05-26 paper/results.tex
Added scalability discussion
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@620 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \label{sec:results}
-
-\subsection{Comparisons to Other Implementations}
 
@@
 input data are shown in Table~\ref{t:rmats}.
-The {\it tiny} problem size (around 8M edges) can typically be used
+The {\it small} problem size (around 8M edges) can typically be used
 on a single processor without significant out-of-core operations.
-The {\it small} problem size (around 134M edges) is a 
+The {\it medium} problem size (around 134M edges) is a 
 problem that can be run
 on a single processor with out-of-core operations; larger processor 
-configurations, however, do not necessary need out-of-core operations.
-The {\it medium} problem size (around 2B
-edges) can be used on a few processors.  The {\it large} problem size
-(around 34B edges) requires most of the machine to run.
+configurations, however, do not necessarily need out-of-core operations.
+The {\it large} problem size (around 2B
+edges) requires out-of-core operations and higher processor counts.
+%The {\it x-large} problem size
+%(around 34B edges) requires most of the machine to run.
 All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.
@@
 \hline
-tiny   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
-small  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
-medium &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
-large  &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
+small   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
+medium  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
+large   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
+%x-large &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
 \hline
@@
 
+\subsection{R-MAT generation results}
 Discuss R-MAT generation times.
 
-Discuss PageRank performance.  Include Trilinos and other
-algorithms and machines?
+\subsection{PageRank results}
 
-Discuss triangle finding times.
-
-Discuss connected component identification times.
-
-Discuss Luby maximally independent sets times.
-
-Discuss SSSP times.
-
-Finally, we compare our MapReduce implementations with a multi-threaded
-implementations
-in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}, and
-distributed-memory matrix-based implementations using the linear
-algebra library Trilinos~\cite{Trilinos-Overview}.
-In the MTGL~\cite{MTGL} implementation,
-rank propagates via adjacency list traversal
-in a compressed sparse-row data structure.
-To maintain scalability, code must
-be written so that a single thread spawns the loop that processes all
-in-neighbors of a given vertex; this detail enables the compiler to generate
-hotspot-free code.
+In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
+compared to a
+distributed-memory matrix-based implementation using the linear
+algebra toolkit Trilinos~\cite{Trilinos-Overview}.
 The matrix-based distributed-memory implementation of PageRank
@@
 
+%We compare our MapReduce implementations with a multi-threaded
+%implementations
+%in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
+%In the MTGL~\cite{MTGL} implementation,
+%rank propagates via adjacency list traversal
+%in a compressed sparse-row data structure.
+%To maintain scalability, code must
+%be written so that a single thread spawns the loop that processes all
+%in-neighbors of a given vertex; this detail enables the compiler to generate
+%hotspot-free code.
+
+%This result is likely due to higher volume of
+%communication in the MapReduce implementation (where all edges are communicated
+%in each iteration), compared to the Trilinos implementation (where only
+%vertex-based data are communicated).
+
+\subsection{Triangle finding results}
+Discuss triangle finding times.
+
+\subsection{Maximally independent set results}
+Discuss Luby maximally independent sets times.
+
+\subsection{Single-source shortest path results}
+Discuss SSSP times.
+
+\subsection{Scalability to large numbers of processors}
+Finally, we demonstrate the scalability of MR-MPI to large numbers of 
+processors.  
+The MR-MPI implementation was run on the Redstorm and Thunderbird
+parallel computers
+Because these systems do not have local disks for each processor, we selected 
+a data set and page sizes that fit in their memory, so out-of-core operations
+were not used.  For these experiments, we used an R-MAT data set with 
+with $2^{25}$ vertices and $2^{28}$ edges, with parameters given in
+Table~\ref{t:rmat}.  We ran both PageRank and Connected Components
+algorithms.
+
 \begin{table}
@@
 \caption{Characteristics of R-MAT input data for PageRank and Connected
-Components experiments.}
+Components scalability experiments.}
 \label{t:rmat}
@@
 
-We performed our experiments using the R-MAT data sets described in 
-Table~\ref{t:rmat}.  
 In Figure~\ref{f:prbig}, we show the performance 
@@
 implementations on distributed memory and multi-threaded architectures.
-The MR-MPI implementation was run on the Redstorm parallel computer;
-for these experiments,
-the data set fit in the memory of Redstorm, so out-of-core operations
-were not used.
 The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
@@
 more execution time than the matrix-based implementations on Redstorm.  
-This result is likely due to higher volume of
-communication in the MapReduce implementation (where all edges are communicated
-in each iteration), compared to the Trilinos implementation (where only
-vertex-based data are communicated).  The distributed memory matrix-based
+The distributed memory matrix-based
 implementations are competitive with the multi-threaded implementation
@@
 \begin{figure}[h!]
-\includegraphics[width=\textwidth]{fig_pagerank.pdf}
+\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
 \caption{Comparison of PageRank implementations using MapReduce,
@@
 \end{figure}
+
+Similar results were obtained for the Connected Components algorithm, as
+we show in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI implementation
+showed good scalability up to 1024 processors, but required significantly
+more time than the Giant Connected Components algorithms using MTGL and
+Trilinos.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_cc_big.pdf}
+\caption{Comparison of Connected Components implementations using MapReduce,
+Trilinos, and MTGL on R-MAT data sets.}
+\label{f:prbig}
+\end{figure}

### 2010-05-26 paper/results.tex
Pagerank results
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@621 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \hline
-small   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
-medium  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
-large   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
-%x-large &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
+RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
+RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
+RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
+%RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
 \hline
@@
 
-In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
+In Figure~\ref{fig:pr}, we show the performance of the MR-MPI PageRank algorithm
 compared to a
 distributed-memory matrix-based implementation using the linear
-algebra toolkit Trilinos~\cite{Trilinos-Overview}.
+algebra toolkit Trilinos~\cite{Trilinos-Overview} 
+and a multi-threaded
+implementation
+in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
 The matrix-based distributed-memory implementation of PageRank
@@
 residuals and norms of $x$ and $y$.
+In the multi-threaded MTGL~\cite{MTGL} implementation,
+rank propagates via adjacency list traversal
+in a compressed sparse-row data structure.
+To maintain scalability, code must
+be written so that a single thread spawns the loop that processes all
+in-neighbors of a given vertex; this detail enables the compiler to generate
+hotspot-free code.
 
-%We compare our MapReduce implementations with a multi-threaded
-%implementations
-%in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
-%In the MTGL~\cite{MTGL} implementation,
-%rank propagates via adjacency list traversal
-%in a compressed sparse-row data structure.
-%To maintain scalability, code must
-%be written so that a single thread spawns the loop that processes all
-%in-neighbors of a given vertex; this detail enables the compiler to generate
-%hotspot-free code.
+Figure~\ref{fig:pr} shows the execution time per PageRank iteration for
+R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  
+Converging the PageRank iterations
+to tolerance $0.002$ requires five or six iterations.
+Several R-MAT matrices of 
+each size were generated for the experiments; the average time over the
+matrices is reported here.  
+The Trilinos implementations show near-perfect
+strong scaling for RMAT-20 and RMAT-24.  The MR-MPI 
+implementations also demonstrate good strong scaling.  However, MR-MPI's
+execution time is at least an order of magnitude greater than the Trilinos
+implementation.  This result is due to two factors:  ({\it i}) a 
+higher volume of
+communication in the MapReduce implementation (where all edges are communicated
+in each iteration), compared to the Trilinos implementation (where only
+vertex-based data are communicated), and ({\it ii}) out-of-core operations
+in MR-MPI were performed because of the page-size restriction, while all
+Trilinos operations were performed in-core.  The benefit of MR-MPI's
+out-of-core implementation is seen, however, with the RMAT-28 data set, 
+which could be solved on smaller processor sets than the Trilinos 
+implementation.  For these experiments, the Trilinos implementation required
+64 processors for the RMAT-28 data set.
 
-%This result is likely due to higher volume of
-%communication in the MapReduce implementation (where all edges are communicated
-%in each iteration), compared to the Trilinos implementation (where only
-%vertex-based data are communicated).
+KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_pagerank.pdf}
+\caption{Comparison of PageRank implementations using MapReduce,
+Trilinos, and MTGL on R-MAT data sets.}
+\label{fig:pr}
+\end{figure}
 
@@
 
-In Figure~\ref{f:prbig}, we show the performance 
+In Figure~\ref{fig:prbig}, we show the performance 
 of the various PageRank
@@
 \includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
-\caption{Comparison of PageRank implementations using MapReduce,
+\caption{Scalability comparison of PageRank implementations using MapReduce,
 Trilinos, and MTGL on R-MAT data sets.}
-\label{f:prbig}
+\label{fig:prbig}
 \end{figure}
@@
 Similar results were obtained for the Connected Components algorithm, as
-we show in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI implementation
+we show in Figure~\ref{fig:ccbig}.  As with PageRank, the MR-MPI implementation
 showed good scalability up to 1024 processors, but required significantly
@@
 \includegraphics[width=\textwidth]{fig_cc_big.pdf}
-\caption{Comparison of Connected Components implementations using MapReduce,
+\caption{Scalability comparison of Connected Components implementations using MapReduce,
 Trilinos, and MTGL on R-MAT data sets.}
-\label{f:prbig}
+\label{fig:ccbig}
 \end{figure}

### 2010-05-27 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@623 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-In Figure~\ref{fig:pr}, we show the performance of the MR-MPI PageRank algorithm
+In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
 compared to a
@@
 
-Figure~\ref{fig:pr} shows the execution time per PageRank iteration for
+Figure~\ref{f:pr} shows the execution time per PageRank iteration for
 R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  
@@
 Trilinos, and MTGL on R-MAT data sets.}
-\label{fig:pr}
+\label{f:pr}
 \end{figure}
@@
 
+KDD NOTE:  Need to add XMT results.
+
+\subsection{Connected Components}
+Discuss Luby maximally independent sets times.
+
+KDD NOTE:  Need to add XMT results.
+
 \subsection{Maximally independent set results}
@@
 
+KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
+
 \subsection{Single-source shortest path results}
-Discuss SSSP times.
+
+The execution times for the single-source shortest path algorithms
+(Figures~\ref{fig:sssp} and \ref{fig:sssp2})
+are shown in Figure~\ref{f:sssp}.  The results show the benefit of
+using the enhanced algorithm, providing at least a 17\% (and often greater)
+reduction in execution time due to reduced communication; only the updated
+distances are communicated throughout most of the enhanced algorithm.
+However, the execution times are still large compared to multi-threaded
+implementations; for example, Madduri et al.~\cite{Madduri07} report
+execution times of only 11 seconds on 40 
+processors for a multithreaded implementation on a Cray MTA-2
+(the predecessor of the XMT) on graphs with one billion edges.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_sssp.pdf}
+\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
+Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
+algorithm (Figure~\ref{fig:sssp2}) are included.}
+\label{f:sssp}
+\end{figure}
+
+KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
+
+To compare our MR-MPI implementation with a wider set of implementations,
+we performed experiments MR-MPI, Hadoop, PBGL (Parallel
+Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
+Cray XMT using two web graphs: {WebGraphA} with 13.2M
+vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
+In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
+using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
+implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
+The XMT and PBGL implementations are based on 
+delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
+over the entire edge list to advance a breadth-first search.  
+We observe that the MR-MPI
+implementation runs in less time than the Hadoop implementation, but requires
+significantly more time than the XMT and PBGL implementations.  
+The Bellman-Ford-style iterations are especially harmful to the MR-MPI and
+Hadoop implementations for {WebGraphB}, which required 109 iterations to 
+complete the search; 
+execution times for this data set are show in Table~\ref{t:ssspB}.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_ssspA.pdf}
+\caption{SSSP using MR-MPI for WebGraphA with
+13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
+are also shown.}
+\label{f:ssspA}
+\end{figure}
+
+\begin{table}
+\begin{tabular}{|l|c|r|}
+\hline
+Implementation & Number of Processes & SSSP Execution Time \\
+\hline
+Hadoop & 48  & 38,925 secs.\\
+MR-MPI & 48 &  13,724 secs.\\
+XMT/C  & 32 &  37 secs.\\
+%Out-of-core MR-MPI & 96 &   8,358 secs.\\
+%Out-of-core MR-MPI & 64 &  12,882 secs.\\
+%Out-of-core MR-MPI & 100 &  6,280 secs.\\
+\hline
+\end{tabular}
+\caption{Execution times for SSSP with {WebGraphB}.}
+\label{t:ssspB}
+\end{table}
 
@@
 
-In Figure~\ref{fig:prbig}, we show the performance 
+In Figure~\ref{f:prbig}, we show the performance 
 of the various PageRank
@@
 Trilinos, and MTGL on R-MAT data sets.}
-\label{fig:prbig}
+\label{f:prbig}
 \end{figure}
@@
 Similar results were obtained for the Connected Components algorithm, as
-we show in Figure~\ref{fig:ccbig}.  As with PageRank, the MR-MPI implementation
+we show in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI implementation
 showed good scalability up to 1024 processors, but required significantly
@@
 Trilinos, and MTGL on R-MAT data sets.}
-\label{fig:ccbig}
+\label{f:ccbig}
 \end{figure}

### 2010-05-28 paper/results.tex
More
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@624 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \subsection{R-MAT generation results}
-Discuss R-MAT generation times.
+
+In Figure~\ref{f:rmat}, we show the scalability of the R-MAT generation 
+algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and RMAT-28.  For
+RMAT-20 and RMAT-24, superlinear speed-up is shown.  This speed-up is due
+to the decreased amount of file I/O needed with greater numbers of 
+processors; with a larger total memory, a fixed-size problem requires
+less I/O with more processors.  For RMAT-28, where most operation require
+significant out-of-core operation, parallel efficiency ranges from 52\% to
+97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
+the original algorithm (Figure~\ref{fig:rmat}) showed approximately 
+10\% reduction in execution time for the enhanced algorithm.
+
+KDD NOTE:  Add XMT or PBGL results here.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_rmat.pdf}
+\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
+\label{f:rmat}
+\end{figure}
 
@@
 In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
-compared to a
+(Figure~\ref{fig:pr2}) compared to a
 distributed-memory matrix-based implementation using the linear
@@
 \subsection{Maximally independent set results}
-Discuss Luby maximally independent sets times.
+
+The execution times for the maximally independent set algorithm
+(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
+generation results, superlinear speed-up of the algorithm occurs for RMAT-20
+and RMAT-24, as more of the graph fits into processor memory and less file
+I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
+operations. In this case, parallel efficiency is nearly perfect going from 
+8 to 32 processors; some loss of efficiency is seen at 64 processors,
+where the parallel efficiency compared to the 8-processor case is ??\%.
 
 KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
+
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{fig_luby.pdf}
+\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
+\label{f:luby}
+\end{figure}
 
@@
 We observe that the MR-MPI
-implementation runs in less time than the Hadoop implementation, but requires
+implementation runs in less time than the Hadoop implementation,but requires
 significantly more time than the XMT and PBGL implementations.  
-The Bellman-Ford-style iterations are especially harmful to the MR-MPI and
-Hadoop implementations for {WebGraphB}, which required 109 iterations to 
-complete the search; 
-execution times for this data set are show in Table~\ref{t:ssspB}.
+In experiments with {WebGraphB},
+the benefit of the enhanced algorithm (Figure~\ref{fig:sssp2}) is clearly
+shown, with a 40\% reduction in execution time compared to the original
+SSSP algorith (Figure~\ref{fig:sssp}).
+But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
+Hadoop implementations for {WebGraphB}, which required 110 iterations to 
+complete; execution times for this data set are show in Table~\ref{t:ssspB}.
 
@@
 Hadoop & 48  & 38,925 secs.\\
-MR-MPI & 48 &  13,724 secs.\\
+MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
+MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
 XMT/C  & 32 &  37 secs.\\

### 2010-05-28 paper/results.tex
More.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@625 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 for out-of-core operations in MR-MPI.  To avoid contention for disk I/O, we 
-ran all experiments with one MPI process per node.
+ran all experiments with one MPI process per node.  For comparisons with
+other implementations, we used either the same cluster or, where noted,
+Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz} processors
+and a 3D-Torus network.
 
@@
 \begin{table}
+\begin{center}
 \begin{tabular}{|l|c|c|c|c|c|c|c|}
@@
 \hline
-RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &   \\
-RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &   \\
-RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$&   \\
+RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &  $\approx 24K$ \\
+RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &  $\approx 147K$ \\
+RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$& $\approx 880K$ \\
 %RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
@@
 \label{t:rmats}
+\end{center}
 \end{table}
@@
 compare the MapReduce algorithm with other parallel implementations.
+We compute parallel efficiency on $p$ processors
+as $({time}_{m} \times m) / ({time}_p \times p)$
+where $m$ is the smallest number of processors on which the experiment
+was run, and ${time}_m$ is the execution time required on $m$ processors.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_rmat.pdf}
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_pagerank.pdf}
@@
 \subsection{Triangle finding results}
-Discuss triangle finding times.
+
+In Figure~\ref{f:tri}, we show the performance of the triangle finding 
+algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm were
+too large to allow the problem to be run with RMAT-28 on our cluster.
+Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
+ranged from 50\% to 120\%.
+
+KDD NOTE:  Need to add XMT results; triangle enumeration is not implemented in PBGL.
+
+\begin{figure}[htb]
+\includegraphics[width=\textwidth]{fig_tri.pdf}
+\caption{Performance of the MR-MPI triangle-finding algorithm (Figure~\ref{fig:tri}).}
+\label{f:tri}
+\end{figure}
+
+\subsection{Connected Components}
+
+We compare our MR-MPI algorithm (Figure~\ref{fig:cc}) 
+with a hybrid ``Giant Connected Component''
+implementation using Trilinos and MR-MPI.  Power law graphs often have one
+or two very large components and many very small components.  
+The hybrid algorithm exploits this feature of the data by using inexpensive
+breadth-first
+search from the vertex with highest degree to identify the largest components,
+followed by a more expensive algorithm to identify the small components in
+the remainder of the graph.  In our hybrid implementation, we use
+Trilinos' matrix-vector multiplication to do the breadth-first search, 
+finding components that include (in total) 70\% or more of the vertices.
+We then apply our MR-MPI algorithm to the remaining graph to identify
+the small components.  This hybrid approach is extremely effective,
+reducing the execution time to identify all components of RMAT-20 and 
+RMAT-24 by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI
+is seen, however, for RMAT-28, where the problem is too large to fit into
+memory for our hybrid algorithm.
 
@@
 
-\subsection{Connected Components}
-Discuss Luby maximally independent sets times.
-
-KDD NOTE:  Need to add XMT results.
+\begin{figure}[htb]
+\includegraphics[width=\textwidth]{fig_cc.pdf}
+\caption{Performance of the MR-MPI connected components algorithm (Figure~\ref{fig:cc}) compared with a hybrid ``Giant Connected Component'' algorithm based
+on Trilinos.}
+\label{f:cc}
+\end{figure}
 
@@
 8 to 32 processors; some loss of efficiency is seen at 64 processors,
-where the parallel efficiency compared to the 8-processor case is ??\%.
+where the parallel efficiency compared to the 8-processor case is KDD\%.
+
+KDD NOTE:  I think I had a bad node for the 64-processor run; I am 
+rerunning it now and soon will have revised RMAT-28 results on 64 processors.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_luby.pdf}
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_sssp.pdf}
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_ssspA.pdf}
@@
 \begin{table}
+\begin{center}
 \begin{tabular}{|l|c|r|}
@@
 \label{t:ssspB}
+\end{center}
 \end{table}
@@
 processors.  
-The MR-MPI implementation was run on the Redstorm and Thunderbird
-parallel computers
+The MR-MPI implementation was run on Sandia's Redstorm and Thunderbird
+parallel computers.  Redstorm is a massively parallel computer with 2+ 
+{GHz} dual/quad-core AMD Opteron processors and 9.6 GB/s link bandwidth.
+Thunderbird is a cluster with 3.6 {GHz} dual-core Intel EM64T processors
+and an Infiniband network.
 Because these systems do not have local disks for each processor, we selected 
-a data set and page sizes that fit in their memory, so out-of-core operations
+a data set and page sizes that fit in memory, so out-of-core operations
 were not used.  For these experiments, we used an R-MAT data set with 
@@
 \label{t:rmat}
+\end{center}
 \end{table}
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_cc_big.pdf}

### 2010-05-28 paper/results.tex
Last
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@626 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 8 to 32 processors; some loss of efficiency is seen at 64 processors,
-where the parallel efficiency compared to the 8-processor case is KDD\%.
+where the parallel efficiency compared to the 8-processor case is 69\%.
 

### 2010-06-07 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@623 3fc257bb-8550-0410-81b3-a5c103da83f7
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 while $M_{\rm remain} > 0$: \\
-\> Map: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
+\> Map: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
            \> \> output Key = $(i,j)$, Value = NULL \\
@@
 
- \caption{MapReduce algorithm for R-MAT graph generation.}
+ \caption{MapReduce algorithm for R-MAT graph generation on $P$ processors.}
 
@@
 
-The {\it map()} will often generate some small number of duplicate edges.
+The {\it map()} often generates some small number of duplicate edges.
 The {\it collate()} and {\it reduce()} operations remove the duplicates.  The
@@
 Time reductions
-in this algorithm result from the reduced number of KV pairs communicated in
+in this algorithm result from the smaller number of KV pairs communicated in
 {\it aggregate()},
@@
 while $M_{\rm remain} > 0$: \\
-\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges on each processor \\
+\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
            \> \> output Key = $(i,j)$, Value = NULL \\
@@
 
- \caption{Enhanced MapReduce algorithm for R-MAT graph generation.}
+ \caption{Enhanced MapReduce algorithm for R-MAT graph generation on $P$ processors.}
 
@@
 \subsection{PageRank}
+\label{subsec:graph_pagerank}
 
-The PageRank algorithm assings a relative numeric rank to each
+The PageRank algorithm assigns a relative numeric rank to each
 vertex in a graph.
@@
 The PageRank algorithm is made more efficient by calling
-{\it aggregate()} on the MapReduce objects representing the matrix $A$, 
-the all-zero matrix rows $MT$, and the vector $x$ once before the PageRank 
+{\it aggregate()} once on the MapReduce objects representing the matrix $A$, 
+the all-zero matrix rows $MT$, and the vector $x$ before the PageRank 
 iterations begin.  Since these three MapReduce
@@
 the PageRank operations become local operations, replacing 
-{\it collate()} operations (which is equivalent to {\it aggregate()} 
+{\it collate()} operations (which are equivalent to {\it aggregate()} 
 followed by {\it convert()}) with only local {\it convert()} calls. 
 In Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
-substituted for {\it convert()} operations are marked with an asterisk.
+substituted for {\it collate()} operations are marked with an asterisk.
 This strategy of pre-aggregating to improve data locality is useful in
@@
 structure of $A$.  In the first pass, {\it convert()} 
-gathers all local row entries $a_{ij}$ with their
-associated $x_i$ entry, and {\it reduce()} computes $a_{ij} x_i$.  
+gathers all local row entries $A_{ij}$ with their
+associated $x_i$ entry, and {\it reduce()} computes $A_{ij} x_i$.  
 In the second pass, 
 {\it collate()} gathers, for each $j$, all contributions to the
-column sum $\sum_i a_{ij} x_i$, which is computed by a second {\it reduce()}. 
+column sum $\sum_i A_{ij} x_i$, which is computed by a second {\it reduce()}. 
 Steps 3 and 4 adjust and scale the 
@@
 %Input: \\
-%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
+%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
 %\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
@@
 %\> \> Reduce $y$: \\
-%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
-%\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
+%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
+%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$. \\
 %\> \> Collate $y$:  Key is column index of $A$. \\
 %\> \> Reduce $y$:  \\
-%\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
-%\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
+%\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
+%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$. \\
 %\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
@@
 Input: \\
-$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, a_{ij}]$ \\
+$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
 $*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
@@
 \> \> Reduce $y$: \\
-\> \> \>  Input: Key = $i$; Multivalue = $i, [j, a_{ij}]$ for nonzeros $a_{ij}$ in $A$ \\
-\> \> \>  Output:  Key = $j$; Value = $a_{ij} y_i$. \\
+\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
+\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$. \\
 \> \> Collate $y$:  Key is column index of $A$. \\
 \> \> Reduce $y$:  \\
-\> \> \> Input:  Key = $j$; Multivalue = $a_{ij} y_i$ for all non-zero entries $a_{ij}$ of column $j$ \\
-\> \> \> Output:  Key = $j$; Value = $\sum_i a_{ij} y_i$. \\
+\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
+\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$. \\
 \> Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
@@
  \caption{MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
-operations to be replaced by with strictly local {\it convert()} operations (as marked by the asterisks).}
+operations to be replaced by strictly local {\it convert()} operations (as marked by the asterisks).}
 
@@
 by $G_0$, then there is a $(V_i,V_j)$ edge in the graph.  Thus all
-vertices in the multivalue are roots of angles which are complete
+vertices in the multivalue are roots of angles that are complete
 triangles and can be emitted as a triplet key.
@@
 \> 1 Map: \> Convert edges to vertices \\
-    \> \> \> input Key = $E_{ij}$, Value = NULL \\
+    \> \> \> input Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
     \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
@@
 vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$.  Each
-iteration will grow the zones, one layer of neighbors at a time.  As
+iteration grows the zones, one layer of neighbors at a time.  As
 zones collide due to shared edges, a winner is chosen (the smaller
 zone ID), and vertices in the losing zone are reassigned to the
-winning zone.  When the iterations complete, each zone will have
+winning zone.  When the iterations complete, each zone has
 become a fully connected component.  The algorithm thus finds all
@@
 assignment together in one multi-value.  The first {\it reduce()}
-operation then re-emits each edge, tagged by the zone assignment of
+operation re-emits each edge, tagged by the zone assignment of
 one of its vertices.
@@
 adds the changing zone assignments to the set of key/value pairs.  The
-third collate() can then collect all the vertices of a zone and zero
+third {\it collate()} can then collect all the vertices of a zone and zero
 or more reassignments for the zone ID.  Since a zone could collide
@@
 zones.  If no zone reassignment value appears in the multi-value, the
-zone ID is unchanged.  The final reduce() a key/value pair for each
+zone ID is unchanged.  The final {\it reduce()} emits a key/value pair for each
 vertex in the zone, with the vertex as a key and the new zone ID as a
@@
 very large key/multi-value (KMV) pairs.  For example, if the graph is
-fully connected, then on the last iteration, a single KMV pair will
-contain all vertices in the graph and be assigned to one processor.
+fully connected, in the last iteration, a single KMV pair 
+contains all vertices in the graph and is assigned to one processor.
 This imbalance in memory and computational work can lead to poor
@@
 During the third {\it map()} operation, if the zone has been
-partitioned, then the vertex is assigned to a random processor and the
+partitioned, the vertex is assigned to a random processor and the
 processor ID bits are added to the zone ID, as indicated by the
@@
 
-This additional partitioning logic means that the third {\it
-collate()} operation, which keys on zone IDs, some of which now
-include processor bits, will collect only a $1/P$ subset of the
+With this additional partitioning logic, the third {\it
+collate()} operation (which keys on zone IDs, some of which now
+include processor bits) collects only $1/P$ of the
 vertices in large zones onto each processor.  But the multivalue on
 each processor contains all the zone-reassignments relevant to the
-unpartitioned zone.  This allows the third {\it reduce()} operation to
+unpartitioned zone.  Thus, the third {\it reduce()} operation can
 change the zone ID (if necessary) in a consistent manner across all
@@
 reduce()} operation emits new zone assignments for each vertex, the
-zone retains its partitioned status, and the partition bit is also
+zone retains its partitioned status; the partition bit is also
 explicitly set if the vertex count exceeds the threshhold for the
@@
 Note that this logic does not guarantee that the partition bits of the
-zone IDs for all the vertices in a single zone will be set
+zone IDs for all the vertices in a single zone are set
 consistently on a given iteration.  For example, an unpartitioned zone
 with a small ID may consume a partitioned zone.  The vertices from the
-partitioned zone will retain their partitioned status, but the
+partitioned zone retain their partitioned status, but the
 original vertices in the small zone may not set the partition bit of
@@
 partitioned and unpartioned zone IDs, to ensure all vertices in the
-zone will know the reassignment information.
+zone know the reassignment information.
 
@@
 An ``independent'' set of vertices in a graph is one where no pair of
-vertices in the set share an edge.  The set is ``maximal'' if no
+vertices in the set shares an edge.  The set is ``maximal'' if no
 additional vertex can be added \footnote{A maximal set is ``maximum''
@@
 help with.}.  Finding a maximal independent set (MIS) is useful in
-several contexts, such as identifying large numbers of indepenent
+several contexts, such as identifying large numbers of independent
 starting vertices for graph traversals.  Luby's algorithm~\cite{Luby86}
@@
 Before the iterations begin, a random value is assigned to each
-vertex, which will be used to compare pairs of vertices.  The vertex
+vertex, which is used to compare pairs of vertices.  The vertex
 ID itself could be used as the random value, but since the vertices
@@
 a {\it reduce()} operation, without need for a {\it collate()} operation.  A
-second empty MapReduce object is also initialized, which will
-accumulate MIS vertices as the iterations proceed.
+second empty MapReduce object is also initialized, which 
+accumulates MIS vertices as the iterations proceed.
 
@@
 In the third {\it reduce()} operation, if $V_i$ was flagged as a loser by
-any edge neighbor who was a winner, then it becomes a losing vertex
+any edge neighbor who was a winner, it becomes a losing vertex
 and emits all its edge vertices with a loser flag.  Otherwise it emits
@@
 
-In the fourth {\it reduce()} operation, some vertices will have all their
+In the fourth {\it reduce()} operation, some vertices have all their
 edge vertices flagged as losers.  These vertices are all winners,
@@
 All edges of each vertex are also emitted, retaining the loser flag if
-they have it.  They are emitted with $E_{ij}$ as their key, with $V_i <
-V_j$.  This is to ensure both copies of the edge come together via the
-fourth {\it collate()} operation.  The multivalue for each edge will now
-have two values, either of which may be NULL or a loser flag.
+they have it;  they are emitted with $E_{ij} = (V_i, V_j)$ 
+as their key, with $V_i < V_j$.  
+This is to ensure both copies of the edge come together via the
+fourth {\it collate()} operation.  The multivalue for each edge now
+has two values, either of which may be NULL or a loser flag.
 
@@
 
-\subsection{Single source shortest path}
+\subsection{Single source shortest path calculation}
 
@@
  \caption{MapReduce algorithm for single-source shortest path (SSSP)
- determination.}
+ calculation.}
 
@@
 vertices and edges are aggregated to processors only once at the beginning
-of the computation.  Then, in each iteration of 
+of the computation.  In each iteration of 
 SSSP, only a small number of new candidate distances are generated.
 The enhanced algorithm uses MapReduce object $U$ to store these updates.
-Only the updates are communicated; each update is aggregated to the 
+Only the updates in $U$ are communicated; each update is aggregated to the 
 same processor that stores the corresponding vertex and its edges.
-Communication is performed using $U$, so the total amount of communication
+Thus, the total amount of communication
 needed is small compared to the original algorithm (which aggregates all
@@
  \caption{Enhanced MapReduce algorithm for single-source shortest path (SSSP)
- determination.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}
+ calculation.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}
 

### 2010-06-07 paper/intro.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 research groups, due to its free availability.  MapReduce programs in
-Hadoop are typically written in Java, though it also supports use of
+Hadoop are typically written in Java, although it also supports use of
 stand-alone map and reduce kernels, which can be written as shell
@@
 For map and reduce operations, the library calls back to user-provided
-map() and reduce() functions.  The use of C++ allows precise control
+{\it map()} and {\it reduce()} functions.  The use of C++ allows precise control
 over the memory and format of data allocated by each processor during
@@
 recent paper~\cite{Hoefler09} also outlines the MapReduce formalism
-from an MPI perspective, though they advocate a more asyncrhonous
+from an MPI perspective, although they advocate a more asynchronous
 approach, using one-way communication of small messages.
 
-\item {Small, portable:} The entire MR-MPI library is a few thousand
+\item {\it Small, portable:} The entire MR-MPI library is a few thousand
 lines of standard C++ code.  For parallel operation, the program is
@@
 embedded in other codes~\cite{Titan} to enable them to
-perform MapReduces.
+perform MapReduce operations.
 
@@
 can be performed using a few such pages.  If the data set fits in a
-single page (per processor), then the library performs its operations
-in-core.  If the data set exceeds the page size, then processors each
-write to temporary disk files (to local disk or a parallel file
-system) as needed and subsequently read from them.  This allows
+single page (per processor), the library performs its operations
+in-core.  If the data set exceeds the page size, processors each
+write to temporary disk files (on local disks or a parallel file
+system) as needed and subsequently read from them.  This out-of-core
+operation allows
 processing of data sets larger than the aggregate memory of all the
@@
 map and reduce operations.  For example, various flags can be stored
-that alter the operation of a map or reduce function, as can richer
-data structures, that accumulate the results.
+that alter the operation of a map or reduce operation, as can richer
+data structures that accumulate results.
 
-\item {\it C++, C, and Python interfaces:} A C++ interface to the
-MR-MPI library means a user program instantiates and then invokes
-methods in one or more MapReduce objects.  A C interface means the
-library can also be called from C or other high-level languages such as
-Fortran.  A C interface also means the library can be easily wrapped
+\item {\it C++, C, and Python interfaces:} The C++ interface to the
+MR-MPI library allows user programs to instantiate and then invoke
+methods in one or more MapReduce objects.  The C interface allows the
+library to be called from C or other high-level languages such as
+Fortran.  The C interface also allows the library to be wrapped easily
 by Python via the Python ``ctypes'' module.  The library can then be
-called from a Python script, allowing the user to write map() and
-reduce() callback functions in Python.  If a machine supports running
+called from a Python script, allowing the user to write {\it map()} and
+{\it reduce()} callback functions in Python.  If a machine supports running
 Python in parallel, a parallel MapReduce can also be run in this mode.
 
-\item {No fault tolerance:} Current MPI implementations do not enable
-easy detection of a dead processor, or retrieval of the data it was
+\item {\it No fault tolerance:} Current MPI implementations do not enable
+easy detection of a dead processor or retrieval of the data it was
 working on.  So like most MPI programs, a parallel program calling the
 MR-MPI library will hang or crash if a processor goes away.  Unlike
-Hadoop, and its HDFS file system which provides for data redundancy,
-the MR-MPI library simply reads and writes simple, flat files.  It can
+Hadoop, and its HDFS file system which provides data redundancy,
+the MR-MPI library reads and writes simple, flat files.  It can
 use local per-processor disks, or a parallel file system, if
@@
 
-The remainder of the paper is organized as follows.  The next two
-sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
+The remainder of the paper is organized as follows.  
+Sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
 out-of-core MapReduce primitives are formulated as MPI-based
@@
 %KDDKDD CHECK SIZES
-ranging in size from 1 million to 2 billion vertices or edges.  In
-this section, we highlight the performance and complexity trade-offs
+ranging in size from 8 million to 2 billion edges;  
+we highlight the performance and complexity trade-offs
 of a MapReduce approach versus other more special-purpose algorithms.

### 2010-06-07 paper/lessons.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 and minimize communication.  To do this they typically require a lot
-of application-specific logic and parallel communication coding, to
+of application-specific logic and parallel communication coding to
 create and maintain a data decomposition, generate ghost versions of
@@
 writing small {\it map()} and {\it reduce()} functions is typically easy.  And
-writing an algorithm that involves complex parallel operations,
+writing an algorithm that involves complex parallel operations
 without actually needing to write application-specific parallel code
-to move data via MPI calls, is often a pleasant surprise.  Moreover,
+to move data via MPI calls is often a pleasant surprise.  Moreover,
 if the MapReduce algorithm is initially coded so that it runs
@@
 model, where data simply exists somewhere in the cloud and Hadoop
-ensures data moves where it needs to and is operated on by some
+ensures data moves where it is needed and is operated on by some
 processor.  The cloud model is a nice data-centric abstraction which
@@
 
-However, since the MPI implementation of MapReduce, at least as
-described in this paper, is processor-centric, one can sometimes
+However, since the MPI implementation of MapReduce as
+described in this paper is processor-centric, one can sometimes
 fruitfully exploit the possibility for processors to maintain

### 2010-06-07 paper/mr.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 etc); they are simply treated as byte strings by the library.  A KV
-pair always has a key; its value may be NULL.  A related data type is
+pair always has a key; a KV's value may be NULL.  A related data type is
 the key/multivalue (KMV) pair, where all values associated with the
@@
 generated by each processor.  Typically, this is done using data read
-in from files, but a {\it map()} may generate data itself or process
+from files, but a {\it map()} may generate data itself or process
 existing KV pairs to create new ones.  The KV pairs produced are
@@
 a 32-bit value whose remainder modulo $P$ processors generates the owning
-processor ID.  Alternatively, the user can provide a hash function
-which converts the key into a processor ID.  Each processor then sends
+processor rank.  Alternatively, the user can provide a hash function
+which converts the key into a processor rank.  Each processor then sends
 each of its KV pairs to the owning processor.
@@
 the list of KMV
-pairs can be created in two passes through the KV pairs, one to populate
-the hash table with needed count and length information, and the
-second pass to copy the key and value datums into the appropriate
+pairs can be created in two passes through the KV pairs. The first 
+pass populates
+the hash table with needed count and length information; the
+second pass copies the key and value datums into the appropriate
 location in a new KMV data structure.  Since the lookup of a key in a
@@
 {\it reduce()} can serve as input to a subsequent {\it map()} or {\it
-collate()}.  KV pairs from mutiple MapReduce objects can be combined
-together to perform new sequences of {\it map()}, {\it collate()}, and
+collate()}.  KV pairs from multiple MapReduce objects can be combined
+to perform new sequences of {\it map()}, {\it collate()}, and
 {\it reduce()} operations.
@@
 The above discussion assumed that the KV or KMV pairs stored by a
-processor fit in its physical memory.  If this is the case, then we
-refer to this as ``in-core'' processing and no disk files are written
+processor fit in its physical memory.  In this case, MR-MPI 
+performs only ``in-core'' processing and no disk files are written
 or read by any of the processors, aside from initial input data if it
@@
 processors throughout the sequence of MapReduce operations.  The use
-of randomized hashing to assign keys to processors is designed for
+of hashing to assign keys to processors is designed for
 such load-balancing.  In the next section, we discuss what happens 

### 2010-06-07 paper/outcore.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 data for the {\it map()} operation is an existing set of KV pairs,
-then those datums are read in, one page at a time, and a pointer to
+those datums are read, one page at a time, and a pointer to
 each KV pair is given to the user function.  Similarly, a {\it
-reduce()} reads one page of KMV pairs from disk, passes a pointer, one
-at a time, to each pair to the user function which typically generates
+reduce()} reads one page of KMV pairs from disk, passes a pointer
+to each pair, one at a time, to the user function that typically generates
 new KV pairs.  The generated pairs fill up a new page, which is
@@
 with a unique key (the component ID), may not fit in one page of
-memory.  The individual values in the multivalue will then be spread
+memory.  The individual values in the multivalue are then spread
 across as many pages as needed.  The user function that processes the
@@
 performed on one page of KV data at a time.  Each processor reads in a page of
-KV pairs, the communication pattern (which processors will receive 
+KV pairs, the communication pattern (which processors receive 
 which datums in that page) is
@@
 one page of KV pairs; the two-page allocation allows for some
-load-imbalance.  If the KV pairs are distributed unevenly so that this
-limit is exceeded, then the all-to-all communication is performed in
+load imbalance.  If the KV pairs are distributed unevenly so that this
+limit is exceeded, the all-to-all communication is performed in
 smaller, multiple passes until the full page contributed by every
@@
 writes out new KV or KMV pages at most three times.  It also uses a
-finite-size in-memory hash table, which stores unique keys as they are
+finite-size in-memory hash table, that stores unique keys as they are
 encountered while looping over the KV pages, as well as auxiliary
@@
 into ``partition'' files, represented by KVp in the figure.  Each
-partition file contains KV pairs whose unique keys will (likely) fit
+partition file contains KV pairs whose unique keys (likely) fit
 in the hash table (HT).  Initially, all KV pairs are assigned to the
@@
 full, e.g. at the point represented by the horizontal line shown on
-the leftmost vertical line as a downward scan is performed, then the
+the leftmost vertical line as a downward scan is performed, the
 fraction of KV pairs read thus far is used to estimate the number of
-additional partition files needed.  As subsequent KV pairs are read
-in, they are assigned to the original partition file if the KV pair's
-key is already in the current HT.  If not, then a subset of bits in
+additional partition files needed.  As subsequent KV pairs are read,
+they are assigned to the original partition file if the KV pair's
+key is already in the current HT.  If not, a subset of bits in
 the hash value of the key is used to assign the KV pair to one of the
@@
 scanned to determine the total size of KMV for keys in the HT.  
-The associated values may create KMV pairs that will span
-$M$ memory pages.  If $M > 1$ the associated partition file of KV
+The associated values create KMV pairs that span
+$M$ memory pages.  If $M > 1$, the associated partition file of KV
 pairs is read again, and each KV pair is assigned to one of $M$
 smaller ``set'' files, represented by KVs in the figure.  Each set
-file contains the KV pairs that will contribute to the KMV pairs that
-will populate one page of KMV output.  If a single KMV pair spans
-multiple pages, because it contains a very large number of values,
-then the corresponding KV pairs will still be written to a single set
+file contains the KV pairs that contribute to the KMV pairs that
+populate one page of KMV output.  If a single KMV pair spans
+multiple pages because it contains a very large number of values,
+the corresponding KV pairs are still written to a single set
 file.  Eventually, this pass reads all partition files, entailing both
@@
 allocated for each file.  The precise number of simultaneously open
-files is difficult to bound, but typically one or two additional pages'
+files is difficult to bound, but typically one or two additional pages of 
 memory suffice for all needed buffers.  Thus the total number of pages required
@@
 Other MR-MPI library calls discussed in the previous section, such as
-{\it compress()}, {\it convert()}, {\it collapse()}, and {\it gather()},
+{\it clone()}, {\it collapse()}, and {\it gather()},
 can also operate in out-of-core mode, typically with one pass

### 2010-06-07 paper/paper.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7
More.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@625 3fc257bb-8550-0410-81b3-a5c103da83f7
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \usepackage[pdftex]{graphicx}
+\usepackage{url}
 
@@
 \author{
-Steven J.~Plimpton, Karen D.~Devine, Others? \\
+Steven J.~Plimpton and Karen D.~Devine \\
 Sandia National Laboratories \\
@@
 explicit parallel code, but instead provides ``map'' and ``reduce''
-functions which operate independently on elements of a data set
+functions that operate independently on elements of a data set
 distributed across processors.  The library performs needed data
@@
 
-\bibliographystyle{aip}
+\bibliographystyle{abbrv}
 \bibliography{paper}

### 2010-06-07 paper/results.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-We ran each of the algorithms on four R-MAT graphs of
+We ran each of the algorithms on three R-MAT graphs of
 different sizes, each on a varying number of processors.  Details of the
 input data are shown in Table~\ref{t:rmats}.
-The {\it small} problem size (around 8M edges) can typically be used
+The {\it small} problem (around 8M edges) can typically be used
 on a single processor without significant out-of-core operations.
-The {\it medium} problem size (around 134M edges) is a 
+The {\it medium} problem (around 134M edges) is a 
 problem that can be run
@@
 configurations, however, do not necessarily need out-of-core operations.
-The {\it large} problem size (around 2B
+The {\it large} problem (around 2B
 edges) requires out-of-core operations and higher processor counts.
-%The {\it x-large} problem size
+%The {\it x-large} problem 
 %(around 34B edges) requires most of the machine to run.
@@
 of fixed size on a growing number of processors.  Where available, we
-compare the MapReduce algorithm with other parallel implementations.
+compare the MapReduce algorithm with other parallel implementations,
+including more traditional distributed-memory algorithms and 
+multi-threaded algorithms
+in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL} on the Cray XMT.
 We compute parallel efficiency on $p$ processors
@@
 processors; with a larger total memory, a fixed-size problem requires
-less I/O with more processors.  For RMAT-28, where most operation require
-significant out-of-core operation, parallel efficiency ranges from 52\% to
+less I/O with more processors.  For RMAT-28, which requires
+significant out-of-core operations, parallel efficiency ranges from 52\% to
 97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
@@
 
-KDD NOTE:  Add XMT or PBGL results here.
+Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
+R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
+threads generate edges independently.  A global hash table is used to remove
+duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
+implementation.  The differences in execution time are greatest on 64
+processors, where MR-MPI needs the least file I/O for out-of-core operations.
 
@@
 \includegraphics[width=\textwidth]{fig_rmat.pdf}
-\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
+\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) and an MTGL implementation.}
 \label{f:rmat}
@@
 \subsection{PageRank results}
+\label{subsec:results_pagerank}
 
@@
 distributed-memory matrix-based implementation using the linear
-algebra toolkit Trilinos~\cite{Trilinos-Overview} 
-and a multi-threaded
-implementation
-in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
+algebra toolkit Trilinos~\cite{Trilinos-Overview}.
 The matrix-based distributed-memory implementation of PageRank
@@
 residuals and norms of $x$ and $y$.
-In the multi-threaded MTGL~\cite{MTGL} implementation,
-rank propagates via adjacency list traversal
-in a compressed sparse-row data structure.
-To maintain scalability, code must
-be written so that a single thread spawns the loop that processes all
-in-neighbors of a given vertex; this detail enables the compiler to generate
-hotspot-free code.
 
@@
 
-KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
-
 \begin{figure}[htb]
 \includegraphics[width=\textwidth]{fig_pagerank.pdf}
-\caption{Comparison of PageRank implementations using MapReduce,
-Trilinos, and MTGL on R-MAT data sets.}
+\caption{Comparison of PageRank implementations using 
+MR-MPI (Figure~\ref{fig:pr2}) and 
+Trilinos' matrix/vector classes on R-MAT data sets.}
 \label{f:pr}
@@
 Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
-ranged from 50\% to 120\%.
-
-KDD NOTE:  Need to add XMT results; triangle enumeration is not implemented in PBGL.
+ranged from 50\% to 120\%.  Comparisons with an MTGL implementation on the 
+Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
+out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
+outperforms the MTGL implementation.  However, when significant out-of-core
+operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.
 
@@
 \includegraphics[width=\textwidth]{fig_tri.pdf}
-\caption{Performance of the MR-MPI triangle-finding algorithm (Figure~\ref{fig:tri}).}
+\caption{Comparison of triangle-finding implementations using
+MR-MPI (Figure~\ref{fig:tri}) and MTGL.}
 \label{f:tri}
@@
 
-KDD NOTE:  Need to add XMT results.
-
 \begin{figure}[htb]
@@
 
-The execution times for the maximally independent set algorithm
+The execution times for the maximal independent set algorithm
 (Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
@@
 operations. In this case, parallel efficiency is nearly perfect going from 
-8 to 32 processors; some loss of efficiency is seen at 64 processors,
-where the parallel efficiency compared to the 8-processor case is 69\%.
-
-KDD NOTE:  I think I had a bad node for the 64-processor run; I am 
-rerunning it now and soon will have revised RMAT-28 results on 64 processors.
-
-KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
+8 to 64 processors.
 
@@
 \includegraphics[width=\textwidth]{fig_luby.pdf}
-\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
+\caption{Performance of the MR-MPI maximal independent set algorithm (Figure~\ref{fig:luby}).}
 \label{f:luby}
@@
 
-KDD NOTE:  Need to add XMT results; PBGL would be nice, too.
-
 To compare our MR-MPI implementation with a wider set of implementations,
@@
 We observe that the MR-MPI
-implementation runs in less time than the Hadoop implementation,but requires
+implementation runs in less time than the Hadoop implementation, but requires
 significantly more time than the XMT and PBGL implementations.  
@@
 shown, with a 40\% reduction in execution time compared to the original
-SSSP algorith (Figure~\ref{fig:sssp}).
+SSSP algorithm (Figure~\ref{fig:sssp}).
 But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
 Hadoop implementations for {WebGraphB}, which required 110 iterations to 
-complete; execution times for this data set are show in Table~\ref{t:ssspB}.
+complete; execution times for this data set are shown in Table~\ref{t:ssspB}.
 
@@
 implementations on distributed memory and multi-threaded architectures.
+The MR-MPI and Trilinos implementations are described in 
+Sections~\ref{subsec:graph_pagerank} and~\ref{subsec:results_pagerank},
+respectively.
+In the multi-threaded MTGL implementation,
+rank propagates via adjacency list traversal
+in a compressed sparse-row data structure.
+To maintain its scalability, code must
+be written so that a single thread spawns the loop that processes all
+in-neighbors of a given vertex; this detail enables the compiler to generate
+hotspot-free code.
+
 The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
@@
 \includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
-\caption{Scalability comparison of PageRank implementations using MapReduce,
-Trilinos, and MTGL on R-MAT data sets.}
+\caption{Scalability comparison of PageRank using MapReduce (MR-MPI),
+matrix-based (Trilinos), and multi-threaded (MTGL) implementation
+on the R-MAT data sets in Table~\ref{t:rmat}.}
 \label{f:prbig}
@@
 \includegraphics[width=\textwidth]{fig_cc_big.pdf}
-\caption{Scalability comparison of Connected Components implementations using MapReduce,
-Trilinos, and MTGL on R-MAT data sets.}
+\caption{Scalability comparison of Connected Components algorithms using 
+MapReduce (MR-MPI),
+matrix-based/MapReduce hybrid (Trilinos/MR-MPI), and MTGL implementations
+on the R-MAT data sets in Table~\ref{t:rmat}.}
 \label{f:ccbig}

### 2010-06-07 paper/results.tex
Removed XMT R-MAT and Triangle results for now. Change Cohen's affiliation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@629 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
-R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
-threads generate edges independently.  A global hash table is used to remove
-duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
-implementation.  The differences in execution time are greatest on 64
-processors, where MR-MPI needs the least file I/O for out-of-core operations.
+%Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
+%R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
+%threads generate edges independently.  A global hash table is used to remove
+%duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
+%implementation.  The differences in execution time are greatest on 64
+%processors, where MR-MPI needs the least file I/O for out-of-core operations.
 
@@
 \includegraphics[width=\textwidth]{fig_rmat.pdf}
-\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) and an MTGL implementation.}
+\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) 
+%and an MTGL implementation.
+}
 \label{f:rmat}
@@
 Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
-ranged from 50\% to 120\%.  Comparisons with an MTGL implementation on the 
-Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
-out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
-outperforms the MTGL implementation.  However, when significant out-of-core
-operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.
+ranged from 50\% to 120\%.  
+%Comparisons with an MTGL implementation on the 
+%Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
+%out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
+%outperforms the MTGL implementation.  However, when significant out-of-core
+%operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.
 
@@
 \includegraphics[width=\textwidth]{fig_tri.pdf}
-\caption{Comparison of triangle-finding implementations using
-MR-MPI (Figure~\ref{fig:tri}) and MTGL.}
+\caption{Performance of the MR-MPI triangle-finding algorithm
+(Figure~\ref{fig:tri}).}
 \label{f:tri}

### 2010-06-07 paper/thanks.tex
Removed XMT R-MAT and Triangle results for now. Change Cohen's affiliation.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@629 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2009-12-09 paper/thanks.tex
@@
 \section{Acknowledgements}
+\label{sec:thanks}
 
-Sandia is a multiprogram laboratory operated by Sandia Corporation, a
-Lockheed Martin Company, for the United States Department of
-Energy's National Nuclear Security Administration under contract
-DE-AC04-94AL85000.
+We thank the following individuals for their contributions to this
+paper: Greg Bayer and Todd Plantenga (Sandia) for explaining Hadoop
+concepts to us, and for the Hadoop implementations and timings of
+Section~\ref{sec:results}; Jon Cohen (DoD) for fruitful discussions
+about his MapReduce graph algorithms~\cite{Cohen09}; Brian Barrett
+(Sandia) for the PBGL results of Section~\ref{sec:results}; Jon Berry
+(Sandia) for the MTGL results of Section~\ref{sec:results}, and for
+his overall support of this work and many useful discussions.
 
-Hadoop folks.
+Sandia National Laboratories is a multi-program laboratory operated by
+Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
+company, for the U.S. Department of Energy's National Nuclear Security
+Administration under contract DE-AC04-94AL85000.

### 2010-06-07 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 randomized graphs, since they are the input for the algorithms
-discussed below.  R-MAT graphs ~\cite{RMAT} are recursively generated
-graphs with power-law degree distributions.  They are commonly used to
-represent web and social networks.  The user specifices six parameters
-that define the graph: the number of vertices $N$ and edges $M$, and
-four parameters $a$, $b$, $c$, $d$ that sum to 1.0 and are discussed
-below.  The algorithm in Figure~\ref{fig:rmat} generates $M$ unique
-non-zero entries in a sparse $N \times N$ matrix $A$, where each entry
-$A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.
+discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
+generated graphs with power-law degree distributions.  They are
+commonly used to represent web and social networks.  The user
+specifices six parameters that define the graph: the number of
+vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
+that sum to 1.0 and are discussed below.  The algorithm in
+Figure~\ref{fig:rmat} generates $M$ unique non-zero entries in a
+sparse $N \times N$ matrix $A$, where each entry $A_{ij}$ represents
+an edge between graph vertices $(V_i,V_j)$.
 
@@
 
- \caption{MapReduce algorithm for R-MAT graph generation on $P$ processors.}
+ \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
+ processors.}
 
@@
 
-The {\it map()} often generates some small number of duplicate edges.
-The {\it collate()} and {\it reduce()} operations remove the duplicates.  The
-entire map-collate-reduce sequence is repeated until the number of
-resulting key/value pairs $N_{kv} = M$.  For reasonably sparse graphs
-this typically takes only a few iterations.
+Though A is very sparse, the {\it map()} typically generates some
+small number of duplicate edges.  The {\it collate()} and {\it
+reduce()} operations remove the duplicates.  The entire
+map-collate-reduce sequence is repeated until the number of resulting
+key/value pairs $N_{kv} = M$.  For reasonably sparse graphs this
+typically takes only a few iterations.
 
@@
 the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
-values is larger than the other three, a highly skewed distribution
-results.  Variants of the above algorithm can be used when $N$ is not
-a power-of-two, to generate graphs with weighted edges (assign a
-numeric value to the $A_{ij}$ edge), graphs without self edges
-(require $i \ne j$), or graphs with undirected edges (require $i < j$).
-Or the general R-MAT matrix can be further processed by MapReduce
-operations to meet these requirements.  For some algorithms below
-(e.g., triangle finding and maximal independent set generation), we
-post-process R-MAT matrices to create upper-triangular R-MAT
-matrices representing undirected graphs.  For each edge $(i,j)$ 
-with $i>j$, we add edge $(j,i)$, essentially symmetrizing the matrix; 
-we then remove duplicate edges and self edges,
-leaving only edges with $i<j$.
+values is larger than the other three, a skewed distribution results.
+Variants of the above algorithm can be used when $N$ is not a
+power-of-two, to generate graphs with weighted edges (assign a numeric
+value to the $A_{ij}$ edge), graphs without self edges (require $i \ne
+j$), or graphs with undirected edges (require $i < j$).  Or the
+general R-MAT matrix can be further processed by MapReduce operations
+to meet these requirements.  For some algorithms below (e.g., triangle
+finding and maximal independent set generation), we post-process R-MAT
+matrices to create upper-triangular R-MAT matrices representing
+undirected graphs.  For each edge $(i,j)$ with $i>j$, we add edge
+$(j,i)$, essentially symmetrizing the matrix; we then remove duplicate
+edges and self edges, leaving only edges with $i<j$.
 
-By using more than one MapReduce object, we can improve the performance
-of the R-MAT generation algorithm, particularly when out-of-core operations
-are needed.  In the enhanced algorithm
-(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce object
-$E_{new}$ and {\it aggregate()} the edges to processors.  
-We then add those edges
-to the MapReduce object $E_{old}$, containing all previously generated
-unique edges; the edges of $E_{old}$ are already aggregated to processors
-using the same mapping as $E_{new}$.  We can then use the {\it compress()}
-function to reduce duplicates locally on each processor.
-Time reductions
-in this algorithm result from the smaller number of KV pairs communicated in
-{\it aggregate()},
-compared to the collate operation in the original algorithm, 
-particularly in later iterations of the generation where few
-edges are created.
+By using more than one MapReduce object, we can improve the
+performance of the R-MAT generation algorithm, particularly when
+out-of-core operations are needed.  In the enhanced algorithm
+(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
+object $E_{new}$ and {\it aggregate()} the edges to processors.  We
+then add those edges to the MapReduce object $E_{old}$, containing all
+previously generated unique edges; the edges of $E_{old}$ are already
+aggregated to processors using the same mapping as $E_{new}$.  We can
+then use the {\it compress()} function to reduce duplicates locally on
+each processor.  The cost of the algorithm is reduced due to the
+smaller number of KV pairs communicated in the {\it aggregate()},
+compared to the {\it collate()} operation of the original algorithm,
+particularly in later iterations of the algorithm where few edges are
+created.
 
@@
            \> \> output Key = $(i,j)$, Value = NULL \\
-\> Aggregate $E_{new}$. \\
-\> Add $E_{new}$ to $E_{old}$. \\
+\> Aggregate $E_{new}$ \\
+\> Add $E_{new}$ to $E_{old}$ \\
 \> Compress $E_{old}$: Remove duplicate edges \\
@@
 
- \caption{Enhanced MapReduce algorithm for R-MAT graph generation on $P$ processors.}
+ \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
+ on $P$ processors.}
 
@@
 
-The PageRank algorithm assigns a relative numeric rank to each
-vertex in a graph.
-It models the web as a directed graph $G(V,E)$, with each vertex $V_i
-\in V$ representing a web page and each edge $(V_i, V_j) \in E$
-representing a hyperlink from $V_i$ to $V_j$.  The probability of
-moving from $V_i$ to another vertex $V_j$ is $\alpha/d_{out}(V_i) +
-(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
-0.8-0.9), $d_{out}(V_i)$ is the outdegree of vertex $V_i$, and $|V|$ is
-the cardinality of $V$.  The first term represents the probability of
-following a given link on page $V_i$; the second represents the
-probability of moving to a random page.  For pages with no outlinks,
-the first term is $\alpha/|V|$, indicating equal likelihood to move to
-any other page.  Equivalently, the graph can be represented by a
-matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
+The PageRank algorithm assigns a relative numeric rank to each vertex
+in a graph.  In an Internet context, it models the web as a directed
+graph $G(V,E)$, with each vertex $V_i \in V$ representing a web page
+and each edge $(V_i, V_j) \in E$ representing a hyperlink from $V_i$
+to $V_j$.  The probability of moving from $V_i$ to another vertex
+$V_j$ is $\alpha/d_{out}(V_i) + (1-\alpha)/|V|$, where $\alpha$ is a
+user-defined parameter (usually 0.8-0.9), $d_{out}(V_i)$ is the
+outdegree of vertex $V_i$, and $|V|$ is the cardinality of $V$.  The
+first term represents the probability of following a given link on
+page $V_i$; the second represents the probability of moving to a
+random page.  For pages with no outlinks, the first term is
+$\alpha/|V|$, indicating equal likelihood to move to any other page.
+Equivalently, the graph can be represented by a matrix
+$A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
 \alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  To maintain the
-sparsity of $A$, terms for random jumps and zero out-degree vertices are
-not explicitly stored in $A$.  Rather, they are computed separately
-as adjustments to the PageRank vector.  The kernel of the PageRank
-algorithm, then, is a power-method iteration consisting of 
-matrix-vector multiplication $A^T x=y$,
-where $x$ is the PageRank vector from the previous iteration, and
-a few dot product and norm computations.
+sparsity of $A$, terms for random jumps and zero out-degree vertices
+are not explicitly stored as part of the matrix.  Rather, they are
+computed separately as adjustments to the PageRank vector.  The kernel
+of the PageRank algorithm is thus a power-method iteration consisting
+of matrix-vector multiplications $A^T x=y$, where $x$ is the PageRank
+vector from the previous iteration.  A few additional dot product and
+norm computations per iteration are also required.
 
-The algorithm in Figure~\ref{fig:pr2} performs these
-iterations.  The nonzeros of matrix $A$ are initialized as described above;
-the PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
-contains the indices of all-zero rows of $A$, corresponding to vertices
-with no outlinks.  
-The PageRank algorithm is made more efficient by calling
-{\it aggregate()} once on the MapReduce objects representing the matrix $A$, 
-the all-zero matrix rows $MT$, and the vector $x$ before the PageRank 
-iterations begin.  Since these three MapReduce
-objects have the same key types, pre-aggregating the objects moves all
-Key-Value pairs with a given key to a single processor.  Thus, many of
-the PageRank operations become local operations, replacing 
-{\it collate()} operations (which are equivalent to {\it aggregate()} 
-followed by {\it convert()}) with only local {\it convert()} calls. 
-In Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
-substituted for {\it collate()} operations are marked with an asterisk.
-This strategy of pre-aggregating to improve data locality is useful in
-many MapReduce algorithms, but assumes that keys are always mapped to
-the same processor.
+The algorithm in Figure~\ref{fig:pr2} performs these iterations.  The
+nonzeros of matrix $A$ are initialized as described above; the
+PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
+contains the indices of all-zero rows of $A$, corresponding to
+vertices with no outlinks.  The PageRank algorithm is made more
+efficient by calling {\it aggregate()} once on the MapReduce objects
+representing the matrix $A$, the all-zero matrix rows $MT$, and the
+vector $x$ before the PageRank iterations begin.  Since these three
+MapReduce objects have the same key types, pre-aggregating the objects
+moves all KV pairs with a given key to a single processor.  Thus, many
+of the PageRank operations become local operations, replacing {\it
+collate()} operations (which are equivalent to an {\it aggregate()}
+followed by {\it convert()}) with {\it convert()} calls.  In
+Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
+substituted for {\it collate()} operations are marked with an
+asterisk.  This strategy of pre-aggregating to improve data locality
+is useful in many MapReduce algorithms, but assumes that specific keys
+are always mapped to the same processor.
 
-In step 1 of a PageRank interation, $MT$ is used 
-in dot products to compute adjustments to the
-PageRank vector to represent uniform probability of going from a leaf page
-to any other page.  Adjustments for random jumps from any page are also
-computed in this step.  Step 2, the matrix-vector product $A^T x$, is the most
-expensive part of the computation, requiring two passes over the nonzero
-structure of $A$.  In the first pass, {\it convert()} 
-gathers all local row entries $A_{ij}$ with their
-associated $x_i$ entry, and {\it reduce()} computes $A_{ij} x_i$.  
-In the second pass, 
-{\it collate()} gathers, for each $j$, all contributions to the
-column sum $\sum_i A_{ij} x_i$, which is computed by a second {\it reduce()}. 
-Steps 3 and 4 adjust and scale the 
-product vector; the work is proportional to $|V|$.  Step 6 computes the
-residual to determine whether the PageRank vector has converged; work is
-again proportional to $|V|$.  Throughout the algorithm, we exploit the
-ability to call MPI\_Allreduce to compute global norms and residuals.
+In Step 1 of a PageRank interation, $MT$ is used in dot products to
+compute adjustments to the PageRank vector to represent the uniform
+probability of going from a leaf page to any other page.  Adjustments
+for random jumps from any page are also computed.  In Step 2, the
+matrix-vector product $A^T x$ is computed.  This is the most expensive
+part of the computation, requiring two passes over the nonzero
+structure of $A$.  In the first pass, a {\it convert()} gathers all
+local row entries $A_{ij}$ with their associated $x_i$ entry, and a
+{\it reduce()} computes $A_{ij} x_i$.  In the second pass, a {\it
+collate()} gathers, for each $j$, all contributions to the column sum
+$\sum_i A_{ij} x_i$, which is computed by a second {\it reduce()}.
+Steps 3 and 4 adjust and scale the product vector; the work is
+proportional to $|V|$.  Step 5 computes the residual to determine
+whether the PageRank vector has converged; work is again proportional
+to $|V|$.  In Step 6, the PageRank vector is overwritten prior to the
+next iteration.  Throughout the algorithm, we exploit the ability to
+call MPI\_Allreduce to compute global norms and residuals.
 
@@
 %\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
-%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
-%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
+%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
+%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
 %While (${residual} > {tolerance}$) \\
-%\> 1. Compute contribution for random jumps and zero-outdegree vertices. \\
-%\> \> Add: $x$ to $MT$. \\
-%\> \> Collate $MT$:  Vertex is Key. \\
+%\> 1. Compute contribution for random jumps and zero-outdegree vertices \\
+%\> \> Add: $x$ to $MT$ \\
+%\> \> Collate $MT$:  Vertex is Key \\
 %\> \> Reduce $MT$:   \\
-%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
-%\> \> \> Compute:  Local contributions $c_{local}$. \\
-%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
-%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
-%\> 2. Compute $y = A^T x$.  \\
-%\> \> Copy $x$ to $y$. \\
-%\> \> Add $A$ to $y$. \\
-%\> \> Collate $y$:  Key is row index of $A$. \\
+%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
+%\> \> \> Compute:  Local contributions $c_{local}$ \\
+%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
+%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
+%\> 2. Compute $y = A^T x$ \\
+%\> \> Copy $x$ to $y$ \\
+%\> \> Add $A$ to $y$ \\
+%\> \> Collate $y$:  Key is row index of $A$ \\
 %\> \> Reduce $y$: \\
 %\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
-%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$. \\
-%\> \> Collate $y$:  Key is column index of $A$. \\
+%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
+%\> \> Collate $y$:  Key is column index of $A$ \\
 %\> \> Reduce $y$:  \\
 %\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
-%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$. \\
-%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
+%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
+%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
 %\> \> Map $y$:  \\
-%\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$. \\
-%\> \> \> Output:  Key = $i$, Value = $y_i$. \\
-%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
-%\> 4. Scale $y$ by $n_{global}$. \\
+%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
+%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$ \\
+%\> \> \> Output:  Key = $i$, Value = $y_i$ \\
+%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
+%\> 4. Scale $y$ by $n_{global}$ \\
 %\> \> Map $y$: \\
-%\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
-%\> 5. Compute residual. \\
-%\> \> Add $y$ to $x$. \\
-%\> \> Collate $x$:  Key is index $i$. \\
+%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
+%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
+%\> 5. Compute residual \\
+%\> \> Add $y$ to $x$ \\
+%\> \> Collate $x$:  Key is index $i$ \\
 %\> \> Reduce $x$: \\
-%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
+%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
 %\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
-%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
+%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
 %\> 6. $x = y$; \\
-%Output:  PageRank vector $x$.
+%Output:  PageRank vector $x$
   %\end{tabbing}
@@
 %
- %\caption{MapReduce algorithm for PageRank vertex ranking.}
+ %\caption{\it MapReduce algorithm for PageRank vertex ranking.}
 %
@@
 $*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
-$*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0. \\
-$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$. \\
+$*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
+$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
 While (${residual} > {tolerance}$) \\
-\> Compute contribution for random jumps and zero-outdegree vertices. \\
-\> \> Add: $x$ to $MT$. \\
-$*$\> \> Convert $MT$:  Vertex is Key. \\
+\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
+\> \> Add: $x$ to $MT$ \\
+$*$\> \> Convert $MT$:  Vertex is Key \\
 \> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$. \\
-\> \> \> Compute:  Local contributions $c_{local}$. \\
-\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0. \\
-\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$. \\
-\> Compute $y = A^T x$.  \\
-\> \> Copy $x$ to $y$. \\
-\> \> Add $A$ to $y$. \\
-$*$\> \> Convert $y$:  Key is row index of $A$. \\
+\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
+\> \> \> Compute:  Local contributions $c_{local}$ \\
+\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
+\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
+\> 2 Compute $y = A^T x$  \\
+\> \> Copy $x$ to $y$ \\
+\> \> Add $A$ to $y$ \\
+$*$\> \> Convert $y$:  Key is row index of $A$ \\
 \> \> Reduce $y$: \\
 \> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
-\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$. \\
-\> \> Collate $y$:  Key is column index of $A$. \\
+\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
+\> \> Collate $y$:  Key is column index of $A$ \\
 \> \> Reduce $y$:  \\
 \> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
-\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$. \\
-\> Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$. \\
+\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
+\> 3 Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
 \> \> Map $y$:  \\
-\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i$. \\
-\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$. \\
-\> Scale $y$ by $n_{global}$. \\
+\> \> \> Input:  Key = $i$, Value = $y_i$ \\
+\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$ \\
+\> \> \> Output:  Key = $i$, Value = $y_i$ \\
+\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
+\> 4 Scale $y$ by $n_{global}$ \\
 \> \> Map $y$: \\
-\> \> \> Input:  Key = $i$, Value = $y_i$. \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$. \\
-\> Compute residual. \\
-\> \> Add $y$ to $x$. \\
-$*$\> \> Convert $x$:  Key is index $i$. \\
+\> \> \> Input:  Key = $i$, Value = $y_i$ \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
+\> 5 Compute residual \\
+\> \> Add $y$ to $x$ \\
+$*$\> \> Convert $x$:  Key is index $i$ \\
 \> \> Reduce $x$:   \\
-\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$. \\
+\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
 \> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
-\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$. \\
-\> $x = y$; \\
-Output:  PageRank vector $x$.
+\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
+\> 6 $x = y$; \\
+Output: PageRank vector $x$
   \end{tabbing}
@@
 
- \caption{MapReduce algorithm for PageRank vertex ranking; pre-aggregating the MapReduce objects allows many communication intensive {\it collate()} 
-operations to be replaced by strictly local {\it convert()} operations (as marked by the asterisks).}
+ \caption{\it MapReduce algorithm for PageRank vertex ranking;
+pre-aggregating the MapReduce objects allows many communication
+intensive {\it collate()} operations to be replaced by strictly local
+{\it convert()} operations (as marked with asterisks).}
 
@@
 
- \caption{MapReduce algorithm for triangle enumeration.}
+ \caption{\it MapReduce algorithm for triangle enumeration.}
 
@@
 The initial step is to store a copy of the graph edges as key/value
-(KV) pairs in an auxiliary MapReduce object $G_0$, for use later in the
-algorithm.  The first {\it map()} operation converts edge keys to
-vertex keys with edge values.  After the {\it collate()}, each vertex
+(KV) pairs in an auxiliary MapReduce object $G_0$, for use later in
+the algorithm.  The first {\it map()} operation converts edge keys to
+vertex keys with edge values.  After a {\it collate()}, each vertex
 has a list of vertices it is connected to; the first {\it reduce()}
@@
 $D_j$ to the other vertex in each edge.  In the third {\it map()},
-only the lower-degree vertex in each edge emits its edges as key/value
-(KV) pairs.  The task of the third {\it reduce()} is to emit
-``angles'' for each of these low-degree vertices.  An ``angle'' is a
-root vertex $V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a
-triangle without the third edge $(V_1,V_2)$.  The {\it reduce()} emits
-a list of all angles of vertex $V_i$, by a double loop over the edges
-of $V_i$.  Note that the aggregate volume of KV pairs emitted at this
-stage is minimized by having only the low-degree vertex in each edge
-generate angles.
+only the lower-degree vertex in each edge emits its edges KV pairs.
+The task of the third {\it reduce()} is to emit ``angles'' for each of
+these low-degree vertices.  An ``angle'' is a root vertex $V_i$, with
+two edges to vertices $V_1$ and $V_2$, i.e. a triangle without the
+third edge $(V_1,V_2)$.  The {\it reduce()} emits a list of all angles
+of vertex $V_i$, by a double loop over the edges of $V_i$.  Note that
+the aggregate volume of KV pairs emitted at this stage is minimized by
+having only the low-degree vertex in each edge generate angles.
 
-In stage 4, the KV pairs in the original graph $G_0$ are added to the
+In stage 4, KV pairs from the original graph $G_0$ are added to the
 current working set of KV pairs.  The KV pairs in $G_0$ are edges that
@@
 follows the connected-component algorithm presented in~\cite{Cohen09},
-with the addition of logic that load-balances data across processors
+with the addition of logic that load-balances data across processors,
 when one or a few giant components exist in the graph.
@@
 
- \caption{MapReduce algorithm for connected component labeling.}
+ \caption{\it MapReduce algorithm for connected component labeling.}
 
@@
 keys, with the edge as a value.  The current zone assignment of each
-vertex is added to the set of key/value pairs.  The first {\it
+vertex is added to the set of key/value (KV) pairs.  The first {\it
 collate()} operation collects all the edges of a vertex and its zone
 assignment together in one multi-value.  The first {\it reduce()}
-operation re-emits each edge, tagged by the zone assignment of
-one of its vertices.
+operation re-emits each edge, tagged by the zone assignment of one of
+its vertices.
 
@@
 If the two zone IDs are different, the second {\it reduce()} operation
-chooses a winner (the smaller of the two IDs), and emits the loser ID as a
-key, with the winning ID as a value.  If no zone ID changes are
-emitted, the algorithm is finished, and the iteration exits.
+chooses a winner (the smaller of the two IDs), and emits the loser ID
+as a key, with the winning ID as a value.  If no zone ID changes are
+emitted, the algorithm is finished, and the iterations cease.
 
-The third {\it map()} operation inverts the vertex/zone key/value
-pairs to become zone/vertex pairs.  The third {\it add()} operation
-adds the changing zone assignments to the set of key/value pairs.  The
-third {\it collate()} can then collect all the vertices of a zone and zero
-or more reassignments for the zone ID.  Since a zone could collide
-with multiple other zones on the same iteration due to shared edges,
-the new zone ID becomes the minimum ID of any of the neighboring
-zones.  If no zone reassignment value appears in the multi-value, the
-zone ID is unchanged.  The final {\it reduce()} emits a key/value pair for each
+The third {\it map()} operation inverts the vertex/zone KV pairs to
+become zone/vertex pairs.  The third {\it add()} operation adds the
+changing zone assignments to the set of KV pairs.  The third {\it
+collate()} can then collect all the vertices of a zone and zero or
+more reassignments for the zone ID.  Since a zone could collide with
+multiple other zones on the same iteration due to shared edges, the
+new zone ID becomes the minimum ID of any of the neighboring zones.
+If no zone reassignment value appears in the multi-value, the zone ID
+is unchanged.  The final {\it reduce()} emits a KV pair for each
 vertex in the zone, with the vertex as a key and the new zone ID as a
@@
 Note that if a graph has only a few components, the third {\it
-collate()} operation, which collates by the zone ID, may generate a few
-very large key/multi-value (KMV) pairs.  For example, if the graph is
-fully connected, in the last iteration, a single KMV pair 
-contains all vertices in the graph and is assigned to one processor.
-This imbalance in memory and computational work can lead to poor
-parallel performance of the overall algorithm.  To counter this
-effect, the various operations of stage 3 include extra logic.  The
-idea is to partition zones whose vertex count exceeds a user-defined
-threshhold into $P$ sub-zones, where $P$ is the number of processors.
-The 64-bit integer that stores the zone ID also stores a bit flag
-indicating the zone has been partitioned and a set of bits that encode
-the processor ID.
+collate()} operation, which collates by the zone ID, may generate a
+few very large key/multi-value (KMV) pairs.  For example, if the
+entire graph is fully connected, in the last iteration, a single KMV
+pair contains all vertices in the graph and is assigned to one
+processor.  This imbalance in memory and computational work can lead
+to poor parallel performance of the overall algorithm.  To counter
+this effect, the various operations of stage 3 include extra logic.
+The idea is to partition zones whose vertex count exceeds a
+user-defined threshhold into $P$ sub-zones, where $P$ is the number of
+processors.  The 64-bit integer that stores the zone ID also stores a
+bit flag indicating the zone has been partitioned and a set of bits
+that encode the processor ID.
 
@@
 partitioned, the third {\it add()} operation emits the zone ID change
-key/value pair ($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In
-this case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times,
-once for each processor, and once as if $Z_i$ had not been partitioned
-(for a reason discussed below).
+KV pair ($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In this case
+($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once for
+each processor, and once as if $Z_i$ had not been partitioned (for a
+reason discussed below).
 
-With this additional partitioning logic, the third {\it
-collate()} operation (which keys on zone IDs, some of which now
-include processor bits) collects only $1/P$ of the
-vertices in large zones onto each processor.  But the multivalue on
-each processor contains all the zone-reassignments relevant to the
-unpartitioned zone.  Thus, the third {\it reduce()} operation can
-change the zone ID (if necessary) in a consistent manner across all
-$P$ multi-values that contain the zone's vertices.  When the {\it
-reduce()} operation emits new zone assignments for each vertex, the
-zone retains its partitioned status; the partition bit is also
-explicitly set if the vertex count exceeds the threshhold for the
-first time.
+With this additional partitioning logic, the third {\it collate()}
+operation (which keys on zone IDs, some of which now include processor
+bits) collects only $1/P$ of the vertices in large zones onto each
+processor.  But the multivalue on each processor contains all the
+zone-reassignments relevant to the unpartitioned zone.  Thus, the
+third {\it reduce()} operation can change the zone ID (if necessary)
+in a consistent manner across all $P$ multi-values that contain the
+zone's vertices.  When the {\it reduce()} operation emits new zone
+assignments for each vertex, the zone retains its partitioned status;
+the partition bit is also explicitly set if the vertex count exceeds
+the threshhold for the first time.
 
 Note that this logic does not guarantee that the partition bits of the
-zone IDs for all the vertices in a single zone are set
-consistently on a given iteration.  For example, an unpartitioned zone
-with a small ID may consume a partitioned zone.  The vertices from the
-partitioned zone retain their partitioned status, but the
-original vertices in the small zone may not set the partition bit of
-their zone IDs.  On subsequent iterations, the third {\it add()}
-operation emits $P+1$ copies of new zone reassignments for both
-partitioned and unpartioned zone IDs, to ensure all vertices in the
-zone know the reassignment information.
+zone IDs for all the vertices in a single zone are set consistently on
+a given iteration.  For example, an unpartitioned zone with a small ID
+may consume a partitioned zone.  The vertices from the partitioned
+zone retain their partitioned status, but the original vertices in the
+small zone may not set the partition bit of their zone IDs.  On
+subsequent iterations, the third {\it add()} operation emits $P+1$
+copies of new zone reassignments for both partitioned and unpartioned
+zone IDs, to ensure all vertices in the zone know the correct
+reassignment information.
 
@@
 
-An ``independent'' set of vertices in a graph is one where no pair of
-vertices in the set shares an edge.  The set is ``maximal'' if no
-additional vertex can be added \footnote{A maximal set is ``maximum''
-if it is the largest maximal set.  Finding a graph's maximum
-independent set is an NP-hard problem, which MapReduce is unlikely to
-help with.}.  Finding a maximal independent set (MIS) is useful in
-several contexts, such as identifying large numbers of independent
-starting vertices for graph traversals.  Luby's algorithm~\cite{Luby86}
-is a well-known parallel method for finding a MIS.  Figure~\ref{fig:luby} 
-outlines a MapReduce version of Luby's algorithm,
-assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
-V_j$ for every edge, i.e. an upper-triangular R-MAT matrix.
+An ``independent'' set of vertices from a graph is one where no pair
+of vertices in the set shares an edge.  The set is ``maximal'' if no
+vertex can be added \footnote{A maximal set is ``maximum'' if it is
+the largest maximal set.  Finding a graph's maximum independent set is
+an NP-hard problem, which MapReduce is unlikely to help with.}.
+Finding a maximal independent set (MIS) is useful in several contexts,
+such as identifying large numbers of independent starting vertices for
+graph traversals.  Luby's algorithm~\cite{Luby86} is a well-known
+parallel method for finding a MIS.  Figure~\ref{fig:luby} outlines a
+MapReduce version of Luby's algorithm, assuming an input graph of
+undirected edges $(V_i,V_j)$ where $V_i < V_j$ for every edge, i.e. an
+upper-triangular R-MAT matrix.
 
@@
 Map: assign random values to each vertex $V_i$ and $V_j$ of edge $E_{ij}$. \\
-Clone: convert edge key/values directly to key/multivalues \\
+Clone: convert edge KV pairs directly to KMV pairs \\
 Create: empty MapReduce object MRv for maximal independent set vertices \\
@@
 
- \caption{MapReduce algorithm for finding a maximal independent set of
- graph vertices via Luby's algorithm.}
+ \caption{\it MapReduce algorithm for finding a maximal independent
+ set of graph vertices via Luby's algorithm.}
 
@@
 Before the iterations begin, a random value is assigned to each
-vertex, which is used to compare pairs of vertices.  The vertex
-ID itself could be used as the random value, but since the vertices
+vertex, which is used to compare pairs of vertices.  The vertex ID
+itself could be used as the random value, but since the vertices
 eventually selected for the MIS depend on the randomization, this may
@@
 used to assign a consistent random value $R_i$ to each vertex in each
-edge (via a {\it map()} operation), which is then carried along with the
-vertex ID through each stage of the algorithm.  In the notation of
+edge (via a {\it map()} operation), which is then carried along with
+the vertex ID through each stage of the algorithm.  In the notation of
 Figure~\ref{fig:luby}, each $V_i$ is then really two quantities, a
@@
 
-The pre-iteration {\it clone()} operation converts the initial list of edge
-key/values one by one to key/multivalue pairs, with the edge as the
-key, and NULL as the value.  Thus, the iterations can begin with
-a {\it reduce()} operation, without need for a {\it collate()} operation.  A
-second empty MapReduce object is also initialized, which 
-accumulates MIS vertices as the iterations proceed.
+The pre-iteration {\it clone()} operation converts the initial list of
+edge key/value (KV) pairs one by one to key/multivalue (KMV) pairs,
+with the edge as the key, and NULL as the value.  Thus, the iterations
+can begin with a {\it reduce()} operation, without need for a {\it
+collate()} operation.  A second empty MapReduce object is also
+initialized, which accumulates MIS vertices as the iterations proceed.
 
@@
 All edges of each vertex are also emitted, retaining the loser flag if
-they have it;  they are emitted with $E_{ij} = (V_i, V_j)$ 
-as their key, with $V_i < V_j$.  
-This is to ensure both copies of the edge come together via the
-fourth {\it collate()} operation.  The multivalue for each edge now
-has two values, either of which may be NULL or a loser flag.
+they have it; they are emitted with $E_{ij} = (V_i, V_j)$ as their
+key, with $V_i < V_j$.  This is to ensure both copies of the edge come
+together via the fourth {\it collate()} operation.  The multivalue for
+each edge now has two values, either of which may be NULL or a loser
+flag.
 
@@
 
-The single-source shortest path (SSSP) algorithm computes, for a directed graph
-with weighted edges,
-the shortest weighted distance from a given source vertex to
-all other vertices in the graph.  This operation is straightforward
-when global graph information is available.  The source vertex is labeled with
-distance zero.  Then in each iteration, edge adjacencies of labeled vertices
-are followed, and adjacent vertices are labeled with updated distances.  When
-no distances change, the iterations are complete.  Most importantly, only
+The single-source shortest path (SSSP) algorithm computes, for a
+directed graph with weighted edges, the shortest weighted distance
+from a chosen source vertex to all other vertices in the graph.  This
+operation is straightforward when global graph information is
+available.  The source vertex is labeled with distance zero.  Then in
+each iteration, edge adjacencies of labeled vertices are followed, and
+adjacent vertices are labeled with updated distances.  When no
+distances change, the iterations are complete.  Most importantly, only
 edges from modified vertices are visited in each iteration.
 
-In MapReduce, global information about edge adjacencies and labeled vertices
-is not available.  Instead, following the Bellman-Ford-style 
-algorithm of~\cite{SSSPMapReduce, Bellman58, Ford62},
-every edge of the graph is visited in each iteration,
-although labels for an edge's vertices are updated only if the breadth-first
-traversal has reached those vertices.  The algorithm begins with a {\it map()}
-function loading edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.
-In each iteration, the edges $(V_i, V_j)$ are collated with respect to
-$V_i$, and a {\it reduce()} operation updates each $V_j$'s distance if
-$V_i$'s distance is non-zero.  The algorithm is fully described in
-Figure~\ref{fig:sssp}.
+For distributed graphs, global information about edge adjacencies and
+labeled vertices is not available.  Instead, following the
+Bellman-Ford-style algorithm of~\cite{SSSPMapReduce, Bellman58,
+Ford62}, every edge of the graph is visited in each iteration,
+although labels for an edge's vertices are updated only if the
+breadth-first traversal has reached those vertices.  A MapReduce
+formulation of this algorithm, described in Figure~\ref{fig:sssp},
+begins with a {\it map()} function loading edges $(V_i, V_j)$ and
+source vertex $V_s$ with distance 0.  In each iteration, the edges
+$(V_i, V_j)$ are collated with respect to $V_i$, and a {\it reduce()}
+operation updates each $V_j$'s distance if $V_i$'s distance is
+non-zero.
 
@@
 Input:  \\
-\> Source vertex $V_s$.\\
+\> Source vertex $V_s$\\
 \> MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$ (initially $\infty$):  \\
-\> \> Key = $V_i$; Value = $\infty$.\\
+\> \> Key = $V_i$; Value = $\infty$\\
 \> MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
 \> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
-Map $V$:  Set distance for source vertex $V_s$ to zero. \\
+Map $V$:  Set distance for source vertex $V_s$ to zero \\
 \> Input:  Key = $V_i$; Value = $\infty$ \\
-\> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$. \\
+\> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$ \\
 while not $done$: \\
 \> $done$ = 1; \\
-\> Add $E$ to $V$. \\
-\> Collate $V$:  Key is vertex $V_i$. \\
-\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
-\> \> Input:  Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$, candidate distances $d_i$. \\
-\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
+\> Add: $E$ to $V$ \\
+\> Collate $V$:  Key is vertex $V_i$ \\
+\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices \\
+\> \> Input:  Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$, candidate distances $d_i$ \\
+\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$  \\
 \> \> \> If $d_{min} \ne d_i$, $done$ = 0; \\
 \> \> Output:  Key = $V_i$; Value = $d_{min}$ (updated distance for $V_i$) \\
-\> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$). \\
-\> MPI\_Allreduce on $done$ with MPI\_MIN. \\
+\> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$) \\
+\> MPI\_Allreduce on $done$ with MPI\_MIN \\
 Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
@@
 
- \caption{MapReduce algorithm for single-source shortest path (SSSP)
- calculation.}
+ \caption{\it MapReduce algorithm for single-source shortest path
+ (SSSP) calculation.}
 
@@
 
-As with PageRank and R-MAT generation, 
-significant savings in communication and execution time
-can be achieved by pre-aggregating certain MapReduce objects and 
-storing data in more than one MapReduce object.  The enhanced algorithm
-is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects storing
-vertices and edges are aggregated to processors only once at the beginning
-of the computation.  In each iteration of 
-SSSP, only a small number of new candidate distances are generated.
-The enhanced algorithm uses MapReduce object $U$ to store these updates.
-Only the updates in $U$ are communicated; each update is aggregated to the 
-same processor that stores the corresponding vertex and its edges.
-Thus, the total amount of communication
-needed is small compared to the original algorithm (which aggregates all
-graph edges in each iteration).  Updated vertex distances are stored back
-in the the MapReduce object for vertices.
+As with PageRank and R-MAT generation, significant savings in
+communication and execution time can be achieved by pre-aggregating
+certain MapReduce objects and storing data in more than one MapReduce
+object.  The enhanced algorithm is shown in Figure~\ref{fig:sssp2}.
+The MapReduce objects storing vertices and edges are aggregated to
+processors only once at the beginning of the computation.  In each
+iteration of SSSP, only a small number of new candidate distances are
+generated.  The enhanced algorithm uses MapReduce object $U$ to store
+these updates.  Only the updates in $U$ are communicated; each update
+is aggregated to the same processor that stores the corresponding
+vertex and its edges.  Thus, the total amount of communication needed
+is smaller than in the original algorithm, which aggregates all graph
+edges in each iteration.  Updated vertex distances are stored in
+MapReduce object $V$.
 
@@
 \> if not $done$ \\
-\> \> Add $U$ to $E$.  Empty $U$. \\
+\> \> Add: $U$ to $E$.  Empty $U$. \\
 \> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
@@
 
- \caption{Enhanced MapReduce algorithm for single-source shortest path (SSSP)
- calculation.  Pre-aggregating the vertices and edges, and later aggregating only updates to vertex distances reduces communication and execution time.}
+ \caption{\it Enhanced MapReduce algorithm for single-source shortest
+ path (SSSP) calculation.  Pre-aggregating the vertices and edges, and
+ later aggregating only updates to vertex distances, reduces
+ communication and execution time.}
 

### 2010-06-07 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 For map and reduce operations, the library calls back to user-provided
-{\it map()} and {\it reduce()} functions.  The use of C++ allows precise control
-over the memory and format of data allocated by each processor during
-a MapReduce.  Library calls for performing a map, reduce, or data
-shuffle, are synchronous, meaning all the processors participate and
-finish the operation before proceeding.  Similarly, the use of MPI
-within the library is the traditional mode of MPI\_Send and
-MPI\_Recv calls between processor pairs using large aggregated
+{\it map()} and {\it reduce()} functions.  The use of C++ allows
+precise control over the memory and format of data allocated by each
+processor during a MapReduce.  Library calls for performing a map,
+reduce, or data shuffle, are synchronous, meaning all the processors
+participate and finish the operation before proceeding.  Similarly,
+the use of MPI within the library is the traditional mode of MPI\_Send
+and MPI\_Recv calls between pairs of processors using large aggregated
 messages to improve bandwidth performance and reduce latency costs.  A
@@
 
-\item {\it Small, portable:} The entire MR-MPI library is a few thousand
-lines of standard C++ code.  For parallel operation, the program is
-linked with MPI, a standard message passing library available on all
-distributed memory machines.  For serial operation, a dummy MPI
-library (provided) can be substituted.  As a library, it can be
-embedded in other codes~\cite{Titan} to enable them to
-perform MapReduce operations.
+\item {\it Small, portable:} The entire MR-MPI library is a few
+thousand lines of standard C++ code.  For parallel operation, the
+program is linked with MPI, a standard message passing library
+available on all distributed memory machines.  For serial operation, a
+dummy MPI library (provided) can be substituted.  As a library, it can
+be embedded in other codes~\cite{Titan} to enable them to perform
+MapReduce operations.
 
-\item {\it In-core or out-of-core operation:} Each MapReduce object that a
-processor defines allocates per-processor ``pages'' of memory, where
-the page size is determined by the user.  Typical MapReduce operations
-can be performed using a few such pages.  If the data set fits in a
-single page (per processor), the library performs its operations
-in-core.  If the data set exceeds the page size, processors each
-write to temporary disk files (on local disks or a parallel file
+\item {\it In-core or out-of-core operation:} Each MapReduce object
+that a processor defines allocates ``pages'' of memory, where the page
+size is determined by the user.  Typical MapReduce operations can be
+performed using a few such pages (per processor).  If the data set
+fits in a single page (per processor), the library performs its
+operations in-core.  If the data set exceeds the page size, processors
+each write to temporary disk files (on local disks or a parallel file
 system) as needed and subsequently read from them.  This out-of-core
-operation allows
-processing of data sets larger than the aggregate memory of all the
-processors, i.e. up to the available aggregate disk space.
+operation allows processing of data sets larger than the aggregate
+memory of all the processors, i.e. up to the available aggregate disk
+space.
 
@@
 The library interface also provides a user data pointer as an argument
-passed back to all callback functions, so it is easy for the user
-program to store ``state'' on each processor, accessible during the
-map and reduce operations.  For example, various flags can be stored
-that alter the operation of a map or reduce operation, as can richer
-data structures that accumulate results.
+passed to all callback functions, so it is easy for the user program
+to store ``state'' on each processor, accessible during the map and
+reduce operations.  For example, various flags can be stored that
+alter the operation of a map or reduce operation, as can richer data
+structures that accumulate results.
 
@@
 results for these algorithms running on a parallel cluster for graphs
-%KDDKDD CHECK SIZES
-ranging in size from 8 million to 2 billion edges;  
-we highlight the performance and complexity trade-offs
-of a MapReduce approach versus other more special-purpose algorithms.
-The latter generally perform better but are harder to implement
-efficiently on distributed memory machines, due to the required
-explicit management of parallelism, particularly for large out-of-core
-data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
-from the implementation and use of our library.
+ranging in size from 8 million to 2 billion edges; we highlight the
+performance and complexity trade-offs of a MapReduce approach versus
+other more special-purpose algorithms.  The latter generally perform
+better but are harder to implement efficiently on distributed memory
+machines, due to the required explicit management of parallelism,
+particularly for large out-of-core data sets.
+Section~\ref{sec:lessons} summarizes some lessons learned from the
+implementation and use of our library.

### 2010-06-07 paper/lessons.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 data across processors, which often intentionally ignores data
-locality.  This translates into maximal data movement (during a
-shuffle) with communication between all pairs of processors.  But the
-benefit is often good load-balance, even for hard-to-balance irregular
-data sets.  By contrast, more traditional distributed-memory parallel
-algorithms, e.g. for matrix operations, or grid-based or
-particle-based simulation codes, tend to work hard to localize data
-and minimize communication.  To do this they typically require a lot
-of application-specific logic and parallel communication coding to
-create and maintain a data decomposition, generate ghost versions of
-nearby spatially-decomposed data, etc.
+locality.  This translates into maximal data movement (during a data
+shuffle or {\it collate()} operation) with communication between all
+pairs of processors.  But the benefit is often good load-balance, even
+for hard-to-balance irregular data sets.  By contrast, more
+traditional distributed-memory parallel algorithms, e.g. for matrix
+operations, or grid-based or particle-based simulation codes,
+typically work hard to localize data and minimize communication.  To
+do this they often require a lot of application-specific logic and
+parallel communication coding to create and maintain a data
+decomposition, generate ghost versions of nearby spatially-decomposed
+data, etc.
 
@@
 seems heretical to intentionally ignore data locality.  However,
-writing small {\it map()} and {\it reduce()} functions is typically easy.  And
-writing an algorithm that involves complex parallel operations
-without actually needing to write application-specific parallel code
-to move data via MPI calls is often a pleasant surprise.  Moreover,
-if the MapReduce algorithm is initially coded so that it runs
-correctly on one processor, it often works out-of-the-box on hundreds
-or thousands of processors, without the need for additional debugging.
+writing small {\it map()} and {\it reduce()} functions is typically
+easy.  And writing an algorithm that involves complex parallel
+operations without needing to write application-specific code to
+communicate data via MPI calls or to move data between out-of-core
+disk files and processor memory is often a pleasant surprise.
+Moreover, if the MapReduce algorithm is initially coded so that it
+runs correctly on one processor, it often works out-of-the-box on
+hundreds or thousands of processors, without the need for additional
+debugging.
 
@@
 a traditional MPI-based parallel machine is a somewhat different
-conceptual model than that of cloud-computing MapReduces using (for
-example) Hadoop.  In the former case, one can potentially control
-which processor owns which data at various stages of an algorithm.
-This is somewhat hidden from the user in a typical cloud-computing
-model, where data simply exists somewhere in the cloud and Hadoop
-ensures data moves where it is needed and is operated on by some
-processor.  The cloud model is a nice data-centric abstraction which
-allows for fault tolerance both to data loss (via redundant storage)
-and to processor failure (via reassignment of work), neither of which
-is typically possible on current MPI-based parallel machines.
+conceptual model than that of performing MapReduce operations on a
+cloud computer using (for example) Hadoop.  In the former case, one
+can potentially control which processor owns which data at various
+stages of an algorithm.  This is somewhat hidden from the user in a
+typical cloud-computing model, where data simply exists somewhere in
+the cloud and Hadoop ensures data moves where it is needed and is
+operated on by some processor.  The cloud model is a nice data-centric
+abstraction which allows for fault tolerance both to data loss (via
+redundant storage) and to processor failure (via reassignment of
+work), neither of which is typically possible on current MPI-based
+parallel machines.
 
-However, since the MPI implementation of MapReduce as
-described in this paper is processor-centric, one can sometimes
-fruitfully exploit the possibility for processors to maintain
-``state'' over the course of multiple map and reduce operations.  By
-controlling where data resides for maps and reduces (e.g. via a
-user-specified hash function), and by assuming that processor will
-always be available, more efficient operations with less data movement
-are sometimes possible.  The discussion of enhanced graph algorithms
-in Section \ref{sec:graph} illustrated this.  To fully exploit this
-idea for large data sets, mechanisms are needed for processors to
-store, retrieve, and efficiently find needed datums on local disk
-(e.g. static edges of a graph), so that archived state can be used
-efficiently during subsequent map and reduce operations.
+However, since the MPI implementation of MapReduce as described in
+this paper is processor-centric, one can sometimes fruitfully exploit
+the possibility for processors to maintain ``state'' over the course
+of multiple map and reduce operations.  By controlling where data
+resides for maps and reduces (e.g. via a user-specified hash
+function), and by assuming that processor will always be available,
+more efficient operations with less data movement are sometimes
+possible.  The discussion of enhanced graph algorithms in Section
+\ref{sec:graph} illustrated this.  To fully exploit this idea for
+large data sets, mechanisms and data structures are needed for
+processors to store, retrieve, and efficiently find needed datums on
+local disk (e.g. static edges of a graph), so that archived state can
+be used efficiently during subsequent map and reduce operations.
 
@@
 MapReduce operations, there is nothing about the MR-MPI library itself
-that is graph centric.  We hope the library can be generally useful on
-large-scale monolithic or cloud-style parallel machines that support
-MPI, for a variety of data-intense or compute-intense problems that
-are amenable to solution using a MapReduce paradigm.
+that is graph specific.  We hope the library can be generally useful
+on large-scale monolithic or cloud-style parallel machines that
+support MPI, for a variety of data-intensive or compute-intensive
+problems that are amenable to solution using a MapReduce paradigm.

### 2010-06-07 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 etc); they are simply treated as byte strings by the library.  A KV
-pair always has a key; a KV's value may be NULL.  A related data type is
-the key/multivalue (KMV) pair, where all values associated with the
+pair always has a key; a KV's value may be NULL.  A related data type
+is the key/multivalue (KMV) pair, where all values associated with the
 same key are collected and stored contiguously as a multivalue, which
-is just a longer byte string with an associated vector of integer
-lengths, one per value.
+is just a longer byte string with an associated vector of lengths, one
+integer length per value.  In this section, we assume the datums
+operated on by each processor fit in local memory.
 
-A typical MR-MPI program makes at least three calls to the MR-MPI library,
-to perform a {\it map()}, {\it collate()}, and {\it reduce()}
-operation.  In a {\it map()}, zero or more key/value pairs are
-generated by each processor.  Typically, this is done using data read
-from files, but a {\it map()} may generate data itself or process
-existing KV pairs to create new ones.  The KV pairs produced are
-stored locally by each processor; a {\it map()} thus requires no
-inter-processor communication.  Users call the library with a count of
-tasks to perform and a pointer to a user function; the MR-MPI {\it
-map()} operation invokes the user function multiple times as a
-callback.  Depending on which variant of {\it map()} is called, the
-user function may be passed a file name, a chunk of bytes from a large
-file, a task ID, or a KV pair.  Options for assigning map tasks to processors
-are specified by the user and include assigning consecutive 
-chunks of tasks to processors, striding the tasks across processors,
-or using a master-slave model that is useful when tasks have widely
-varying workloads.
+A typical MR-MPI program makes at least three calls to the MR-MPI
+library, to perform {\it map()}, {\it collate()}, and {\it reduce()}
+operations.  In a {\it map()}, zero or more key/value pairs are
+generated by each processor.  Often this is done using data read from
+files, but a {\it map()} may generate data itself or process existing
+KV pairs to create new ones.  The KV pairs produced are stored locally
+by each processor; a {\it map()} thus requires no inter-processor
+communication.  Users call the library with a count of tasks to
+perform and a pointer to a user function; the MR-MPI {\it map()}
+operation invokes the user function multiple times as a callback.
+Depending on which variant of {\it map()} is called, the user function
+may be passed a file name, a chunk of bytes from a large file, a task
+ID, or a KV pair.  Options for assigning map tasks to processors are
+specified by the user and include assigning consecutive chunks of
+tasks to processors, striding the tasks across processors, or using a
+master-slave model that is useful when tasks have widely varying
+workloads.
 
@@
 by any processor.  Each processor hashes each of its keys to determine
-which processor will ``own'' it.  The $k$-byte length key is hashed into
-a 32-bit value whose remainder modulo $P$ processors generates the owning
-processor rank.  Alternatively, the user can provide a hash function
-which converts the key into a processor rank.  Each processor then sends
-each of its KV pairs to the owning processor.
+which processor will ``own'' it.  The $k$-byte length key is hashed
+into a 32-bit value whose remainder modulo $P$ processors is the
+owning processor rank.  Alternatively, the user can provide a hash
+function which converts the key into a processor rank.  Each processor
+then sends each of its KV pairs to the owning processor.
 
@@
 owns.  This is done using a hash table, rather than a sort.  A sort
-requires $\log(N)$ passes through the $N$ KV pairs.  With hashing,
-the list of KMV
-pairs can be created in two passes through the KV pairs. The first 
-pass populates
-the hash table with needed count and length information; the
-second pass copies the key and value datums into the appropriate
-location in a new KMV data structure.  Since the lookup of a key in a
-well-formed hash table is a constant-time $O(1)$ operation, the cost
-of the data reorganization is $O(N)$, rather than $O(N\log(N))$ for full
-sorting.  It also has the added benefit that the number of values in each
-KMV pair is known and is passed to the user function during a
-reduce.  For some reduce operations, this count is all the information a
-reduce requires; thus the values need not be looped over.  This feature is not
-available in Hadoop-style data reorganizations, which sort the values; 
-with sorting, the number of values associated with the same key is not known 
-without interating over the values.
+scales as $N\log_2(N)$, i.e. it requires $\log_2(N)$ passes through
+the $N$ KV pairs.  With hashing, the list of KMV pairs can be created
+in two passes.  The first pass populates the hash table with needed
+count and length information; the second pass copies the key and value
+datums into the appropriate location in a new KMV data structure.
+Since the cost to lookup a key in a well-formed hash table is a
+constant-time $O(1)$ operation, the cost of the data reorganization is
+also $O(N)$.  This methodology has the added benefit that the number
+of values in each KMV pair is known and can be passed to the user
+function during a reduce.  For some reduce operations, this count is
+all the information a reduce requires; the values need not be looped
+over.  The count is not available in Hadoop-style data shuffles, which
+sort the values; counting the values associated with the same key
+requires iterating over the values.
 
@@
 all-to-all communication (each processor sends and receives data from
-every other processor) driven by a distributed hash table.  The
+every other processor) using a distributed hash table.  The
 communication can either be done via a MPI\_Alltoall() library call,
-or by a custom routine that aggregates messages and invokes pointwise
-MPI\_Send() and MPI\_IRecv() calls.
+or by a custom routine that aggregates messages and invokes
+point-to-point MPI\_Send() and MPI\_IRecv() calls.
 
@@
 
-Several related MapReduce operations are provided by the library and
-exposed in the library interface.  For example, the {\it collate()} function
-described above
-calls two other functions:  {\it aggregate()}, which performs the all-to-all
-communication, and {\it convert()}, which turns a list of KV pairs into
-KMV pairs.  
-Both {\it aggregate()} and {\it convert()} are exposed in the MR-MPI
-interface for direct access by the user.
-The {\it compress()} function allows on-processor operations to be 
-performed; it is equivalent to a {\it convert()} with on-processor KVs as
-input, followed by a {\it reduce()}.  
-The {\it clone()} operation turns a list of KV pairs into
-KMV pairs, with one value per key.  The {\it collapse()} operation turns
-$N$ KV pairs into one KMV pair, with the keys and values of the KV
-pairs becoming $2N$ values of a single multivalue assigned to a new
-key on the KMV.  The {\it gather()} operation collects KV pairs from
-all processors to a subset of processors; it is useful for doing
-output from a single or subset of processors.  Library calls for
-sorting datums by key or value or within multivalues are also
-provided.  These routines invoke the C-library {\it quicksort()} function to
-compute the sorted ordering of the KV pairs (or values within a
-multivalue), using a user-provided comparison function.  The KV pairs
-(or values in a multivalue) are then copied one-by-one into a new data
-structure in sorted order.
+Several related MapReduce operations are provided by the library.  For
+example, the {\it collate()} function described above calls two other
+functions: {\it aggregate()}, which performs the all-to-all
+communication, and {\it convert()}, which turns a list of KV pairs
+into KMV pairs.  Both functions can be called directly.  The {\it
+compress()} function allows on-processor operations to be performed;
+it is equivalent to a {\it convert()} with on-processor KVs as input,
+followed by a {\it reduce()}.  The {\it clone()} function turns a list
+of KV pairs into KMV pairs, with one value per key.  The {\it
+collapse()} function turns $N$ KV pairs into one KMV pair, with the
+keys and values of the KV pairs becoming $2N$ values of a single
+multivalue assigned to a new key.  The {\it gather()} function
+collects KV pairs from all processors to a subset of processors; it is
+useful for doing output from one or a few processors.  Library calls
+for sorting datums by key or value or for sorting the values within
+each multivalue are also provided.  These routines invoke the
+C-library {\it quicksort()} function to compute the sorted ordering of
+the KV pairs (or values within a multivalue), using a user-provided
+comparison function.  The KV pairs (or values in a multivalue) are
+then copied into a new data structure in sorted order.
 
@@
 The above discussion assumed that the KV or KMV pairs stored by a
-processor fit in its physical memory.  In this case, MR-MPI 
-performs only ``in-core'' processing and no disk files are written
-or read by any of the processors, aside from initial input data if it
-exists or final output data if it is generated.  Note that the
-aggregate physical memory of large parallel machines can be multiple
-terabytes, which allows for large data sets to be processed in-core,
-assuming the KV and KMV pairs remain evenly distributed across
-processors throughout the sequence of MapReduce operations.  The use
-of hashing to assign keys to processors is designed for
-such load-balancing.  In the next section, we discuss what happens 
-when data sets do not fit in available memory.
+processor fit in its physical memory.  In this case, MR-MPI performs
+only ``in-core'' processing and no disk files are written or read by
+any of the processors, aside from initial input data if it exists or
+final output data if it is generated.  Note that the aggregate
+physical memory of large parallel machines can be multiple terabytes,
+which allows for large data sets to be processed in-core, assuming the
+KV and KMV pairs remain evenly distributed across processors
+throughout the sequence of MapReduce operations.  The use of hashing
+to assign keys to processors typically provides for such
+load-balancing.  In the next section, we discuss what happens when
+data sets do not fit in available memory.

### 2010-06-07 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 specified, which defaults to 64 Mbytes.  As described below, each
-MR-MPI operation is constrained to use no more than a handful of
-these pages.  The {\it pagesize} setting can be as small as 1 Mbyte or
-as large as desired, though the user should ensure the allocated pages
-fit in physical memory; otherwise, the MR-MPI library may allocate slow
+MR-MPI operation is constrained to use no more than a handful of these
+pages.  The {\it pagesize} setting can be as small as 1 Mbyte or as
+large as desired, though the user should ensure the allocated pages
+fit in physical memory; otherwise a processor may allocate slow
 virtual memory.  The {\it pagesize} is also the typical size of
@@
 new processors in an all-to-all fashion.  In out-of-core mode, this
-operation is
-performed on one page of KV data at a time.  Each processor reads in a page of
-KV pairs, the communication pattern (which processors receive 
-which datums in that page) is
-determined, and all-to-all communication is
-performed.  Each processor allocates a two-page chunk of memory to
-receive incoming KV pairs.  On average each processor should receive
-one page of KV pairs; the two-page allocation allows for some
-load imbalance.  If the KV pairs are distributed unevenly so that this
-limit is exceeded, the all-to-all communication is performed in
-smaller, multiple passes until the full page contributed by every
-processor has been communicated.  Performing an MPI\_Alltoall(), or
-using the custom all-to-all routines provided by the MR-MPI library,
-requires allocation of auxiliary arrays that store processor indices,
-datum pointers, and datum lengths.  For the case of many tiny KV
-datums, the total number of memory pages required to perform the
-all-to-all communication, including the source and target KV pages is
-at most seven.
+operation is performed on one page of KV data at a time.  Each
+processor reads in a page of KV pairs, the communication pattern
+(which processors receive which datums in that page) is determined,
+and all-to-all communication is performed.  Each processor allocates a
+two-page chunk of memory to receive incoming KV pairs.  On average
+each processor should receive one page of KV pairs; the two-page
+allocation allows for some load imbalance.  If the KV pairs are
+distributed unevenly so that this limit is exceeded, the all-to-all
+communication is performed in smaller, multiple passes until the full
+page contributed by every processor has been communicated.  Performing
+an MPI\_Alltoall(), or using the custom all-to-all routines provided
+by the MR-MPI library, requires allocation of auxiliary arrays that
+store processor indices, datum pointers, and datum lengths.  For the
+case of many tiny KV datums, the total number of memory pages required
+to perform the all-to-all communication, including the source and
+target KV pages is at most seven.
 
@@
 out-of-core KV file, that need to be reorganized into a set of KMV
-pages, stored in a new out-of-core KMV file.  In principle, 
+pages, likewise stored in a new out-of-core KMV file.  In principle,
 creating one page of KMV pairs would require all KV pages be scanned
@@
 arbitrarily large for data sets with many unique keys.  We need an
-algorithm for generating KMV pairs that operates within a small
-number of memory pages and performs a minimal number of passes through
-the KV disk file.  An algorithm that meets these goals is diagrammed
-in Figure \ref{fig:collate}; it reads the KV pages at most four times, and
-writes out new KV or KMV pages at most three times.  It also uses a
-finite-size in-memory hash table, that stores unique keys as they are
-encountered while looping over the KV pages, as well as auxiliary
+algorithm for generating KMV pairs that operates within a small number
+of memory pages and performs a minimal number of passes through the KV
+disk file.  An algorithm that meets these goals is diagrammed in
+Figure \ref{fig:collate}; it reads the KV pages at most four times,
+and writes out new KV or KMV pages at most three times.  It also uses
+a finite-size in-memory hash table, that stores unique keys as they
+are encountered while looping over the KV pages, as well as auxiliary
 information needed to construct the output pages of KMV pairs.
@@
 \includegraphics[width=\textwidth,angle=-90]{fig_collate2.pdf}
-\caption{Multi-pass algorithm for converting KV data to KMV data.  The
-vertical lines represent out-of-core data sets.  KV is key/value
+\caption{\it Multi-pass algorithm for converting KV data to KMV data.
+The vertical lines represent out-of-core data sets.  KV is key/value
 pairs; HT is an in-memory hash table; KMV is key/multivalue pairs.}
@@
 is somewhat complex, but requires only a small, constant number of
-passes through the data.  The KV datums are read from disk at most four
-times, and written to disk three times.  The first two write passes
-reorganize KV pairs into partition and set files; the final write pass
-creates the KMV data set.  Depending on the size and characteristics
-of the KV datums (e.g., the number of unique keys), some of
-these passes may not be needed.  By contrast, a full out-of-core sort
-of the KV pairs, performed via a merge sort as discussed below,
-requires $O(\log{M})$ read/write passes through the KV data, where $M$
-is the number of pages of KV pairs, which can be a large number for
-big data sets.
+passes through the data.  The KV datums are read from disk at most
+four times, and written to disk three times.  The first two write
+passes reorganize KV pairs into partition and set files; the final
+write pass creates the KMV data set.  Depending on the size and
+characteristics of the KV datums (e.g., the number of unique keys),
+some of these passes may not be needed.  By contrast, a full
+out-of-core sort of the KV pairs, performed via a merge sort as
+discussed below is still an $O(N\log_2{N})$ operation.  The number of
+read/write passes through the KV data files depends on the amount of
+KV data that can be held in memory, but can be a large number for big
+data sets.
 
-The memory page requirements for the above algorithm are as follows.
-Two contiguous pages of memory are used for the hash table.
-Intermediate passes 2 and 3 can require numerous partition or set
-files to be opened simultaneously and written to.  To avoid small
-writes of individual KV datums, a buffer of at least 16K bytes is
-allocated for each file.  The precise number of simultaneously open
-files is difficult to bound, but typically one or two additional pages of 
-memory suffice for all needed buffers.  Thus the total number of pages required
-for the data reorganization stage of the {\it collate()} operation is
-less than the seven used by the all-to-all communication stage.
+The memory page requirements for the out-of-core {\it collate()}
+operation are as follows.  Two contiguous pages of memory are used for
+the hash table.  Intermediate passes 2 and 3 can require numerous
+partition or set files to be opened simultaneously and written to.  To
+avoid small writes of individual KV datums, a buffer of at least 16K
+bytes is allocated for each file.  The precise number of
+simultaneously open files is difficult to bound, but typically one or
+two additional pages of memory suffice for all needed buffers.  Thus
+the total number of pages required for the data reorganization stage
+of the {\it collate()} operation is less than the seven used by the
+all-to-all communication stage.
 
 Other MR-MPI library calls discussed in the previous section, such as
-{\it clone()}, {\it collapse()}, and {\it gather()},
-can also operate in out-of-core mode, typically with one pass
-through the KV or KMV data and the use of one or two memory pages.
-Sorting KV pairs, by key or value, is an exception.  An out-of-core
-merge sort is performed as follows.  Two pages of KV datums are read
-from disk.  Each is sorted using the C-library quicksort() function,
-as discussed in the previous section.  The two pages are then scanned
-and merged into a new file which is written to disk as its associated
-memory page fills up.  This process requires five memory pages, which
-includes vectors of KV datum pointers and lengths needed by the
-in-memory quicksort() operation.  Once all pairs of pages have been
-merged, the sort continues in a recursive fashion, merging pairs of
-files into a new third file, without the need to perform additional
-in-memory sorts.  Thus the overall memory requirement is five pages.  The
-number of read/write passes through the KV data set for the merge sort
-is $O(\log{M})$, where $M$ is the number of pages of KV pairs.
+{\it clone()}, {\it collapse()}, and {\it gather()}, can also operate
+in out-of-core mode, typically with one pass through the KV or KMV
+data and the use of one or two memory pages.  Sorting KV pairs, by key
+or value, is an exception.  An out-of-core merge sort is performed as
+follows.  Two pages of KV datums are read from disk.  Each is sorted
+using the C-library quicksort() function, as discussed in the previous
+section.  The two pages are then scanned and merged into a new file
+which is written to disk as its associated memory page fills up.  This
+process requires five memory pages, which includes vectors of KV datum
+pointers and lengths needed by the in-memory quicksort() operation.
+Once all pairs of pages have been merged, the sort continues in a
+recursive fashion, merging pairs of files into a new third file,
+without the need to perform additional in-memory sorts.  Thus the
+overall memory requirement is five pages.  The number of read/write
+passes through the KV data set for the merge sort is $O(\log_2{M})$,
+where $M$ is the number of pages of KV pairs.  The number of passes
+could be reduced at the cost of allocating more in-memory pages.

### 2010-06-07 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@630 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 graph algorithms of the preceeding section, implemented as small C++
-programs calling our MR-MPI library.
-The benchmarks were run on a medium-sized Linux cluster made up of 2 GHz
-dual-core AMD Opteron processors with a Myrinet network.  Most importantly
-for MR-MPI, each node of the cluster has one local disk, which is used
-for out-of-core operations in MR-MPI.  To avoid contention for disk I/O, we 
-ran all experiments with one MPI process per node.  For comparisons with
-other implementations, we used either the same cluster or, where noted,
-Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz} processors
-and a 3D-Torus network.
+programs calling our MR-MPI library.  The benchmarks were run on a
+medium-sized Linux cluster of 2 GHz dual-core AMD Opteron processors
+connected via a Myrinet network.  Most importantly for MR-MPI, each
+node of the cluster has one local disk, which is used for out-of-core
+operations in MR-MPI.  To avoid contention for disk I/O, we ran all
+experiments with one MPI process per node.  For comparisons with other
+implementations, we used either the same cluster or, where noted,
+Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz}
+processors and a 3D-Torus network.
 
-We ran each of the algorithms on three R-MAT graphs of
-different sizes, each on a varying number of processors.  Details of the
-input data are shown in Table~\ref{t:rmats}.
-The {\it small} problem (around 8M edges) can typically be used
-on a single processor without significant out-of-core operations.
-The {\it medium} problem (around 134M edges) is a 
-problem that can be run
-on a single processor with out-of-core operations; larger processor 
-configurations, however, do not necessarily need out-of-core operations.
-The {\it large} problem (around 2B
-edges) requires out-of-core operations and higher processor counts.
+We ran each of the algorithms on three R-MAT graphs of different
+sizes, each on a varying number of processors.  Details of the input
+data are shown in Table~\ref{t:rmats}.  The {\it small} problem
+(around 8M edges) can typically be run on a single processor without
+incurring out-of-core operations.  The {\it medium} problem (around
+134M edges) can be run on a single processor with out-of-core
+operations; larger processor configurations, however, do not
+necessarily require out-of-core operations.  The {\it large} problem
+(around 2B edges) requires out-of-core operations on our 64-node
+cluster.
+
 %The {\it x-large} problem 
 %(around 34B edges) requires most of the machine to run.
-All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.
+
+All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
+0.05)$ and generated 8 edges per vertex (on average).  These values
+create a highly skewed degree distribution, as indicated by the
+maximumn vertex degree in Table~\ref{t:rmats}.
 
@@
 \caption{Characteristics of R-MAT input data for graph algorithm
-experiments.}
+benchmarks.}
 \label{t:rmats}
@@
 including more traditional distributed-memory algorithms and
-multi-threaded algorithms
-in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL} on the Cray XMT.
-We compute parallel efficiency on $p$ processors
-as $({time}_{m} \times m) / ({time}_p \times p)$
-where $m$ is the smallest number of processors on which the experiment
-was run, and ${time}_m$ is the execution time required on $m$ processors.
+multi-threaded algorithms in the Multi-Threaded Graph Library
+(MTGL)~\cite{MTGL} on the Cray XMT.  We compute parallel efficiency on
+$P$ processors as $({time}_{M} \times M) / ({time}_P \times P)$ where
+$M$ is the smallest number of processors on which the experiment was
+run, and ${time}_I$ is the execution time required on $I$ processors.
 
@@
 
-In Figure~\ref{f:rmat}, we show the scalability of the R-MAT generation 
-algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and RMAT-28.  For
-RMAT-20 and RMAT-24, superlinear speed-up is shown.  This speed-up is due
-to the decreased amount of file I/O needed with greater numbers of 
-processors; with a larger total memory, a fixed-size problem requires
-less I/O with more processors.  For RMAT-28, which requires
-significant out-of-core operations, parallel efficiency ranges from 52\% to
-97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
-the original algorithm (Figure~\ref{fig:rmat}) showed approximately 
-10\% reduction in execution time for the enhanced algorithm.
+In Figure~\ref{f:rmat}, we show the scalability of the R-MAT
+generation algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and
+RMAT-28.  For RMAT-20 and RMAT-24, superlinear speed-up is shown.
+This speed-up is due to the decreased amount of file I/O needed with
+greater numbers of processors; with a larger total memory, a
+fixed-size problem requires less I/O with more processors.  For
+RMAT-28, which requires significant out-of-core operations, parallel
+efficiency ranges from 52\% to 97\%.  Comparisons between the enhanced
+algorithm (Figure~\ref{fig:rmat2}) and the original algorithm
+(Figure~\ref{fig:rmat}) showed approximately 10\% reduction in
+execution time for the enhanced algorithm.
 
@@
 
-In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
-(Figure~\ref{fig:pr2}) compared to a
-distributed-memory matrix-based implementation using the linear
-algebra toolkit Trilinos~\cite{Trilinos-Overview}.
-The matrix-based distributed-memory implementation of PageRank
-uses Trilinos' Epetra matrix/vector classes to represent the graph and
-PageRank vector.
-Rows of matrix $A$ and the associated entries of the PageRank vector $x$
-are uniquely assigned to processors; a random permutation of the input
-matrix achieves processor load balance.
-Interprocessor communication gathers $x$ values for matrix-vector
-multiplication and sums partial products into the $y$ vector.
-Most communication is point-to-point communication,
-but some global communication is needed for computing
-residuals and norms of $x$ and $y$.
+In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank
+algorithm (Figure~\ref{fig:pr2}) compared to a distributed-memory
+matrix-based implementation using the linear algebra toolkit
+Trilinos~\cite{Trilinos-Overview}.  The matrix-based
+distributed-memory implementation of PageRank uses Trilinos Epetra
+matrix/vector classes to represent the graph and PageRank vector.
+Rows of matrix $A$ and the associated entries of the PageRank vector
+$x$ are uniquely assigned to processors; a random permutation of the
+input matrix effectively load balances the non-zero matrix entries
+across processors.  Interprocessor communication gathers $x$ values
+for matrix-vector multiplication and sums partial products into the
+$y$ vector.  Most communication is point-to-point communication, but
+some global communication is needed for computing residuals and norms
+of $x$ and $y$.
 
 Figure~\ref{f:pr} shows the execution time per PageRank iteration for
-R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  
-Converging the PageRank iterations
-to tolerance $0.002$ requires five or six iterations.
-Several R-MAT matrices of 
-each size were generated for the experiments; the average time over the
-matrices is reported here.  
-The Trilinos implementations show near-perfect
-strong scaling for RMAT-20 and RMAT-24.  The MR-MPI 
-implementations also demonstrate good strong scaling.  However, MR-MPI's
-execution time is at least an order of magnitude greater than the Trilinos
-implementation.  This result is due to two factors:  ({\it i}) a 
-higher volume of
-communication in the MapReduce implementation (where all edges are communicated
-in each iteration), compared to the Trilinos implementation (where only
-vertex-based data are communicated), and ({\it ii}) out-of-core operations
-in MR-MPI were performed because of the page-size restriction, while all
-Trilinos operations were performed in-core.  The benefit of MR-MPI's
-out-of-core implementation is seen, however, with the RMAT-28 data set, 
-which could be solved on smaller processor sets than the Trilinos 
-implementation.  For these experiments, the Trilinos implementation required
-64 processors for the RMAT-28 data set.
+R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  Converging the PageRank
+iterations to tolerance $0.002$ requires five or six iterations.
+Several R-MAT matrices of each size were generated for the
+experiments; the average time over the matrices is reported here.  The
+Trilinos implementations show near-perfect strong scaling for RMAT-20
+and RMAT-24.  The MR-MPI implementations also demonstrate good strong
+scaling.  However, MR-MPI's execution time is at least an order of
+magnitude greater than the Trilinos implementation.  This result is
+due to two factors: ({\it i}) a higher volume of communication in the
+MapReduce implementation (where all edges are communicated in each
+iteration), compared to the Trilinos implementation (where only
+vertex-based data is communicated), and ({\it ii}) out-of-core
+operations in MR-MPI were performed because of the page-size
+restriction, while all Trilinos operations were performed in-core.
+The benefit of MR-MPI's out-of-core implementation is seen, however,
+with the RMAT-28 data set, which could be solved on smaller processor
+sets than the Trilinos implementation.  For these experiments, the
+Trilinos implementation required 64 processors for the RMAT-28 data
+set, since it will not operate on matrices that exceed the aggregate
+memory of the processors.
 
@@
 In Figure~\ref{f:tri}, we show the performance of the triangle finding
-algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm were
-too large to allow the problem to be run with RMAT-28 on our cluster.
-Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
-ranged from 50\% to 120\%.  
+algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm
+were too large to allow the problem to be easily run with RMAT-28 on
+our cluster.  Parallel efficiencies for RMAT-20 ranged from 80\% to
+140\%; for RMAT-24, they ranged from 50\% to 120\%.
+
 %Comparisons with an MTGL implementation on the 
@@
 
-We compare our MR-MPI algorithm (Figure~\ref{fig:cc}) 
-with a hybrid ``Giant Connected Component''
-implementation using Trilinos and MR-MPI.  Power law graphs often have one
-or two very large components and many very small components.  
-The hybrid algorithm exploits this feature of the data by using inexpensive
-breadth-first
-search from the vertex with highest degree to identify the largest components,
-followed by a more expensive algorithm to identify the small components in
-the remainder of the graph.  In our hybrid implementation, we use
-Trilinos' matrix-vector multiplication to do the breadth-first search, 
-finding components that include (in total) 70\% or more of the vertices.
-We then apply our MR-MPI algorithm to the remaining graph to identify
-the small components.  This hybrid approach is extremely effective,
-reducing the execution time to identify all components of RMAT-20 and 
-RMAT-24 by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI
-is seen, however, for RMAT-28, where the problem is too large to fit into
-memory for our hybrid algorithm.
+In Figure~\ref{fig:cc}, we show the performance of the connected
+component identification algorithm (Figure~\ref{fig:cc}).  A
+comparison is made with a hybrid ``Giant Connected Component''
+implementation using both Trilinos and MR-MPI.  Power law graphs often
+have one or two very large components and many very small components.
+The hybrid algorithm exploits this feature of the data by using
+inexpensive breadth-first search (BFS) from the vertex with highest
+degree to identify the largest components, followed by a more
+expensive algorithm to identify the small components in the remainder
+of the graph.  In our hybrid implementation, we intially perform
+matrix-vector multiplications in Trilinos to perform a BFS, finding
+components that include (in total) 70\% or more of the vertices.  We
+then apply our MR-MPI algorithm to the remaining graph to identify the
+small components.  This hybrid approach is quite effective, reducing
+the execution time to identify all components of RMAT-20 and RMAT-24
+by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI is
+seen, however, for RMAT-28, where the graph is too large to fit into
+memory, and thus Trilinos cannot perform the initial BFS for our
+hybrid algorithm.
 
@@
 The execution times for the maximal independent set algorithm
-(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
-generation results, superlinear speed-up of the algorithm occurs for RMAT-20
-and RMAT-24, as more of the graph fits into processor memory and less file
-I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
-operations. In this case, parallel efficiency is nearly perfect going from 
-8 to 64 processors.
+(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
+R-MAT generation results, superlinear speed-up of the algorithm occurs
+for RMAT-20 and RMAT-24, as more of the graph fits into processor
+memory and less file I/O is needed.  For RMAT-28, the algorithm
+requires significant out-of-core operations. In this case, parallel
+efficiency is nearly perfect going from 8 to 64 processors.
 
@@
 The execution times for the single-source shortest path algorithms
-(Figures~\ref{fig:sssp} and \ref{fig:sssp2})
-are shown in Figure~\ref{f:sssp}.  The results show the benefit of
-using the enhanced algorithm, providing at least a 17\% (and often greater)
-reduction in execution time due to reduced communication; only the updated
-distances are communicated throughout most of the enhanced algorithm.
-However, the execution times are still large compared to multi-threaded
-implementations; for example, Madduri et al.~\cite{Madduri07} report
-execution times of only 11 seconds on 40 
-processors for a multithreaded implementation on a Cray MTA-2
-(the predecessor of the XMT) on graphs with one billion edges.
+(Figures~\ref{fig:sssp} and \ref{fig:sssp2}) are shown in
+Figure~\ref{f:sssp}.  The results show the benefit of using the
+enhanced algorithm, providing at least a 17\% (and often greater)
+reduction in execution time due to reduced communication; only the
+updated distances are communicated throughout most of the enhanced
+algorithm.  However, the execution times are still large compared to
+multi-threaded implementations; for example, Madduri et
+al.~\cite{Madduri07} report execution times of only 11 seconds on 40
+processors for a multithreaded implementation on a Cray MTA-2 (the
+predecessor of the XMT) on graphs with one billion edges.
 
@@
 
-To compare our MR-MPI implementation with a wider set of implementations,
-we performed experiments MR-MPI, Hadoop, PBGL (Parallel
-Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
-Cray XMT using two web graphs: {WebGraphA} with 13.2M
-vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
-In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
-using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
-implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
-The XMT and PBGL implementations are based on 
-delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
-over the entire edge list to advance a breadth-first search.  
-We observe that the MR-MPI
-implementation runs in less time than the Hadoop implementation, but requires
-significantly more time than the XMT and PBGL implementations.  
-In experiments with {WebGraphB},
-the benefit of the enhanced algorithm (Figure~\ref{fig:sssp2}) is clearly
-shown, with a 40\% reduction in execution time compared to the original
-SSSP algorithm (Figure~\ref{fig:sssp}).
-But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
-Hadoop implementations for {WebGraphB}, which required 110 iterations to 
-complete; execution times for this data set are shown in Table~\ref{t:ssspB}.
+To compare our MR-MPI implementation with a wider set of algorithms,
+we performed experiments comparing MR-MPI, Hadoop, PBGL (Parallel
+Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation
+on the Cray XMT using two web graphs: {WebGraphA} with 13.2M vertices
+and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M
+edges.  In Figure~\ref{f:ssspA}, we show execution times for the SSSP
+algorithm using {WebGraphA}.  Like our MR-MPI implementation, the
+Hadoop implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}
+algorithm.  The XMT and PBGL implementations are based on
+delta-stepping~\cite{MeyerSanders98}, and do not require full
+iterations over the entire edge list to advance a breadth-first
+search.  We observe that the MR-MPI implementation runs in less time
+than the Hadoop implementation, but requires significantly more time
+than the XMT and PBGL implementations.  In experiments with
+{WebGraphB}, the benefit of the enhanced algorithm
+(Figure~\ref{fig:sssp2}) is clearly shown, with a 40\% reduction in
+execution time compared to the original SSSP algorithm
+(Figure~\ref{fig:sssp}).  But the Bellman-Ford-style iterations are
+especially harmful to the MR-MPI and Hadoop implementations for
+{WebGraphB}, which required 110 iterations to complete; execution
+times for this data set are shown in Table~\ref{t:ssspB}.
 
@@
 \subsection{Scalability to large numbers of processors}
-Finally, we demonstrate the scalability of MR-MPI to large numbers of 
-processors.  
-The MR-MPI implementation was run on Sandia's Redstorm and Thunderbird
-parallel computers.  Redstorm is a massively parallel computer with 2+ 
-{GHz} dual/quad-core AMD Opteron processors and 9.6 GB/s link bandwidth.
-Thunderbird is a cluster with 3.6 {GHz} dual-core Intel EM64T processors
-and an Infiniband network.
-Because these systems do not have local disks for each processor, we selected 
-a data set and page sizes that fit in memory, so out-of-core operations
-were not used.  For these experiments, we used an R-MAT data set with 
-with $2^{25}$ vertices and $2^{28}$ edges, with parameters given in
-Table~\ref{t:rmat}.  We ran both PageRank and Connected Components
+
+Finally, we demonstrate the scalability of our MR-MPI library to large
+numbers of processors.  The library was used on Sandia's Redstorm and
+Thunderbird parallel computers.  Redstorm is a large Cray XT3 with 2+
+{GHz} dual/quad-core AMD Opteron processors and a custom interconnect
+providing 9.6 GB/s of interprocessor bandwidth.  Thunderbird is a
+large Linux cluster with 3.6 {GHz} dual-core Intel EM64T processors
+connected by an Infiniband network.  Because these systems do not have
+local disks for each processor, we selected a data set and page sizes
+that fit in memory, so out-of-core operations were not needed.  For
+these experiments, we used an R-MAT data set with with $2^{25}$
+vertices and $2^{28}$ edges, with parameters given in
+Table~\ref{t:rmat}.  We ran both the PageRank and Connected Components
 algorithms.
@@
 
-In Figure~\ref{f:prbig}, we show the performance 
-of the various PageRank
-implementations on distributed memory and multi-threaded architectures.
-The MR-MPI and Trilinos implementations are described in 
-Sections~\ref{subsec:graph_pagerank} and~\ref{subsec:results_pagerank},
-respectively.
-In the multi-threaded MTGL implementation,
-rank propagates via adjacency list traversal
-in a compressed sparse-row data structure.
-To maintain its scalability, code must
-be written so that a single thread spawns the loop that processes all
-in-neighbors of a given vertex; this detail enables the compiler to generate
-hotspot-free code.
+Figure~\ref{f:prbig}, shows the performance of the various PageRank
+implementations on distributed memory and multi-threaded
+architectures.  The MR-MPI and Trilinos implementations are described
+in Sections~\ref{subsec:graph_pagerank}
+and~\ref{subsec:results_pagerank}, respectively.  In the
+multi-threaded MTGL implementation, rank propagates via adjacency list
+traversal in a compressed sparse-row data structure.  To maintain its
+scalability, code must be written so that a single thread spawns the
+loop that processes all in-neighbors of a given vertex; this detail
+enables the compiler to generate hotspot-free code.
 
-The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
-however, it required an order-of-magnitude
-more execution time than the matrix-based implementations on Redstorm.  
-The distributed memory matrix-based
-implementations are competitive with the multi-threaded implementation
-in MTGL on the Cray XMT.
+The MR-MPI implementation demonstrated good scalability up to 1024
+processors; however, as before, it required an order-of-magnitude more
+execution time than the matrix-based implementations on Redstorm.  The
+distributed memory matrix-based implementations are competitive with
+the multi-threaded implementation in MTGL on the Cray XMT.
 
@@
 \caption{Scalability comparison of PageRank using MapReduce (MR-MPI),
-matrix-based (Trilinos), and multi-threaded (MTGL) implementation
-on the R-MAT data sets in Table~\ref{t:rmat}.}
+matrix-based (Trilinos), and multi-threaded (MTGL) implementation on
+the R-MAT data sets in Table~\ref{t:rmat}.}
 \label{f:prbig}
@@
 
-Similar results were obtained for the Connected Components algorithm, as
-we show in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI implementation
-showed good scalability up to 1024 processors, but required significantly
-more time than the Giant Connected Components algorithms using MTGL and
-Trilinos.
+Similar results were obtained for the Connected Components algorithm,
+as shown in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI
+implementation showed good scalability up to 1024 processors, but
+required significantly more time than the hybrid algorithm using
+Trilinos and MR-MPI or the MTGL algorithm.
 

### 2010-06-07 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@632 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 due to two factors: ({\it i}) a higher volume of communication in the
-MapReduce implementation (where all edges are communicated in each
-iteration), compared to the Trilinos implementation (where only
-vertex-based data is communicated), and ({\it ii}) out-of-core
-operations in MR-MPI were performed because of the page-size
-restriction, while all Trilinos operations were performed in-core.
-The benefit of MR-MPI's out-of-core implementation is seen, however,
-with the RMAT-28 data set, which could be solved on smaller processor
-sets than the Trilinos implementation.  For these experiments, the
-Trilinos implementation required 64 processors for the RMAT-28 data
-set, since it will not operate on matrices that exceed the aggregate
-memory of the processors.
+MapReduce implementation (which comunicates one datum per graph edge
+in each iteration), compared to the Trilinos implementation which only
+communicates a volume of data proportional to the number of vertices,
+and ({\it ii}) out-of-core operations in MR-MPI were performed because
+of the page-size restriction, while all Trilinos operations were
+performed in-core.  The benefit of MR-MPI's out-of-core implementation
+is seen, however, with the RMAT-28 data set, which could be solved on
+smaller processor sets than the Trilinos implementation.  For these
+experiments, the Trilinos implementation required 64 processors for
+the RMAT-28 data set, since it will not operate on matrices that
+exceed the aggregate memory of the processors.
 

### 2010-10-25 programs/cc_driver.cpp
Mostly minor changes made during paper writing.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@644 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -19,6 +19,8 @@
 #include "cc_stats.h"
 #include "mapreduce.h"
 
+#include "localdisks.hpp"
+
 using MAPREDUCE_NS::MapReduce;
 
 struct Params {
@@ -41,6 +43,12 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD,&me);
   MPI_Comm_size(MPI_COMM_WORLD,&nprocs);
 
+  greetings();
+  if (nprocs < 10)  {
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+  }
   // parse command-line args
   
   Params in;
@@ -49,8 +57,8 @@ int main(int narg, char **args)
   // MRE for matrix edges
 
   MapReduce *mre = new MapReduce(MPI_COMM_WORLD);
-  mre->verbosity = 1;
-  mre->timer = 1;
+  mre->verbosity = 0;
+  mre->timer = 0;
 
   // generate RMAT matrix
 
@@ -61,11 +69,11 @@ int main(int narg, char **args)
   // MRV and MRZ for CC algorithm
 
   MapReduce *mrv = new MapReduce(MPI_COMM_WORLD);
-  mrv->verbosity = 1;
-  mrv->timer = 1;
+  mrv->verbosity = 0;
+  mrv->timer = 0;
   MapReduce *mrz = new MapReduce(MPI_COMM_WORLD);
-  mrz->verbosity = 1;
-  mrz->timer = 1;
+  mrz->verbosity = 0;
+  mrz->timer = 0;
 
   // find CCs
 

### 2010-10-25 programs/luby_driver.cpp

@@ -16,6 +16,8 @@
 #include "luby_find.h"
 #include "mapreduce.h"
 
+#include "localdisks.hpp"
+
 using MAPREDUCE_NS::MapReduce;
 
 struct Params {
@@ -37,6 +39,13 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD,&me);
   MPI_Comm_size(MPI_COMM_WORLD,&nprocs);
 
+  greetings();
+  if (nprocs < 10)  {
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+  }
+
   // parse command-line args
   
   Params in;
@@ -45,8 +54,8 @@ int main(int narg, char **args)
   // MRE for matrix edges
 
   MapReduce *mre = new MapReduce(MPI_COMM_WORLD);
-  mre->verbosity = 1;
-  mre->timer = 1;
+  mre->verbosity = 0;
+  mre->timer = 0;
 
   // generate RMAT matrix, make it upper triangular with no diagonal elements
 
@@ -62,8 +71,8 @@ int main(int narg, char **args)
   // perform Luby's algorithm
 
   MapReduce *mrv = new MapReduce(MPI_COMM_WORLD);
-  mre->verbosity = 1;
-  mre->timer = 1;
+  mre->verbosity = 0;
+  mre->timer = 0;
 
   niterate = 0;
   uint64_t nset;

### 2010-10-25 programs/mrmatrix2.hpp

@@ -169,7 +169,11 @@ MRMatrix<IDTYPE>::MRMatrix(
   transposeFlag = transpose;
   scaleFactor = alpha;
   // Create matrix MapReduce object mr.  
-  mr = mredge->copy();
+  // KDDKDD Probably should do as a copy to prevent side effects on mredge, 
+  // KDDKDD but to save execution time, we'll copy its pointer and modify
+  // KDDKDD mredge.
+  // mr = mredge->copy();
+  mr = mredge;
   mr->memsize = pagesize;
   mr->set_fpath(fpath);
 

### 2010-10-25 programs/rmat.hpp

@@ -192,10 +192,12 @@ void GenerateRMAT::run(
   // Loop until desired number of unique nonzero entries
 
   if (me == 0) cout << "Generating edges..." << endl;
+  MPI_Barrier(MPI_COMM_WORLD);
+  double tstarte = MPI_Wtime();
 
   MapReduce *mrnew = new MapReduce(MPI_COMM_WORLD);
-  mrnew->timer = 1;
-  mrnew->verbosity = 1;
+//  mrnew->timer = 1;
+//  mrnew->verbosity = 1;
   MapReduce *mredge = NULL;
 
   int niterate = 0;
@@ -214,13 +216,12 @@ void GenerateRMAT::run(
     if (me == 0) {cout << "  add..." << endl; flush(cout);}
     if (niterate > 1) {
       mredge->add(mrnew);
-      mrnew->map(1, rmat_do_nothing, NULL);  // Empty mrnew.
     }
     else {  // First iteration -- no need to add; just set the pointer.
       mredge = mrnew;
       mrnew = new MapReduce(MPI_COMM_WORLD);
-      mrnew->timer = 1;
-      mrnew->verbosity = 1;
+//      mrnew->timer = 1;
+//      mrnew->verbosity = 1;
     }
     if (me == 0) {cout << "  convert..." << endl; flush(cout);}
     uint64_t nunique = mredge->convert();
@@ -234,10 +235,12 @@ void GenerateRMAT::run(
                   flush(cout);}
   }
   delete mrnew;
-  mredge->timer = 0;
-  mredge->verbosity = 0;
+//  mredge->timer = 0;
+//  mredge->verbosity = 0;
 
   // output matrix if requested
+  MPI_Barrier(MPI_COMM_WORLD);
+  double tstope = MPI_Wtime();
 
   if (outfile) {
     if (me == 0) {cout << "Writing output file..." << endl; flush(cout);}
@@ -260,6 +263,9 @@ void GenerateRMAT::run(
   if (me == 0) {
     std::cout << order << " rows in matrix" << std::endl; flush(std::cout);
     std::cout << ntotal << " nonzeros in matrix" << std::endl; flush(std::cout);
+    std::cout << tstope-tstarte << " secs to generate edges on " << nprocs
+              << " procs in " << niterate << " iterations" << std::endl;
+    flush(std::cout);
   }
  
   if (printstats > 0) {
@@ -337,7 +343,7 @@ void GenerateRMAT::run(
   double tstop = MPI_Wtime();
 
   if (me == 0) {
[…]

### 2010-10-25 programs/sssp.cpp

@@ -24,7 +24,7 @@
 using namespace std;
 using namespace MAPREDUCE_NS;
 
-#define MAX_NUM_EXPERIMENTS 60
+#define MAX_NUM_EXPERIMENTS 1
 
 #define FBFILE 0
 #define MMFILE 1

### 2010-10-25 programs/sssp_v2.cpp

@@ -26,7 +26,7 @@
 using namespace std;
 using namespace MAPREDUCE_NS;
 
-#define MAX_NUM_EXPERIMENTS 10
+#define MAX_NUM_EXPERIMENTS 1
 
 #define FBFILE 0
 #define MMFILE 1

### 2010-10-25 programs/test_pagerank.cpp

@@ -218,7 +218,6 @@ KDDallzero += (MPI_Wtime() - KDDtmp);
 MPI_Barrier(MPI_COMM_WORLD);
 KDDtmp = MPI_Wtime();
 #endif //KDDTIME
-
     // Compute global adjustment.
     A->MatVec(mr, x, y, 1, storage_aware);
 
@@ -460,8 +459,8 @@ int main(int narg, char **args)
   }
 
   MapReduce *mr = new MapReduce(MPI_COMM_WORLD);
-  mr->verbosity = 0;
-  mr->timer = 0;
+  mr->verbosity = 2;
+  mr->timer = 2;
   mr->mapstyle = 1;  // mapstyle == 0 does not work for this code.
   mr->memsize = pagesize;
 

### 2010-10-25 programs/test_pagerank2.cpp

@@ -170,13 +170,19 @@ MRVector<IDXTYPE> *pagerank(
 )
 {
   int me = A->mr->my_proc();
-  if (me == 0) {cout << "Initializing vectors..." << endl; flush(cout);}
+if (me == 0) {cout << "Initializing vectors..." << endl; flush(cout);}
   MRVector<IDXTYPE> *x = new MRVector<IDXTYPE>(A->NumRows(),
                                              A->mr->memsize, A->mr->fpath);
   MRVector<IDXTYPE> *y = new MRVector<IDXTYPE>(x->GlobalLen(), 
                                              A->mr->memsize, A->mr->fpath);
   MRVector<IDXTYPE> *zerovec = new MRVector<IDXTYPE>(x->GlobalLen(), 
                                              A->mr->memsize, A->mr->fpath);
+  A->mr->timer = 0;
+  y->mr->timer = 0;
+  x->mr->timer = 0;
+  A->mr->verbosity = 0;
+  y->mr->verbosity = 0;
+  x->mr->verbosity = 0;
 
   double randomlink = (1.-alpha)/(double)(x->GlobalLen());
   int iter; 
@@ -184,6 +190,7 @@ MRVector<IDXTYPE> *pagerank(
 
   // If pre-scaled A by alpha in the constructor, don't need to do it here.
   // Would like to avoid the additional pass over A here and at end.
+if (me == 0) cout << "KDDKDD initializing..." << endl;
   if (alpha != A->scaleFactor) {
     if (me == 0)
       cout << "Scaling A by " << alpha << " " << A->scaleFactor << endl;
@@ -194,9 +201,20 @@ MRVector<IDXTYPE> *pagerank(
   MPI_Barrier(MPI_COMM_WORLD);
   double tstart = MPI_Wtime();
 
+#define KDDTIME
+#ifdef KDDTIME
+double step_tmp;
+double KDD_step[6] = {0., 0., 0., 0., 0., 0.};
+#endif
+
   if (me == 0) {cout << "Beginning iterations..." << endl; flush(cout);}
   // PageRank iteration
   for (iter = 0; iter < maxniter; iter++) {
+
+#ifdef KDDTIME
+if (me == 0) cout << "KDDKDD compute adjustment..." << endl;
+step_tmp = MPI_Wtime();
+#endif
 
[…]

### 2010-10-25 programs/tri_driver.cpp

@@ -16,6 +16,8 @@
 #include "tri_find.h"
 #include "mapreduce.h"
 
+#include "localdisks.hpp"
+
 using MAPREDUCE_NS::MapReduce;
 
 struct Params {
@@ -37,6 +39,12 @@ int main(int narg, char **args)
   MPI_Comm_rank(MPI_COMM_WORLD,&me);
   MPI_Comm_size(MPI_COMM_WORLD,&nprocs);
 
+  greetings();
+  if (nprocs < 10)  {
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+    greetings();  // Odin skips some early program output; we'll output extra
+  }
   // parse command-line args
   
   Params in;
@@ -45,8 +53,8 @@ int main(int narg, char **args)
   // MRE for matrix edges
 
   MapReduce *mre = new MapReduce(MPI_COMM_WORLD);
-  mre->verbosity = 1;
-  mre->timer = 1;
+  mre->verbosity = 0;
+  mre->timer = 0;
 
   // generate RMAT matrix, make it upper triangular with no diagonal elements
 
@@ -62,8 +70,8 @@ int main(int narg, char **args)
   // enumerate triangles
 
   MapReduce *mrt = new MapReduce(MPI_COMM_WORLD);
-  mrt->verbosity = 1;
-  mrt->timer = 1;
+  mrt->verbosity = 0;
+  mrt->timer = 0;
 
   uint64_t ntri;
   TriFind tri;

### 2010-10-29 paper/graph.tex
Fixed a typo in pagerank algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@643 3fc257bb-8550-0410-81b3-a5c103da83f7
Added discussion of improvements that could be made to PageRank.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@645 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-The algorithm in Figure~\ref{fig:pr2} performs these iterations.  The
+The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
+linear-algebra concepts implemented in a MapReduce framework.  The
 nonzeros of matrix $A$ are initialized as described above; the
@@
 call MPI\_Allreduce to compute global norms and residuals.
+
+This linear-algebra approach is attractive because it allows construction
+of the MapReduce PageRank algorithm from simple operations (matrix-vector
+multiplication, dot products) on commonly used data types (matrices, vectors).
+One can build a variety of algorithms based on these primitives;
+indeed, linear-algebra toolkits such as Trilinos~\cite{Trilinos-Overview} and 
+PETSc~\cite{petsc-efficient} use this exact design.  In parallel, these toolkits
+share pre-computed data layouts and communication maps among distributed
+objects to enable efficient operations between the objects.
+For MapReduce, however, 
+separating the data into vectors and matrices results in additional overhead 
+to gather key-value pairs with the same key into a key-multivalue pair.  
+This overhead can be reduced in PageRank, for example, by storing the
+$x$ and $y$ vectors in a single MapReduce object, with key $i$ and value 
+$(x_i, y_i)$.  Doing so eliminates the copy operation in Step 2 of 
+Figure~\ref{fig:pr2}, as well as the add and convert operations in Step 5.
+A further performance improvement would store an entire row of $A$ along with
+$x$ and $y$ in a single MapReduce object, with key $i$ and value 
+$(x_i, y_i, d_{out}, [j_1, A_{ij_1}], [j_2, A_{ij_2}], \dots, [j_{d_{out}}, 
+A_ij_{d_{out}}])$ (that is, storing together, for each vertex, its old and
+new PageRank value along with all out-going edges incident to the vertex).  
+This data layout may further reduce time spent in converting from key-value 
+pairs to key-multivalue pairs, but at the cost of code reuse and scalability.  
+While this data layout may be efficient for PageRank, it is more complex
+and less likely to be useful for additional algorithms.  In addition, since
+each key-value pair must fit within a single page in MR-MPI, storing entire
+matrix rows in a single key-value pair limits the size of the the matrices
+on which the algorithm can be applied.  The more general approach of storing
+non-zeros separately enables arbitrarily large problems to be solved.
+The detailed timing break-downs in section~\ref{subsec:results_pagerank} 
+allow one to estimate the benefit of these implementations relative to the
+code complexity and scalability.
 
@@
 \> \> Reduce $y$: \\
-\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
+\> \> \>  Input: Key = $i$; Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
 \> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\

### 2010-10-29 paper/results.tex
Adding tables of results; need to add the numbers and some text still.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@647 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2010-06-07 paper/results.tex
@@
 scaling.  However, MR-MPI's execution time is at least an order of
-magnitude greater than the Trilinos implementation.  This result is
-due to two factors: ({\it i}) a higher volume of communication in the
-MapReduce implementation (where all edges are communicated in each
-iteration), compared to the Trilinos implementation (where only
-vertex-based data is communicated), and ({\it ii}) out-of-core
-operations in MR-MPI were performed because of the page-size
-restriction, while all Trilinos operations were performed in-core.
-The benefit of MR-MPI's out-of-core implementation is seen, however,
-with the RMAT-28 data set, which could be solved on smaller processor
-sets than the Trilinos implementation.  For these experiments, the
-Trilinos implementation required 64 processors for the RMAT-28 data
-set, since it will not operate on matrices that exceed the aggregate
-memory of the processors.
+magnitude greater than the Trilinos implementation.  
+The benefit of MR-MPI's out-of-core implementation
+is seen, however, with the RMAT-28 data set, which could be solved on
+smaller processor sets than the Trilinos implementation.  For these
+experiments, the Trilinos implementation required 64 processors for
+the RMAT-28 data set, since it will not operate on matrices that
+exceed the aggregate memory of the processors.
 
@@
 \end{figure}
+
+A more detailed analysis of the performance of the MR-MPI PageRank is shown
+below.  We perform the analysis using RMAT-20, which is small enough to
+fit in memory and thus does not need out-of-core operations, and RMAT-28,
+which requires out-of-core operation in MR-MPI.  Table~\ref{table:prdetail}
+shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
+the MR-MPI implementation and the Trilinos implementation.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{|l|r|r|r|r|}
+\hline 
+ & \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
+ & MR-MPI & Trilinos & MR-MPI & Trilinos  \\
+\hline
+Step 1 (Compute adjustment) & & & &  \\
+Step 2 (Multiply $A^T x$) & & & &  \\
+Step 3 (Apply adjustment) & & & &  \\
+Step 4 (Scale $y$)  & & & & \\
+Step 5 (Compute residual)  & & & & \\
+\hline
+\end{tabular}
+\caption{Detailed execution times in each stage of the PageRank algorithm
+described in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
+\label{table:prdetail}
+\end{center}
+\end{table}
+
+In Tables~\ref{table:prdetail2},
+we break down the operation times further for the most expensive 
+step of the PageRank computation, Step 2.
+In the table, we divide the collate operation into its
+two parts:  an aggregate that involves parallel communication and a convert
+that gathers local key-value pairs with matching keys into a key-multivalue
+pair.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{|l|r|r|}
+\hline 
+ & RMAT-20 & RMAT-28 \\
+\hline
+Copy $x$ to $y$ & &  \\
+Add $A$ to $y$ & &  \\
+Convert $y$ & &  \\
+Reduce $y$ & & \\
+Aggregate $y$ (first part of collate) & & \\
+Convert $y$ (second part of collate) & & \\
+Reduce $y$ & & \\
+\hline
+\end{tabular}
+\caption{Detailed execution times of MR-MPI in Step 2 of 
+Figure~\ref{fig:pr2}.}
+\label{table:prdetial2}
+\end{center}
+\end{table}
+
+Similarly, in Table~\ref{table:prdetail5}, we show the execution times 
+required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
+residual computation.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{|l|r|r|}
+\hline 
+ & RMAT-20 & RMAT-28 \\
+\hline
+Add $y$ to $x$ & &  \\
+Convert $x$ & &  \\
+Reduce $x$ & & \\
+MPI\_Allreduce & & \\
+\hline
+\end{tabular}
+\caption{Detailed execution times of MR-MPI in Step 5 of 
+Figure~\ref{fig:pr2}.}
+\label{table:prdetial2}
+\end{center}
+\end{table}
+%This result is
+%due to two factors: ({\it i}) a higher volume of communication in the
+%MapReduce implementation (which comunicates one datum per graph edge
+%in each iteration), compared to the Trilinos implementation which only
+%communicates a volume of data proportional to the number of vertices,
+%and ({\it ii}) out-of-core operations in MR-MPI were performed because
+%of the page-size restriction, while all Trilinos operations were
+%performed in-core.  
+%
 

### 2010-10-29 paper/results.tex
Updated with RMAT-20 results; still waiting for RMAT-28 runs.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@648 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-A more detailed analysis of the performance of the MR-MPI PageRank is shown
+A more detailed analysis of the performance of the MR-MPI PageRank 
+algorithm is shown
 below.  We perform the analysis using RMAT-20, which is small enough to
@@
 shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
-the MR-MPI implementation and the Trilinos implementation.
+the MR-MPI implementation and the Trilinos implementation.  We see that for
+both implementation, the matrix-vector multiplication (Step 2) is the most
+costly step.  However, the MR-MPI algorithm is consistently slower than
+the Trilinos implementation in each step.  This effect is more pronounced
+for RMAT-28, where MR-MPI performs out-of-core operations.
 
@@
 \hline
-Step 1 (Compute adjustment) & & & &  \\
-Step 2 (Multiply $A^T x$) & & & &  \\
-Step 3 (Apply adjustment) & & & &  \\
-Step 4 (Scale $y$)  & & & & \\
-Step 5 (Compute residual)  & & & & \\
+Step 1 (Compute adjustment) & 0.0203 & 0.00030 &
+&  \\
+Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 &
+&  \\
+Step 3 (Apply adjustment) & 0.0015 & 0.00015 &
+&  \\
+Step 4 (Scale $y$)  & 0.0013 & 0.00014 &
+& \\
+Step 5 (Compute residual)  & 0.0207 & 0.00014 &
+& \\
 \hline
 \end{tabular}
-\caption{Detailed execution times in each stage of the PageRank algorithm
-described in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
+\caption{Detailed execution times (in seconds) per PageRank iteration
+for each stage of the algorithm
+in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
 \label{table:prdetail}
@@
 
-In Tables~\ref{table:prdetail2},
+To understand where MR-MPI spends its effort during the PageRank computation,
 we break down the operation times further for the most expensive 
-step of the PageRank computation, Step 2.
+step of the PageRank computation, Step 2; the results are in 
+Table~\ref{table:prdetail2}.
 In the table, we divide the collate operation into its
@@
 that gathers local key-value pairs with matching keys into a key-multivalue
-pair.
+pair.  All interprocessor communication occurs in the aggregate operation.
+While the time spent in aggregate is significant, it is small compared to
+the time spent in the two convert operations.  Further experimentation
+is needed to determine whether reorganizing the MapReduce data 
+(as described in Section~\ref{subsec:graph_pagerank}) would reduce the
+execution time of Step 2.  Organizing the data such that both the matrix
+$A$ and vectors $x$ and $y$ are in the same MapReduce object would eliminate
+the copy and add functions, as well as the first convert.  However, to update
+the $y$ values at the completion of Step 2, either an additional convert
+or an aggregate with greater amounts of data would be needed.
+
+STILL NEED RMAT-28 RESULTS AND DISCUSSION OF I/O EFFECT.
 
@@
 \hline 
- & RMAT-20 & RMAT-28 \\
+PageRank Step 2 & RMAT-20 & RMAT-28 \\
 \hline
-Copy $x$ to $y$ & &  \\
-Add $A$ to $y$ & &  \\
-Convert $y$ & &  \\
-Reduce $y$ & & \\
-Aggregate $y$ (first part of collate) & & \\
-Convert $y$ (second part of collate) & & \\
-Reduce $y$ & & \\
+Copy $x$ to $y$ & 1.1\% &  \\
+Add $A$ to $y$ & 3.0\% &  \\
+Convert $y$ & 23.5\% &  \\
+Reduce $y$ & 7.7\% & \\
+Aggregate $y$ (first part of collate) & 24.9\% & \\
+Convert $y$ (second part of collate) & 38.2\% & \\
+Reduce $y$ & 1.6\% & \\
 \hline
 \end{tabular}
-\caption{Detailed execution times of MR-MPI in Step 2 of 
-Figure~\ref{fig:pr2}.}
-\label{table:prdetial2}
+\caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
+PageRank algorithm in Figure~\ref{fig:pr2}.}
+\label{table:prdetail2}
 \end{center}
@@
 required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
-residual computation.
+residual computation.  Clearly the time to convert key-value pairs to 
+key-multivalues dominates this computation.  Reorganization of the vectors
+$x$ and $y$ into a single MapReduce object (as described in 
+Section~\ref{subsec:graph_pagerank}) would benefit this step of the computation,
+as the add, convert, and reduce operations could be replaced by a single 
+map operation, with execution time comparable to the vector-scaling operation
+in Step 4.
 
@@
 \hline 
- & RMAT-20 & RMAT-28 \\
+PageRank Step 5 & RMAT-20 & RMAT-28 \\
 \hline
-Add $y$ to $x$ & &  \\
-Convert $x$ & &  \\
-Reduce $x$ & & \\
-MPI\_Allreduce & & \\
+Add $y$ to $x$ & 2.2\% &  \\
+Convert $x$ & 95.1\% &  \\
+Reduce $x$ & 2.5\% & \\
+MPI\_Allreduce & 0.2\% & \\
 \hline
 \end{tabular}
-\caption{Detailed execution times of MR-MPI in Step 5 of 
-Figure~\ref{fig:pr2}.}
-\label{table:prdetial2}
+\caption{Percentage of execution time in each operation of Step 5 of the MR-MPI 
+PageRank algorithm in  Figure~\ref{fig:pr2}.}
+\label{table:prdetail5}
 \end{center}

### 2010-10-29 paper/Response.tex
Fixed one typo caught by reviewers. Started a response document Response.tex.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@649 3fc257bb-8550-0410-81b3-a5c103da83f7

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{url}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.25in}

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

{\it
October 29, 2010\\
Response to reviewers:  \\
We thank the reviewers for their careful reading of our paper.
Their suggestions definitely strengthen the paper and sparked new ideas
for improving our algorithms.
Below we list the Reviewers' comments verbatim.
Our responses are shown in italics. \\
Steve Plimpton and 
Karen Devine
}

\vspace{0.5 in}

Reviewer \#1: The submitted paper features a description and analysis of a
new MapReduce implementation based on MPI.  The features and limitations
of the implementation are covered with respect to other MapReduce
implementations.  The implementation is applied to several
graph-theoretic algorithms, and the performance of these applications
is compared with applications from an established library (Trilinos).

The introductory and supporting text are, overall, very good.  The
paper places MapReduce in its historical context and motivates its
usage.  The outline of strengths is helpful- important features
include support for out-of-core operations that other libraries (e.g.,
Trilinos) often cannot handle and the ability to use MPI directly in
addition to the MapReduce abstraction.  The paper then provides a good
overview of the composition of operations in a MapReduce application
and a summary of the out-of-core mechanisms available.  The figure
(Figure 1) used to provide a high-level schematic of MapReduce data
flows could be enhanced- for example, it should indicate node-local
and communication intensive operations.

All performance results are shown as scaling studies for a range of
problem sizes for the six graph algorithms.  The results show that the
MapReduce applications do not compete well with the reference
implementation (Trilinos), but the authors make clear the limitations
involved in using this high-level library which trades some
performance for usability.  Additionally, the implementation supports
very large data sets which can only be processed using its out-of-core
feature set.  The results are presented well.

Since improved performance for small and mid-range problems is not the
contribution of the paper, it could be improved by stressing other
aspects of the system.  Real-world out-of-core problems could be
surveyed.  Code size and complexity could be considered with usability
as a motivation.  Is the MapReduce-based graph algorithm library
easier to use than the Trilinos library?  Should higher-level user
applications be considered?  Can the MapReduce-based inefficiencies be
isolated more concisely?

The paper does not have a concise "Conclusion", only a "Lessons Learned"
section.

\vspace{0.5 in}

Reviewer \#2: This is a nice paper addressing an important current topic
with a good quality open-source implementation.  The overall presentation
is good: the description of algorithms and implementations is clear and the
performance evaluation is detailed. As the authors pointed out, the
performance of the MR-MPI implementations of some graph algorithms falls
behind the matrix-based counterparts so it would be interesting to further
investigate where the performance descrepency comes from and if it's
possible to optimize the algorithm and implementation to bridge the
performance gap.

{\it We instrumented our MR-MPI and Trilinos implementations of PageRank 
to show where time is spent during the PageRank iterations.  These results
are in Table 2.  We further instrumented the two most expensive steps of
the PageRank iterations, as shown in Tables 3 and 4.  These experiments
showed that much time is spent in reorganizing data from key-value pairs
to key-multivalue pairs.  In sections 4.1 and 5.2, we describe alternative
data layouts that can reduce the number of times the data must be 
reorganized to improve performance.}

\vspace{0.5 in}

Reviewer \#3: I appreciated the clear exposition of your map/reduce
framework, its decomposition into a number of independent and composable
stages, and the explanation of how out-of-core computations are handled.

I also liked that you presented the map/reduce implementation of a number
of different graph algorithms. However, I thought the presentation of these
algorithms could be improved in two ways:
- The text tends to just explaining what the code does, not why, nor why
  this solves the specific problem (4.2 was the worst culprit here - f I
understood correctly, the vertices get labeled by edge counts that are used
to give the edges a direction that helps identify triangles: if this is
correct a little diagram showing this would help...)
- The detailed algorithms (Figures 4-9) are a little too imprecise to
  follow at times. In particular, when multiple map/reduces are involved
the lack of names for the map/reduce operations and their inputs makes
things confusing. Please  be very explicit (after all the code itself is),
and Figure 4 at least actually had names for the map-reduce objects.

Your performance is quite low when compared to non map/reduce
implementations of the same algorithm. It would be interesting to know the
breakdown of execution time between map, shuffle and reduce phases to get a
better understanding of where your system spends its time (not necessarily
for every single performance measurement, but some representative or
overall measurements would be interesting).

{\it Reviewer \#2 made this same suggestion; please see the response above.}

Minor:
- 5.4: In Figure *13* (not 6)...

{\it Fixed; good catch.}
\end{document}

### 2010-10-30 paper/results.tex
Fixed one typo caught by reviewers. Started a response document Response.tex.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@649 3fc257bb-8550-0410-81b3-a5c103da83f7
Some RMAT-28 results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@650 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \hline
-Step 1 (Compute adjustment) & 0.0203 & 0.00030 &
-&  \\
-Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 &
-&  \\
-Step 3 (Apply adjustment) & 0.0015 & 0.00015 &
-&  \\
-Step 4 (Scale $y$)  & 0.0013 & 0.00014 &
-& \\
-Step 5 (Compute residual)  & 0.0207 & 0.00014 &
-& \\
+Step 1 (Compute adjustment) & 0.0203 & 0.00030 & 17.53 & 0.0838 \\
+Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 & 341.26 & 2.9239 \\
+Step 3 (Apply adjustment) & 0.0015 & 0.00015 & 2.14 & 0.0494 \\
+Step 4 (Scale $y$)  & 0.0013 & 0.00014 & 1.74 & 0.0359 \\
+Step 5 (Compute residual)  & 0.0207 & 0.00014 & 21.20 & 0.0457 \\
 \hline
@@
 While the time spent in aggregate is significant, it is small compared to
-the time spent in the two convert operations.  Further experimentation
+the time spent in the two convert operations.  When out-of-core operations
+are used for RMAT-28, the communication cost in the aggregate operation is
+a smaller percentage of total time for Step 2, 
+as other data manipulation operations such as convert and add 
+take a relatively larger percentage of the total time.
+Further experimentation
 is needed to determine whether reorganizing the MapReduce data 
@@
 
-STILL NEED RMAT-28 RESULTS AND DISCUSSION OF I/O EFFECT.
-
 \begin{table}[htb]
@@
 \hline
-Copy $x$ to $y$ & 1.1\% &  \\
-Add $A$ to $y$ & 3.0\% &  \\
-Convert $y$ & 23.5\% &  \\
-Reduce $y$ & 7.7\% & \\
-Aggregate $y$ (first part of collate) & 24.9\% & \\
-Convert $y$ (second part of collate) & 38.2\% & \\
-Reduce $y$ & 1.6\% & \\
+Copy $x$ to $y$                       &  1.1\% &  0.4\%\\
+Add $A$ to $y$                        &  3.0\% & 14.3\% \\
+Convert $y$                           & 23.5\% & 43.6\% \\
+Reduce $y$                            &  7.7\% &  5.8\% \\
+Aggregate $y$ (first part of collate) & 24.9\% &  6.8\% \\
+Convert $y$ (second part of collate)  & 38.2\% & 27.4\% \\
+Reduce $y$                            &  1.6\% &  1.7\% \\
 \hline
@@
 \hline
-Add $y$ to $x$ & 2.2\% &  \\
-Convert $x$ & 95.1\% &  \\
-Reduce $x$ & 2.5\% & \\
-MPI\_Allreduce & 0.2\% & \\
+Add $y$ to $x$ &  2.2\% &  9.8\% \\
+Convert $x$    & 95.1\% & 84.9\% \\
+Reduce $x$     &  2.5\% &  5.3\% \\
+MPI\_Allreduce &  0.2\% &  0.0\% \\
 \hline
@@
 
-In Figure~\ref{fig:cc}, we show the performance of the connected
+In Figure~\ref{f:cc}, we show the performance of the connected
 component identification algorithm (Figure~\ref{fig:cc}).  A

### 2010-11-03 paper/Response.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-{\it
 October 29, 2010\\
+
 Response to reviewers:  \\
-We thank the reviewers for their careful reading of our paper.
-Their suggestions definitely strengthen the paper and sparked new ideas
-for improving our algorithms.
-Below we list the Reviewers' comments verbatim.
-Our responses are shown in italics. \\
-Steve Plimpton and 
-Karen Devine
-}
+
+We thank the reviewers for their careful reading of our paper.  Their
+suggestions definitely strengthen the paper and sparked new ideas for
+improving our algorithms.  Below we list the Reviewers' comments
+verbatim.  Our responses are shown in italics. \\
+
+Steve Plimpton and Karen Devine
 
@@
 
+{\it Figure 1 is actually describing an on-processor rearrangement of
+data that each processor performs on its local data it owns.  Thus
+there is no inter-processor communication involved.}
+
 All performance results are shown as scaling studies for a range of
@@
 
+{\it We have made some of these comparison points more
+clearly in the final ``Lessons Learned'' section.}
+
+%comment on MR vs Trilinos, even if libraries already exist
+%  cannot use linear algebra lib for all graph algorithms described
+%  here - PR is an exception since can be cast as matvec
+%added line counts of algorithms
+
 \vspace{0.5 in}
@@
 
-{\it We instrumented our MR-MPI and Trilinos implementations of PageRank 
-to show where time is spent during the PageRank iterations.  These results
-are in Table 2.  We further instrumented the two most expensive steps of
-the PageRank iterations, as shown in Tables 3 and 4.  These experiments
-showed that much time is spent in reorganizing data from key-value pairs
-to key-multivalue pairs.  In sections 4.1 and 5.2, we describe alternative
-data layouts that can reduce the number of times the data must be 
-reorganized to improve performance.}
+{\it We instrumented our MR-MPI and Trilinos implementations of
+PageRank to show where time is spent during the PageRank iterations.
+These results were added in Table 2.  We further instrumented the two
+most expensive steps of the PageRank iterations, as shown in Tables 3
+and 4.  These experiments showed that considerable time is spent in
+reorganizing data from key-value pairs to key-multivalue pairs.  In
+sections 4.1 and 5.2, we describe alternative data layouts that could
+reduce the number of times the data must be reorganized to improve
+performance.}
 
@@
 
-I also liked that you presented the map/reduce implementation of a number
-of different graph algorithms. However, I thought the presentation of these
-algorithms could be improved in two ways:
-- The text tends to just explaining what the code does, not why, nor why
+I also liked that you presented the map/reduce implementation of a
+number of different graph algorithms. However, I thought the
+presentation of these algorithms could be improved in two ways: - The
+text tends to just explaining what the code does, not why, nor why
 this solves the specific problem (4.2 was the worst culprit here - f I
-understood correctly, the vertices get labeled by edge counts that are used
-to give the edges a direction that helps identify triangles: if this is
-correct a little diagram showing this would help...)
-- The detailed algorithms (Figures 4-9) are a little too imprecise to
-  follow at times. In particular, when multiple map/reduces are involved
-the lack of names for the map/reduce operations and their inputs makes
-things confusing. Please  be very explicit (after all the code itself is),
-and Figure 4 at least actually had names for the map-reduce objects.
+understood correctly, the vertices get labeled by edge counts that are
+used to give the edges a direction that helps identify triangles: if
+this is correct a little diagram showing this would help...)  - The
+detailed algorithms (Figures 4-9) are a little too imprecise to follow
+at times. In particular, when multiple map/reduces are involved the
+lack of names for the map/reduce operations and their inputs makes
+things confusing. Please be very explicit (after all the code itself
+is), and Figure 4 at least actually had names for the map-reduce
+objects.
+
+{\it We significantly re-worked and standardized the notation used
+in Section 4 to describe all the graph algorithms.  In particular
+we noted which operations are performed with which MapReduce
+objects and their associated key/value data, so that the flow
+of data through each algorithm should be more clear.  A discussion
+of the MR-MPI library's ability to allow creation of multiple
+MapReduce objects was added to Section 2.}
 
@@
 
-{\it Fixed; good catch.}
+{\it Fixed this; good catch.}
+
 \end{document}

### 2010-11-03 paper/graph.tex
Minor change
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@646 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 that sum to 1.0 and are discussed below.  The algorithm in
-Figure~\ref{fig:rmat} generates $M$ unique non-zero entries in a
-sparse $N \times N$ matrix $A$, where each entry $A_{ij}$ represents
-an edge between graph vertices $(V_i,V_j)$.
+Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
+unique non-zero entries in a sparse $N \times N$ matrix $A$, where
+each entry $A_{ij}$ represents an edge between graph vertices
+$(V_i,V_j)$.
 
@@
  \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
- processors.}
+ processors, using a single MapReduce object.}
 
@@
  \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
- on $P$ processors.}
+ on $P$ processors, using two MapReduce objects $E_{new}$ and
+ $E_{old}$.}
 
@@
 The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
-linear-algebra concepts implemented in a MapReduce framework.  The
-nonzeros of matrix $A$ are initialized as described above; the
+linear-algebra concepts implemented in a MapReduce framework with data
+stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$ stores
+the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are vectors.
+The nonzeros of matrix $A$ are initialized as described above; the
 PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
@@
 collate()} operations (which are equivalent to an {\it aggregate()}
-followed by {\it convert()}) with {\it convert()} calls.  In
-Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
+followed by {\it convert()}) with simply {\it convert()} calls.  In
+Figure~\ref{fig:pr2}, steps where {\it convert()} operations were
 substituted for {\it collate()} operations are marked with an
@@
 is useful in many MapReduce algorithms, but assumes that specific keys
-are always mapped to the same processor.
+can always be mapped to the same processor.
 
@@
 for random jumps from any page are also computed.  In Step 2, the
-matrix-vector product $A^T x$ is computed.  This is the most expensive
-part of the computation, requiring two passes over the nonzero
-structure of $A$.  In the first pass, a {\it convert()} gathers all
-local row entries $A_{ij}$ with their associated $x_i$ entry, and a
-{\it reduce()} computes $A_{ij} x_i$.  In the second pass, a {\it
-collate()} gathers, for each $j$, all contributions to the column sum
-$\sum_i A_{ij} x_i$, which is computed by a second {\it reduce()}.
-Steps 3 and 4 adjust and scale the product vector; the work is
-proportional to $|V|$.  Step 5 computes the residual to determine
-whether the PageRank vector has converged; work is again proportional
-to $|V|$.  In Step 6, the PageRank vector is overwritten prior to the
-next iteration.  Throughout the algorithm, we exploit the ability to
-call MPI\_Allreduce to compute global norms and residuals.
+matrix-vector product $A^T x$ is computed and the result storey in
+$y$.  This is the most expensive part of the computation, requiring
+two passes over the nonzero structure of $A$.  In the first pass, a
+{\it convert()} gathers all local row entries $A_{ij}$ with their
+associated $x_i$ entry, and a {\it reduce()} computes $A_{ij} x_i$.
+In the second pass, a {\it collate()} gathers, for each $j$, all
+contributions to the column sum $\sum_i A_{ij} x_i$, which is computed
+by a second {\it reduce()}.  Steps 3 and 4 adjust and scale the
+product vector; the work is proportional to $|V|$.  Step 5 computes
+the residual to determine whether the PageRank vector has converged;
+work is again proportional to $|V|$.  In Step 6, the PageRank vector
+is overwritten prior to the next iteration.  Throughout the algorithm,
+we exploit the ability to call MPI\_Allreduce to compute global norms
+and residuals.
 
-This linear-algebra approach is attractive because it allows construction
-of the MapReduce PageRank algorithm from simple operations (matrix-vector
-multiplication, dot products) on commonly used data types (matrices, vectors).
-One can build a variety of algorithms based on these primitives;
-indeed, linear-algebra toolkits such as Trilinos~\cite{Trilinos-Overview} and 
-PETSc~\cite{petsc-efficient} use this exact design.  In parallel, these toolkits
-share pre-computed data layouts and communication maps among distributed
-objects to enable efficient operations between the objects.
-For MapReduce, however, 
-separating the data into vectors and matrices results in additional overhead 
-to gather key-value pairs with the same key into a key-multivalue pair.  
-This overhead can be reduced in PageRank, for example, by storing the
-$x$ and $y$ vectors in a single MapReduce object, with key $i$ and value 
-$(x_i, y_i)$.  Doing so eliminates the copy operation in Step 2 of 
-Figure~\ref{fig:pr2}, as well as the add and convert operations in Step 5.
-A further performance improvement would store an entire row of $A$ along with
-$x$ and $y$ in a single MapReduce object, with key $i$ and value 
-$(x_i, y_i, d_{out}, [j_1, A_{ij_1}], [j_2, A_{ij_2}], \dots, [j_{d_{out}}, 
-A_ij_{d_{out}}])$ (that is, storing together, for each vertex, its old and
-new PageRank value along with all out-going edges incident to the vertex).  
-This data layout may further reduce time spent in converting from key-value 
-pairs to key-multivalue pairs, but at the cost of code reuse and scalability.  
-While this data layout may be efficient for PageRank, it is more complex
-and less likely to be useful for additional algorithms.  In addition, since
-each key-value pair must fit within a single page in MR-MPI, storing entire
-matrix rows in a single key-value pair limits the size of the the matrices
-on which the algorithm can be applied.  The more general approach of storing
+This linear-algebra approach is attractive because it allows
+construction of the MapReduce PageRank algorithm from simple
+operations (matrix-vector multiplication, dot products) on commonly
+used data types (matrices, vectors).  One can build a variety of
+algorithms based on these primitives; indeed, linear-algebra toolkits
+such as Trilinos~\cite{Trilinos-Overview} and
+PETSc~\cite{petsc-efficient} exploit this idea.  In parallel, these
+toolkits share pre-computed data layouts and communication maps among
+distributed objects to enable efficient operations between the
+objects.  For MapReduce, however, separating the data into vectors and
+matrices results in additional overhead to gather KV pairs with the
+same key into a key-multivalue (KMV) pair.  This overhead can be
+reduced in PageRank, for example, by storing the $x$ and $y$ vectors
+in a single MapReduce object, with key $i$ and value $(x_i, y_i)$.
+Doing so eliminates the {\it copy()} operation in Step 2 of
+Figure~\ref{fig:pr2}, as well as the {\it add()} and {\it convert()}
+operations in Step 5.  A further performance improvement would store
+an entire row of $A$ along with $x$ and $y$ in a single MapReduce
+object, with key $i$ and value $(x_i, y_i, d_{out}, [j_1, A_{ij_1}],
+[j_2, A_{ij_2}], \dots, [j_{d_{out}}, A_ij_{d_{out}}])$ (that is,
+storing together, for each vertex, its old and new PageRank value
+along with all out-going edges incident to the vertex).  This data
+layout eliminates an additional {it convert()} operation, but at the
+cost of code reuse and scalability.  While this data layout may be
+efficient for PageRank, it is more complex and less likely to be
+useful for additional algorithms.  In addition, since each KV pair
+must fit within a single page in MR-MPI, storing entire matrix rows in
+a single KV pair limits the size of the the matrices on which the
+algorithm can be applied.  The more general approach of storing
 non-zeros separately enables arbitrarily large problems to be solved.
-The detailed timing break-downs in section~\ref{subsec:results_pagerank} 
-allow one to estimate the benefit of these implementations relative to the
-code complexity and scalability.
+The detailed timing break-downs in
+section~\ref{subsec:results_pagerank} allow one to estimate the
+benefit of these implementations relative to code complexity and
+scalability issues.
 
@@
 %  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
-%Input: \\
+
 %\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
@@
 %
- %\caption{\it MapReduce algorithm for PageRank vertex ranking.}
+ %\caption{\it MapReduce algorithm for PageRank vertex ranking, using
+ %4 MapReduce objects $A$, $MT$, $x$, and $y$.}
 %
@@
 
-Input: \\
-$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
-$*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
-$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
+$*$ Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
+$*$ Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
+$*$ Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
+
 While (${residual} > {tolerance}$) \\
@@
 
- \caption{\it MapReduce algorithm for PageRank vertex ranking;
-pre-aggregating the MapReduce objects allows many communication
-intensive {\it collate()} operations to be replaced by strictly local
-{\it convert()} operations (as marked with asterisks).}
+ \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
+ MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
+ of the MapReduce objects allows many communication intensive {\it
+ collate()} operations to be replaced by local {\it convert()}
+ operations (marked with asterisks).}
 
@@
 \ref{fig:tri} outlines a MapReduce algorithm that enumerates all
-triangles, assuming an input graph of undirected edges $(V_i,V_j)$
+triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
 where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
@@
 
+MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
+
 1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
-1 Map: \> Convert edges to vertices \\
+1 Map $G$: \> Convert edges to vertices \\
            \> \> input Key = $(V_i,V_j)$, Value = NULL \\
@@
            \> \> output Key = $V_j$, Value = $V_i$ \\
-1 Collate \\
-1 Reduce: \> Add first degree to one vertex in edge \\
+1 Collate $G$ \\
+1 Reduce $G$: \> Add first degree to one vertex in edge \\
               \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
@@
               \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
-2 Collate \\
-2 Reduce: \> Add second degree to other vertex in edge \\
+2 Collate $G$ \\
+2 Reduce $G$: \> Add second degree to other vertex in edge \\
               \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
               \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
-3 Map: \> Low degree vertex emits edges \\
+3 Map $G$: \> Low degree vertex emits edges \\
            \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
@@
            \> \> else: output Key = $V_i$, Value = $V_j$ \\
-3 Collate \\
-3 Reduce: \> Emit angles of each vertex \\
+3 Collate $G$ \\
+3 Reduce $G$: \> Emit angles of each vertex \\
               \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
@@
 	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
-4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs \\
-4 Collate \\
-4 Reduce: \> Emit triangles \\
+4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs in $G$ \\
+4 Collate $G$ \\
+4 Reduce $G$: \> Emit triangles \\
               \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
@@
 
- \caption{\it MapReduce algorithm for triangle enumeration.}
+ \caption{\it MapReduce algorithm for triangle enumeration, using 2 MapReduce objects $G$ and $G_0$}
 
@@
 (KV) pairs in an auxiliary MapReduce object $G_0$, for use later in
-the algorithm.  The first {\it map()} operation converts edge keys to
-vertex keys with edge values.  After a {\it collate()}, each vertex
-has a list of vertices it is connected to; the first {\it reduce()}
-can thus flag one vertex $V_i$ in each edge with a degree count $D_i$.
-The second {\it collate()} and {\it reduce()} assign a degree count
-$D_j$ to the other vertex in each edge.  In the third {\it map()},
-only the lower-degree vertex in each edge emits its edges KV pairs.
-The task of the third {\it reduce()} is to emit ``angles'' for each of
-these low-degree vertices.  An ``angle'' is a root vertex $V_i$, with
-two edges to vertices $V_1$ and $V_2$, i.e. a triangle without the
-third edge $(V_1,V_2)$.  The {\it reduce()} emits a list of all angles
-of vertex $V_i$, by a double loop over the edges of $V_i$.  Note that
-the aggregate volume of KV pairs emitted at this stage is minimized by
-having only the low-degree vertex in each edge generate angles.
+the algorithm.  The first {\it map()} operation on $G$ converts edge
+keys to vertex keys with edge values.  After a {\it collate()}, each
+vertex has a list of vertices it is connected to; the first {\it
+reduce()} can thus flag one vertex $V_i$ in each edge with a degree
+count $D_i$.  The second {\it collate()} and {\it reduce()} assign a
+degree count $D_j$ to the other vertex in each edge.  In the third
+{\it map()}, only the lower-degree vertex in each edge emits its edges
+KV pairs.  The task of the third {\it reduce()} is to emit ``angles''
+for each of these low-degree vertices.  An ``angle'' is a root vertex
+$V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a triangle
+without the third edge $(V_1,V_2)$.  The {\it reduce()} emits a list
+of all angles of vertex $V_i$, by a double loop over the edges of
+$V_i$.  Note that the aggregate volume of KV pairs emitted at this
+stage is minimized by having only the low-degree vertex in each edge
+generate angles.
 
 In stage 4, KV pairs from the original graph $G_0$ are added to the
-current working set of KV pairs.  The KV pairs in $G_0$ are edges that
-complete triangles for the angle KV pairs just generated.  After the
-fourth {\it collate()}, a pair of vertices $(V_i,V_j)$ is the key, and
-the multivalue is the list of all root vertices in angles that contain
-$V_i$ and $V_j$.  If the multivalue also contains a NULL, contributed
-by $G_0$, then there is a $(V_i,V_j)$ edge in the graph.  Thus all
-vertices in the multivalue are roots of angles that are complete
-triangles and can be emitted as a triplet key.
+current working set of KV pairs in MapReduce object $G$.  The KV pairs
+in $G_0$ are edges that complete triangles for the angle KV pairs just
+generated.  After the fourth {\it collate()}, a pair of vertices
+$(V_i,V_j)$ is the key, and the multivalue is the list of all root
+vertices in angles that contain $V_i$ and $V_j$.  If the multivalue
+also contains a NULL, contributed by $G_0$, then there is a
+$(V_i,V_j)$ edge in the graph.  Thus all vertices in the multivalue
+are roots of angles that are complete triangles and can be emitted as
+a triplet key.
 
@@
 component ID.  All vertices in the same component are labelled with
-the same ID, which is the ID of a vertex in the component.  We assume
-an input graph of undirected edges $(V_i,V_j)$.  This exposition also
-follows the connected-component algorithm presented in~\cite{Cohen09},
-with the addition of logic that load-balances data across processors,
-when one or a few giant components exist in the graph.
+the same ID, which is the ID of some vertex in the component.  We
+assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
+exposition also follows the connected-component algorithm presented
+in~\cite{Cohen09}, with the addition of logic that load-balances data
+across processors, which is important when one or a few giant
+components exist in the graph.
 
@@
 
+MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
+MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
+MapReduce object $W$ = workspace
+
 Iterate: \\
-\> 1 Map: \> Convert edges to vertices \\
+
+\> 1 Map $W$ using $E$ as input: \> Convert edges to vertices \\
     \> \> \> input Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
@@
     \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
-\> 1 Add: \> Zone assignment of each vertex \\
-    \> \> \> output Key = $V_i$, Value = $Z_i$ \\
-\> 1 Collate: \> Vertex as key \\
-\> 1 Reduce: \> Emit edges of each vertex with zone of vertex \\
+\> 1 Add $V$ to $W$ \\
+\> 1 Collate $W$: \> Vertex as key \\
+\> 1 Reduce $W$: \> Emit edges of each vertex with zone of vertex \\
        \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
@@
 
-\> 2 Collate: \> Edge as key \\
-\> 2 Reduce: \> Emit zone re-assignments \\
+\> 2 Collate $W$: \> Edge as key \\
+\> 2 Reduce $W$: \> Emit zone re-assignments \\
        \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
@@
       \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
-\> 2 Exit: \> if no output by Reduce 2 \\
-\> 3 Map: \> Invert vertex/zone pairs \\
+\> 2 Exit: \> if $W$ is empty, i.e. no output by Reduce 2 \\
+\> 3 Map $V$: \> Invert vertex/zone pairs \\
     \> \> \> input Key = $V_i$, Value = $Z_i$ \\
@@
    \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
-\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
+\> 4 Map $V$ using $W$ as input: \> Add zone reassignments in $W$ to $V$ \\
     \> \> \> if $Z_i$ is not partitioned: \\
@@
    \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
-\> 3 Collate: \> Zone ID as key \\
-\> 3 Reduce: \> Emit new zone assignment of each vertex \\
+\> 4 Collate $V$: \> Zone ID as key \\
+\> 4 Reduce $V$: \> Emit new zone assignment of each vertex \\
        \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
@@
 
- \caption{\it MapReduce algorithm for connected component labeling.}
+ \caption{\it MapReduce algorithm for connected component labeling,
+ using 3 MpaReduce objects $E$, $V$, and $W$.}
 
@@
 The algorithm begins (before the iteration loop) by assigning each
-vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$.  Each
-iteration grows the zones, one layer of neighbors at a time.  As
-zones collide due to shared edges, a winner is chosen (the smaller
-zone ID), and vertices in the losing zone are reassigned to the
-winning zone.  When the iterations complete, each zone has
-become a fully connected component.  The algorithm thus finds all
-connected components in the graph simultaneously.  The number of
-iterations required depends on the largest diameter of any component
-in the graph.
+vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$ and
+storing this assignment in MapReduce object $V$.  Each iteration grows
+the zones, one layer of neighbors at a time.  As zones collide due to
+shared edges, a winner is chosen (the smaller zone ID), and vertices
+in the losing zone are reassigned to the winning zone.  When the
+iterations complete, each zone has become a fully connected component
+and the MR object $V$ holds that zone assignment for every vertex.
+The algorithm thus finds all connected components in the graph
+simultaneously.  The number of iterations required depends on the
+largest diameter of any component in the graph.
 
-The first {\it map()} operation emits the vertices in each edge as
+The first {\it map()} operation uses MapReduce object $E$ as input to
+the empty MapReduce object $W$ to emit the vertices in each edge as
 keys, with the edge as a value.  The current zone assignment of each
-vertex is added to the set of key/value (KV) pairs.  The first {\it
-collate()} operation collects all the edges of a vertex and its zone
-assignment together in one multi-value.  The first {\it reduce()}
-operation re-emits each edge, tagged by the zone assignment of one of
-its vertices.
+vertex in $V$ is added to the set of key/value (KV) pairs in $W$.  The
+first {\it collate()} operation collects all the edges of a vertex and
+its zone assignment together in one multi-value.  The first {\it
+reduce()} operation re-emits each edge, tagged by the zone assignment
+of one of its vertices.
 
 Since each edge was emitted twice, the second {\it collate()}
-operation collects the zone assignments for its two vertices together.
-If the two zone IDs are different, the second {\it reduce()} operation
-chooses a winner (the smaller of the two IDs), and emits the loser ID
-as a key, with the winning ID as a value.  If no zone ID changes are
-emitted, the algorithm is finished, and the iterations cease.
+operation on $W$ collects the zone assignments for its two vertices
+together.  If the two zone IDs are different, the second {\it
+reduce()} operation chooses a winner (the smaller of the two IDs), and
+emits the loser ID as a key, with the winning ID as a value.  If no
+zone ID changes are emitted, the algorithm is finished, and the
+iterations cease.
 
 The third {\it map()} operation inverts the vertex/zone KV pairs to
-become zone/vertex pairs.  The third {\it add()} operation adds the
-changing zone assignments to the set of KV pairs.  The third {\it
-collate()} can then collect all the vertices of a zone and zero or
-more reassignments for the zone ID.  Since a zone could collide with
-multiple other zones on the same iteration due to shared edges, the
-new zone ID becomes the minimum ID of any of the neighboring zones.
-If no zone reassignment value appears in the multi-value, the zone ID
-is unchanged.  The final {\it reduce()} emits a KV pair for each
-vertex in the zone, with the vertex as a key and the new zone ID as a
-value.
+become zone/vertex KV pairs.  The partitioning logic described in
+Figure~\ref{fig:cc} for this step (and subsequent ones) is described
+below.  The fourth {\it map()} operation adds the changed zone
+assignments in $W$ to the set of KV pairs in $V$.  The fourth {\it
+collate()} can now collect all the vertices in each zone along with
+zero or more reassignments for that zone ID.  Since a zone could
+collide with multiple other zones on the same iteration due to shared
+edges, the new zone ID for a zone becomes the minimum ID of any of its
+neighboring zones.  If no zone reassignment values appear in the
+multi-value resulting from the fourth {\it collate()}, the zone ID is
+unchanged.  The final {\it reduce()} emits a KV pair for each vertex
+in the zone, with the vertex as a key and the new zone ID as a value.
+This is updated version of MapReduce object $V$, ready to be used as
+input for the next iteration.
 
-Note that if a graph has only a few components, the third {\it
-collate()} operation, which collates by the zone ID, may generate a
-few very large key/multi-value (KMV) pairs.  For example, if the
-entire graph is fully connected, in the last iteration, a single KMV
-pair contains all vertices in the graph and is assigned to one
-processor.  This imbalance in memory and computational work can lead
-to poor parallel performance of the overall algorithm.  To counter
-this effect, the various operations of stage 3 include extra logic.
+Note that if a graph has only a few components, the fourth {\it
+collate()} operation, which keys on the zone ID, may generate a few
+very large key/multi-value (KMV) pairs.  For example, if the entire
+graph is fully connected, in the last iteration, a single KMV pair
+assigned to one processor will result, containing all vertices in the
+graph.  This imbalance in memory and computational work can lead to
+poor parallel performance of the overall algorithm.  To counter this
+effect, the various operations in stages 3 and 4 include extra logic.
 The idea is to partition zones whose vertex count exceeds a
@@
 $Z_i^+$ notation in Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been
-partitioned, the third {\it add()} operation emits the zone ID change
-KV pair ($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In this case
-($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once for
-each processor, and once as if $Z_i$ had not been partitioned (for a
-reason discussed below).
+partitioned, the fourth {\it map()} operation emits the KV pair for
+the zone ID as ($Z_i^+$,$Z_{winner}$) instead of as
+($Z_i$,$Z_{winner}$).  In this case ($Z_i$,$Z_{winner}$), is emitted
+not once, but $P+1$ times, once for each processor, and once as if
+$Z_i$ had not been partitioned (for a reason discussed below).
 
-With this additional partitioning logic, the third {\it collate()}
+With this additional partitioning logic, the fourth {\it collate()}
 operation (which keys on zone IDs, some of which now include processor
 bits) collects only $1/P$ of the vertices in large zones onto each
-processor.  But the multivalue on each processor contains all the
-zone-reassignments relevant to the unpartitioned zone.  Thus, the
-third {\it reduce()} operation can change the zone ID (if necessary)
+processor.  Yet the multivalue on each processor still contains all
+the zone-reassignments relevant to the unpartitioned zone.  Thus, the
+fourth {\it reduce()} operation can change the zone ID (if necessary)
 in a consistent manner across all $P$ multi-values that contain the
-zone's vertices.  When the {\it reduce()} operation emits new zone
-assignments for each vertex, the zone retains its partitioned status;
-the partition bit is also explicitly set if the vertex count exceeds
-the threshhold for the first time.
+zone's vertices.  When the fourth {\it reduce()} operation emits new
+zone assignments for each vertex, the zone retains its partitioned
+status; the partition bit is also explicitly set if the vertex count
+exceeds the threshhold for the first time.
 
@@
 small zone may not set the partition bit of their zone IDs.  On
-subsequent iterations, the third {\it add()} operation emits $P+1$
+subsequent iterations, the fourth {\it map()} operation emits $P+1$
 copies of new zone reassignments for both partitioned and unpartioned
@@
 An ``independent'' set of vertices from a graph is one where no pair
-of vertices in the set shares an edge.  The set is ``maximal'' if no
-vertex can be added \footnote{A maximal set is ``maximum'' if it is
-the largest maximal set.  Finding a graph's maximum independent set is
-an NP-hard problem, which MapReduce is unlikely to help with.}.
-Finding a maximal independent set (MIS) is useful in several contexts,
-such as identifying large numbers of independent starting vertices for
-graph traversals.  Luby's algorithm~\cite{Luby86} is a well-known
-parallel method for finding a MIS.  Figure~\ref{fig:luby} outlines a
-MapReduce version of Luby's algorithm, assuming an input graph of
-undirected edges $(V_i,V_j)$ where $V_i < V_j$ for every edge, i.e. an
-upper-triangular R-MAT matrix.
+of vertices in the set is connected by an edge.  The set is
+``maximal'' if no vertex can be added to it \footnote{A maximal set is
+``maximum'' if it is the largest maximal set.  Finding a graph's
+maximum independent set is an NP-hard problem, which MapReduce is
+unlikely to help with.}.  Finding a maximal independent set (MIS) is
+useful in several contexts, such as identifying large numbers of
+independent starting vertices for graph traversals.  Luby's
+algorithm~\cite{Luby86} is a well-known parallel method for finding a
+MIS.  Figure~\ref{fig:luby} outlines a MapReduce version of Luby's
+algorithm, assuming an input graph of undirected edges $(V_i,V_j)$
+where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
+matrix.
 
@@
 
-Map: assign random values to each vertex $V_i$ and $V_j$ of edge $E_{ij}$. \\
-Clone: convert edge KV pairs directly to KMV pairs \\
-Create: empty MapReduce object MRv for maximal independent set vertices \\
+MapReduce object $V$ = maximal independent set vertices, initially empty \\
+MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
+Map $E$: assign random values to each vertex $V_i$ and $V_j$ in $E$ \\
+Clone $E$: convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
 \\
-while $N_{edges} > 0$: \\
-\> 1 Reduce: \> Determine WINNER/LOSER vertex of each edge \\
-       \> \> \> input Key = $E_{ij}$ with $V_i < V_j$, Multivalue = NULL or LOSER \\
-       \> \> \> if either value is LOSER, emit nothing \\
+while $N_{edges}$ in $E$ > 0: \\
+\> 1 Reduce $E$: \> Determine WINNER/LOSER vertex of each edge \\
+       \> \> \> input Key = $E_{ij}$, Multivalue = NULL or LOSER \\
+       \> \> \> if multivalue = LOSER, emit nothing \\
        \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
-       \> \> \> output Key = $V_w$, Value = $V_l$ WINNER \\
-       \> \> \> output Key = $V_l$, Value = $V_w$ LOSER \\
-\> 1 Collate: \> Vertex as key \\
-\> 2 Reduce: \> Find WINNER vertices \\
+       \> \> \> output Key = $V_w$, Value = $V_l$, WINNER \\
+       \> \> \> output Key = $V_l$, Value = $V_w$, LOSER \\
+\> 1 Collate $E$: \> Vertex as key \\
+\> 2 Reduce $E$: \> Find WINNER vertices \\
        \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if all $V$ are WINNERs: \\
+       \> \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
       \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ LOSER \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, WINNER \\
        \> \> \> else: \\
@@
      \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
-\> 2 Collate: \> Vertex as key \\
-\> 3 Reduce: \> Find LOSER vertices \\
+\> 2 Collate $E$: \> Vertex as key \\
+\> 3 Reduce $E$: \> Find LOSER vertices \\
        \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if any $V$ is LOSER: \\
+       \> \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
       \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ LOSER \\
+     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, LOSER \\
        \> \> \> else: \\
@@
      \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
-\> 3 Collate: \> Vertex as key \\
-\> 4 Reduce: \> Emit WINNER vertices and LOSER edges \\
+\> 3 Collate $E$: \> Vertex as key \\
+\> 4 Reduce $E$: \> Emit WINNER vertices and LOSER edges \\
        \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if all $V$ are LOSERs: \\
-      \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
+       \> \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
+      \> \> \> \> output to $V$: Key = $V_i$, Value = NULL \\
        \> \> \> for each $V_j$ in Multivalue: \\
-      \> \> \> \> if $V_j$ is LOSER: \\
-     \> \> \> \> \> output Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
+      \> \> \> \> if $V_j$ is flagged with LOSER: \\
+     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
       \> \> \> \> else: \\
-     \> \> \> \> \> output Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
-\> 4 Collate: \> Edge as key \\
+     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
+\> 4 Collate $E$: \> Edge as key \\
 
@@
  \caption{\it MapReduce algorithm for finding a maximal independent
- set of graph vertices via Luby's algorithm.}
+ set of graph vertices via Luby's algorithm, using two MapReduce
+ objects $E$ and $V$.}
 
@@
 
-Before the iterations begin, a random value is assigned to each
-vertex, which is used to compare pairs of vertices.  The vertex ID
-itself could be used as the random value, but since the vertices
-eventually selected for the MIS depend on the randomization, this may
-introduce an unwanted bias.  Instead a random number generator can be
-used to assign a consistent random value $R_i$ to each vertex in each
-edge (via a {\it map()} operation), which is then carried along with
-the vertex ID through each stage of the algorithm.  In the notation of
+Before the iterations begin, a random value is assigned to each vertex
+(stored consistently in multiple edges in $E$), which is used to
+compare pairs of vertices.  The vertex ID itself could be used as the
+random value, but since the vertices eventually selected for the MIS
+depend on the randomization, this may introduce an unwanted bias.
+Instead a random number generator can be used to assign a consistent
+random value $R_i$ to each vertex in each edge (via a {\it map()}
+operation), which is then carried along with the vertex ID through
+each stage of the algorithm.  In the notation of
 Figure~\ref{fig:luby}, each $V_i$ is then really two quantities, a
@@
 vertex ID, and the vertices in the final MIS could be remapped to
-recover their original IDs.  These operations could be performed with
-a few extra MapReduce steps, outside the iteration loop.
+recover their original IDs.  These operations can be performed with a
+few extra MapReduce steps, outside the iteration loop.
 
 The pre-iteration {\it clone()} operation converts the initial list of
-edge key/value (KV) pairs one by one to key/multivalue (KMV) pairs,
-with the edge as the key, and NULL as the value.  Thus, the iterations
-can begin with a {\it reduce()} operation, without need for a {\it
-collate()} operation.  A second empty MapReduce object is also
-initialized, which accumulates MIS vertices as the iterations proceed.
+edge key/value (KV) pairs stored in MapReduce object $E$ to
+key/multivalue (KMV) pairs, with the edge as the key, and NULL as the
+value.  Thus, the iterations can begin with a {\it reduce()}
+operation, without need for a {\it collate()} operation.  A second
+empty MapReduce object $V$ is also created, which will accumulate MIS
+vertices as the iterations proceed.
 
-At each iteration, the current set of edges is examined to identify
-winning vertices.  Vertices adjacent to winners are flagged as losers,
-and all edges adjacent to losers are flagged for removal at the start
-of the next iteration.  Vertices with all losing edges are flagged as
-winners.  At each iteration edges are removed from the graph, and
-winning vertices are added to the MIS.  When the graph is empty, the
-MIS is complete.
+At each iteration, the current set of edges in $E$ is examined to
+identify winning vertices.  Vertices adjacent to winners are flagged
+as losers, and all edges which contain a loser vertex are marked for
+removal at the start of the next iteration.  Vertices with all their
+edges marked as losers are also flagged as winners.  At each iteration
+marked edges are removed from the graph, and winning vertices are
+added to the MIS.  When the graph is empty, the MIS is complete.
 
-The first {\it reduce()} operation flags the two vertices in each edge as a
-potential winner or loser.  This is done by comparing the two random
-values for the vertices (or the randomized vertex IDs as explained
-above).  The first {\it collate()} operation thus produces a multivalue for
-each vertex $V_i$ which contains all the vertices it shares an edge
-with, and associated winner/loser flags.  In the second {\it reduce()}
-operation, if $V_i$ won the comparison with all its edge neighbors,
-then it is a ``winner'' vertex.  All vertices connected to the winner
-are emitted with a loser flag.  If $V_i$ is not an overall winner, its
-edge vertices are emitted without the loser flag.  The second
-{\it collate()} operation collects this new information for each vertex.
+The first {\it reduce()} operation flags the two vertices in each edge
+as a potential winner or loser.  This is done by comparing the two
+random values for the vertices (or the randomized vertex IDs as
+explained above).  The first {\it collate()} operation thus produces a
+multivalue for each vertex $V_i$ which contains all the vertices it
+shares an edge with, and associated winner/loser flags.
 
-In the third {\it reduce()} operation, if $V_i$ was flagged as a loser by
-any edge neighbor who was a winner, it becomes a losing vertex
-and emits all its edge vertices with a loser flag.  Otherwise it emits
-all its edge vertices without the flag.
+In the second {\it reduce()} operation, if $V_i$ won the comparison
+with all its neighbor vertices, then it becomes a ``winner'' vertex.
+All vertices connected to the winner are emitted with $V_i$ tagged
+with a winner flag.  If $V_i$ is not an overall winner, its neighbor
+vertices are emitted without the winner flag on $V_i$..  The second
+{\it collate()} operation collects this new information for each
+vertex.
+
+In the third {\it reduce()} operation, if any neighbor of $V_i$ was a
+winner in the previoue {\it reduce()} operation (indicated by a winner
+flag), $V_i$ becomes a losing vertex and emits all its neighbor
+vertices with $V_i$ tagged with a loser flag.  Otherwise it emits all
+its neighbor vertices without the loser flag on $V_i$.
 
 In the fourth {\it reduce()} operation, some vertices have all their
-edge vertices flagged as losers.  These vertices are all winners,
-either the ones identified in the second {\it reduce()} operation, or
-additional vertices who would become singletons (having no edges) when
-edges containing a loser vertex are removed (on the next iteration).
-All of these winning vertices are emitted to the MapReduce object that
-stores the accumulating MIS.
+neighbor vertices flagged as losers; these vertices become winners.
+They are either the vertices identified as winners in the second {\it
+reduce()} operation, or vertices who would become singletons (having
+no edges) when edges containing a loser vertex are removed (on the
+next iteration).  These winning vertices are emitted to the MapReduce
+object $V$ that stores the accumulating MIS.
 
@@
 together via the fourth {\it collate()} operation.  The multivalue for
-each edge now has two values, either of which may be NULL or a loser
+each edge now has two values, each of which is either NULL or a loser
 flag.
 
-At the first stage of the next iteration, any edge flagged by either
-value as a loser does not compare its two vertices; it is effectively
-deleted from the graph.  The Luby iterations end when all edges have
-been removed from the graph.  At this point, the second MapReduce
-object contains a complete MIS of vertices.
+In the first {\it reduce()} of the next iteration, any edge with a
+loser flag in its multivalue does not compare its two vertices; it is
+effectively deleted from the graph.  The Luby iterations end when all
+edges have been removed from the graph.  At this point, the MapReduce
+object $V$ contains a complete MIS of vertices.
 
@@
 from a chosen source vertex to all other vertices in the graph.  This
-operation is straightforward when global graph information is
+calculation is straightforward when global graph information is
 available.  The source vertex is labeled with distance zero.  Then in
 each iteration, edge adjacencies of labeled vertices are followed, and
-adjacent vertices are labeled with updated distances.  When no
+adjacent vertices are relabeled with updated distances.  When no
 distances change, the iterations are complete.  Most importantly, only
-edges from modified vertices are visited in each iteration.
+edges adjacent to modified vertices need be visited in each iteration.
 
-For distributed graphs, global information about edge adjacencies and
+For a distributed graph, global information about edge adjacencies and
 labeled vertices is not available.  Instead, following the
@@
 although labels for an edge's vertices are updated only if the
-breadth-first traversal has reached those vertices.  A MapReduce
-formulation of this algorithm, described in Figure~\ref{fig:sssp},
-begins with a {\it map()} function loading edges $(V_i, V_j)$ and
-source vertex $V_s$ with distance 0.  In each iteration, the edges
-$(V_i, V_j)$ are collated with respect to $V_i$, and a {\it reduce()}
-operation updates each $V_j$'s distance if $V_i$'s distance is
-non-zero.
+breadth-first traversal from the source vertex has reached those
+vertices.  A MapReduce formulation of this algorithm, described in
+Figure~\ref{fig:sssp}, begins with a {\it map()} function loading
+edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.  In each
+iteration, the edges $(V_i, V_j)$ are collated with respect to $V_i$,
+and a {\it reduce()} operation updates each $V_j$'s distance if
+$V_i$'s distance is non-zero.
 
@@
 
-Input:  \\
-\> Source vertex $V_s$\\
-\> MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$ (initially $\infty$):  \\
-\> \> Key = $V_i$; Value = $\infty$\\
-\> MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
-\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
-Map $V$:  Set distance for source vertex $V_s$ to zero \\
-\> Input:  Key = $V_i$; Value = $\infty$ \\
-\> Output:  Key = $V_i$; Value = 0 if $V_i = V_s$, otherwise Value = $\infty$ \\
+MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
+\> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
+MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
+$done$ = 0 \\
+
 while not $done$: \\
-\> $done$ = 1; \\
-\> Add: $E$ to $V$ \\
-\> Collate $V$:  Key is vertex $V_i$ \\
-\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices \\
-\> \> Input:  Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$, candidate distances $d_i$ \\
-\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$  \\
-\> \> \> If $d_{min} \ne d_i$, $done$ = 0; \\
-\> \> Output:  Key = $V_i$; Value = $d_{min}$ (updated distance for $V_i$) \\
-\> \> \> and if $d_{min} \ne d_i$, Key = $V_j$; Value = $d_{min} + w_{ij}$ for each edge ($V_i, V_j$) \\
-\> MPI\_Allreduce on $done$ with MPI\_MIN \\
-Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
+\> $done$ = 1 \\
+\> Map $V$ using $E$ as input: add edge weights in $E$ to $V$ \\
+\> \> input Key = $E_{ij}$, Value = $w_{ij}$ \\
+\> \> output Key = $V_i$, Value = $[V_j, w_{ij}]$ \\
+\> Collate $V$: Vertex is key \\
+\> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
+\> \> Input: Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$ and candidate distances $d_i$ \\
+\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
+\> \> If $d_{min} \ne d_i$: \\
+\> \> \>  $done$ = 0 \\
+\> \> \>  Output: Key = $V_j$, Value = $d_{min} + w_{ij}$ for each edge vertex $V_j$ \\
+\> \>  Output:  Key = $V_i$, Value = $d_{min}$ \\
+\> $done$ = MPI\_Allreduce($done$,MPI\_MIN)
 
@@
  \caption{\it MapReduce algorithm for single-source shortest path
- (SSSP) calculation.}
+ (SSSP) calculation, using two MapReduce objects $V$ and $E$.}
 
@@
 object.  The enhanced algorithm is shown in Figure~\ref{fig:sssp2}.
-The MapReduce objects storing vertices and edges are aggregated to
-processors only once at the beginning of the computation.  In each
-iteration of SSSP, only a small number of new candidate distances are
-generated.  The enhanced algorithm uses MapReduce object $U$ to store
-these updates.  Only the updates in $U$ are communicated; each update
-is aggregated to the same processor that stores the corresponding
-vertex and its edges.  Thus, the total amount of communication needed
-is smaller than in the original algorithm, which aggregates all graph
-edges in each iteration.  Updated vertex distances are stored in
-MapReduce object $V$.
+The MapReduce objects $V$ and $E$ storing vertices and edges are
+aggregated across processors only once at the beginning of the
+computation.  In each iteration of SSSP, only a small number of new
+candidate distances are generated.  The enhanced algorithm uses
+MapReduce object $U$ to store these updates.  Only the updates in $U$
+are communicated; each update is aggregated to the same processor that
+stores the corresponding vertex and its edges.  Thus, the total amount
+of communication needed is smaller than in the original algorithm,
+which aggregated all graph edges in each iteration.  Updated vertex
+distances are stored in MapReduce object $V$, as before.
 
@@
 
-Input:  \\
-\> Source vertex $V_s$.\\
-\> Pre-aggregated MapReduce object $V$ containing graph vertices $V_i$ with distances $d_i$ from $V_s$:  \\
-\> \> Key = $V_i$; Value = $\infty$.\\
-\> Pre-aggregated MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
-\> \>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
-$U$ = new MapReduce object to store updates to distances. \\
+Pre-aggregated MapReduce object $V$ = graph vertices $V_i$ with distances $d_i$ from $V_s$:  \\
+\> Key = $V_i$; Value = $\infty$.\\
+Pre-aggregated MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
+\>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
+MapReduce object $U$ to store updates to distances = empty \\
 Map $U$:  Add $V_s$ with distance $d_s = 0$ to $U$. \\
@@
 \> Output:  Key = $V_s$; Value = 0. \\
+done = 0 \\
 while not $done$: \\
-\> $done$ = 1; \\
-\> Aggregate $U$. \\
-\> Add $U$ to $V$. Empty $U$.\\
-\> Compress $V$:  Pick best candidate distance for each vertex. \\
-\> \> Input:  Key = $V_i$; Multivalue = candidate distances $d_i$. \\
-\> \> Compute:  Find smallest candidate distance $d_{min}$ for $V_i$.  \\
-\> \> \> If $d_{min} \ne d_i$, $done$ = 0; \\
-\> \> Output to $V$:  Key = $V_i$; Value = $d_{min}$. \\
-\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$. \\
-\> MPI\_Allreduce on $done$ with MPI\_MIN. \\
-\> if not $done$ \\
-\> \> Add: $U$ to $E$.  Empty $U$. \\
+\> $done$ = 1 \\
+\> Aggregate $U$ \\
+\> Add $U$ to $V$, empty $U$ \\
+\> Compress $V$:  Pick best candidate distance for each vertex \\
+\> \> Input:  Key = $V_i$; Multivalue = candidate distances $d_i$ \\
+\> \> Find smallest candidate distance $d_{min}$ for $V_i$ \\
+\> \> \> If $d_{min} \ne d_i$, $done$ = 0 \\
+\> \> Output to $V$:  Key = $V_i$; Value = $d_{min}$ \\
+\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$ \\
+\> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
+\> if not $done$: \\
+\> \> Add $U$ to $E$.  Empty $U$. \\
 \> \> Compress $E$:  Generate candidate distances for neighbors of changed vertices. \\
@@
 \> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
-\> \> \> \> Key = $V_j$; Value = $d_i+w_{ij}$ for each edge ($V_i, V_j$). \\
-Output: MapReduce object $V$ with Key = $V_i$, Value = $d_i$.
+\> \> \> \> Key = $V_j$, Value = $d_i+w_{ij}$ for each edge ($V_i, V_j$) \\
 
@@
  \caption{\it Enhanced MapReduce algorithm for single-source shortest
- path (SSSP) calculation.  Pre-aggregating the vertices and edges, and
- later aggregating only updates to vertex distances, reduces
- communication and execution time.}
+ path (SSSP) calculation, with an additional MapReduce object $U$.
+ Pre-aggregating the vertices $V$ and edges $E$, and later aggregating
+ only updates $U$ to vertex distances, reduces communication and
+ execution time.}
 

### 2010-11-03 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 program is linked with MPI, a standard message passing library
-available on all distributed memory machines.  For serial operation, a
-dummy MPI library (provided) can be substituted.  As a library, it can
-be embedded in other codes~\cite{Titan} to enable them to perform
-MapReduce operations.
+available on all distributed-memory machines and many shared-memory
+parallel machines.  For serial operation, a dummy MPI library
+(provided) can be substituted.  As a library, MR-MPI can be embedded
+in other codes~\cite{Titan} to enable them to perform MapReduce
+operations.
 
 \item {\it In-core or out-of-core operation:} Each MapReduce object
-that a processor defines allocates ``pages'' of memory, where the page
+that a processor creates allocates ``pages'' of memory, where the page
 size is determined by the user.  Typical MapReduce operations can be
@@
 system) as needed and subsequently read from them.  This out-of-core
-operation allows processing of data sets larger than the aggregate
-memory of all the processors, i.e. up to the available aggregate disk
-space.
+mode of operation allows processing of data sets larger than the
+aggregate memory of all the processors, i.e. up to the aggregate
+available disk space.
 
 \item {\it Flexible programmability:} An advantage of writing a
-MapReduce program on top of MPI, is that the user program can invoke
-MPI calls directly, if desired.  For example, one-line calls to
+MapReduce program on top of MPI, is that the user program can also
+invoke MPI calls directly, if desired.  For example, one-line calls to
 MPI\_Allreduce are often useful in determining the status of an
@@
 methods in one or more MapReduce objects.  The C interface allows the
-library to be called from C or other high-level languages such as
-Fortran.  The C interface also allows the library to be wrapped easily
-by Python via the Python ``ctypes'' module.  The library can then be
-called from a Python script, allowing the user to write {\it map()} and
-{\it reduce()} callback functions in Python.  If a machine supports running
-Python in parallel, a parallel MapReduce can also be run in this mode.
+same operations to be called from C or other high-level languages such
+as Fortran.  The C interface also allows the library to be easily
+wrapped by Python via the Python ``ctypes'' module.  The library can
+then be called from a Python script, allowing the user to write {\it
+map()} and {\it reduce()} callback functions in Python.  If a machine
+supports running Python in parallel, a parallel MapReduce can also be
+run in this mode.
 
-\item {\it No fault tolerance:} Current MPI implementations do not enable
-easy detection of a dead processor or retrieval of the data it was
-working on.  So like most MPI programs, a parallel program calling the
-MR-MPI library will hang or crash if a processor goes away.  Unlike
-Hadoop, and its HDFS file system which provides data redundancy,
-the MR-MPI library reads and writes simple, flat files.  It can
-use local per-processor disks, or a parallel file system, if
-available, but these typically provide no data redundancy.
+\item {\it No fault tolerance or data redundancy:} Current MPI
+implementations do not enable easy detection of a dead processor or
+retrieval of the data it was working on.  So like most MPI programs, a
+parallel program calling the MR-MPI library will hang or crash if a
+processor goes away.  Unlike Hadoop, and its HDFS file system which
+provides data redundancy, the MR-MPI library reads and writes simple,
+flat files.  It can use local per-processor disks, or a parallel file
+system, if available, but these typically provide no data redundancy.
 

### 2010-11-03 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 etc); they are simply treated as byte strings by the library.  A KV
-pair always has a key; a KV's value may be NULL.  A related data type
-is the key/multivalue (KMV) pair, where all values associated with the
-same key are collected and stored contiguously as a multivalue, which
-is just a longer byte string with an associated vector of lengths, one
-integer length per value.  In this section, we assume the datums
-operated on by each processor fit in local memory.
+pair always has a key; its value may be NULL.  KV pairs are stored
+within a MapReduce (MR) object.  A user program may create one or more
+MR objects to implement an algorithm.  Various MapReduce operations
+(map, reduce, etc) are invoked on an object and its KV pairs; KV pairs
+can also be passed between and combined across objects.
+
+A related data type in the MR-MPI library is the key/multivalue (KMV)
+pair, where all values associated with the same key are collected and
+stored contiguously as a multivalue, which is just a longer byte
+string with an associated vector of lengths, one integer length per
+value.  In this section, we assume the KV and KVM datums operated on
+by each processor fit in local memory.
 
@@
 library, to perform {\it map()}, {\it collate()}, and {\it reduce()}
-operations.  In a {\it map()}, zero or more key/value pairs are
-generated by each processor.  Often this is done using data read from
-files, but a {\it map()} may generate data itself or process existing
-KV pairs to create new ones.  The KV pairs produced are stored locally
-by each processor; a {\it map()} thus requires no inter-processor
-communication.  Users call the library with a count of tasks to
-perform and a pointer to a user function; the MR-MPI {\it map()}
-operation invokes the user function multiple times as a callback.
-Depending on which variant of {\it map()} is called, the user function
-may be passed a file name, a chunk of bytes from a large file, a task
-ID, or a KV pair.  Options for assigning map tasks to processors are
-specified by the user and include assigning consecutive chunks of
-tasks to processors, striding the tasks across processors, or using a
-master-slave model that is useful when tasks have widely varying
-workloads.
+operations on a MR object it creates.  In a {\it map()}, zero or more
+key/value pairs are generated by each processor.  Often this is done
+using data read from files, but a {\it map()} may generate data itself
+or process existing KV pairs to create new ones.  The KV pairs
+produced are stored locally by each processor; a {\it map()} thus
+requires no inter-processor communication.  Users call the library
+with a count of tasks to perform and a pointer to a user function; the
+MR-MPI {\it map()} operation invokes the user function multiple times
+as a callback.  Depending on which variant of {\it map()} is called,
+the user function may be passed a file name, a chunk of bytes from a
+large file, a task ID, or a KV pair.  Options for assigning map tasks
+to processors are specified by the user and include assigning
+consecutive chunks of tasks to processors, striding the tasks across
+processors, or using a master-slave model that is useful when tasks
+have widely varying workloads.
 
@@
 
-The interface to various low-level operations is provided so
-that a user's program can string them together in various ways to
-produce interesting MapReduce algorithms.  For example, output from a
-{\it reduce()} can serve as input to a subsequent {\it map()} or {\it
-collate()}.  KV pairs from multiple MapReduce objects can be combined
-to perform new sequences of {\it map()}, {\it collate()}, and
-{\it reduce()} operations.
+The interface to various low-level operations is provided so that a
+user's program can string them together in various ways to construct
+interesting MapReduce algorithms.  For example, output from a {\it
+reduce()} can serve as input to a subsequent {\it map()} or {\it
+collate()}.  Likewise, data can be partitioned across multiple
+MapReduce (MR) objects, each with their own set of KV pairs.  KV pairs
+from multiple MR objects can be combined to perform new sequences of
+{\it map()}, {\it collate()}, and {\it reduce()} operations.
 

### 2010-11-03 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 these files can reside on disks local to each processor, on the front
-end (typically a NSF-mounted file system for a parallel machine), or
-on a parallel file system/disk array.
+end (typically a NFS-mounted file system for a parallel machine), or
+on a parallel file system and its associated disk array.
 
@@
 data for the {\it map()} operation is an existing set of KV pairs,
-those datums are read, one page at a time, and a pointer to
-each KV pair is given to the user function.  Similarly, a {\it
-reduce()} reads one page of KMV pairs from disk, passes a pointer
-to each pair, one at a time, to the user function that typically generates
-new KV pairs.  The generated pairs fill up a new page, which is
-written to disk when full, just as with the {\it map()} operation.
-Thus for both a {\it map()} and {\it reduce()}, out-of-core disk files
-are read and written sequentially, one page at a time, which requires
-at most two pages of memory.
+those datums are read, one page at a time, and a pointer to each KV
+pair is given to the user function.  Similarly, a {\it reduce()} reads
+one page of KMV pairs from disk, and passes a pointer to each pair,
+one at a time, to the user function that typically generates new KV
+pairs.  The generated pairs fill up a new page, which is written to
+disk when full, just as with the {\it map()} operation.  Thus for both
+a {\it map()} and {\it reduce()}, out-of-core disk files are read and
+written sequentially, one page at a time, which requires at most two
+pages of memory.
 

### 2010-11-03 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 library was to enable graph algorithms to be written as MapReduce
-operations, allowing processing of Terabyte-scale data sets.  We
+operations, allowing processing of terabyte-scale data sets.  We
 outline MapReduce versions of several such algorithms: vertex ranking

### 2010-11-03 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@657 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-The single-source shortest path (SSSP) algorithm computes, for a
-directed graph with weighted edges, the shortest weighted distance
-from a chosen source vertex to all other vertices in the graph.  This
+For a directed graph with weighted edges, the single-source shortest
+path (SSSP) algorithm computes the shortest weighted distance from a
+chosen source vertex to all other vertices in the graph.  This
 calculation is straightforward when global graph information is
@@
 Bellman-Ford-style algorithm of~\cite{SSSPMapReduce, Bellman58,
-Ford62}, every edge of the graph is visited in each iteration,
-although labels for an edge's vertices are updated only if the
-breadth-first traversal from the source vertex has reached those
-vertices.  A MapReduce formulation of this algorithm, described in
-Figure~\ref{fig:sssp}, begins with a {\it map()} function loading
-edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.  In each
-iteration, the edges $(V_i, V_j)$ are collated with respect to $V_i$,
-and a {\it reduce()} operation updates each $V_j$'s distance if
-$V_i$'s distance is non-zero.
+Ford62}, every edge of the graph is visited in each iteration and
+vertex distancs are updated.  A MapReduce formulation of this
+algorithm, described in Figure~\ref{fig:sssp}, stores directed edges
+in $E$ and begins by labeling the distance to all vertices as
+inifinte, except a distance of 0.0 to the source vertex $V_s$.  At
+each iteration, a {\it map()} operation is performed using $E$ as
+input to add edges and their weights to each vertex in $V$.  After a
+{\it collate()} operation, each vertex $V_i$ knows its outgoing edges
+and their weights, as well as its own current distance.  It may also
+have one or more distances assigned to it from a previous iteration
+(i.e. from an inbound edge).  In the {\it reduce()} operation, each
+vertex resets its distance to the minimum of these new distances and
+its current distance.  If its distance changed, it emits new distance
+values $d_j$ for the vertices associated with its outgoing edges,
+where $d_j = d_i + w_{ij}$.  It also unsets a ``done'' flag, stored by
+each processor.  When the reduce is complete, an MPI\_Allreduce
+operation is performed to check the ``done'' flag status on all
+processors.  If any distance was changed, the iterations continue, and
+the distance updates propagate through the graph.  If no distances
+changed, the algorithm is done.
 
@@
 \> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
-\> \> Input: Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$ and candidate distances $d_i$ \\
+\> \> Input: Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$ and candidate distances for $d_i$ \\
 \> \> Find smallest new distance $d_{min}$ for $V_i$ \\
@@
 communication and execution time can be achieved by pre-aggregating
-certain MapReduce objects and storing data in more than one MapReduce
-object.  The enhanced algorithm is shown in Figure~\ref{fig:sssp2}.
-The MapReduce objects $V$ and $E$ storing vertices and edges are
-aggregated across processors only once at the beginning of the
-computation.  In each iteration of SSSP, only a small number of new
-candidate distances are generated.  The enhanced algorithm uses
-MapReduce object $U$ to store these updates.  Only the updates in $U$
-are communicated; each update is aggregated to the same processor that
-stores the corresponding vertex and its edges.  Thus, the total amount
-of communication needed is smaller than in the original algorithm,
-which aggregated all graph edges in each iteration.  Updated vertex
+the data stored in MapReduce objects and by splitting the data
+appropriately across more than one MapReduce object.  An enhanced SSSP
+algorithm is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects
+$V$ and $E$ which store vertices and edges are aggregated across
+processors only once at the beginning of the computation, which means
+the communication portion of the {\it collate()} operation is
+performed to assign each key to a processor.  The {\it collate()}
+operation of Figure~\ref{fig:sssp} can then be replaced with a {\it
+compress()} operation which requires no communication.
+
+In each iteration of SSSP, typically only a small number of new
+distances are generated.  The enhanced algorithm uses MapReduce object
+$U$ to store these updates.  Only the updates in $U$ are communicated;
+which aggregates them to the same processor that permanently stores
+the corresponding vertex and its edges.  Thus, the total amount of
+required communication is smaller than in the original algorithm,
+which communicated all graph edges at each iteration.  Updated vertex
 distances are stored in MapReduce object $V$, as before.
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-Pre-aggregated MapReduce object $V$ = graph vertices $V_i$ with distances $d_i$ from $V_s$:  \\
-\> Key = $V_i$; Value = $\infty$.\\
-Pre-aggregated MapReduce object $E$ containing graph edges ($V_i, V_j$) with weights $w_{ij}$: \\
-\>  Key = $V_i$; Value = $[V_j, w_{ij}]$ \\
-MapReduce object $U$ to store updates to distances = empty \\
-Map $U$:  Add $V_s$ with distance $d_s = 0$ to $U$. \\
-\> Input:  $V_s$ \\
-\> Output:  Key = $V_s$; Value = 0. \\
+Pre-aggregated MapReduce object $V$ = graph vertices $V_i$ with distances from sourece vertex $V_s$:  \\
+\> Key = $V_i$; Value = $\infty$, except Value for $V_s$ = 0 \\
+Pre-aggregated MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
+MapReduce object $U$ to store updates to distances, initially empty \\
 done = 0 \\
+
 while not $done$: \\
@@
 \> Add $U$ to $V$, empty $U$ \\
-\> Compress $V$:  Pick best candidate distance for each vertex \\
-\> \> Input:  Key = $V_i$; Multivalue = candidate distances $d_i$ \\
-\> \> Find smallest candidate distance $d_{min}$ for $V_i$ \\
-\> \> \> If $d_{min} \ne d_i$, $done$ = 0 \\
+\> Compress $V$:  Compute best distance for each vertex \\
+\> \> Input:  Key = $V_i$; Multivalue = candidate distances for $d_i$ \\
+\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
+\> \> If $d_{min} \ne d_i$: \\
+\> \> \> $done$ = 0 \\
+\> \> \> Output to $U$: Key = $V_i$; Value = $d_{min}$ \\
 \> \> Output to $V$: Key = $V_i$; Value = $d_{min}$ \\
-\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $V_i$; Value = $d_{min}$ \\
 \> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
 \> if not $done$: \\
-\> \> Add $U$ to $E$.  Empty $U$. \\
+\> \> Add $U$ to $E$, empty $U$ \\
 \> \> Compress $E$: Generate candidate distances for neighbors of changed vertices. \\
-\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ for edges ($V_i, V_j$) \\
-\> \> \> \> and distance $d_i$ if updated in previous step. \\
-\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
-\> \> \> \> Key = $V_j$, Value = $d_i+w_{ij}$ for each edge ($V_i, V_j$) \\
+\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ and candidate distance $d_i$ if updated in previous step \\
+\> \> \> If updated distance $d_i$ exists: \\
+\> \> \> \> Output to $U$: Key = $V_j$, Value = $d_i+w_{ij}$ for each edge vertex $V_j$) \\
 
@@
  path (SSSP) calculation, with an additional MapReduce object $U$.
- Pre-aggregating the vertices $V$ and edges $E$, and later aggregating
- only updates $U$ to vertex distances, reduces communication and
- execution time.}
+ Pre-aggregating the vertices $V$ and edges $E$, and storing only
+ changed distances in $U$, reduces communication and execution time.}
 

### 2010-11-03 paper/lessons.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@657 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 pairs of processors.  But the benefit is often good load-balance, even
-for hard-to-balance irregular data sets.  By contrast, more
+for hard-to-balance irregular data sets, such as those derived from
+sparse graphs with power-law degree distributions.  By contrast, more
 traditional distributed-memory parallel algorithms, e.g. for matrix
@@
 typically work hard to localize data and minimize communication.  To
-do this they often require a lot of application-specific logic and
-parallel communication coding to create and maintain a data
-decomposition, generate ghost versions of nearby spatially-decomposed
-data, etc.
+do this they often require application-specific logic and parallel
+communication coding to create and maintain a data decomposition,
+generate ghost copies of nearby spatially-decomposed data, etc.
 
@@
 communicate data via MPI calls or to move data between out-of-core
-disk files and processor memory is often a pleasant surprise.
-Moreover, if the MapReduce algorithm is initially coded so that it
-runs correctly on one processor, it often works out-of-the-box on
-hundreds or thousands of processors, without the need for additional
-debugging.
+disk files and processor memory is often a pleasant surprise.  By
+contrast, extending a library like Trilinos (or other in-core MPI
+applications) to enable it to perform its suite of linear algebra
+operations in out-of-core mode on large data sets would be a
+considerable effort.  Moreover, if a MapReduce algorithm is initially
+coded so that it runs correctly on one processor, it often works
+out-of-the-box on hundreds or thousands of processors, without the
+need for additional debugging.
 
@@
 a traditional MPI-based parallel machine is a somewhat different
-conceptual model than that of performing MapReduce operations on a
-cloud computer using (for example) Hadoop.  In the former case, one
-can potentially control which processor owns which data at various
-stages of an algorithm.  This is somewhat hidden from the user in a
-typical cloud-computing model, where data simply exists somewhere in
-the cloud and Hadoop ensures data moves where it is needed and is
-operated on by some processor.  The cloud model is a nice data-centric
-abstraction which allows for fault tolerance both to data loss (via
-redundant storage) and to processor failure (via reassignment of
-work), neither of which is typically possible on current MPI-based
-parallel machines.
+conceptual model than performing MapReduce operations on a cloud
+computer using (for example) Hadoop.  In the former case, one can
+potentially control which processor owns which data at various stages
+of an algorithm.  This is somewhat hidden from the user in a typical
+cloud-computing model, where data simply exists somewhere in the cloud
+and Hadoop ensures data moves where it is needed and is operated on by
+some processor.  The cloud model is a nice data-centric abstraction
+which allows for fault tolerance both to data loss (via redundant
+storage) and to processor failure (via reassignment of work), neither
+of which is typically possible on current MPI-based parallel machines.
 

### 2010-11-03 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@659 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 post-processing analysis of simulation data~\cite{Tu08}, graph
-algorithmics~\cite{Cohen09}, and linear algebra operations~\cite{Ekanayake09}.
-The paper by Tu et al.~\cite{Tu08} was particularly insightful to us,
-because it described how MapReduce could be implemented on top of the
-ubiquitous distributed-memory message-passing interface (MPI), and how
-the intermediate data-shuffle operation is conceptually identical to
-the familiar MPI\_Alltoall operation.  Their implementation of
-MapReduce was within a Python wrapper to simplify the writing of user
-programs.  The paper motivated us to develop our own C++ library built
-on top of MPI for use in graph analytics, which we initially released
-as open-source software in mid-2009~\cite{MRMPI}.  We have since
-worked to optimize several of the library's underlying algorithms and
-to enable its operation in out-of-core mode on larger data sets.
-These algorithmic improvements are described in this paper and are
-part of the current downloadable version~\cite{MRMPI}.
+algorithmics~\cite{Cohen09}, and linear algebra
+operations~\cite{Ekanayake09}.  The paper by Tu et al.~\cite{Tu08} was
+particularly insightful to us, because it described how MapReduce
+could be implemented on top of the ubiquitous distributed-memory
+message-passing interface (MPI), and how the intermediate data-shuffle
+operation is conceptually identical to the familiar MPI\_Alltoall
+operation.  Their implementation of MapReduce was within a Python
+wrapper to simplify the writing of user programs.  The paper motivated
+us to develop our own C++ library built on top of MPI for use in graph
+analytics, which we initially released as open-source software in
+mid-2009.  We have since worked to optimize several of the library's
+underlying algorithms and to enable its operation in out-of-core mode
+for larger data sets.  These algorithmic improvements are described in
+this paper and are part of the current downloadable version; see
+Section ~\ref{sec:thanks} for details.
 
@@
 that a processor creates allocates ``pages'' of memory, where the page
-size is determined by the user.  Typical MapReduce operations can be
-performed using a few such pages (per processor).  If the data set
-fits in a single page (per processor), the library performs its
-operations in-core.  If the data set exceeds the page size, processors
-each write to temporary disk files (on local disks or a parallel file
-system) as needed and subsequently read from them.  This out-of-core
-mode of operation allows processing of data sets larger than the
-aggregate memory of all the processors, i.e. up to the aggregate
-available disk space.
+size is determined by the user.  MapReduce operations can be performed
+using a few such pages (per processor).  If the data set fits in a
+single page (per processor), the library performs its operations
+in-core.  If the data set exceeds the page size, processors each write
+to temporary disk files (on local disks or a parallel file system) as
+needed and subsequently read from them.  This out-of-core mode of
+operation allows processing of data sets larger than the aggregate
+memory of all the processors, i.e. up to the aggregate available disk
+space.
 
@@
 parallel program calling the MR-MPI library will hang or crash if a
-processor goes away.  Unlike Hadoop, and its HDFS file system which
+processor goes away.  Unlike Hadoop, and its file system (HDFS) which
 provides data redundancy, the MR-MPI library reads and writes simple,

### 2010-11-03 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@659 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 Note that the first portion of the {\it collate()} operation involves
-all-to-all communication (each processor sends and receives data from
-every other processor) using a distributed hash table.  The
-communication can either be done via a MPI\_Alltoall() library call,
-or by a custom routine that aggregates messages and invokes
-point-to-point MPI\_Send() and MPI\_IRecv() calls.
+all-to-all communication since each processor exchanges KV pairs with
+every other processor.  The communication can either be done via a
+MPI\_Alltoall() library call, or by a custom routine that aggregates
+datums into large messages and invokes point-to-point MPI\_Send() and
+MPI\_IRecv() calls.
 
@@
 functions: {\it aggregate()}, which performs the all-to-all
-communication, and {\it convert()}, which turns a list of KV pairs
-into KMV pairs.  Both functions can be called directly.  The {\it
-compress()} function allows on-processor operations to be performed;
-it is equivalent to a {\it convert()} with on-processor KVs as input,
-followed by a {\it reduce()}.  The {\it clone()} function turns a list
-of KV pairs into KMV pairs, with one value per key.  The {\it
-collapse()} function turns $N$ KV pairs into one KMV pair, with the
-keys and values of the KV pairs becoming $2N$ values of a single
+communication, and {\it convert()}, which reorganizes a set of KV
+pairs into KMV pairs.  Both functions can be called directly.  The
+{\it compress()} function allows on-processor operations to be
+performed; it is equivalent to a {\it convert()} with on-processor KVs
+as input, followed by a {\it reduce()}.  The {\it clone()} function
+turns a list of KV pairs into KMV pairs, with one value per key.  The
+{\it collapse()} function turns $N$ KV pairs into one KMV pair, with
+the keys and values of the KV pairs becoming $2N$ values of a single
 multivalue assigned to a new key.  The {\it gather()} function
@@
 user's program can string them together in various ways to construct
-interesting MapReduce algorithms.  For example, output from a {\it
+usefule MapReduce algorithms.  For example, output from a {\it
 reduce()} can serve as input to a subsequent {\it map()} or {\it
 collate()}.  Likewise, data can be partitioned across multiple
-MapReduce (MR) objects, each with their own set of KV pairs.  KV pairs
-from multiple MR objects can be combined to perform new sequences of
-{\it map()}, {\it collate()}, and {\it reduce()} operations.
+MapReduce (MR) objects, each with their own set of KV pairs.  For
+example, as will be illustrated in Section ~\ref{sec:graph}, one MR
+object can store the edges of a graph, and another its vertices.  KV
+pairs from multiple MR objects can be combined to perform new
+sequences of {\it map()}, {\it collate()}, and {\it reduce()}
+operations.
 
 The above discussion assumed that the KV or KMV pairs stored by a
-processor fit in its physical memory.  In this case, MR-MPI performs
-only ``in-core'' processing and no disk files are written or read by
-any of the processors, aside from initial input data if it exists or
-final output data if it is generated.  Note that the aggregate
-physical memory of large parallel machines can be multiple terabytes,
-which allows for large data sets to be processed in-core, assuming the
-KV and KMV pairs remain evenly distributed across processors
-throughout the sequence of MapReduce operations.  The use of hashing
-to assign keys to processors typically provides for such
-load-balancing.  In the next section, we discuss what happens when
-data sets do not fit in available memory.
+processor fit in its physical memory.  In this case, the MR-MPI
+library performs only ``in-core'' processing and no disk files are
+written or read by any of the processors, aside from initial input
+data (if it exists) or final output data (if it is generated).  Note
+that the aggregate physical memory of large parallel machines can be
+multiple terabytes, which allows for large data sets to be processed
+in-core, assuming the KV and KMV pairs remain evenly distributed
+across processors throughout the sequence of MapReduce operations.
+The use of hashing to assign keys to processors typically provides for
+such load-balancing.  In the next section, we discuss what happens
+when data sets do not fit in available memory.

### 2010-11-03 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@659 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 operations are relatively simple.  As a {\it map()} generates KV pairs
-via the user function, a page of memory fills up, one KV pair at a
-time.  When the page is full, it is written to disk.  If the source of
-data for the {\it map()} operation is an existing set of KV pairs,
-those datums are read, one page at a time, and a pointer to each KV
-pair is given to the user function.  Similarly, a {\it reduce()} reads
-one page of KMV pairs from disk, and passes a pointer to each pair,
-one at a time, to the user function that typically generates new KV
-pairs.  The generated pairs fill up a new page, which is written to
-disk when full, just as with the {\it map()} operation.  Thus for both
-a {\it map()} and {\it reduce()}, out-of-core disk files are read and
-written sequentially, one page at a time, which requires at most two
-pages of memory.
+via the user callback function, a page of memory fills up, one KV pair
+at a time.  When the page is full, it is written to disk.  If the
+source of data for the {\it map()} operation is an existing set of KV
+pairs, those datums are read, one page at a time, and a pointer to
+each KV pair is given to the user function.  Similarly, a {\it
+reduce()} reads one page of KMV pairs from disk, and passes a pointer
+to each pair, one at a time, to the user callback function that
+typically generates new KV pairs.  The generated pairs fill up a new
+page, which is written to disk when full, just as with the {\it map()}
+operation.  Thus for both a {\it map()} and {\it reduce()},
+out-of-core disk files are read and written sequentially, one page at
+a time, which requires at most two pages of memory.
 
@@
 Recall that the operation occurs in two stages.  First, keys are
-hashed to ``owning'' processors and the KV datums are communicated to
+hashed to ``owning'' processors and the KV pairs are communicated to
 new processors in an all-to-all fashion.  In out-of-core mode, this
-operation is performed on one page of KV data at a time.  Each
-processor reads in a page of KV pairs, the communication pattern
-(which processors receive which datums in that page) is determined,
-and all-to-all communication is performed.  Each processor allocates a
-two-page chunk of memory to receive incoming KV pairs.  On average
-each processor should receive one page of KV pairs; the two-page
-allocation allows for some load imbalance.  If the KV pairs are
-distributed unevenly so that this limit is exceeded, the all-to-all
-communication is performed in smaller, multiple passes until the full
-page contributed by every processor has been communicated.  Performing
-an MPI\_Alltoall(), or using the custom all-to-all routines provided
-by the MR-MPI library, requires allocation of auxiliary arrays that
-store processor indices, datum pointers, and datum lengths.  For the
-case of many tiny KV datums, the total number of memory pages required
-to perform the all-to-all communication, including the source and
-target KV pages is at most seven.
+operation is performed on one page of KV data (per processor) at a
+time.  Each processor reads in a page of KV pairs, the communication
+pattern is determined (which processors receive which datums in that
+page), and all-to-all communication is performed.  Each processor
+allocates a two-page chunk of memory to receive incoming KV pairs.  On
+average each processor should receive one page of KV pairs; the
+two-page allocation allows for some load imbalance.  If the KV pairs
+are distributed unevenly so that this limit is exceeded, the
+all-to-all communication is performed in smaller, multiple passes
+until the full page contributed by every processor has been
+communicated.  Performing an MPI\_Alltoall(), or using the custom
+all-to-all routines provided by the MR-MPI library, requires
+allocation of auxiliary arrays that store processor indices, datum
+pointers, and datum lengths.  For the case of many tiny KV pairs, the
+total number of memory pages required to perform the all-to-all
+communication, including the source and target KV pages is at most
+seven.
 
@@
 follows.  Two pages of KV datums are read from disk.  Each is sorted
-using the C-library quicksort() function, as discussed in the previous
-section.  The two pages are then scanned and merged into a new file
-which is written to disk as its associated memory page fills up.  This
-process requires five memory pages, which includes vectors of KV datum
-pointers and lengths needed by the in-memory quicksort() operation.
-Once all pairs of pages have been merged, the sort continues in a
-recursive fashion, merging pairs of files into a new third file,
-without the need to perform additional in-memory sorts.  Thus the
-overall memory requirement is five pages.  The number of read/write
-passes through the KV data set for the merge sort is $O(\log_2{M})$,
-where $M$ is the number of pages of KV pairs.  The number of passes
-could be reduced at the cost of allocating more in-memory pages.
+in local memory using the C-library quicksort() function, as discussed
+in the previous section.  The two pages are then scanned and merged
+into a new file which is written to disk as its associated memory page
+fills up.  This process requires five memory pages, which includes
+vectors of KV datum pointers and lengths needed by the in-memory
+quicksort() operation.  Once all pairs of pages have been merged, the
+sort continues in a recursive fashion, merging pairs of files into a
+new third file, without the need to perform additional in-memory
+sorts.  Thus the overall memory requirement is five pages.  The number
+of read/write passes through the KV data set for the merge sort is
+$O(\log_2{M})$, where $M$ is the number of pages of KV pairs.  The
+number of passes could be reduced at the cost of allocating more
+in-memory pages.

### 2010-11-03 paper/Response.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@660 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \usepackage{url}
+
+\usepackage{color}
+\newcommand{\revise}[1]{\textcolor{red}{#1}}
 
@@
 
-October 29, 2010\\
+October 29, 2010
 
-Response to reviewers:  \\
+Response to reviewers:
 
 We thank the reviewers for their careful reading of our paper.  Their
-suggestions definitely strengthen the paper and sparked new ideas for
-improving our algorithms.  Below we list the Reviewers' comments
-verbatim.  Our responses are shown in italics. \\
+suggestions helped us improve the clarity of what we discuss, and even
+sparked some new ideas for improving our algorithms.  Below we list
+the Reviewers' comments verbatim.  Our responses are shown in red.
 
@@
 
-{\it Figure 1 is actually describing an on-processor rearrangement of
-data that each processor performs on its local data it owns.  Thus
-there is no inter-processor communication involved.}
+\revise{Figure 1 is actually describing an on-processor rearrangement
+of data that each processor performs on its local data.  Thus there is
+no inter-processor communication involved.}
 
@@
 
+\revise{We added some of these comparison points to the final
+``Lessons Learned'' section.  In particular we highlight the
+relatively small line count of the MapReduce algorithms and the
+natural out-of-core ability it provides, as compared for example to
+Trilinos.  The comment with respect to MR inefficiencies is addressed
+below.}
+
 The paper does not have a concise "Conclusion", only a "Lessons Learned"
 section.
-
-{\it We have made some of these comparison points more
-clearly in the final ``Lessons Learned'' section.}
-
-%comment on MR vs Trilinos, even if libraries already exist
-%  cannot use linear algebra lib for all graph algorithms described
-%  here - PR is an exception since can be cast as matvec
-%added line counts of algorithms
 
@@
 
-{\it We instrumented our MR-MPI and Trilinos implementations of
+\revise{We instrumented our MR-MPI and Trilinos implementations of
 PageRank to show where time is spent during the PageRank iterations.
@@
 and 4.  These experiments showed that considerable time is spent in
-reorganizing data from key-value pairs to key-multivalue pairs.  In
-sections 4.1 and 5.2, we describe alternative data layouts that could
-reduce the number of times the data must be reorganized to improve
-performance.}
+reorganizing data from key-value pairs to key-multivalue pairs.  We
+added text in sections 4.1 and 5.2 to describe alternative data
+layouts that could reduce the number of times the data needs to be
+reorganized to improve performance.}
 
@@
 
-{\it We significantly re-worked and standardized the notation used
-in Section 4 to describe all the graph algorithms.  In particular
-we noted which operations are performed with which MapReduce
-objects and their associated key/value data, so that the flow
-of data through each algorithm should be more clear.  A discussion
-of the MR-MPI library's ability to allow creation of multiple
-MapReduce objects was added to Section 2.}
+\revise{We agree that our notation in Section 4 was deficient.  I
+couldn't understand some of the deteails when I re-read it!  So we
+significantly re-worked and standardized the notation used in Section
+4 to hopefully describe all the graph algorithms more precisely.  In
+particular we noted which operations are performed with which
+MapReduce objects and their associated key/value data, so that the
+flow of data through each algorithm should be more clear.  A
+discussion of the MR-MPI library's ability to allow creation of
+multiple MapReduce objects was added to Section 2.}
 
@@
 
-{\it Reviewer \#2 made this same suggestion; please see the response above.}
+\revise{Reviewer \#2 made this same suggestion; please see the
+response above.}
 
@@
 
-{\it Fixed this; good catch.}
+\revise{Fixed this; good catch.}
 

### 2010-11-04 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@662 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
-unique non-zero entries in a sparse $N \times N$ matrix $A$, where
-each entry $A_{ij}$ represents an edge between graph vertices
-$(V_i,V_j)$.
+unique non-zero entries in a sparse $N \times N$ matrix $G$, where
+each entry $G_{ij}$ represents an edge between graph vertices
+$(V_i,V_j)$, with $V_i$ an integer from $1$ to $N$.
 
@@
 
+Input: \\
+\> MapReduce object $G$ = edges of graph, initially empty \\
+\\
 $M_{\rm remain} = M$ \\
 while $M_{\rm remain} > 0$: \\
-\> Map: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
-           \> \> output Key = $(i,j)$, Value = NULL \\
-\> Collate \\
-\> Reduce: \> Remove duplicate edges \\
-           \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
-           \> \> output Key = $(i,j)$, Value = NULL \\
-\> $M_{\rm remain} = M - N_{kv}$
+\> Map $G$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
+\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Collate $G$ \\
+\> Reduce $G$: \> Remove duplicate edges \\
+\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $M_{\rm remain} = M - N_{kv}$ \\
+\\
+Output: \\
+\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
 
@@
  \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
- processors, using a single MapReduce object.}
+ processors, using a single MapReduce object $G$.}
 
@@
 In the {\it map()} operation, each of $P$ processors generates a $1/P$
-fraction of the desired edges.  A single random edge ($i,j$) is computed
-recursively as follows.  Pick a random quadrant of the $A$ matrix with
-relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
-quadrant as a sub-matrix and select a random quadrant within it, in
-the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
-the end of the recursion, the final ``quadrant'' is non-zero matrix
-element $A_{ij}$.
+fraction of the desired edges.  A single random edge $(V_i,V_j)$ is
+computed recursively as follows.  Pick a random quadrant of the $G$
+matrix with relative probabilities $a$, $b$, $c$, and $d$.  Treat the
+chosen quadrant as a sub-matrix and select a random quadrant within
+it, in the same manner.  Repeat this process $n$ times where $N =
+2^n$.  At the end of the recursion, the final ``quadrant'' is a single
+non-zero matrix element $G_{ij}$.
 
-Though A is very sparse, the {\it map()} typically generates some
+Though $G$ is very sparse, the {\it map()} typically generates some
 small number of duplicate edges.  The {\it collate()} and {\it
@@
 map-collate-reduce sequence is repeated until the number of resulting
-key/value pairs $N_{kv} = M$.  For reasonably sparse graphs this
+key/value (KV) pairs $N_{kv} = M$.  For reasonably sparse graphs this
 typically takes only a few iterations.
@@
 power-of-two, to generate graphs with weighted edges (assign a numeric
-value to the $A_{ij}$ edge), graphs without self edges (require $i \ne
+value to the $G_{ij}$ edge), graphs without self edges (require $i \ne
 j$), or graphs with undirected edges (require $i < j$).  Or the
@@
 matrices to create upper-triangular R-MAT matrices representing
-undirected graphs.  For each edge $(i,j)$ with $i>j$, we add edge
-$(j,i)$, essentially symmetrizing the matrix; we then remove duplicate
-edges and self edges, leaving only edges with $i<j$.
+undirected graphs.  For each edge $(V_i,V_j)$ with $i > j$, we add
+edge $(V_j,V_i)$, essentially symmetrizing the matrix; we then remove
+duplicate edges and self edges, leaving only edges with $i < j$.
 
-By using more than one MapReduce object, we can improve the
+By using an additional MapReduce object, we can improve the
 performance of the R-MAT generation algorithm, particularly when
@@
 (Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
-object $E_{new}$ and {\it aggregate()} the edges to processors.  We
-then add those edges to the MapReduce object $E_{old}$, containing all
-previously generated unique edges; the edges of $E_{old}$ are already
-aggregated to processors using the same mapping as $E_{new}$.  We can
-then use the {\it compress()} function to reduce duplicates locally on
-each processor.  The cost of the algorithm is reduced due to the
-smaller number of KV pairs communicated in the {\it aggregate()},
+object $G_{new}$ and {\it aggregate()} the edges to processors.  We
+then add those edges to the MapReduce object $G$, containing all
+previously generated unique edges; the edges of $G$ are already
+aggregated to processors using the same mapping as $G_{new}$.  We can
+then use the {\it compress()} operation to reduce duplicates locally
+on each processor.  The cost of the enhanced algorithm is less due to
+the smaller number of KV pairs communicated in the {\it aggregate()},
 compared to the {\it collate()} operation of the original algorithm,
-particularly in later iterations of the algorithm where few edges are
-created.
+particularly in later iterations of the algorithm where few new edges
+are generated.
 
@@
 
+Input: \\
+\> MapReduce object $G$ = edges of graph, initially empty \\
+\> MapReduce object $G_{new}$ = new edges of graph, initially empty \\
+\\
 $M_{\rm remain} = M$ \\
 while $M_{\rm remain} > 0$: \\
-\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
-           \> \> output Key = $(i,j)$, Value = NULL \\
-\> Aggregate $E_{new}$ \\
-\> Add $E_{new}$ to $E_{old}$ \\
-\> Compress $E_{old}$: Remove duplicate edges \\
-            \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
-            \> \> output Key = $(i,j)$, Value = NULL \\
-\> $M_{\rm remain} = M - N_{kv}$
+\> Map $G_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
+\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
+\> Aggregate $G_{new}$ \\
+\> Add $G_{new}$ to $G$, empty $G_{new}$ \\
+\> Compress $G$: Remove duplicate edges \\
+\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
+\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
+\> $M_{\rm remain} = M - N_{kv}$ \\
+\\
+Output: \\
+\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
 
@@
  \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
- on $P$ processors, using two MapReduce objects $E_{new}$ and
- $E_{old}$.}
+ on $P$ processors, using two MapReduce objects $G$ and $G_{new}$.}
 

### 2010-11-04 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@663 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 values is larger than the other three, a skewed distribution results.
-Variants of the above algorithm can be used when $N$ is not a
-power-of-two, to generate graphs with weighted edges (assign a numeric
-value to the $G_{ij}$ edge), graphs without self edges (require $i \ne
-j$), or graphs with undirected edges (require $i < j$).  Or the
-general R-MAT matrix can be further processed by MapReduce operations
-to meet these requirements.  For some algorithms below (e.g., triangle
-finding and maximal independent set generation), we post-process R-MAT
-matrices to create upper-triangular R-MAT matrices representing
-undirected graphs.  For each edge $(V_i,V_j)$ with $i > j$, we add
-edge $(V_j,V_i)$, essentially symmetrizing the matrix; we then remove
-duplicate edges and self edges, leaving only edges with $i < j$.
+Variants of the algorithm can be used when $N$ is not a power-of-two,
+to generate graphs with weighted edges (assign a numeric value to the
+edge), graphs without self edges (require $i \ne j$), or graphs with
+undirected edges (require $i < j$).  Or the general R-MAT matrix can
+be further processed by MapReduce operations to meet these
+requirements.  For some algorithms below (e.g., triangle finding and
+maximal independent set generation), we post-process R-MAT matrices to
+create upper-triangular R-MAT matrices representing undirected graphs.
+For each edge $(V_i,V_j)$ with $i > j$, we add edge $(V_j,V_i)$,
+essentially symmetrizing the matrix; we then remove duplicate edges
+and self edges, leaving only edges with $i < j$.
 
@@
 aggregated to processors using the same mapping as $G_{new}$.  We can
-then use the {\it compress()} operation to reduce duplicates locally
-on each processor.  The cost of the enhanced algorithm is less due to
-the smaller number of KV pairs communicated in the {\it aggregate()},
-compared to the {\it collate()} operation of the original algorithm,
-particularly in later iterations of the algorithm where few new edges
-are generated.
+use the {\it compress()} operation to reduce duplicates locally on
+each processor.  The enhanced algorithm runs faster due to the smaller
+number of KV pairs communicated in the {\it aggregate()}, compared to
+the {\it collate()} operation of the original algorithm, particularly
+in later iterations of the algorithm where few new edges are
+generated.
 

### 2010-11-08 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@664 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-In the {\it map()} operation, each of $P$ processors generates a $1/P$
-fraction of the desired edges.  A single random edge $(V_i,V_j)$ is
-computed recursively as follows.  Pick a random quadrant of the $G$
-matrix with relative probabilities $a$, $b$, $c$, and $d$.  Treat the
-chosen quadrant as a sub-matrix and select a random quadrant within
-it, in the same manner.  Repeat this process $n$ times where $N =
-2^n$.  At the end of the recursion, the final ``quadrant'' is a single
-non-zero matrix element $G_{ij}$.
+In the {\it map()}, each of $P$ processors generates a $1/P$ fraction
+of the desired edges.  A single random edge $(V_i,V_j)$ is computed
+recursively as follows.  Pick a random quadrant of the $G$ matrix with
+relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
+quadrant as a sub-matrix and select a random quadrant within it, in
+the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
+the end of the recursion, the final ``quadrant'' is a single non-zero
+matrix element $G_{ij}$.
 
@@
 small number of duplicate edges.  The {\it collate()} and {\it
-reduce()} operations remove the duplicates.  The entire
-map-collate-reduce sequence is repeated until the number of resulting
-key/value (KV) pairs $N_{kv} = M$.  For reasonably sparse graphs this
-typically takes only a few iterations.
+reduce()} remove the duplicates.  The entire map-collate-reduce
+sequence is repeated until the number of resulting key/value (KV)
+pairs $N_{kv} = M$.  For reasonably sparse graphs this typically takes
+only a few iterations.
 
@@
 aggregated to processors using the same mapping as $G_{new}$.  We can
-use the {\it compress()} operation to reduce duplicates locally on
-each processor.  The enhanced algorithm runs faster due to the smaller
-number of KV pairs communicated in the {\it aggregate()}, compared to
-the {\it collate()} operation of the original algorithm, particularly
-in later iterations of the algorithm where few new edges are
-generated.
+use a {\it compress()} to reduce duplicates locally on each processor.
+The enhanced algorithm runs faster due to the smaller number of KV
+pairs communicated in the {\it aggregate()}, compared to the {\it
+collate()} of the original algorithm, particularly in later iterations
+of the algorithm where few new edges are generated.
 
@@
 \> Aggregate $G_{new}$ \\
-\> Add $G_{new}$ to $G$, empty $G_{new}$ \\
-\> Compress $G$: Remove duplicate edges \\
+\> Add $G_{new}$ to $G$, $G_{new}$ = empty \\
+\> Compress $G$: \> Remove duplicate edges \\
 \> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
@@
 page $V_i$; the second represents the probability of moving to a
-random page.  For pages with no outlinks, the first term is
+random page.  For pages with no outlinks, the first term is changed to
 $\alpha/|V|$, indicating equal likelihood to move to any other page.
@@
 $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
-\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  To maintain the
-sparsity of $A$, terms for random jumps and zero out-degree vertices
-are not explicitly stored as part of the matrix.  Rather, they are
-computed separately as adjustments to the PageRank vector.  The kernel
-of the PageRank algorithm is thus a power-method iteration consisting
-of matrix-vector multiplications $A^T x=y$, where $x$ is the PageRank
+\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  To keep $A$
+sparse, terms for random jumps and zero out-degree vertices are not
+explicitly stored as part of the matrix.  Rather, they are computed
+separately as adjustments to the PageRank vector.  The kernel of the
+PageRank algorithm is thus a power-method iteration consisting of
+matrix-vector multiplications $A^T x=y$, where $x$ is the PageRank
 vector from the previous iteration.  A few additional dot product and
@@
 The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
-linear-algebra concepts implemented in a MapReduce framework with data
-stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$ stores
-the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are vectors.
-The nonzeros of matrix $A$ are initialized as described above; the
-PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
-contains the indices of all-zero rows of $A$, corresponding to
-vertices with no outlinks.  The PageRank algorithm is made more
-efficient by calling {\it aggregate()} once on the MapReduce objects
-representing the matrix $A$, the all-zero matrix rows $MT$, and the
-vector $x$ before the PageRank iterations begin.  Since these three
-MapReduce objects have the same key types, pre-aggregating the objects
-moves all KV pairs with a given key to a single processor.  Thus, many
-of the PageRank operations become local operations, replacing {\it
-collate()} operations (which are equivalent to an {\it aggregate()}
-followed by {\it convert()}) with simply {\it convert()} calls.  In
-Figure~\ref{fig:pr2}, steps where {\it convert()} operations were
-substituted for {\it collate()} operations are marked with an
-asterisk.  This strategy of pre-aggregating to improve data locality
-is useful in many MapReduce algorithms, but assumes that specific keys
-can always be mapped to the same processor.
+linear-algebra concepts implemented in a MapReduce framework, with
+data stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$
+stores the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are
+vectors of length $|V|$.  The nonzeros of matrix $A$ are initialized
+as described above; the PageRank vector $x$ is initialized uniformly.
+A MapReduce object $MT$ contains the indices of all-zero rows of $A$,
+corresponding to vertices with no outlinks.  The PageRank algorithm is
+made more efficient by invoking the {\it aggregate()} operation once
+on the MapReduce objects representing the matrix $A$, the all-zero
+matrix rows $MT$, and the vector $x$, before the PageRank iterations
+begin.  Since these 3 MapReduce objects are all keyed by vertices,
+pre-aggregating them moves all KV pairs with a given key to the same
+processor.  Many of the subsequent PageRank operations then become
+local operations, replacing {\it collate()} (which is equivalent to an
+{\it aggregate()} followed by {\it convert()}) with simply a {\it
+convert()}.  In Figure~\ref{fig:pr2}, steps where {\it convert()} is
+substituted for {\it collate()} are marked with an asterisk.  This
+strategy of pre-aggregating to improve data locality is useful in many
+MapReduce algorithms, but assumes that identical keys in different
+MapReduce objects can be mapped to the same processor.
+
+\begin{figure}[htb]
+ \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
+  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+
+Input: \\
+\> Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
+\> Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = NULL \\
+\> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
+\> MapReduce object $y$ = vector, initially empty \\
+\\
+while $({residual} > {tolerance})$ \\
+\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
+\> \> Add $x$ to $MT$ \\
+$*$\> \> Convert $MT$:  Key is vertex $i$ \\
+\> \> Reduce $MT$:   \\
+\> \> \> Input:  Key = $i$; MultiValue = $(x_i, NULL)$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
+\> \> \> $c_{local} = c_{local} + foo$ \\
+\> \> \> Output:  If ${nvalues} = 2$: Key = $i$, Value = NULL \\
+\> \> $c_{global}$ = MPI\_Allreduce($c_{local}$,MPI\_SUM) \\
+\> 2 Compute $y = A^T x$  \\
+\> \> Copy $x$ to $y$ \\
+\> \> Add $A$ to $y$ \\
+$*$\> \> Convert $y$:  Key is row index of $A$ \\
+\> \> Reduce $y$: \\
+\> \> \>  Input: Key = $i$, Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
+\> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
+\> \> Collate $y$:  Key is column index of $A$ \\
+\> \> Reduce $y$: \\
+\> \> \> Input: Key = $j$, Multivalue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
+\> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
+\> 3 Add $c_{global}$ to $y$; compute $n_{global}$ = max element in $y$ \\
+\> \> Map $y$:  \\
+\> \> \> Input: Key = $i$, Value = $y_i$ \\
+\> \> \> $y_i = y_i + c_{global}$; $n_{local} = {\rm max}(y_i,n_{local})$ \\
+\> \> \> Output:  Key = $i$, Value = $y_i$ \\
+\> \> $n_{global}$ = MPI\_Allreduce($n_{local}$,MPI\_SUM) \\
+\> 4 Scale $y$ by $n_{global}$ \\
+\> \> Map $y$: \\
+\> \> \> Input:  Key = $i$, Value = $y_i$ \\
+\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
+\> 5 Compute residual \\
+\> \> Add $y$ to $x$ \\
+$*$\> \> Convert $x$:  Key is vertex $i$ \\
+\> \> Reduce $x$:  \\
+\> \> \> Input: Key = $i$, Multivalue = $x_i, y_i$ \\
+\> \> \> $r_{local} = {\rm max}(|x_i - y_i|,r_{local})$ \\
+\> \> $residual$ = MPI\_Allreduce($r_{local}$,MPI\_MAX) \\
+\> 6 Copy $y$ to $x$ \\
+\\
+Output: \\
+\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank \\
+
+  \end{tabbing}
+ \end{minipage}}\end{center}
+
+ \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
+ MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
+ of the MapReduce objects allows several communication intensive {\it
+ collate()} operations to be replaced by local {\it convert()}
+ operations (marked with asterisks).}
+
+ \label{fig:pr2}
+\end{figure}
 
@@
 for random jumps from any page are also computed.  In Step 2, the
-matrix-vector product $A^T x$ is computed and the result storey in
+matrix-vector product $A^T x$ is computed and the result stored in
 $y$.  This is the most expensive part of the computation, requiring
@@
 This linear-algebra approach is attractive because it allows
-construction of the MapReduce PageRank algorithm from simple
-operations (matrix-vector multiplication, dot products) on commonly
-used data types (matrices, vectors).  One can build a variety of
-algorithms based on these primitives; indeed, linear-algebra toolkits
-such as Trilinos~\cite{Trilinos-Overview} and
-PETSc~\cite{petsc-efficient} exploit this idea.  In parallel, these
-toolkits share pre-computed data layouts and communication maps among
-distributed objects to enable efficient operations between the
-objects.  For MapReduce, however, separating the data into vectors and
-matrices results in additional overhead to gather KV pairs with the
-same key into a key-multivalue (KMV) pair.  This overhead can be
-reduced in PageRank, for example, by storing the $x$ and $y$ vectors
-in a single MapReduce object, with key $i$ and value $(x_i, y_i)$.
-Doing so eliminates the {\it copy()} operation in Step 2 of
+construction of the MapReduce PageRank algorithm from simple kernels
+(matrix-vector multiplication, dot products) on commonly used data
+types (matrices, vectors).  One can build a variety of algorithms
+based on these primitives; indeed, linear-algebra toolkits such as
+Trilinos~\cite{Trilinos-Overview} and PETSc~\cite{petsc-efficient}
+naturally exploit this idea.  In parallel, these toolkits share
+pre-computed data layouts and communication maps among distributed
+objects to enable efficient operations on the objects.  In a MapReduce
+context, however, separating the data into vectors and matrices
+results in additional overhead to gather KV pairs with the same key
+into a key-multivalue (KMV) pair.  This overhead could be reduced in
+the PageRank algorithm, for example, by storing the $x$ and $y$
+vectors in a single MapReduce object, with key $i$ and value $(x_i,
+y_i)$.  Doing this would eliminate the {\it copy()} in Step 2 of
 Figure~\ref{fig:pr2}, as well as the {\it add()} and {\it convert()}
-operations in Step 5.  A further performance improvement would store
-an entire row of $A$ along with $x$ and $y$ in a single MapReduce
-object, with key $i$ and value $(x_i, y_i, d_{out}, [j_1, A_{ij_1}],
-[j_2, A_{ij_2}], \dots, [j_{d_{out}}, A_ij_{d_{out}}])$ (that is,
-storing together, for each vertex, its old and new PageRank value
-along with all out-going edges incident to the vertex).  This data
-layout eliminates an additional {it convert()} operation, but at the
-cost of code reuse and scalability.  While this data layout may be
-efficient for PageRank, it is more complex and less likely to be
-useful for additional algorithms.  In addition, since each KV pair
-must fit within a single page in MR-MPI, storing entire matrix rows in
-a single KV pair limits the size of the the matrices on which the
-algorithm can be applied.  The more general approach of storing
-non-zeros separately enables arbitrarily large problems to be solved.
+in Step 5.  A further performance improvement would store an entire
+row of $A$ along with $x$ and $y$ in a single MapReduce object, with
+key $i$ and value $(x_i, y_i, d_{out}, [j_1, A_{ij_1}], [j_2,
+A_{ij_2}], \dots, [j_{d_{out}}, A_ij_{d_{out}}])$ (that is, storing
+together, for each vertex, its old and new PageRank value along with
+all out-going edges incident to the vertex).  This data layout would
+eliminate an additional {it convert()}, but at the cost of code
+reusability.  This data layout would be efficient for PageRank, but is
+more complex and less likely to be useful for additional algorithms.
+In addition, since each KV pair must fit within a single page of
+memory (in MR-MPI), storing entire matrix rows in a single KV pair
+limits the size of matrices to which the algorithm can be applied.
+The more general approach used in Figure~\ref{fig:pr2} of storing
+non-zeros separately, enables arbitrarily large problems to be solved.
 The detailed timing break-downs in
-section~\ref{subsec:results_pagerank} allow one to estimate the
-benefit of these implementations relative to code complexity and
-scalability issues.
-
-%\begin{figure}[htb]
-% \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-%  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
-
-%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
-%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
-%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
-%While (${residual} > {tolerance}$) \\
-%\> 1. Compute contribution for random jumps and zero-outdegree vertices \\
-%\> \> Add: $x$ to $MT$ \\
-%\> \> Collate $MT$:  Vertex is Key \\
-%\> \> Reduce $MT$:   \\
-%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
-%\> \> \> Compute:  Local contributions $c_{local}$ \\
-%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
-%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
-%\> 2. Compute $y = A^T x$ \\
-%\> \> Copy $x$ to $y$ \\
-%\> \> Add $A$ to $y$ \\
-%\> \> Collate $y$:  Key is row index of $A$ \\
-%\> \> Reduce $y$: \\
-%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
-%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
-%\> \> Collate $y$:  Key is column index of $A$ \\
-%\> \> Reduce $y$:  \\
-%\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
-%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
-%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
-%\> \> Map $y$:  \\
-%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
-%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$ \\
-%\> \> \> Output:  Key = $i$, Value = $y_i$ \\
-%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
-%\> 4. Scale $y$ by $n_{global}$ \\
-%\> \> Map $y$: \\
-%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
-%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
-%\> 5. Compute residual \\
-%\> \> Add $y$ to $x$ \\
-%\> \> Collate $x$:  Key is index $i$ \\
-%\> \> Reduce $x$: \\
-%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
-%\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
-%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
-%\> 6. $x = y$; \\
-%Output:  PageRank vector $x$
-  %\end{tabbing}
- %\end{minipage}}\end{center}
-%
- %\caption{\it MapReduce algorithm for PageRank vertex ranking, using
- %4 MapReduce objects $A$, $MT$, $x$, and $y$.}
-%
- %\label{fig:pr}
-%\end{figure}
-
-\begin{figure}[htb]
- \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
-
-$*$ Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
-$*$ Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
-$*$ Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
-
-While (${residual} > {tolerance}$) \\
-\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
-\> \> Add: $x$ to $MT$ \\
-$*$\> \> Convert $MT$:  Vertex is Key \\
-\> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
-\> \> \> Compute:  Local contributions $c_{local}$ \\
-\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
-\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
-\> 2 Compute $y = A^T x$  \\
-\> \> Copy $x$ to $y$ \\
-\> \> Add $A$ to $y$ \\
-$*$\> \> Convert $y$:  Key is row index of $A$ \\
-\> \> Reduce $y$: \\
-\> \> \>  Input: Key = $i$; Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
-\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
-\> \> Collate $y$:  Key is column index of $A$ \\
-\> \> Reduce $y$:  \\
-\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
-\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
-\> 3 Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
-\> \> Map $y$:  \\
-\> \> \> Input:  Key = $i$, Value = $y_i$ \\
-\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$ \\
-\> \> \> Output:  Key = $i$, Value = $y_i$ \\
-\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
-\> 4 Scale $y$ by $n_{global}$ \\
-\> \> Map $y$: \\
-\> \> \> Input:  Key = $i$, Value = $y_i$ \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
-\> 5 Compute residual \\
-\> \> Add $y$ to $x$ \\
-$*$\> \> Convert $x$:  Key is index $i$ \\
-\> \> Reduce $x$:   \\
-\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
-\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
-\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
-\> 6 $x = y$; \\
-Output: PageRank vector $x$
-  \end{tabbing}
- \end{minipage}}\end{center}
-
- \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
- MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
- of the MapReduce objects allows many communication intensive {\it
- collate()} operations to be replaced by local {\it convert()}
- operations (marked with asterisks).}
-
- \label{fig:pr2}
-\end{figure}
+section~\ref{subsec:results_pagerank} allow estimation of the benefit
+of these possible enhancements.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxxxxxxxxxx\=xxx\=xxx\=xxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxxx\=xxxxxx\=xxx\=xxx\=xxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
-
-1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
-1 Map $G$: \> Convert edges to vertices \\
-           \> \> input Key = $(V_i,V_j)$, Value = NULL \\
-           \> \> output Key = $V_i$, Value = $V_j$ \\
-           \> \> output Key = $V_j$, Value = $V_i$ \\
+Input: \\
+\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
+\> MapReduce object $G_0$ = edges of graph, initially empty \\
+\\
+1 Copy $G$ to $G_0$ \\
+1 Map $G$: \> \> Convert edges to vertices \\
+           \> \> \> Input: Key = $(V_i,V_j)$, Value = NULL \\
+           \> \> \> Output: Key = $V_i$, Value = $V_j$ \\
+           \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
 1 Collate $G$ \\
-1 Reduce $G$: \> Add first degree to one vertex in edge \\
-              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
-	      \> \> for each V in MultiValue: \\
-              \> \> \> if $V_i < V$: output Key = $(V_i,V)$, Value = $(D_i,0)$ \\
-              \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
+1 Reduce $G$: \> \> Add first degree to one vertex in edge \\
+              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> \> for each V in MultiValue: \\
+              \> \> \> \> if $V_i < V$: Output: Key = $(V_i,V)$, Value = $(D_i,0)$ \\
+              \> \> \> \> else: Output: Key = $(V,V_i)$, Value = $(0,D_i)$ \\
 2 Collate $G$ \\
-2 Reduce $G$: \> Add second degree to other vertex in edge \\
-              \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
-              \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
-3 Map $G$: \> Low degree vertex emits edges \\
-           \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
-           \> \> else if $D_j < D_i$: output Key = $V_j$, Value = $V_i$ \\
-           \> \> else: output Key = $V_i$, Value = $V_j$ \\
+2 Reduce $G$: \> \> Add second degree to other vertex in edge \\
+              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
+              \> \> \> Output: Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
+3 Map $G$: \> \> Low degree vertex emits edges \\
+           \> \> \> if $D_i < D_j$: Output: Key = $V_i$, Value = $V_j$ \\
+           \> \> \> else if $D_j < D_i$: Output: Key = $V_j$, Value = $V_i$ \\
+           \> \> \> else: Output: Key = $V_i$, Value = $V_j$ \\
 3 Collate $G$ \\
-3 Reduce $G$: \> Emit angles of each vertex \\
-              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
-	      \> \> for each $V_1$ in MultiValue: \\
-	      \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
-	      \> \> \> \> if $V_1 < V_2$: output Key = $(V_1,V_2)$, Value = $V_i$ \\
-	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
-4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs in $G$ \\
+3 Reduce $G$: \> \> Emit angles of each vertex \\
+              \> \> \> Input: Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
+	      \> \> \> for each $V_1$ in MultiValue: \\
+	      \> \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
+	      \> \> \> \> \> if $V_1 < V_2$: Output: Key = $(V_1,V_2)$, Value = $V_i$ \\
+	      \> \> \> \> \> else: Output: Key = $(V_2,V_1)$, Value = $V_i$ \\
+4 Add $G_0$ to $G$ \\
 4 Collate $G$ \\
-4 Reduce $G$: \> Emit triangles \\
-              \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
-              \> \> if NULL exists in MultiValue: \\
-	      \> \> \> for each non-NULL V in MultiValue: \\
-	      \> \> \> \> output Key = $(V_i,V_j,V)$, Value = NULL
+4 Reduce $G$: \> \> Emit triangles \\
+              \> \> \> Input: Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
+              \> \> \> if NULL exists in MultiValue: \\
+	      \> \> \> \> for each non-NULL V in MultiValue: \\
+	      \> \> \> \> \> Output: Key = $(V_i,V_j,V)$, Value = NULL \\
+\\
+Output: \\
+\> MapReduce object $G$ = triangles in graph: Key = $(V_i,V_j,V_k)$, Value = NULL \\
 
@@
 
-The initial step is to store a copy of the graph edges as key/value
-(KV) pairs in an auxiliary MapReduce object $G_0$, for use later in
-the algorithm.  The first {\it map()} operation on $G$ converts edge
-keys to vertex keys with edge values.  After a {\it collate()}, each
-vertex has a list of vertices it is connected to; the first {\it
-reduce()} can thus flag one vertex $V_i$ in each edge with a degree
-count $D_i$.  The second {\it collate()} and {\it reduce()} assign a
-degree count $D_j$ to the other vertex in each edge.  In the third
-{\it map()}, only the lower-degree vertex in each edge emits its edges
-KV pairs.  The task of the third {\it reduce()} is to emit ``angles''
-for each of these low-degree vertices.  An ``angle'' is a root vertex
-$V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a triangle
-without the third edge $(V_1,V_2)$.  The {\it reduce()} emits a list
-of all angles of vertex $V_i$, by a double loop over the edges of
-$V_i$.  Note that the aggregate volume of KV pairs emitted at this
-stage is minimized by having only the low-degree vertex in each edge
-generate angles.
+The initial step is to store a copy of the graph edges as KV pairs in
+an auxiliary MapReduce object $G_0$, for use later in the algorithm.
+The first {\it map()} on $G$ converts edge keys to vertex keys with
+edge values.  After a {\it collate()}, each vertex has the list of
+vertices it is connected to; the first {\it reduce()} can thus flag
+one vertex $V_i$ in each edge with a degree count $D_i$.  The second
+{\it collate()} and {\it reduce()} assign a degree count $D_j$ to the
+other vertex in each edge.  In the third {\it map()}, only the
+lower-degree vertex in each edge emits its edges as KV pairs.  The
+task of the third {\it reduce()} is to emit ``angles'' for each of
+these low-degree vertices.  An ``angle'' is a root vertex $V_i$, with
+two edges to vertices $V_1$ and $V_2$, i.e. a triangle without the
+third edge $(V_1,V_2)$.  The {\it reduce()} emits a list of all angles
+of vertex $V_i$, by a double loop over the edges of $V_i$.  Note that
+the aggregate number of KV pairs emitted at this stage is minimized by
+having only the low-degree vertex in each edge generate angles.
 
@@
 current working set of KV pairs in MapReduce object $G$.  The KV pairs
-in $G_0$ are edges that complete triangles for the angle KV pairs just
-generated.  After the fourth {\it collate()}, a pair of vertices
+in $G_0$ contain edges that complete triangles for the angle KV pairs
+just generated.  After the fourth {\it collate()}, a pair of vertices
 $(V_i,V_j)$ is the key, and the multivalue is the list of all root
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
-MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
-MapReduce object $W$ = workspace
-
+Input: \\
+\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
+\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
+\> MapReduce object $W$ = workspace
+\\
 Iterate: \\
-
-\> 1 Map $W$ using $E$ as input: \> Convert edges to vertices \\
-    \> \> \> input Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
-    \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
-    \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
+\> 1 Map $W$ using $E$ as input: Convert edges to vertices \\
+    \> \> Input: Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
+    \> \> Output: Key = $V_i$, Value = $E_{ij}$ \\
+    \> \> Output: Key = $V_j$, Value = $E_{ij}$ \\
 \> 1 Add $V$ to $W$ \\
-\> 1 Collate $W$: \> Vertex as key \\
-\> 1 Reduce $W$: \> Emit edges of each vertex with zone of vertex \\
-       \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
-       \> \> \> for each $E$ in MultiValue: \\
-      \> \> \> \> output Key = $E_{ij}$, Value = $Z_i$ \\
-
-\> 2 Collate $W$: \> Edge as key \\
-\> 2 Reduce $W$: \> Emit zone re-assignments \\
-       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
-       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
-       \> \> \> if $Z_i$ and $Z_j$ are different: \\
-      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
-\> 2 Exit: \> if $W$ is empty, i.e. no output by Reduce 2 \\
-\> 3 Map $V$: \> Invert vertex/zone pairs \\
-    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
-    \> \> \> if $Z_i$ is not partitioned: \\
-   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
-    \> \> \> else: \\
-   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
-\> 4 Map $V$ using $W$ as input: \> Add zone reassignments in $W$ to $V$ \\
-    \> \> \> if $Z_i$ is not partitioned: \\
-   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
-    \> \> \> else: \\
-   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
-   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
-\> 4 Collate $V$: \> Zone ID as key \\
-\> 4 Reduce $V$: \> Emit new zone assignment of each vertex \\
-       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
-       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
-       \> \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
-       \> \> \> for each $V$ in MultiValue: \\
-      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$
+\> 1 Collate $W$: Vertex as key \\
+\> 1 Reduce $W$: Emit edges of each vertex with zone of vertex \\
+    \> \> Input: Key = $V_i$, MultiValue = $E E E E ... Z$ \\
+    \> \> for each $E$ in MultiValue: \\
+    \> \> \> Output: Key = $E_{ij}$, Value = $Z_i$ \\
+\> 2 Collate $W$: Edge as key \\
+\> 2 Reduce $W$: Emit zone re-assignments \\
+       \> \> Input: Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
+       \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
+       \> \> if $Z_i$ and $Z_j$ are different: \\
+       \> \> \> Output: Key = $Z_{loser}$, Value = $Z_{winner}$ \\
+\> 2 Exit: if $W$ is empty (no output by Reduce 2) \\
+\> 3 Map $V$: Invert vertex/zone pairs \\
+    \> \> Input: Key = $V_i$, Value = $Z_i$ \\
+    \> \> if $Z_i$ is not partitioned: \\
+    \> \> \> Output: Key = $Z_i$, Value = $V_i$ \\
+    \> \> else: \\
+    \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
+\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to $V$ \\
+    \> \> if $Z_i$ is not partitioned: \\
+    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
+    \> \> else: \\
+    \> \> \> Output: Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
+    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
+\> 4 Collate $V$: Zone ID as key \\
+\> 4 Reduce $V$: Emit new zone assignment of each vertex \\
+       \> \> Input: Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
+       \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
+       \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
+       \> \> for each $V$ in MultiValue: \\
+       \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
+\\
+Output: \\
+\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
 
@@
  \caption{\it MapReduce algorithm for connected component labeling,
- using 3 MpaReduce objects $E$, $V$, and $W$.}
+ using 3 MapReduce objects $E$, $V$, and $W$.}
 
@@
 iterations complete, each zone has become a fully connected component
-and the MR object $V$ holds that zone assignment for every vertex.
-The algorithm thus finds all connected components in the graph
+and the MR object $V$ holds the zone assignment for every vertex.  The
+algorithm thus finds all connected components in the graph
 simultaneously.  The number of iterations required depends on the
@@
 
-The first {\it map()} operation uses MapReduce object $E$ as input to
-the empty MapReduce object $W$ to emit the vertices in each edge as
-keys, with the edge as a value.  The current zone assignment of each
-vertex in $V$ is added to the set of key/value (KV) pairs in $W$.  The
-first {\it collate()} operation collects all the edges of a vertex and
-its zone assignment together in one multi-value.  The first {\it
-reduce()} operation re-emits each edge, tagged by the zone assignment
-of one of its vertices.
+The first {\it map()} uses MapReduce object $E$ as input to the empty
+MapReduce object $W$ to emit the vertices in each edge as keys, with
+the edge as a value.  The current zone assignment of each vertex in
+$V$ is added to the set of KV pairs in $W$.  The first {\it collate()}
+collects all the edges of a vertex and its zone assignment together in
+one multivalue.  The first {\it reduce()} re-emits each edge, tagged
+by the zone assignment of one of its vertices.
 
-Since each edge was emitted twice, the second {\it collate()}
-operation on $W$ collects the zone assignments for its two vertices
-together.  If the two zone IDs are different, the second {\it
-reduce()} operation chooses a winner (the smaller of the two IDs), and
-emits the loser ID as a key, with the winning ID as a value.  If no
-zone ID changes are emitted, the algorithm is finished, and the
-iterations cease.
+Since each edge was emitted twice, the second {\it collate()} on $W$
+collects the zone assignments for its two vertices together.  If the
+two zone IDs are different, the second {\it reduce()} chooses a winner
+(the smaller of the two IDs), and emits the loser ID as a key, with
+the winning ID as a value.  If no zone ID changes are emitted, the
+algorithm is finished, and the iterations cease.
 
-The third {\it map()} operation inverts the vertex/zone KV pairs to
-become zone/vertex KV pairs.  The partitioning logic described in
+The third {\it map()} inverts the vertex/zone KV pairs to become
+zone/vertex KV pairs.  The partitioning logic described in
 Figure~\ref{fig:cc} for this step (and subsequent ones) is described
-below.  The fourth {\it map()} operation adds the changed zone
-assignments in $W$ to the set of KV pairs in $V$.  The fourth {\it
-collate()} can now collect all the vertices in each zone along with
-zero or more reassignments for that zone ID.  Since a zone could
-collide with multiple other zones on the same iteration due to shared
-edges, the new zone ID for a zone becomes the minimum ID of any of its
-neighboring zones.  If no zone reassignment values appear in the
-multi-value resulting from the fourth {\it collate()}, the zone ID is
-unchanged.  The final {\it reduce()} emits a KV pair for each vertex
-in the zone, with the vertex as a key and the new zone ID as a value.
-This is updated version of MapReduce object $V$, ready to be used as
-input for the next iteration.
+below.  The fourth {\it map()} adds the changed zone assignments in
+$W$ to the set of KV pairs in $V$.  The fourth {\it collate()} can now
+collect all the vertices in each zone along with any reassignments of
+that zone ID.  Since a zone could collide with multiple other zones on
+the same iteration due to shared edges, the new zone ID for a zone
+becomes the minimum ID of any of its neighboring zones.  If no zone
+reassignment values appear in the multivalue resulting from the fourth
+{\it collate()}, the zone ID is unchanged.  The final {\it reduce()}
+emits a KV pair for each vertex in the zone, with the vertex as a key
+and the new zone ID as a value.  This is an updated version of
+MapReduce object $V$, ready to be used as input for the next
+iteration.
 
 Note that if a graph has only a few components, the fourth {\it
-collate()} operation, which keys on the zone ID, may generate a few
-very large key/multi-value (KMV) pairs.  For example, if the entire
-graph is fully connected, in the last iteration, a single KMV pair
-assigned to one processor will result, containing all vertices in the
-graph.  This imbalance in memory and computational work can lead to
-poor parallel performance of the overall algorithm.  To counter this
-effect, the various operations in stages 3 and 4 include extra logic.
-The idea is to partition zones whose vertex count exceeds a
-user-defined threshhold into $P$ sub-zones, where $P$ is the number of
-processors.  The 64-bit integer that stores the zone ID also stores a
-bit flag indicating the zone has been partitioned and a set of bits
-that encode the processor ID.
+collate()}, which keys on the zone ID, may generate a few very large
+key/multivalue (KMV) pairs.  For example, if the entire graph is fully
+connected, in the last iteration, a single KMV pair assigned to one
+processor will result, containing all vertices in the graph.  This
+imbalance in memory and computational work can lead to poor parallel
+performance of the overall algorithm.  To counter this effect, various
+operations in stages 3 and 4 include extra logic.  The idea is to
+partition zones whose vertex count exceeds a user-defined threshhold
+into $P$ sub-zones, where $P$ is the number of processors.  The 64-bit
+integer that stores the zone ID also stores a bit flag indicating the
+zone has been partitioned and a set of bits that encode the processor
+ID.
 
-During the third {\it map()} operation, if the zone has been
-partitioned, the vertex is assigned to a random processor and the
-processor ID bits are added to the zone ID, as indicated by the
-$Z_i^+$ notation in Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been
-partitioned, the fourth {\it map()} operation emits the KV pair for
-the zone ID as ($Z_i^+$,$Z_{winner}$) instead of as
-($Z_i$,$Z_{winner}$).  In this case ($Z_i$,$Z_{winner}$), is emitted
-not once, but $P+1$ times, once for each processor, and once as if
-$Z_i$ had not been partitioned (for a reason discussed below).
+During the third {\it map()}, if the zone has been partitioned, the
+vertex is assigned to a random processor and the processor ID bits are
+added to the zone ID, as indicated by the $Z_i^+$ notation in
+Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
+fourth {\it map()} emits the KV pair for the zone ID as
+($Z_i^+$,$Z_{winner}$) instead of as ($Z_i$,$Z_{winner}$).  In this
+case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once
+for each processor, and once as if $Z_i$ had not been partitioned (for
+a reason discussed below).
 
 With this additional partitioning logic, the fourth {\it collate()}
-operation (which keys on zone IDs, some of which now include processor
-bits) collects only $1/P$ of the vertices in large zones onto each
+(which keys on zone IDs, some of which now include processor bits)
+collects only $1/P$ of the vertices in partitioned zones onto each
 processor.  Yet the multivalue on each processor still contains all
 the zone-reassignments relevant to the unpartitioned zone.  Thus, the
-fourth {\it reduce()} operation can change the zone ID (if necessary)
-in a consistent manner across all $P$ multi-values that contain the
-zone's vertices.  When the fourth {\it reduce()} operation emits new
-zone assignments for each vertex, the zone retains its partitioned
-status; the partition bit is also explicitly set if the vertex count
-exceeds the threshhold for the first time.
+fourth {\it reduce()} can change the zone ID (if necessary) in a
+consistent manner across all $P$ multivalues that contain the zone's
+vertices.  When the fourth {\it reduce()} emits new zone assignments
+for each vertex, the zone retains its partitioned status; the
+partition bit is also explicitly set if the vertex count exceeds the
+threshhold for the first time.
 
@@
 small zone may not set the partition bit of their zone IDs.  On
-subsequent iterations, the fourth {\it map()} operation emits $P+1$
-copies of new zone reassignments for both partitioned and unpartioned
-zone IDs, to ensure all vertices in the zone know the correct
-reassignment information.
+subsequent iterations, the fourth {\it map()} emits $P+1$ copies of
+new zone reassignments for both partitioned and unpartioned zone IDs,
+to ensure all vertices in the zone will know the correct reassignment
+information.
 
@@
  \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
-  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
+  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
 
-MapReduce object $V$ = maximal independent set vertices, initially empty \\
-MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
-Map $E$: assign random values to each vertex $V_i$ and $V_j$ in $E$ \\
-Clone $E$: convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
+Input: \\
+\> MapReduce object $V$ = maximal independent set vertices, initially empty \\
+\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
+Map $E$: Assign random values consistently to each vertex $V_i$ and $V_j$ in $E$ \\
+Clone $E$: Convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
 \\
-while $N_{edges}$ in $E$ > 0: \\
-\> 1 Reduce $E$: \> Determine WINNER/LOSER vertex of each edge \\
-       \> \> \> input Key = $E_{ij}$, Multivalue = NULL or LOSER \\
-       \> \> \> if multivalue = LOSER, emit nothing \\
-       \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
-       \> \> \> output Key = $V_w$, Value = $V_l$, WINNER \\
-       \> \> \> output Key = $V_l$, Value = $V_w$, LOSER \\
-\> 1 Collate $E$: \> Vertex as key \\
-\> 2 Reduce $E$: \> Find WINNER vertices \\
-       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
-      \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, WINNER \\
+while $N_{edges}$ in $E > 0$: \\
+\> 1 Reduce $E$: Determine WINNER/LOSER vertex of each edge \\
+       \> \> Input: Key = $E_{ij}$, Multivalue = NULL or LOSER \\
+       \> \> if multivalue = LOSER, emit nothing \\
+       \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
+       \> \> Output: Key = $V_w$, Value = ($V_l$, WINNER) \\
+       \> \> Output: Key = $V_l$, Value = ($V_w$, LOSER) \\
+\> 1 Collate $E$: Vertex as key \\
+\> 2 Reduce $E$: Find WINNER vertices \\
+       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
+       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, WINNER) \\
+       \> \> else: \\
+       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
+\> 2 Collate $E$: Vertex as key \\
+\> 3 Reduce $E$: Find LOSER vertices \\
+       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
+       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, LOSER) \\
+       \> \> else: \\
+       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
+\> 3 Collate $E$: Vertex as key \\
+\> 4 Reduce $E$: Emit WINNER vertices and LOSER edges \\
+       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
+       \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
+       \> \> \> Output: to $V$: Key = $V_i$, Value = NULL \\
+       \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> if $V_j$ is flagged with LOSER: \\
+       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
        \> \> \> else: \\
-      \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
-\> 2 Collate $E$: \> Vertex as key \\
-\> 3 Reduce $E$: \> Find LOSER vertices \\
-       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
-      \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, LOSER \\
-       \> \> \> else: \\
-      \> \> \> \> for each $V_j$ in Multivalue: \\
-     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
-\> 3 Collate $E$: \> Vertex as key \\
-\> 4 Reduce $E$: \> Emit WINNER vertices and LOSER edges \\
-       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
-      \> \> \> \> output to $V$: Key = $V_i$, Value = NULL \\
-       \> \> \> for each $V_j$ in Multivalue: \\
-      \> \> \> \> if $V_j$ is flagged with LOSER: \\
-     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
-      \> \> \> \> else: \\
-     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
-\> 4 Collate $E$: \> Edge as key \\
+       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
+\> 4 Collate $E$: Edge as key \\
+\\
+Output: \\
+\> MapReduce object $V$ = maximal independent set vertices: Key = $V_i$, Value = NULL \\
 
@@
 Before the iterations begin, a random value is assigned to each vertex
-(stored consistently in multiple edges in $E$), which is used to
-compare pairs of vertices.  The vertex ID itself could be used as the
-random value, but since the vertices eventually selected for the MIS
-depend on the randomization, this may introduce an unwanted bias.
-Instead a random number generator can be used to assign a consistent
-random value $R_i$ to each vertex in each edge (via a {\it map()}
-operation), which is then carried along with the vertex ID through
-each stage of the algorithm.  In the notation of
-Figure~\ref{fig:luby}, each $V_i$ is then really two quantities, a
-vertex ID (1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used
-to relabel the vertices in a random manner, assigning each a new,
-unique vertex ID from 1 to $N$.  In this case, $V_i$ is simply the new
-vertex ID, and the vertices in the final MIS could be remapped to
-recover their original IDs.  These operations can be performed with a
-few extra MapReduce steps, outside the iteration loop.
+(stored consistently for all edges in $E$ containing $V_i$), which is
+used to compare pairs of vertices.  The vertex ID itself could be used
+as the random value, but since the vertices eventually selected for
+the MIS depend on the randomization, this may introduce an unwanted
+bias.  Instead, a random number generator can be used to assign a
+consistent random value $R_i$ to each vertex in each edge (via a {\it
+map()}), which is then carried along with the vertex ID through each
+stage of the algorithm.  In the notation of Figure~\ref{fig:luby},
+each $V_i$ is then really two quantities, a vertex ID (1 to $N$) and
+$R_i$.  Alternatively, the $R_i$ can be used to relabel the vertices
+in a random manner, assigning each a new, unique vertex ID from 1 to
+$N$.  In this case, $V_i$ is simply the new vertex ID, and the
+vertices in the final MIS could be remapped to recover their original
+IDs.  These operations can be performed with a few extra MapReduce
+steps, outside the iteration loop.
 
-The pre-iteration {\it clone()} operation converts the initial list of
-edge key/value (KV) pairs stored in MapReduce object $E$ to
-key/multivalue (KMV) pairs, with the edge as the key, and NULL as the
-value.  Thus, the iterations can begin with a {\it reduce()}
-operation, without need for a {\it collate()} operation.  A second
-empty MapReduce object $V$ is also created, which will accumulate MIS
-vertices as the iterations proceed.
+The pre-iteration {\it clone()} converts the initial list of edge
+key/value (KV) pairs stored in MapReduce object $E$ to key/multivalue
+(KMV) pairs, with the edge as the key, and NULL as the value.  Thus,
+the iterations can begin with a {\it reduce()}, without need for a
+{\it collate()}.  A second empty MapReduce object $V$ is also created,
+which will accumulate MIS vertices as the iterations proceed.
 
@@
 
-The first {\it reduce()} operation flags the two vertices in each edge
-as a potential winner or loser.  This is done by comparing the two
-random values for the vertices (or the randomized vertex IDs as
-explained above).  The first {\it collate()} operation thus produces a
-multivalue for each vertex $V_i$ which contains all the vertices it
-shares an edge with, and associated winner/loser flags.
+The first {\it reduce()} flags the two vertices in each edge as a
+potential winner or loser.  This is done by comparing the two random
+values for the vertices (or the randomized vertex IDs as explained
+above).  The first {\it collate()} thus produces a multivalue for each
+vertex $V_i$ which contains all the vertices it shares an edge with,
+and associated winner/loser flags.
 
-In the second {\it reduce()} operation, if $V_i$ won the comparison
-with all its neighbor vertices, then it becomes a ``winner'' vertex.
-All vertices connected to the winner are emitted with $V_i$ tagged
-with a winner flag.  If $V_i$ is not an overall winner, its neighbor
-vertices are emitted without the winner flag on $V_i$..  The second
-{\it collate()} operation collects this new information for each
-vertex.
+In the second {\it reduce()}, if $V_i$ won the comparison with all its
+neighbor vertices, then it becomes a ``winner'' vertex.  All vertices
+connected to the winner are emitted with $V_i$ tagged with a winner
+flag.  If $V_i$ is not an overall winner, its neighbor vertices are
+emitted without the winner flag on $V_i$..  The second {\it collate()}
+collects this new information for each vertex.
 
-In the third {\it reduce()} operation, if any neighbor of $V_i$ was a
-winner in the previoue {\it reduce()} operation (indicated by a winner
-flag), $V_i$ becomes a losing vertex and emits all its neighbor
-vertices with $V_i$ tagged with a loser flag.  Otherwise it emits all
-its neighbor vertices without the loser flag on $V_i$.
+In the third {\it reduce()}, if any neighbor of $V_i$ was a winner in
+the previoue {\it reduce()} (indicated by a winner flag), $V_i$
+becomes a ``loser'' vertex and emits all its neighbor vertices with
+$V_i$ tagged with a loser flag.  Otherwise it emits all its neighbor
+vertices without the loser flag on $V_i$.
 
-In the fourth {\it reduce()} operation, some vertices have all their
-neighbor vertices flagged as losers; these vertices become winners.
-They are either the vertices identified as winners in the second {\it
-reduce()} operation, or vertices who would become singletons (having
-no edges) when edges containing a loser vertex are removed (on the
-next iteration).  These winning vertices are emitted to the MapReduce
+In the fourth {\it reduce()}, some vertices have all their neighbor
+vertices flagged as losers; these vertices become winners.  They are
+either the vertices identified as winners in the second {\it
+reduce()}, or vertices who would become singletons (having no edges)
+when edges containing a loser vertex are removed (on the next
+iteration).  These winning vertices are emitted to the MapReduce
 object $V$ that stores the accumulating MIS.
@@
 key, with $V_i < V_j$.  This is to ensure both copies of the edge come
-together via the fourth {\it collate()} operation.  The multivalue for
-each edge now has two values, each of which is either NULL or a loser
-flag.
+together via the fourth {\it collate()}.  The multivalue for each edge
+now has two values, each of which is either NULL or a loser flag.
 
@@
 loser flag in its multivalue does not compare its two vertices; it is
-effectively deleted from the graph.  The Luby iterations end when all
-edges have been removed from the graph.  At this point, the MapReduce
-object $V$ contains a complete MIS of vertices.
+effectively deleted from the graph.  The iterations end when all edges
+have been removed from the graph.  At this point, the MapReduce object
+$V$ contains a complete MIS of vertices.
 
@@
 algorithm, described in Figure~\ref{fig:sssp}, stores directed edges
-in $E$ and begins by labeling the distance to all vertices as
-inifinte, except a distance of 0.0 to the source vertex $V_s$.  At
-each iteration, a {\it map()} operation is performed using $E$ as
-input to add edges and their weights to each vertex in $V$.  After a
-{\it collate()} operation, each vertex $V_i$ knows its outgoing edges
-and their weights, as well as its own current distance.  It may also
-have one or more distances assigned to it from a previous iteration
-(i.e. from an inbound edge).  In the {\it reduce()} operation, each
-vertex resets its distance to the minimum of these new distances and
-its current distance.  If its distance changed, it emits new distance
-values $d_j$ for the vertices associated with its outgoing edges,
-where $d_j = d_i + w_{ij}$.  It also unsets a ``done'' flag, stored by
-each processor.  When the reduce is complete, an MPI\_Allreduce
-operation is performed to check the ``done'' flag status on all
+in $E$ and begins by labeling the distance to all vertices as infinte,
+except a distance of 0 to the source vertex $V_s$.  At each iteration,
+a {\it map()} is performed using $E$ as input to add edges and their
+weights to each vertex in $V$.  After a {\it collate()}, each vertex
+$V_i$ knows its outgoing edges and their weights, as well as its own
+current distance.  It may also have one or more distances assigned to
+it from a previous iteration (from an inbound edge).  In the {\it
+reduce()}, each vertex resets its distance to the minimum of these new
+distances and its current distance.  If its distance changed, it emits
+new distance values $d_j$ for the vertices associated with its
+outgoing edges, where $d_j = d_i + w_{ij}$.  It also unsets a ``done''
+flag, stored by each processor.  When the reduce is complete, an
+MPI\_Allreduce is performed to check the ``done'' flag status on all
 processors.  If any distance was changed, the iterations continue, and
@@
 
-MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
-\> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
-MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
-$done$ = 0 \\
-
+Input: \\
+\> MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
+\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
+\> \> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
+\\
 while not $done$: \\
 \> $done$ = 1 \\
-\> Map $V$ using $E$ as input: add edge weights in $E$ to $V$ \\
-\> \> input Key = $E_{ij}$, Value = $w_{ij}$ \\
-\> \> output Key = $V_i$, Value = $[V_j, w_{ij}]$ \\
+\> Map $V$ using $E$ as input: Add edge weights in $E$ to $V$ \\
+\> \> Input: Key = $E_{ij}$, Value = $w_{ij}$ \\
+\> \> Output: Key = $V_i$, Value = $(V_j, w_{ij})$ \\
 \> Collate $V$: Vertex is key \\
 \> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
-\> \> Input: Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$ and candidate distances for $d_i$ \\
-\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
+\> \> Input: Key = $V_i$, Multivalue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\
+\> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
 \> \> If $d_{min} \ne d_i$: \\
@@
 \> \>  Output:  Key = $V_i$, Value = $d_{min}$ \\
-\> $done$ = MPI\_Allreduce($done$,MPI\_MIN)
+\> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
+\\
+Output: \\
+\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
+\> \> Key = $V_i$, Value = $d_i$ \\
 
@@
 processors only once at the beginning of the computation, which means
-the communication portion of the {\it collate()} operation is
-performed to assign each key to a processor.  The {\it collate()}
-operation of Figure~\ref{fig:sssp} can then be replaced with a {\it
-compress()} operation which requires no communication.
+the communication portion of the {\it collate()} is performed to
+assign each key to a processor.  The {\it collate()} of
+Figure~\ref{fig:sssp} can then be replaced with a {\it compress()}
+which requires no communication.
 
@@
 
-Pre-aggregated MapReduce object $V$ = graph vertices $V_i$ with distances from sourece vertex $V_s$:  \\
-\> Key = $V_i$; Value = $\infty$, except Value for $V_s$ = 0 \\
-Pre-aggregated MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
-MapReduce object $U$ to store updates to distances, initially empty \\
-done = 0 \\
-
+Input: \\
+\> Pre-aggregated MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
+\> Pre-aggregated MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
+\> \> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
+\> MapReduce object $U$ = updates to distances, initially empty \\
+\\
 while not $done$: \\
@@
 \> Aggregate $U$ \\
-\> Add $U$ to $V$, empty $U$ \\
+\> Add $U$ to $V$, $U$ = empty \\
 \> Compress $V$: Compute best distance for each vertex \\
-\> \> Input:  Key = $V_i$; Multivalue = candidate distances for $d_i$ \\
-\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
+\> \> Input: Key = $V_i$, Multivalue = $d_i$, candidate distances $d_j$ \\
+\> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
 \> \> If $d_{min} \ne d_i$: \\
 \> \> \> $done$ = 0 \\
-\> \> \> Output to $U$: Key = $V_i$; Value = $d_{min}$ \\
-\> \> Output to $V$: Key = $V_i$; Value = $d_{min}$ \\
+\> \> \> Output to $U$: Key = $V_i$, Value = $d_{min}$ \\
+\> \> Output to $V$: Key = $V_i$, Value = $d_{min}$ \\
 \> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
 \> if not $done$: \\
-\> \> Add $U$ to $E$, empty $U$ \\
+\> \> Add $U$ to $E$, $U$ = empty \\
 \> \> Compress $E$: Generate candidate distances for neighbors of changed vertices. \\
-\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ and candidate distance $d_i$ if updated in previous step \\
-\> \> \> If updated distance $d_i$ exists: \\
-\> \> \> \> Output to $U$: Key = $V_j$, Value = $d_i+w_{ij}$ for each edge vertex $V_j$) \\
+\> \> \> Input: Key = $V_i$, Multivalue = edges $(V_j, w_{ij})$ and $d_i$ if updated in previous step \\
+\> \> \> If updated $d_i$ exists: \\
+\> \> \> \> Output to $U$: Key = $V_j$, Value = $d_i+w_{ij}$ for each edge vertex $V_j$ \\
+\\
+Output: \\
+\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
+\> \> Key = $V_i$, Value = $d_i$ \\
 

### 2010-11-12 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@665 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 made more efficient by invoking the {\it aggregate()} operation once
-on the MapReduce objects representing the matrix $A$, the all-zero
-matrix rows $MT$, and the vector $x$, before the PageRank iterations
-begin.  Since these 3 MapReduce objects are all keyed by vertices,
-pre-aggregating them moves all KV pairs with a given key to the same
-processor.  Many of the subsequent PageRank operations then become
-local operations, replacing {\it collate()} (which is equivalent to an
-{\it aggregate()} followed by {\it convert()}) with simply a {\it
+on each of these MapReduce objects before the PageRank iterations
+begin.  Since the 3 objects are all keyed by vertex, pre-aggregating
+them moves all KV pairs with a given key to the same processor.  Many
+of the subsequent PageRank operations then become local operations,
+replacing a {\it collate()} (which is equivalent to an {\it
+aggregate()} followed by {\it convert()}) with simply a {\it
 convert()}.  In Figure~\ref{fig:pr2}, steps where {\it convert()} is
@@
 \> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $i$; MultiValue = $(x_i, NULL)$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
-\> \> \> $c_{local} = c_{local} + foo$ \\
-\> \> \> Output:  If ${nvalues} = 2$: Key = $i$, Value = NULL \\
-\> \> $c_{global}$ = MPI\_Allreduce($c_{local}$,MPI\_SUM) \\
+\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$ else MultiValue = $x_i$ \\
+\> \> \> If ${nvalues} = 2$: $c_{dot} = c_{dot} + x_i$ \\
+\> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
+\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$,MPI\_SUM) \\
 \> 2 Compute $y = A^T x$  \\
@@
 \> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
-\> 3 Add $c_{global}$ to $y$; compute $n_{global}$ = max element in $y$ \\
+\> 3 Add $c_{dot}$ to $y$; compute $y_{max}$ = max element in $y$ \\
 \> \> Map $y$:  \\
 \> \> \> Input: Key = $i$, Value = $y_i$ \\
-\> \> \> $y_i = y_i + c_{global}$; $n_{local} = {\rm max}(y_i,n_{local})$ \\
+\> \> \> $y_i = y_i + c_{dot}$; $y_{max} = {\rm max}(y_i,y_{max})$ \\
 \> \> \> Output:  Key = $i$, Value = $y_i$ \\
-\> \> $n_{global}$ = MPI\_Allreduce($n_{local}$,MPI\_SUM) \\
-\> 4 Scale $y$ by $n_{global}$ \\
+\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$,MPI\_MAX) \\
+\> 4 Scale $y$ by $y_{max}$ \\
 \> \> Map $y$: \\
 \> \> \> Input:  Key = $i$, Value = $y_i$ \\
-\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
+\> \> \> Output:  Key = $i$, Value = $y_i / y_{max}$ \\
 \> 5 Compute residual \\
@@
 \> \> \> Input: Key = $i$, Multivalue = $x_i, y_i$ \\
-\> \> \> $r_{local} = {\rm max}(|x_i - y_i|,r_{local})$ \\
-\> \> $residual$ = MPI\_Allreduce($r_{local}$,MPI\_MAX) \\
+\> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
+\> \> $residual$ = MPI\_Allreduce($residual$,MPI\_MAX) \\
 \> 6 Copy $y$ to $x$ \\
@@
 
-In Step 1 of a PageRank interation, $MT$ is used in dot products to
-compute adjustments to the PageRank vector to represent the uniform
-probability of going from a leaf page to any other page.  Adjustments
-for random jumps from any page are also computed.  In Step 2, the
+In Step 1 of a PageRank interation, $MT$ is effectively dotted with
+$x$ to compute an adjustment $c_{dot}$ to the PageRank vector that
+represents the uniform probability of going from a leaf page to any
+other page.  $MT$ is unchanged by the operation.  Adjustments for
+random jumps from any page are also computed.  In Step 2, the
 matrix-vector product $A^T x$ is computed and the result stored in
@@
 product vector; the work is proportional to $|V|$.  Step 5 computes
-the residual to determine whether the PageRank vector has converged;
-work is again proportional to $|V|$.  In Step 6, the PageRank vector
-is overwritten prior to the next iteration.  Throughout the algorithm,
-we exploit the ability to call MPI\_Allreduce to compute global norms
-and residuals.
+the residual as an infinity norm to determine whether the PageRank
+vector has converged; work is again proportional to $|V|$.  Note that
+$y$ and $x$ are both keyed by vertex at this point, so they can be
+joined via an {\it add()} and {\it convert()} without need for the
+communication required by a full {\it collate()}.  In Step 6, the
+PageRank vector is overwritten prior to the next iteration.
+Throughout the algorithm, we exploit the ability to call
+MPI\_Allreduce to compute global dot products and norms.
 
@@
 vertex distancs are updated.  A MapReduce formulation of this
-algorithm, described in Figure~\ref{fig:sssp}, stores directed edges
-in $E$ and begins by labeling the distance to all vertices as infinte,
-except a distance of 0 to the source vertex $V_s$.  At each iteration,
-a {\it map()} is performed using $E$ as input to add edges and their
-weights to each vertex in $V$.  After a {\it collate()}, each vertex
-$V_i$ knows its outgoing edges and their weights, as well as its own
-current distance.  It may also have one or more distances assigned to
-it from a previous iteration (from an inbound edge).  In the {\it
-reduce()}, each vertex resets its distance to the minimum of these new
-distances and its current distance.  If its distance changed, it emits
-new distance values $d_j$ for the vertices associated with its
-outgoing edges, where $d_j = d_i + w_{ij}$.  It also unsets a ``done''
-flag, stored by each processor.  When the reduce is complete, an
+algorithm, described in Figure~\ref{fig:sssp}, begins with a graph $E$
+of directed edges $(V_i,V_j)$.  As generated by the R-MAT procedure
+disussed earlier, this may include self-edges $(V_i,V_i)$ and loop
+edges (both $(V_i,V_j)$ and $(V_j,V_i)$).  These are handled as
+special cases within the algorithm to avoid spurious contributions to
+the shortest path distances.  MapReduce object $V$ stores the distance
+of each vertex $V_i$ from the source vertex $V_s$.
+
+To begin, the distance to all vertices is set to infinity, except a
+distance of 0 for $V_s$ itself.  At each iteration, a {\it map()} is
+performed using $E$ as input to add edges and their weights to each
+vertex in $V$.  After a {\it collate()}, each vertex $V_i$ knows its
+outgoing edges and their weights, as well as its own current distance.
+It may also have one or more distances assigned to it from a previous
+iteration (from an inbound edge).  In the {\it reduce()}, each vertex
+resets its distance to the minimum of these new distances and its
+current distance.  If its distance changed, it emits new distance
+values $d_j$ for the vertices associated with its outgoing edges,
+where $d_j = d_i + w_{ij}$.  It also unsets a ``done'' flag, stored by
+each processor.  When the {\it reduce()} is complete, an
 MPI\_Allreduce is performed to check the ``done'' flag status on all
 processors.  If any distance was changed, the iterations continue, and
-the distance updates propagate through the graph.  If no distances
-changed, the algorithm is done.
+the distance updates propagate through the graph.  Otherwise the
+algorithm is done.
 

### 2010-11-12 paper/lessons.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@665 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 for hard-to-balance irregular data sets, such as those derived from
-sparse graphs with power-law degree distributions.  By contrast, more
-traditional distributed-memory parallel algorithms, e.g. for matrix
-operations, or grid-based or particle-based simulation codes,
-typically work hard to localize data and minimize communication.  To
-do this they often require application-specific logic and parallel
-communication coding to create and maintain a data decomposition,
-generate ghost copies of nearby spatially-decomposed data, etc.
+sparse graphs with power-law degree distributions.  By contrast,
+MPI-based parallel algorithms, e.g. for matrix operations, or
+grid-based or particle-based simulation codes, typically work hard to
+localize data and minimize communication.  To do this they often
+require application-specific logic and parallel communication coding
+to create and maintain a data decomposition, generate ghost copies of
+nearby spatially-decomposed data, etc.
 
@@
 support MPI, for a variety of data-intensive or compute-intensive
-problems that are amenable to solution using a MapReduce paradigm.
+problems that are amenable to solution using a MapReduce approach.

### 2010-11-12 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@665 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-In this section, we present performance results for the MapReduce
-graph algorithms of the preceeding section, implemented as small C++
+In this section, we present performance results for the graph
+algorithms of the preceeding section, implemented as small C++
 programs calling our MR-MPI library.  The benchmarks were run on a
@@
 node of the cluster has one local disk, which is used for out-of-core
-operations in MR-MPI.  To avoid contention for disk I/O, we ran all
-experiments with one MPI process per node.  For comparisons with other
+operations.  To avoid contention for disk I/O, we ran all experiments
+with one MPI process per node.  For comparisons with other
 implementations, we used either the same cluster or, where noted,
@@
 
-We ran each of the algorithms on three R-MAT graphs of different
-sizes, each on a varying number of processors.  Details of the input
-data are shown in Table~\ref{t:rmats}.  The {\it small} problem
-(around 8M edges) can typically be run on a single processor without
-incurring out-of-core operations.  The {\it medium} problem (around
-134M edges) can be run on a single processor with out-of-core
-operations; larger processor configurations, however, do not
-necessarily require out-of-core operations.  The {\it large} problem
-(around 2B edges) requires out-of-core operations on our 64-node
-cluster.
+We ran the algorithms on three R-MAT graphs of different sizes, each
+for a varying number of processors.  Details of the input data are
+shown in Table~\ref{t:rmats}.  The {\it small} problem (around 8M
+edges) can typically be run on a single processor without incurring
+out-of-core operations.  The {\it medium} problem (around 134M edges)
+can be run on a single processor with out-of-core operations; larger
+processor configurations, however, do not necessarily require
+out-of-core operations.  The {\it large} problem (around 2B edges)
+requires out-of-core operations on our 64-node cluster.  These are
+edge counts before pre-processing steps used in some of the
+algorithms, e.g. to eliminate self-edges or convert the matrix to
+upper-triangular form.
 
@@
 All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
-0.05)$ and generated 8 edges per vertex (on average).  These values
-create a highly skewed degree distribution, as indicated by the
-maximumn vertex degree in Table~\ref{t:rmats}.
+0.05)$ and generated 8 edges per vertex (on average).  These
+parameters create a highly skewed degree distribution, as indicated by
+the maximumn vertex degree in Table~\ref{t:rmats}.
 
@@
 compare the MapReduce algorithm with other parallel implementations,
-including more traditional distributed-memory algorithms and
+including more traditional distributed-memory algorithms and also
 multi-threaded algorithms in the Multi-Threaded Graph Library
@@
 $P$ processors as $({time}_{M} \times M) / ({time}_P \times P)$ where
-$M$ is the smallest number of processors on which the experiment was
+$M$ is the smallest number of processors on which the same graph was
 run, and ${time}_I$ is the execution time required on $I$ processors.
@@
 generation algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and
-RMAT-28.  For RMAT-20 and RMAT-24, superlinear speed-up is shown.
-This speed-up is due to the decreased amount of file I/O needed with
+RMAT-28.  For RMAT-20 and RMAT-24, the algorithm exhibits superlinear
+speed-up.  This is due to the decreased amount of file I/O needed with
 greater numbers of processors; with a larger total memory, a
-fixed-size problem requires less I/O with more processors.  For
+fixed-size problem requires less I/O as processors are added.  For
 RMAT-28, which requires significant out-of-core operations, parallel
-efficiency ranges from 52\% to 97\%.  Comparisons between the enhanced
-algorithm (Figure~\ref{fig:rmat2}) and the original algorithm
-(Figure~\ref{fig:rmat}) showed approximately 10\% reduction in
-execution time for the enhanced algorithm.
+efficiency ranges from 52\% to 97\%.  For these problems the enhanced
+algorithm of Figure~\ref{fig:rmat2} ran about 10\% faster than the
+algorithm of Figure~\ref{fig:rmat}.
 
@@
 
-In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank
+In Figure~\ref{f:pr}, we show the performance of the PageRank
 algorithm (Figure~\ref{fig:pr2}) compared to a distributed-memory
 matrix-based implementation using the linear algebra toolkit
-Trilinos~\cite{Trilinos-Overview}.  The matrix-based
-distributed-memory implementation of PageRank uses Trilinos Epetra
-matrix/vector classes to represent the graph and PageRank vector.
-Rows of matrix $A$ and the associated entries of the PageRank vector
-$x$ are uniquely assigned to processors; a random permutation of the
-input matrix effectively load balances the non-zero matrix entries
-across processors.  Interprocessor communication gathers $x$ values
-for matrix-vector multiplication and sums partial products into the
-$y$ vector.  Most communication is point-to-point communication, but
-some global communication is needed for computing residuals and norms
-of $x$ and $y$.
+Trilinos~\cite{Trilinos-Overview}.  The Trilinos implementation of
+PageRank uses Epetra matrix/vector classes to represent the graph and
+PageRank vectors and requires the matrix and associated vectors fit in
+the aggregate memory of the processors used.  Rows of matrix $A$ and
+the associated entries of the PageRank vector $x$ are uniquely
+assigned to processors; a random permutation of the input matrix
+effectively load balances the non-zero matrix entries across
+processors.  Interprocessor communication gathers $x$ values for
+matrix-vector multiplication and sums partial products into the $y$
+vector.  Most communication is point-to-point communication, but some
+global communication is needed for computing residuals and norms of
+$x$ and $y$.
 
@@
 Several R-MAT matrices of each size were generated for the
-experiments; the average time over the matrices is reported here.  The
-Trilinos implementations show near-perfect strong scaling for RMAT-20
-and RMAT-24.  The MR-MPI implementations also demonstrate good strong
-scaling.  However, MR-MPI's execution time is at least an order of
-magnitude greater than the Trilinos implementation.  
-The benefit of MR-MPI's out-of-core implementation
-is seen, however, with the RMAT-28 data set, which could be solved on
-smaller processor sets than the Trilinos implementation.  For these
-experiments, the Trilinos implementation required 64 processors for
-the RMAT-28 data set, since it will not operate on matrices that
-exceed the aggregate memory of the processors.
+experiments; the average time over the set of matrices is reported
+here.  The Trilinos implementations show near-perfect strong scaling
+for RMAT-20 and RMAT-24.  The MR-MPI implementations also demonstrate
+good strong scaling.  However, MR-MPI's execution time is more than an
+order of magnitude slower than the Trilinos implementation.  The
+benefit of MR-MPI's out-of-core implementation is seen, however, with
+the RMAT-28 data set, which could be run on smaller processor sets
+than the Trilinos implementation, which required 64 processors for
+RMAT-28.
 
@@
 A more detailed analysis of the performance of the MR-MPI PageRank
-algorithm is shown
-below.  We perform the analysis using RMAT-20, which is small enough to
-fit in memory and thus does not need out-of-core operations, and RMAT-28,
-which requires out-of-core operation in MR-MPI.  Table~\ref{table:prdetail}
-shows the amount of time spent in each step of Figure~\ref{fig:pr2} for 
-the MR-MPI implementation and the Trilinos implementation.  We see that for
-both implementation, the matrix-vector multiplication (Step 2) is the most
-costly step.  However, the MR-MPI algorithm is consistently slower than
-the Trilinos implementation in each step.  This effect is more pronounced
-for RMAT-28, where MR-MPI performs out-of-core operations.
+algorithm is shown below.  We perform the analysis using RMAT-20,
+which is small enough to fit in memory and thus does not need
+out-of-core operations, and RMAT-28, which requires out-of-core
+operation in MR-MPI.  Table~\ref{table:prdetail} shows the amount of
+time spent in each step of Figure~\ref{fig:pr2} for the MR-MPI
+implementation and the equivalent Trilinos implementation.  For both
+implementations, the matrix-vector multiplication (Step 2) is the most
+costly step.  However, the MR-MPI algorithm is consistently slower
+than the Trilinos implementation in each step.  This effect is more
+pronounced for RMAT-28, where MR-MPI performs out-of-core operations.
 
@@
 
-To understand where MR-MPI spends its effort during the PageRank computation,
-we break down the operation times further for the most expensive 
-step of the PageRank computation, Step 2; the results are in 
-Table~\ref{table:prdetail2}.
-In the table, we divide the collate operation into its
-two parts:  an aggregate that involves parallel communication and a convert
-that gathers local key-value pairs with matching keys into a key-multivalue
-pair.  All interprocessor communication occurs in the aggregate operation.
-While the time spent in aggregate is significant, it is small compared to
-the time spent in the two convert operations.  When out-of-core operations
-are used for RMAT-28, the communication cost in the aggregate operation is
-a smaller percentage of total time for Step 2, 
-as other data manipulation operations such as convert and add 
-take a relatively larger percentage of the total time.
-Further experimentation
-is needed to determine whether reorganizing the MapReduce data 
-(as described in Section~\ref{subsec:graph_pagerank}) would reduce the
-execution time of Step 2.  Organizing the data such that both the matrix
-$A$ and vectors $x$ and $y$ are in the same MapReduce object would eliminate
-the copy and add functions, as well as the first convert.  However, to update
-the $y$ values at the completion of Step 2, either an additional convert
-or an aggregate with greater amounts of data would be needed.
+To understand where MR-MPI spends its effort during the PageRank
+computation, we break down the operation times further for the most
+expensive Step 2 of the PageRank computation; the results are in
+Table~\ref{table:prdetail2}.  In the table, we divide the {\it
+collate()} operation into its two parts: an {\it aggregate()} that
+involves parallel communication and a {\it convert()} that gathers
+local key/value (KV) pairs with matching keys into a key-multivalue
+(KMV) pair.  All interprocessor communication occurs in the {\it
+aggregate()}.  While the time spent in {\it aggregate()} is
+significant, it is small compared to the time spent in the two {\it
+convert()} operations.  When out-of-core operations are required for
+RMAT-28, the communication cost in the {\it aggregate()} is a smaller
+percentage of total time for Step 2, as other data manipulation
+operations such as {\it convert()} and {\it add()} require a
+relatively larger percentage of the total time.  Further
+experimentation is needed to determine whether reorganizing the
+MapReduce data, as described in Section~\ref{subsec:graph_pagerank}.
+would reduce the execution time of Step 2.  Organizing the data such
+that both the matrix $A$ and vectors $x$ and $y$ are in the same
+MapReduce object would eliminate the {\it copy()} and {\it add()}
+functions, as well as the first {\it convert()}.  However, to update
+the $y$ values at the completion of Step 2, either an additional {\it
+convert()} or {\it aggregate()} with greater amounts of data would be
+needed.
 
@@
 Similarly, in Table~\ref{table:prdetail5}, we show the execution times
-required for each operation in Step 5 of Figure~\ref{fig:pr2}, the 
-residual computation.  Clearly the time to convert key-value pairs to 
-key-multivalues dominates this computation.  Reorganization of the vectors
-$x$ and $y$ into a single MapReduce object (as described in 
-Section~\ref{subsec:graph_pagerank}) would benefit this step of the computation,
-as the add, convert, and reduce operations could be replaced by a single 
-map operation, with execution time comparable to the vector-scaling operation
-in Step 4.
+required for each operation in Step 5 of Figure~\ref{fig:pr2}, for the
+residual computation.  Clearly the time to convert KV pairs to KMV
+pairs dominates this computation.  Reorganization of the vectors $x$
+and $y$ into a single MapReduce object (as described in
+Section~\ref{subsec:graph_pagerank}) would benefit this step of the
+computation, as the {\it add()}, {\it convert()}, and {\it reduce()}
+could be replaced by a single {\it map()}, with execution time
+comparable to the vector-scaling operation of Step 4.
 
@@
 \end{tabular}
-\caption{Percentage of execution time in each operation of Step 5 of the MR-MPI 
-PageRank algorithm in  Figure~\ref{fig:pr2}.}
+\caption{Percentage of execution time in each operation of Step 5 of
+the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}.}
 \label{table:prdetail5}
@@
 \end{table}
+
 %This result is
@@
 our cluster.  Parallel efficiencies for RMAT-20 ranged from 80\% to
-140\%; for RMAT-24, they ranged from 50\% to 120\%.
+140\%; for RMAT-24, they ranged from 50\% to 120\%.  The RMAT-20
+matrix contained about 100 million triangles; the RMAT-24 contained
+2.1 billion.
 
@@
 have one or two very large components and many very small components.
-The hybrid algorithm exploits this feature of the data by using
-inexpensive breadth-first search (BFS) from the vertex with highest
-degree to identify the largest components, followed by a more
+The hybrid algorithm exploits this feature by using an inexpensive
+breadth-first search (BFS) from the vertex with highest degree to
+identify the largest components one at a time, followed by a more
 expensive algorithm to identify the small components in the remainder
 of the graph.  In our hybrid implementation, we intially perform
-matrix-vector multiplications in Trilinos to perform a BFS, finding
-components that include (in total) 70\% or more of the vertices.  We
-then apply our MR-MPI algorithm to the remaining graph to identify the
-small components.  This hybrid approach is quite effective, reducing
-the execution time to identify all components of RMAT-20 and RMAT-24
-by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI is
-seen, however, for RMAT-28, where the graph is too large to fit into
-memory, and thus Trilinos cannot perform the initial BFS for our
-hybrid algorithm.
+matrix-vector multiplications in Trilinos to perform a BFS, finding a
+handful of components that include 70\% or more of the vertices (for
+these R-MAT matrices).  We then apply our MapReduce algorithm to the
+remaining graph to identify the small components.  This hybrid
+approach is quite effective, reducing the execution time to identify
+all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
+Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
+into memory, Trilinos cannot perform the initial BFS, so only the
+MapReduce algorithm was used.  The total number of components
+identified were A, B, C for the 3 increasing R-MAT sizes; the number
+of iterations required by the algorithm to find them was X, Y, Z
+respectively.
 
@@
 
-The execution times for the maximal independent set algorithm
+The execution times for the maximal independent set (MIS) algorithm
 (Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
@@
 requires significant out-of-core operations. In this case, parallel
-efficiency is nearly perfect going from 8 to 64 processors.
+efficiency is nearly perfect going from 8 to 64 processors.  The
+number of vertices in the resulting MIS were A, B, C for the 3
+increasing R-MAT sizes; the number of iterations required by the
+algorithm to find them was X, Y, Z respectively.
 
@@
 
-The execution times for the single-source shortest path algorithms
-(Figures~\ref{fig:sssp} and \ref{fig:sssp2}) are shown in
-Figure~\ref{f:sssp}.  The results show the benefit of using the
-enhanced algorithm, providing at least a 17\% (and often greater)
-reduction in execution time due to reduced communication; only the
-updated distances are communicated throughout most of the enhanced
-algorithm.  However, the execution times are still large compared to
-multi-threaded implementations; for example, Madduri et
-al.~\cite{Madduri07} report execution times of only 11 seconds on 40
-processors for a multithreaded implementation on a Cray MTA-2 (the
-predecessor of the XMT) on graphs with one billion edges.
+The execution times for the SSSP algorithms of Figures~\ref{fig:sssp}
+and \ref{fig:sssp2} are shown in Figure~\ref{f:sssp}.  The results
+show the benefit of using the enhanced algorithm, which gave at least
+a 17\% (and often greater) reduction in execution time due to reduced
+communication; only the updated distances are communicated throughout
+most of the enhanced algorithm.  However, the execution times are
+still large compared to multi-threaded implementations; for example,
+Madduri et al.~\cite{Madduri07} report execution times of only 11
+seconds on 40 processors for a multithreaded implementation on a Cray
+MTA-2 (the predecessor of the XMT) on graphs with one billion edges.
+The number of iterations required in the MapReduce algorithms to find
+the SSSP to all vertices for the 3 increasing R-MAT sizes was X, Y, Z
+respectively.
 
@@
 (Figure~\ref{fig:sssp}).  But the Bellman-Ford-style iterations are
-especially harmful to the MR-MPI and Hadoop implementations for
+especially costly in the MR-MPI and Hadoop implementations for
 {WebGraphB}, which required 110 iterations to complete; execution
@@
 that fit in memory, so out-of-core operations were not needed.  For
-these experiments, we used an R-MAT data set with with $2^{25}$
+these experiments, we used two R-MAT data sets with with $2^{25}$
 vertices and $2^{28}$ edges, with parameters given in
-Table~\ref{t:rmat}.  We ran both the PageRank and Connected Components
-algorithms.
+Table~\ref{t:rmat}.  The ``nice'' parameters produce a less-skewed
+sparsity pattern in the matrix than the ``nasty'' parameters, as
+indicated by the maximum degree metric.  We ran both the PageRank and
+Connected Components algorithms on these matrices.
 
@@
 execution time than the matrix-based implementations on Redstorm.  The
-distributed memory matrix-based implementations are competitive with
-the multi-threaded implementation in MTGL on the Cray XMT.
+Trilinos implementation is competitive with the multi-threaded
+implementation in MTGL on the Cray XMT, when both are run on the same
+number of processors.
 

### 2010-11-12 paper/thanks.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@659 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@665 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \label{sec:thanks}
+
+The MR-MPI library is open-source software, which can be downloaded
+from \url{http://www.sandia.gov/~sjplimp/mapreduce.html}.  It is
+freely available under the terms of a BSD license.  Benchmark programs
+which call the library to implement the algorithms of and which
+produced the timing results of Section~\ref{sec:results} are included
+in the distribution.
 

### 2010-11-19 paper/graph.tex
Minor fixes to pagerank.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@666 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 linear-algebra concepts implemented in a MapReduce framework, with
-data stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$
-stores the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are
+data stored in four MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$
+stores the matrix; $MT$ stores the row index of ``empty'' rows of $A$; 
+and $x$ and $y$ are
 vectors of length $|V|$.  The nonzeros of matrix $A$ are initialized
@@
 on each of these MapReduce objects before the PageRank iterations
-begin.  Since the 3 objects are all keyed by vertex, pre-aggregating
+begin.  Since the objects are all keyed by vertex, pre-aggregating
 them moves all KV pairs with a given key to the same processor.  Many
@@
 \> MapReduce object $y$ = vector, initially empty \\
+\> User-defined parameter $\alpha$ \\
 \\
 while $({residual} > {tolerance})$ \\
-\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
+\> 1 Compute adjustment for random jumps and zero-outdegree vertices \\
 \> \> Add $x$ to $MT$ \\
@@
 \> \> Reduce $MT$:   \\
-\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$ else MultiValue = $x_i$ \\
-\> \> \> If ${nvalues} = 2$: $c_{dot} = c_{dot} + x_i$ \\
+\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$; else MultiValue = $x_i$ \\
+\> \> \> If (${nvalues} = 2$) $c_{dot} = c_{dot} + x_i$; else $c_{dot} = c_{dot} + \alpha$ \\
 \> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
-\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$,MPI\_SUM) \\
+\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$, MPI\_SUM) / $|V|$ \\
 \> 2 Compute $y = A^T x$  \\
@@
 \> \> \> Output:  Key = $i$, Value = $y_i$ \\
-\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$,MPI\_MAX) \\
+\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$, MPI\_MAX) \\
 \> 4 Scale $y$ by $y_{max}$ \\
@@
 \> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
-\> \> $residual$ = MPI\_Allreduce($residual$,MPI\_MAX) \\
+\> \> $residual$ = MPI\_Allreduce($residual$, MPI\_MAX) \\
 \> 6 Copy $y$ to $x$ \\
@@
 Output: \\
-\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank \\
+\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank
 
@@
 
-In Step 1 of a PageRank interation, $MT$ is effectively dotted with
+Step 1 of a PageRank interation is, effectively, a dot-product of $MT$ and
 $x$ to compute an adjustment $c_{dot}$ to the PageRank vector that
-represents the uniform probability of going from a leaf page to any
-other page.  $MT$ is unchanged by the operation.  Adjustments for
-random jumps from any page are also computed.  In Step 2, the
+represents random jumps and the uniform probability of going from a 
+leaf page to any
+other page.  $MT$ is unchanged by the operation.  
+In Step 2, the
 matrix-vector product $A^T x$ is computed and the result stored in
@@
 Trilinos~\cite{Trilinos-Overview} and PETSc~\cite{petsc-efficient}
-naturally exploit this idea.  In parallel, these toolkits share
-pre-computed data layouts and communication maps among distributed
+naturally exploit this idea.  In parallel, each of these toolkits shares
+pre-computed data layouts and communication maps among its distributed
 objects to enable efficient operations on the objects.  In a MapReduce

### 2010-11-19 paper/results.tex
Updated counts of iterations.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@667 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 MapReduce algorithm was used.  The total number of components
-identified were A, B, C for the 3 increasing R-MAT sizes; the number
-of iterations required by the algorithm to find them was X, Y, Z
-respectively.
+identified were roughly 498K, 9.37M, and 169M for the three increasing 
+R-MAT sizes; five or six iterations were 
+required by the algorithm.
 
@@
 efficiency is nearly perfect going from 8 to 64 processors.  The
-number of vertices in the resulting MIS were A, B, C for the 3
+number of vertices in the resulting MIS were 368K, 5.18M, and 72.4M
+for the three
 increasing R-MAT sizes; the number of iterations required by the
-algorithm to find them was X, Y, Z respectively.
+algorithm ranged from five for RMAT-20 to nine for RMAT-28.
 
@@
 MTA-2 (the predecessor of the XMT) on graphs with one billion edges.
-The number of iterations required in the MapReduce algorithms to find
-the SSSP to all vertices for the 3 increasing R-MAT sizes was X, Y, Z
-respectively.
+Seven to ten iterations were required in the MapReduce algorithms to find
+the SSSP to all vertices.
 

### 2010-11-19 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@669 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
+\subsection{Implementation effort}
+
+To give a rough sense of the effort involved to implement the various
+MapReduce algorithms described in this section, we list the line
+counts of the C++ codes used to perform the benchmarks.  Each code
+makes calls to the MR-MPI library and implements the algorithm in the
+indicated figure.  The line counts exclude comments and blank lines
+which typically add about 30\% to the total.  They also exclude the
+routines used to generate the R-MAT matrix input (except in the case
+of the two R-MAT algorithms themselves) or process it into
+upper-triangular form, since that is common to several of the
+benchmarks.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
+\hline
+Algorithm & R-MAT & R-MAT & PageRank & Tri & CC & MIS & SSSP & SSSP \\
+\hline
+Line count & 208 & 216 & 694 & 367 & 453 & 393 & 505 & 569 \\
+Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
+\hline
+\end{tabular}
+\caption{Line counts of codes that implement each of the graph
+  algorithms in the associated figures.}
+\label{t:rmat}
+\end{center}
+\end{table}
+
+For comparison, we wrote a parallel distributed-memory R-MAT
+generation code with explicit MPI calls using a different
+(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
+\ref{fig:rmat2}.  It required 1842 lines of code, mainly because it
+must implement the parallelism explicitly whereas the codes of this
+section relegate that task to the MR-MPI library.  Our collaborators
+(see Acknowledgements) have a Hadoop implementation of the SSSP
+algorithm of Figure \ref{fig:sssp}.  It required XXX lines of Java
+code.
+
 \subsection{Scalability to large numbers of processors}
@@
 
-Figure~\ref{f:prbig}, shows the performance of the various PageRank
+Figure~\ref{f:prbig} shows the performance of the various PageRank
 implementations on distributed memory and multi-threaded

### 2010-11-19 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@670 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@671 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \hline
-Algorithm & R-MAT & R-MAT & PageRank & Tri & CC & MIS & SSSP & SSSP \\
+Algorithm & R-MAT & R-MAT++ & PageRank & Tri & CC & MIS & SSSP & SSSP++ \\
 \hline
-Line count & 208 & 216 & 694 & 367 & 453 & 393 & 505 & 569 \\
+Lines of code & 208 & 216 & 694 & 367 & 453 & 393 & 505 & 569 \\
 Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
@@
 \end{tabular}
-\caption{Line counts of codes that implement each of the graph
+\caption{Lines of code to implement each of the graph
   algorithms in the associated figures.}
@@
 (non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
-\ref{fig:rmat2}.  It required 1842 lines of code, mainly because it
+\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
 must implement the parallelism explicitly whereas the codes of this
-section relegate that task to the MR-MPI library.  Our collaborators
-(see Acknowledgements) have a Hadoop implementation of the SSSP
-algorithm of Figure \ref{fig:sssp}.  It required XXX lines of Java
-code.
+section relegate that task to the MR-MPI library.  A similar all-MPI
+implementation of the PageRank algorithm required 1206 lines.  By
+contrast an implementation of PageRank using the Trilinos parallel
+linear algebra library (discussed in Section
+\ref{subsec:results_pagerank}) required only 153 lines of code, since
+an entire matrix-vector multiply can be invoked by a single call to
+Trilinos.  Our collaborators (see Acknowledgements) have a Hadoop
+implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
+required XXX lines of Java code.
 

### 2010-11-19 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 commonly used to represent web and social networks.  The user
-specifices six parameters that define the graph: the number of
+specifies six parameters that define the graph: the number of
 vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
@@
 \> \> Reduce $y$: \\
-\> \> \>  Input: Key = $i$, Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
+\> \> \>  Input: Key = $i$, MultiValue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
 \> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
@@
 \> \> Reduce $y$: \\
-\> \> \> Input: Key = $j$, Multivalue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
+\> \> \> Input: Key = $j$, MultiValue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
 \> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
@@
 \> \> Reduce $x$:  \\
-\> \> \> Input: Key = $i$, Multivalue = $x_i, y_i$ \\
+\> \> \> Input: Key = $i$, MultiValue = $x_i, y_i$ \\
 \> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
@@
 
-Step 1 of a PageRank interation is, effectively, a dot-product of $MT$ and
+Step 1 of a PageRank iteration is, effectively, a dot-product of $MT$ and
 $x$ to compute an adjustment $c_{dot}$ to the PageRank vector that
@@
 a MapReduce algorithm that labels each vertex in a graph with a
-component ID.  All vertices in the same component are labelled with
+component ID.  All vertices in the same component are labeled with
 the same ID, which is the ID of some vertex in the component.  We
@@
        \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
-       \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
+       \> \> partition $Z_{new}$ if number of $V >$ threshold \\
        \> \> for each $V$ in MultiValue: \\
@@
 operations in stages 3 and 4 include extra logic.  The idea is to
-partition zones whose vertex count exceeds a user-defined threshhold
+partition zones whose vertex count exceeds a user-defined threshold
 into $P$ sub-zones, where $P$ is the number of processors.  The 64-bit
@@
 partition bit is also explicitly set if the vertex count exceeds the
-threshhold for the first time.
+threshold for the first time.
 
@@
 subsequent iterations, the fourth {\it map()} emits $P+1$ copies of
-new zone reassignments for both partitioned and unpartioned zone IDs,
-to ensure all vertices in the zone will know the correct reassignment
-information.
+new zone reassignments for both partitioned and unpartitioned zone
+IDs, to ensure all vertices in the zone will know the correct
+reassignment information.
 
@@
 \> 1 Reduce $E$: Determine WINNER/LOSER vertex of each edge \\
-       \> \> Input: Key = $E_{ij}$, Multivalue = NULL or LOSER \\
+       \> \> Input: Key = $E_{ij}$, MultiValue = NULL or LOSER \\
        \> \> if multivalue = LOSER, emit nothing \\
@@
        \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
-       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> for each $V_j$ in MultiValue: \\
        \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, WINNER) \\
        \> \> else: \\
-       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> for each $V_j$ in MultiValue: \\
        \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
@@
        \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
-       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> for each $V_j$ in MultiValue: \\
        \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, LOSER) \\
        \> \> else: \\
-       \> \> \> for each $V_j$ in Multivalue: \\
+       \> \> \> for each $V_j$ in MultiValue: \\
        \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
@@
        \> \> \> Output: to $V$: Key = $V_i$, Value = NULL \\
-       \> \> for each $V_j$ in Multivalue: \\
+       \> \> for each $V_j$ in MultiValue: \\
        \> \> \> if $V_j$ is flagged with LOSER: \\
@@
 In the third {\it reduce()}, if any neighbor of $V_i$ was a winner in
-the previoue {\it reduce()} (indicated by a winner flag), $V_i$
+the previous {\it reduce()} (indicated by a winner flag), $V_i$
 becomes a ``loser'' vertex and emits all its neighbor vertices with
@@
 Ford62}, every edge of the graph is visited in each iteration and
-vertex distancs are updated.  A MapReduce formulation of this
+vertex distance are updated.  A MapReduce formulation of this
 algorithm, described in Figure~\ref{fig:sssp}, begins with a graph $E$
 of directed edges $(V_i,V_j)$.  As generated by the R-MAT procedure
-disussed earlier, this may include self-edges $(V_i,V_i)$ and loop
+discussed earlier, this may include self-edges $(V_i,V_i)$ and loop
 edges (both $(V_i,V_j)$ and $(V_j,V_i)$).  These are handled as
@@
 \> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
-\> \> Input: Key = $V_i$, Multivalue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\
+\> \> Input: Key = $V_i$, MultiValue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\
 \> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
@@
 \> Compress $V$: Compute best distance for each vertex \\
-\> \> Input: Key = $V_i$, Multivalue = $d_i$, candidate distances $d_j$ \\
+\> \> Input: Key = $V_i$, MultiValue = $d_i$, candidate distances $d_j$ \\
 \> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
@@
 \> \> Compress $E$: Generate candidate distances for neighbors of changed vertices. \\
-\> \> \> Input: Key = $V_i$, Multivalue = edges $(V_j, w_{ij})$ and $d_i$ if updated in previous step \\
+\> \> \> Input: Key = $V_i$, MultiValue = edges $(V_j, w_{ij})$ and $d_i$ if updated in previous step \\
 \> \> \> If updated $d_i$ exists: \\

### 2010-11-19 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-The basic datums stored and operated on by any MapReduce framweork are
+The basic datums stored and operated on by any MapReduce framework are
 key/value (KV) pairs.  In the MR-MPI library, individual keys or
@@
 string with an associated vector of lengths, one integer length per
-value.  In this section, we assume the KV and KVM datums operated on
+value.  In this section, we assume the KV and KMV datums operated on
 by each processor fit in local memory.
@@
 user's program can string them together in various ways to construct
-usefule MapReduce algorithms.  For example, output from a {\it
+useful MapReduce algorithms.  For example, output from a {\it
 reduce()} can serve as input to a subsequent {\it map()} or {\it

### 2010-11-19 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 movement between processors.  We describe how typical MapReduce
-funtionality can be implemented in an MPI context, and also in an
+functionality can be implemented in an MPI context, and also in an
 out-of-core manner for data sets that do not fit within the aggregate

### 2010-11-19 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@672 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 In this section, we present performance results for the graph
-algorithms of the preceeding section, implemented as small C++
+algorithms of the preceding section, implemented as small C++
 programs calling our MR-MPI library.  The benchmarks were run on a
@@
 parameters create a highly skewed degree distribution, as indicated by
-the maximumn vertex degree in Table~\ref{t:rmats}.
+the maximum vertex degree in Table~\ref{t:rmats}.
 
@@
 expensive algorithm to identify the small components in the remainder
-of the graph.  In our hybrid implementation, we intially perform
+of the graph.  In our hybrid implementation, we initially perform
 matrix-vector multiplications in Trilinos to perform a BFS, finding a
@@
 identified were roughly 498K, 9.37M, and 169M for the three increasing
-R-MAT sizes; five or six iterations were 
-required by the algorithm.
+R-MAT sizes; five or six iterations were required by the algorithm.
 
@@
 \includegraphics[width=\textwidth]{fig_ssspA.pdf}
-\caption{SSSP using MR-MPI for WebGraphA with
-13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
-are also shown.}
+\caption{SSSP using MR-MPI for WebGraphA with 13.2M vertices and 31.9M
+  edges.  Run times using Hadoop and PBGL are also shown.}
 \label{f:ssspA}
@@
 an entire matrix-vector multiply can be invoked by a single call to
-Trilinos.  Our collaborators (see Acknowledgements) have a Hadoop
+Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
 implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It

### 2010-11-22 paper/graph.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 randomized graphs, since they are the input for the algorithms
-discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
+discussed below.  R-MAT matrices~\cite{RMAT} are recursively
 generated graphs with power-law degree distributions.  They are
@@
 Output: \\
-\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
-
+\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL 
   \end{tabbing}
@@
 quadrant as a sub-matrix and select a random quadrant within it, in
-the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
+the same manner.  Perform this process a total of $n$ times where $N = 2^n$.  At
 the end of the recursion, the final ``quadrant'' is a single non-zero
@@
 sequence is repeated until the number of resulting key/value (KV)
-pairs $N_{kv} = M$.  For reasonably sparse graphs this typically takes
-only a few iterations.
+pairs $N_{kv} = M$.  For reasonably sparse graphs,
+only a few iterations are typically needed.
 
@@
 \> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank 
-
   \end{tabbing}
@@
 
- \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
+ \caption{\it MapReduce algorithm for PageRank vertex ranking, using four
  MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
@@
 triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
-where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
+where $V_i < V_j$ for every edge, i.e., an upper-triangular R-MAT
 matrix.  This exposition follows the triangle-finding algorithm
@@
 Output: \\
-\> MapReduce object $G$ = triangles in graph: Key = $(V_i,V_j,V_k)$, Value = NULL \\
-
+\> MapReduce object $G$ = triangles in graph: Key = $(V_i,V_j,V_k)$, Value = NULL 
   \end{tabbing}
@@
 
- \caption{\it MapReduce algorithm for triangle enumeration, using 2 MapReduce objects $G$ and $G_0$}
+ \caption{\it MapReduce algorithm for triangle enumeration, using two MapReduce objects $G$ and $G_0$}
 
@@
 assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
-exposition also follows the connected-component algorithm presented
+exposition follows the connected-component algorithm presented
 in~\cite{Cohen09}, with the addition of logic that load-balances data
@@
 Output: \\
-\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
-
+\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ 
   \end{tabbing}
@@
 The third {\it map()} inverts the vertex/zone KV pairs to become
-zone/vertex KV pairs.  The partitioning logic described in
+zone/vertex KV pairs.  The partitioning logic in
 Figure~\ref{fig:cc} for this step (and subsequent ones) is described
@@
 ($Z_i^+$,$Z_{winner}$) instead of as ($Z_i$,$Z_{winner}$).  In this
-case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once
+case, ($Z_i$,$Z_{winner}$) is emitted not once, but $P+1$ times, once
 for each processor, and once as if $Z_i$ had not been partitioned (for
@@
 Output: \\
-\> MapReduce object $V$ = maximal independent set vertices: Key = $V_i$, Value = NULL \\
-
+\> MapReduce object $V$ = maximal independent set vertices: Key = $V_i$, Value = NULL 
   \end{tabbing}
@@
 identify winning vertices.  Vertices adjacent to winners are flagged
-as losers, and all edges which contain a loser vertex are marked for
+as losers, and all edges that contain a loser vertex are marked for
 removal at the start of the next iteration.  Vertices with all their
@@
 flag.  If $V_i$ is not an overall winner, its neighbor vertices are
-emitted without the winner flag on $V_i$..  The second {\it collate()}
+emitted without the winner flag on $V_i$.  The second {\it collate()}
 collects this new information for each vertex.
@@
 Ford62}, every edge of the graph is visited in each iteration and
-vertex distance are updated.  A MapReduce formulation of this
+vertex distances are updated.  A MapReduce formulation of this
 algorithm, described in Figure~\ref{fig:sssp}, begins with a graph $E$
@@
 \> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
-\> \> Key = $V_i$, Value = $d_i$ \\
-
+\> \> Key = $V_i$, Value = $d_i$ 
   \end{tabbing}
@@
 $U$ to store these updates.  Only the updates in $U$ are communicated;
-which aggregates them to the same processor that permanently stores
+they are aggregated to the same processor that permanently stores
 the corresponding vertex and its edges.  Thus, the total amount of

### 2010-11-22 paper/intro.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 this paper and are part of the current downloadable version; see
-Section ~\ref{sec:thanks} for details.
+Section~\ref{sec:thanks} for details.
 
@@
 flat files.  It can use local per-processor disks, or a parallel file
-system, if available, but these typically provide no data redundancy.
+system, if available, but these typically do not provide data redundancy.
 
@@
 results for these algorithms running on a parallel cluster for graphs
-ranging in size from 8 million to 2 billion edges; we highlight the
+ranging in size from 8 million to 2 billion edges. We highlight the
 performance and complexity trade-offs of a MapReduce approach versus

### 2010-11-22 paper/mr.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 owning processor rank.  Alternatively, the user can provide a hash
-function which converts the key into a processor rank.  Each processor
+function that converts the key into a processor rank.  Each processor
 then sends each of its KV pairs to the owning processor.
@@
 collate()}.  Likewise, data can be partitioned across multiple
-MapReduce (MR) objects, each with their own set of KV pairs.  For
-example, as will be illustrated in Section ~\ref{sec:graph}, one MR
+MapReduce (MR) objects, each with its own set of KV pairs.  For
+example, as will be illustrated in Section~\ref{sec:graph}, one MR
 object can store the edges of a graph, and another its vertices.  KV

### 2010-11-22 paper/outcore.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 
-Performing an our-of-core {\it collate()} operation is more complex.
+Performing an out-of-core {\it collate()} operation is more complex.
 Recall that the operation occurs in two stages.  First, keys are
@@
 are distributed unevenly so that this limit is exceeded, the
-all-to-all communication is performed in smaller, multiple passes
+all-to-all communication is performed in multiple smaller passes
 until the full page contributed by every processor has been
@@
 \includegraphics[width=\textwidth,angle=-90]{fig_collate2.pdf}
-\caption{\it Multi-pass algorithm for converting KV data to KMV data.
+\caption{\it Multi-pass algorithm for converting KV data to KMV data within
+a processor.
 The vertical lines represent out-of-core data sets.  KV is key/value

### 2010-11-22 paper/paper.tex
Minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@673 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 library was to enable graph algorithms to be written as MapReduce
-operations, allowing processing of terabyte-scale data sets.  We
+operations, allowing processing of terabyte-scale data sets on 
+traditional MPI-based clusters.  We
 outline MapReduce versions of several such algorithms: vertex ranking

### 2010-11-22 paper/graph.tex
A few notation fixes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@674 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
-\> MapReduce object $y$ = vector, initially empty \\
-\> User-defined parameter $\alpha$ \\
+\> MapReduce object $y$ = vector, initially empty; user-defined parameter $\alpha$ \\
 \\
@@
 \> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
-\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
+\> MapReduce object $V$ = zone assignments of vertices: Key = $V_i$, Value = $Z_i=V_i$ \\
 \> MapReduce object $W$ = workspace
@@
     \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
-\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to $V$ \\
+\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to data in $V$ \\
     \> \> if $Z_i$ is not partitioned: \\
@@
        \> \> partition $Z_{new}$ if number of $V >$ threshold \\
-       \> \> for each $V$ in MultiValue: \\
+       \> \> for each $V_i$ in MultiValue: \\
        \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
@@
 Map $E$: Assign random values consistently to each vertex $V_i$ and $V_j$ in $E$ \\
-Clone $E$: Convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
+Clone $E$: Convert KV pairs in $E$ directly to KMV pairs with MultiValue = NULL \\
 \\
@@
        \> \> Input: Key = $E_{ij}$, MultiValue = NULL or LOSER \\
-       \> \> if multivalue = LOSER, emit nothing \\
+       \> \> if MultiValue = LOSER, emit nothing \\
        \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
@@
        \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
+       \> \> if all $V_j$ in MultiValue are flagged with WINNER, $V_i$ is a WINNER \\
        \> \> \> for each $V_j$ in MultiValue: \\
@@
        \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
+       \> \> if any $V_j$ in MultiValue is flagged with WINNER, $V_i$ is a LOSER \\
        \> \> \> for each $V_j$ in MultiValue: \\
@@
        \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
-       \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
+       \> \> if all $V_j$ in MultiValue are flagged with LOSER, $V_i$ is a WINNER \\
        \> \> \> Output: to $V$: Key = $V_i$, Value = NULL \\
@@
 \> $done$ = 1 \\
-\> Map $V$ using $E$ as input: Add edge weights in $E$ to $V$ \\
+\> Map $V$ using $E$ as input: Add edge weights in $E$ to data in $V$ \\
 \> \> Input: Key = $E_{ij}$, Value = $w_{ij}$ \\
@@
 \> Collate $V$: Vertex is key \\
-\> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
+\> Reduce $V$:  Loop over edges of vertex to update distance from $V_s$ \\
 \> \> Input: Key = $V_i$, MultiValue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\

### 2010-11-22 paper/results.tex
Clarified the conditions for the detailed timing tests. Added HADOOP line count. Other minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@675 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
-0.05)$ and generated 8 edges per vertex (on average).  These
+0.05)$ and generated eight edges per vertex (on average).  These
 parameters create a highly skewed degree distribution, as indicated by
@@
 time spent in each step of Figure~\ref{fig:pr2} for the MR-MPI
-implementation and the equivalent Trilinos implementation.  For both
+implementation and the equivalent Trilinos implementation on 64 processors.  For both
 implementations, the matrix-vector multiplication (Step 2) is the most
@@
 \caption{Detailed execution times (in seconds) per PageRank iteration
+on 64 processors
 for each stage of the algorithm
@@
 To understand where MR-MPI spends its effort during the PageRank
-computation, we break down the operation times further for the most
-expensive Step 2 of the PageRank computation; the results are in
+computation, we break down the operation times further for Step 2, the most
+expensive step of the PageRank computation; the results are in
 Table~\ref{table:prdetail2}.  In the table, we divide the {\it
@@
 experimentation is needed to determine whether reorganizing the
-MapReduce data, as described in Section~\ref{subsec:graph_pagerank}.
+MapReduce data, as described in Section~\ref{subsec:graph_pagerank},
 would reduce the execution time of Step 2.  Organizing the data such
@@
 \caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
-PageRank algorithm in Figure~\ref{fig:pr2}.}
+PageRank algorithm in Figure~\ref{fig:pr2}; experiments used 64 processors.}
 \label{table:prdetail2}
@@
 \caption{Percentage of execution time in each operation of Step 5 of
-the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}.}
+the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}; experiments
+used 64 processors.}
 \label{table:prdetail5}
@@
 \ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
-must implement the parallelism explicitly whereas the codes of this
+must implement the parallelism explicitly, whereas the codes of this
 section relegate that task to the MR-MPI library.  A similar all-MPI
@@
 \ref{subsec:results_pagerank}) required only 153 lines of code, since
-an entire matrix-vector multiply can be invoked by a single call to
+an entire matrix-vector multiplication can be invoked by a single call to
 Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
 implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
-required XXX lines of Java code.
+required 697 lines of Java code.
 

### 2010-11-23 paper/Response.tex
Minor addition
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@676 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@678 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 sparked some new ideas for improving our algorithms.  Below we list
-the Reviewers' comments verbatim.  Our responses are shown in red.
+the Reviewers' comments; our responses are shown in red.
 
@@
 of data that each processor performs on its local data.  Thus there is
-no inter-processor communication involved.}
+no inter-processor communication involved.  We added text to the caption
+to clarify that all operations are done on-processor.}
 
@@
 
-\revise{We added some of these comparison points to the final
-``Lessons Learned'' section.  In particular we highlight the
-relatively small line count of the MapReduce algorithms and the
-natural out-of-core ability it provides, as compared for example to
-Trilinos.  The comment with respect to MR inefficiencies is addressed
-below.}
+\revise{We added some of these comparison points to a new section 5.7
+  and to the final ``Lessons Learned'' section.  In particular we
+  highlight the relatively small line count of the MapReduce
+  algorithms and the natural out-of-core ability it provides, as
+  compared for example to Trilinos.  The comment with respect to MR
+  inefficiencies is addressed below.}
 
@@
 \revise{We agree that our notation in Section 4 was deficient.  I
-couldn't understand some of the deteails when I re-read it!  So we
-significantly re-worked and standardized the notation used in Section
-4 to hopefully describe all the graph algorithms more precisely.  In
+  couldn't understand some of the details when I re-read it!  So we
+  re-worked and standardized the notation used in Section 4 to
+  hopefully describe all the graph algorithms more precisely.  In
   particular we noted which operations are performed with which

### 2010-11-29 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@677 3fc257bb-8550-0410-81b3-a5c103da83f7
More accurate line counts.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@680 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \hline
-Lines of code & 208 & 216 & 694 & 367 & 453 & 393 & 505 & 569 \\
+Lines of code & 208 & 216 & 603 & 367 & 453 & 393 & 394 & 448 \\
 Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
@@
 (non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
-\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
+\ref{fig:rmat2}.  It required 845 lines of code, mainly because it
 must implement the parallelism explicitly, whereas the codes of this
 section relegate that task to the MR-MPI library.  A similar all-MPI
-implementation of the PageRank algorithm required 1206 lines.  By
+implementation of the PageRank algorithm required 718 lines.  By
 contrast an implementation of PageRank using the Trilinos parallel
@@
 \ref{subsec:results_pagerank}) required only 153 lines of code, since
-an entire matrix-vector multiplication can be invoked by a single call to
+entire dot-products and matrix-vector multiplications can be invoked by single function calls to
 Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
 implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
-required 697 lines of Java code.
+required 697 lines of Java.
 

### 2011-01-06 src/irregular.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@700 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -15,6 +15,7 @@
 #include "stdlib.h"
 #include "string.h"
 #include "irregular.h"
+#include "mrtype.h"
 #include "memory.h"
 #include "error.h"
 
@@ -107,7 +108,7 @@ int Irregular::setup(int n, int *proclist, int *sizes, int *reorder,
   // error return if any proc's send total > INTMAX
 
   uint64_t sendtotalmax;
-  MPI_Allreduce(&sendtotal,&sendtotalmax,1,MPI_UNSIGNED_LONG,MPI_MAX,comm);
+  MPI_Allreduce(&sendtotal,&sendtotalmax,1,MRMPI_BIGINT,MPI_MAX,comm);
   if (sendtotalmax > INTMAX) {
     fraction = ((double) INTMAX) / sendtotal;
     return 0;
@@ -127,7 +128,7 @@ int Irregular::setup(int n, int *proclist, int *sizes, int *reorder,
   // error return if any proc's recv total > min(recvlimit,INTMAX)
   
   uint64_t recvtotalmax;
-  MPI_Allreduce(&recvtotal,&recvtotalmax,1,MPI_UNSIGNED_LONG,MPI_MAX,comm);
+  MPI_Allreduce(&recvtotal,&recvtotalmax,1,MRMPI_BIGINT,MPI_MAX,comm);
   if (recvtotalmax > recvlimit) {
     fraction = ((double) recvlimit) / recvtotal;
     return 0;

### 2011-01-06 src/mapreduce.cpp

@@ -20,6 +20,7 @@
 #include "sys/types.h"
 #include "sys/stat.h"
 #include "mapreduce.h"
+#include "mrtype.h"
 #include "keyvalue.h"
 #include "keymultivalue.h"
 #include "spool.h"
@@ -221,9 +222,9 @@ void MapReduce::defaults()
     error->all("Not compiled for 8-byte integers and pointers");
 
   int mpisize;
-  MPI_Type_size(MPI_UNSIGNED_LONG,&mpisize);
+  MPI_Type_size(MRMPI_BIGINT,&mpisize);
   if (mpisize != 8)
-    error->all("MPI_UNSIGNED_LONG is not 8-byte data type");
+    error->all("MRMPI_BIGINT is not 8-byte data type: edit mrtype.h");
 }
 
 /* ----------------------------------------------------------------------
@@ -324,7 +325,7 @@ uint64_t MapReduce::add(MapReduce *mr)
   stats("Add",0);
 
   uint64_t nkeyall;
-  MPI_Allreduce(&kv->nkv,&nkeyall,1,MPI_UNSIGNED_LONG,MPI_SUM,comm);
+  MPI_Allreduce(&kv->nkv,&nkeyall,1,MRMPI_BIGINT,MPI_SUM,comm);
   return nkeyall;
 }
 
@@ -488,7 +489,7 @@ uint64_t MapReduce::aggregate(int (*hash)(char *, int))
   stats("Aggregate",0);
 
   uint64_t nkeyall;
-  MPI_Allreduce(&kv->nkv,&nkeyall,1,MPI_UNSIGNED_LONG,MPI_SUM,comm);
+  MPI_Allreduce(&kv->nkv,&nkeyall,1,MRMPI_BIGINT,MPI_SUM,comm);
   return nkeyall;
 }
 
@@ -510,7 +511,7 @@ uint64_t MapReduce::broadcast(int root)
   if (nprocs == 1) {
     stats("Broadcast",0);
     uint64_t nkeyall;
-    MPI_Allreduce(&kv->nkv,&nkeyall,1,MPI_UNSIGNED_LONG,MPI_SUM,comm);
[…]

### 2011-01-06 src/spool.cpp

@@ -226,7 +226,7 @@ void Spool::write_page()
   int nwrite = fwrite(page,pages[npage].filesize,1,fp);
   mr->wsize += pages[npage].filesize;
 
-  if (nwrite != 1) {
+  if (nwrite != 1 && pages[npage].filesize) {
     char str[128];
     sprintf(str,"Bad SP fwrite: %d %u",nwrite,pages[npage].filesize);
     error->warning(str);
@@ -247,7 +247,7 @@ void Spool::read_page(int ipage)
   int nread = fread(page,pages[ipage].filesize,1,fp);
   mr->rsize += pages[ipage].filesize;
 
-  if (nread != 1 || ferror(fp)) {
+  if ((nread != 1 && pages[ipage].filesize) || ferror(fp)) {
     char str[128];
     sprintf(str,"Bad SP fread: %d %u",nread,pages[ipage].filesize);
     error->warning(str);

### 2011-01-06 mpistubs/mpi.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@701 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -82,8 +82,8 @@ void MPI_Type_size(MPI_Datatype datatype, int *size)
   else if (datatype == MPI_DOUBLE) *size =sizeof(double);
   else if (datatype == MPI_CHAR) *size = sizeof(char);
   else if (datatype == MPI_BYTE) *size = sizeof(char);
-  else if (datatype == MPI_DOUBLE_INT) *size = sizeof(double_int);
   else if (datatype == MPI_UNSIGNED_LONG) *size = sizeof(uint64_t);
+  else if (datatype == MPI_DOUBLE_INT) *size = sizeof(double_int);
 }
 
 /* ---------------------------------------------------------------------- */
@@ -250,8 +250,8 @@ void MPI_Allreduce(void *sendbuf, void *recvbuf, int count,
   else if (datatype == MPI_DOUBLE) n = count*sizeof(double);
   else if (datatype == MPI_CHAR) n = count*sizeof(char);
   else if (datatype == MPI_BYTE) n = count*sizeof(char);
-  else if (datatype == MPI_DOUBLE_INT) n = count*sizeof(double_int);
   else if (datatype == MPI_UNSIGNED_LONG) n = count*sizeof(uint64_t);
+  else if (datatype == MPI_DOUBLE_INT) n = count*sizeof(double_int);
 
   memcpy(recvbuf,sendbuf,n);
 }
@@ -267,8 +267,8 @@ void MPI_Scan(void *sendbuf, void *recvbuf, int count,
   else if (datatype == MPI_DOUBLE) n = count*sizeof(double);
   else if (datatype == MPI_CHAR) n = count*sizeof(char);
   else if (datatype == MPI_BYTE) n = count*sizeof(char);
-  else if (datatype == MPI_DOUBLE_INT) n = count*sizeof(double_int);
   else if (datatype == MPI_UNSIGNED_LONG) n = count*sizeof(uint64_t);
+  else if (datatype == MPI_DOUBLE_INT) n = count*sizeof(double_int);
 
   memcpy(recvbuf,sendbuf,n);
 }
@@ -287,8 +287,8 @@ void MPI_Allgather(void *sendbuf, int sendcount, MPI_Datatype sendtype,
   else if (sendtype == MPI_DOUBLE) n = sendcount*sizeof(double);
[…]

### 2011-01-06 mpistubs/mpi.h

@@ -23,8 +23,9 @@
 #define MPI_DOUBLE 3
 #define MPI_CHAR 4
 #define MPI_BYTE 5
-#define MPI_DOUBLE_INT 6
-#define MPI_UNSIGNED_LONG 7
+#define MPI_UNSIGNED_LONG 6
+#define MPI_UNSIGNED_LONG_LONG 6
+#define MPI_DOUBLE_INT 7
 
 #define MPI_SUM 1
 #define MPI_MAX 2

### 2011-01-06 programs/blockmacros.h

@@ -1,12 +1,12 @@
 // Code to simplify use of multi-page KMVs
 // Karen Devine, March 2010
 
-#ifndef _BLOCKMACROS_H
-#define _BLOCKMACROS_H
+#ifndef _BLOCKMACROS_HPP
+#define _BLOCKMACROS_HPP
 
-// Macros to loop over blocks when reduce multivalues may span more than 1 block
-// Use CHECK_FOR_BLOCKS initially to get # of blocks in the multivalue
-// Enclose code for each block between BEGIN_BLOCK_LOOP and END_BLOCK_LOOP
+// macros to loop over blocks when reduce multivalues may span more than 1 block
+// use CHECK_FOR_BLOCKS initially to get # of blocks in the multivalue
+// enclose code for each block between BEGIN_BLOCK_LOOP and END_BLOCK_LOOP
 // NOTE: DO NOT put a semicolon afer these macros
 
 #define CHECK_FOR_BLOCKS(multivalue, valuebytes, nvalues, totalnvalues)  \
@@ -25,7 +25,5 @@
                                              &(multivalue),&(valuebytes)); 
 
 #define END_BLOCK_LOOP } 
-
-#define BREAK_BLOCK_LOOP break;
 
 #endif

### 2011-01-06 programs/cc_driver.cpp

@@ -84,7 +84,7 @@ int main(int narg, char **args)
 
   if (me == 0)
     printf("CC find: %g secs, %d iterations on RMAT with "
-	   "%u verts, %u edges on %d procs\n",time,niterate,
+	   "%lu verts, %u edges on %d procs\n",time,niterate,
 	   in.nvert,in.nedge,nprocs);
 
   // CC stats

### 2011-01-06 programs/cc_stats.cpp

@@ -86,6 +86,6 @@ void CCStats::map_print(uint64_t itask, char *key, int keybytes,
 {
   uint64_t nvert = *(uint64_t *) key;
   uint64_t ncc = *(uint64_t *) value;
-  printf("  %u CCs with %u vertices\n",ncc,nvert);
+  printf("  %lu CCs with %lu vertices\n",ncc,nvert);
 }
 

### 2011-01-06 programs/luby_driver.cpp

@@ -81,8 +81,8 @@ int main(int narg, char **args)
   double time = luby.run(mre,mrv,niterate,nset);
 
   if (me == 0)
-    printf("MIS find: %g secs, %u vertices, %d iterations on RMAT with "
-	   "%u verts, %u edges on %d procs\n",time,nset,niterate,
+    printf("MIS find: %g secs, %lu vertices, %d iterations on RMAT with "
+	   "%lu verts, %lu edges on %d procs\n",time,nset,niterate,
 	   in.nvert,newedge,nprocs);
 
   delete mre;

### 2011-01-06 programs/matrix_stats.cpp

@@ -38,12 +38,12 @@ double MatrixStats::run(MapReduce *mr, uint64_t nrow)
   int me;
   MPI_Comm_rank(MPI_COMM_WORLD,&me);
   if (me == 0)
-    printf("Stats for matrix with %u nrows, %u non-zeroes:\n",nrow,nonzero);
+    printf("Stats for matrix with %lu nrows, %lu non-zeroes:\n",nrow,nonzero);
   uint64_t sum = 0;
   mr->verbosity = 0;
   mr->timer = 0;
   mr->map(mr,map_print,&sum);
-  if (me == 0) printf("  %u rows with 0 elements\n",nrow-sum);
+  if (me == 0) printf("  %lu rows with 0 elements\n",nrow-sum);
 
   MPI_Barrier(MPI_COMM_WORLD);
   double tstop = MPI_Wtime();
@@ -100,7 +100,7 @@ void MatrixStats::map_print(uint64_t itask, char *key, int keybytes,
 {
   uint64_t ncount = *(uint64_t *) key;
   uint64_t nrow = *(uint64_t *) value;
-  printf("  %u rows with %u elements\n",nrow,ncount);
+  printf("  %lu rows with %lu elements\n",nrow,ncount);
   uint64_t *sum = (uint64_t *) ptr;
   *sum += nrow;
 }

### 2011-01-06 programs/rmat_driver.cpp

@@ -56,7 +56,7 @@ int main(int narg, char **args)
 
   if (me == 0)
     printf("RMAT creation: %g secs, %d iterations with "
-	   "%u verts, %u edges on %d procs\n",time,niterate,
+	   "%lu verts, %lu edges on %d procs\n",time,niterate,
 	   in.nvert,in.nedge,nprocs);
 
   // RMAT matrix stats

### 2011-01-06 programs/sgi_check.cpp

@@ -75,7 +75,7 @@ int main(int narg, char **args)
   combine(nvert,verts,nedge,edges);
 
   //for (int i = 0; i < nedge; i++)
-  //  printf("EDGE %u %u: %d %d %d\n",
+  //  printf("EDGE %lu %lu: %d %d %d\n",
   //	   edges[i].vi,edges[i].vj,edges[i].fij,edges[i].wi,edges[i].wj);
 
   // recursive find of each SGI match
@@ -96,7 +96,7 @@ int main(int narg, char **args)
 
   // stats
 
-  printf("SGI find: %g secs, %u SG in big graph with "
+  printf("SGI find: %g secs, %d SG in big graph with "
 	 "%d verts, %d edges\n",time,nsgi,nvert,nedge);
 
   // clean up
@@ -122,7 +122,7 @@ void find(int itour, int nvert, VERT *verts, int nedge, EDGE *edges,
 
   if (itour == ntour) {
     for (int i = 0; i < ntour; i++)
-      fprintf(fp,"%u ",tour[i]);
+      fprintf(fp,"%lu ",tour[i]);
     fprintf(fp,"\n");
     nsgi++;
     return;

### 2011-01-06 programs/sgi_driver.cpp

@@ -108,9 +108,6 @@ int main(int narg, char **args)
   mre->verbosity = 0;
   mre->timer = 0;
 
-  //mrv->memsize = 1;
-  //mre->memsize = 1;
-
   // populate MRV,MVE with labeled graph
   // MRE = Eij : Fij
   // MRV = Vi : Wi
@@ -189,6 +186,9 @@ int main(int narg, char **args)
     mry->verbosity = 0;
     mry->timer = 0;
 
+    //mrs->timer = mrx->timer = mry->timer = 1;
+    //mrs->verbosity = mrx->verbosity = mry->verbosity = 1;
+
     //mrs->memsize = 1;
     //mrx->memsize = 1;
     //mry->memsize = 1;
@@ -198,8 +198,8 @@ int main(int narg, char **args)
     double time = sgi.run(mrv,mre,mrs,mrx,mry,nsgi);
 
     if (me == 0)
-      printf("SGI enumerate: %g secs, %u SG in big graph with "
-	     "%u verts, %u edges on %d procs\n",time,nsgi,
+      printf("SGI enumerate: %g secs, %lu SG in big graph with "
+	     "%lu verts, %lu edges on %d procs\n",time,nsgi,
 	     in.nvert,in.nedge,nprocs);
 
     delete mrs;
@@ -207,30 +207,45 @@ int main(int narg, char **args)
     delete mry;
 
   } else {
-    MapReduce *mrs = new MapReduce(MPI_COMM_WORLD);
-    mrs->verbosity = 0;
-    mrs->timer = 0;
-    MapReduce *mrx = new MapReduce(MPI_COMM_WORLD);
-    mrx->verbosity = 0;
-    mrx->timer = 0;
-    MapReduce *mry = new MapReduce(MPI_COMM_WORLD);
-    mry->verbosity = 0;
-    mry->timer = 0;
+    MapReduce **mro = new MapReduce*[in.ntour-1];
+    for (int k = 0; k < in.ntour-1; k++) {
+      mro[k] = new MapReduce(MPI_COMM_WORLD);
+      mro[k]->verbosity = 0;
+      mro[k]->timer = 0;
+      mro[k]->memsize = 8;
+    }
+    MapReduce **mri = new MapReduce*[in.ntour];
+    for (int k = 1; k < in.ntour; k++) {
+      mri[k] = new MapReduce(MPI_COMM_WORLD);
[…]

### 2011-01-06 programs/sgi_enumerate.cpp

@@ -107,7 +107,7 @@ double SGIEnumerate::run(MapReduce *mrv, MapReduce *mre,
     // debug
     //mrs->map(mrs,tprint,this,1);
 
-    if (me == 0) printf("%u sub-graphs after edge %d\n",nsgi,k);
+    if (me == 0) printf("%lu sub-graphs after edge %d\n",nsgi,k);
   }
 
   MPI_Barrier(world);
@@ -187,8 +187,8 @@ void SGIEnumerate::map4(uint64_t itask, char *key, int keybytes,
   VERTEX *vlast = (VERTEX *) key;
   VERTEX *vlist = (VERTEX *) value;
 
-  for (int i = 0; i < ntourm1; i++) fprintf(fp,"%u ",vlist[i]);
-  fprintf(fp,"%u\n",*vlast);
+  for (int i = 0; i < ntourm1; i++) fprintf(fp,"%lu ",vlist[i]);
+  fprintf(fp,"%lu\n",*vlast);
 
   kv->add(key,keybytes,value,valuebytes);
 }
@@ -479,6 +479,7 @@ void SGIEnumerate::reduce3(char *key, int keybytes,
 }
 
 /* ---------------------------------------------------------------------- */
+/* ---------------------------------------------------------------------- */
 
 void SGIEnumerate::x1print(uint64_t itask, char *key, int keybytes, 
 			   char *value, int valuebytes, 
@@ -489,10 +490,10 @@ void SGIEnumerate::x1print(uint64_t itask, char *key, int keybytes,
     X1VALUE *x = (X1VALUE *) value;
     VERTEX vj = x->vj;
     int fij = x->fij;
-    printf("MR X1a: %d %d: %u %u %d\n",keybytes,valuebytes,vi,vj,fij);
+    printf("MR X1a: %d %d: %lu %lu %d\n",keybytes,valuebytes,vi,vj,fij);
   } else {
     VERTEX vj = *(VERTEX *) value;
-    printf("MR X1b: %d %d: %u %u\n",keybytes,valuebytes,vi,vj);
+    printf("MR X1b: %d %d: %lu %lu\n",keybytes,valuebytes,vi,vj);
   }
 }
 
@@ -509,10 +510,11 @@ void SGIEnumerate::x2print(uint64_t itask, char *key, int keybytes,
     VERTEX vi = x->vi;
     int wi = x->wi;
[…]

### 2011-01-06 programs/sgi_sample.cpp

@@ -49,80 +49,125 @@ SGISample::SGISample(int msample_in, int ntour_in,
 /* ---------------------------------------------------------------------- */
 
 double SGISample::run(MapReduce *mrv, MapReduce *mre,
-		      MapReduce *mrs, MapReduce *mrx, MapReduce *mry,
+		      MapReduce **mro_in, MapReduce **mri_in,
+		      MapReduce *mreprime_in, MapReduce *mrc_in,
 		      uint64_t &nsgi)
 {
   MPI_Barrier(world);
   double tstart = MPI_Wtime();
 
-  // pre-calculate X = Vi : Vj Wi Wj Fij, two for each edge in big graph
+  mro = mro_in;
+  mri = mri_in;
+  mreprime = mreprime_in;
+  mrc = mrc_in;
+
+  // pre-calculate X = Eij : Wi Wj Fij, one for each edge in big graph
 
   mrv->aggregate(NULL);
-  mrx->map(mre,map1,this);
-  mrx->aggregate(NULL);
-  mrx->add(mrv);
+  mrc->map(mre,map1,this);
+  mrc->aggregate(NULL);
+  mrc->add(mrv);
+  mrc->convert();
+  mrc->reduce(reduce1a,this);
+  mrc->aggregate(NULL);
+  mrc->add(mrv);
+  mrc->convert();
+  mrc->reduce(reduce1b,this);
 
   // debug
-  //mrx->map(mrx,x1print,this,1);
+  //mrc->map(mrc,x3print,this,1);
 
-  mrx->convert();
-  mrx->reduce(reduce1a,this);
-  mrx->aggregate(NULL);
-  mrx->add(mrv);
+  // create outbound and inbound BPG (bipartite graph) layers
+  // create mro and mri lists of MapReduce objects
 
-  // debug
-  //mrx->map(mrx,x2print,this,1);
+  for (int k = 0; k < ntour-1; k++) mro[k]->open();
+  for (int k = 1; k < ntour; k++) mri[k]->open();
+  
+  mrc->map(mrc,map2,this);
 
-  mrx->convert();
-  mrx->reduce(reduce1b,this);
-  mrx->aggregate(NULL);
-
-  // debug
-  //mrx->map(mrx,x3print,this,1);
-
-  // initial S = Vi : NULL
-
-  mrs->map(mrv,map2,this);
-  mrs->aggregate(NULL);
-
-  // debug
-  //mrs->map(mrs,sprint,this,1);
[…]

### 2011-01-06 programs/sgi_sample.h

@@ -14,8 +14,9 @@ class SGISample {
  public:
   SGISample(int, int, int *, int *, int *, MPI_Comm);
   double run(MAPREDUCE_NS::MapReduce *, MAPREDUCE_NS::MapReduce *, 
+	     MAPREDUCE_NS::MapReduce **, MAPREDUCE_NS::MapReduce **, 
 	     MAPREDUCE_NS::MapReduce *, MAPREDUCE_NS::MapReduce *, 
-	     MAPREDUCE_NS::MapReduce *, uint64_t &);
+	     uint64_t &);
 
  private:
   int me;
@@ -26,6 +27,7 @@ class SGISample {
   char *outfile;
   FILE *fp;
 
+  typedef uint64_t ULONG;
   typedef uint64_t VERTEX;
   typedef int LABEL;
   typedef struct {
@@ -40,10 +42,21 @@ class SGISample {
     LABEL wi,fij;
   } X2VALUE;
   typedef struct {
-    VERTEX vj;
     LABEL wi,wj;
     LABEL fij;
   } X3VALUE;
+  typedef struct {
+    VERTEX vi,vj;
+    LABEL k;
+  } X4VALUE;
+  typedef struct {
+    ULONG count,dummy;
+  } COUNT;
+
+  MAPREDUCE_NS::MapReduce **mro,**mri;
+  MAPREDUCE_NS::MapReduce *mreprime,*mrc;
+  int eflag,kindex;
+  ULONG tlocal;
 
   static void map1(uint64_t, char *, int, char *, int, 
 		   MAPREDUCE_NS::KeyValue *, void *);
@@ -51,13 +64,25 @@ class SGISample {
 		   MAPREDUCE_NS::KeyValue *, void *);
   static void map3(uint64_t, char *, int, char *, int,
 		   MAPREDUCE_NS::KeyValue *, void *);
-  static void map4(uint64_t, char *, int, char *, int,
+  static void map5(uint64_t, char *, int, char *, int,
 		   MAPREDUCE_NS::KeyValue *, void *);
+
   static void reduce1a(char *, int, char *, int, int *,
 		      MAPREDUCE_NS::KeyValue *, void *);
   static void reduce1b(char *, int, char *, int, int *,
 		      MAPREDUCE_NS::KeyValue *, void *);
   static void reduce3(char *, int, char *, int, int *,
+		      MAPREDUCE_NS::KeyValue *, void *);
+  static void reduce4a(char *, int, char *, int, int *,
[…]

### 2011-01-06 programs/stress_test1.cpp

@@ -311,8 +311,8 @@ void checksort(uint64_t itask, char *key, int keybytes,
     int i1 = atoi(old);
     int i2 = atoi(value);
     if (i1 > i2)
-      printf("BAD SORT %u %d: %d %d\n",ME,itask,i1,i2);
-    //if (ME == 0) printf("AAA %u %s %s\n",itask,old,value);
+      printf("BAD SORT %d %d: %d %d\n",ME,itask,i1,i2);
+    //if (ME == 0) printf("AAA %lu %s %s\n",itask,old,value);
   }
   strcpy(old,value);
   //printf("AAA %d %d %d %s %d\n",itask,*(int *)key,keybytes,value,valuebytes);

### 2011-01-06 programs/tri_driver.cpp

@@ -80,8 +80,8 @@ int main(int narg, char **args)
   double time = tri.run(mre,mrt,ntri);
 
   if (me == 0)
-    printf("Triangle find: %g secs, %u triangles in RMAT with "
-	   "%u verts, %u edges on %d procs\n",time,ntri,
+    printf("Triangle find: %g secs, %lu triangles in RMAT with "
+	   "%lu verts, %lu edges on %d procs\n",time,ntri,
 	   in.nvert,newedge,nprocs);
 
   delete mre;

### 2011-01-06 python/mrmpi.py

@@ -50,8 +50,8 @@ class mrmpi:
     MAPFUNC = CFUNCTYPE(c_void_p,c_int,c_void_p,c_void_p)
     self.map_def = MAPFUNC(self.map_callback)
     
-    MAP_FILE_LIST_FUNC = CFUNCTYPE(c_void_p,c_int,c_char_p,c_void_p,c_void_p)
-    self.map_file_list_def = MAP_FILE_LIST_FUNC(self.map_file_list_callback)
+    MAP_FILE_FUNC = CFUNCTYPE(c_void_p,c_int,c_char_p,c_void_p,c_void_p)
+    self.map_file_def = MAP_FILE_FUNC(self.map_file_callback)
     
     MAP_FILE_STR_FUNC = CFUNCTYPE(c_void_p,c_int,POINTER(c_char),c_int,
                                  c_void_p,c_void_p)
@@ -183,13 +183,24 @@ class mrmpi:
     self.map_argcount = map.func_code.co_argcount
     self.map_ptr = ptr
     if not addflag:
-      n = self.lib.MR_map_file_list(self.mr,file,self.map_file_list_def,None)
+      n = self.lib.MR_map_file_list(self.mr,file,self.map_file_def,None)
     else:
-      n = self.lib.MR_map_file_list_add(self.mr,file,self.map_file_list_def,
+      n = self.lib.MR_map_file_list_add(self.mr,file,self.map_file_def,
                                         None,addflag)
     return n
 
-  def map_file_list_callback(self,itask,file,kv,dummy):
+  def map_dir(self,dir,oneflag,map,ptr=None,addflag=0):
+    self.map_caller = map
+    self.map_argcount = map.func_code.co_argcount
+    self.map_ptr = ptr
+    if not addflag:
+      n = self.lib.MR_map_dir(self.mr,dir,oneflag,self.map_file_def,None)
+    else:
+      n = self.lib.MR_map_dir_add(self.mr,dir,oneflag,self.map_file_def,
+                                  None,addflag)
+    return n
+
+  def map_file_callback(self,itask,file,kv,dummy):
     self.kv = kv
     if self.map_argcount == 3: self.map_caller(itask,file,self)
     else: self.map_caller(itask,file,self,self.map_ptr)

### 2011-01-06 src/cmapreduce.cpp

@@ -158,6 +158,26 @@ uint64_t MR_map_file_list_add(void *MRptr, char *file,
   return mr->map(file,appmap,APPptr,addflag);
 }
 
+uint64_t MR_map_dir(void *MRptr, char *dir, int oneflag,
+		    void (*mymap)(int, char *, void *, void *),
+		    void *APPptr)
+{
+  typedef void (MapFunc)(int, char *, KeyValue *, void *);
+  MapReduce *mr = (MapReduce *) MRptr;
+  MapFunc *appmap = (MapFunc *) mymap;
+  return mr->map(dir,oneflag,appmap,APPptr);
+}
+
+uint64_t MR_map_dir_add(void *MRptr, char *dir, int oneflag,
+			      void (*mymap)(int, char *, void *, void *),
+			      void *APPptr, int addflag)
+{
+  typedef void (MapFunc)(int, char *, KeyValue *, void *);
+  MapReduce *mr = (MapReduce *) MRptr;
+  MapFunc *appmap = (MapFunc *) mymap;
+  return mr->map(dir,oneflag,appmap,APPptr,addflag);
+}
+
 uint64_t MR_map_file_char(void *MRptr, int nmap, int nfiles, char **files,
 			  char sepchar, int delta,
 			  void (*mymap)(int, char *, int, void *, void *),

### 2011-01-06 src/cmapreduce.h

@@ -56,6 +56,14 @@ uint64_t MR_map_file_list_add(void *MRptr, char *file,
 			      void (*mymap)(int, char *, 
 					    void *KVptr, void *APPptr),
 			      void *APPptr, int addflag);
+uint64_t MR_map_dir(void *MRptr, char *dir, int oneflag,
+		    void (*mymap)(int, char *, 
+				  void *KVptr, void *APPptr),
+		    void *APPptr);
+uint64_t MR_map_dir_add(void *MRptr, char *dir, int oneflag,
+			void (*mymap)(int, char *, 
+				      void *KVptr, void *APPptr),
+			void *APPptr, int addflag);
 uint64_t MR_map_file_char(void *MRptr, int nmap, int nfiles, char **files,
 			  char sepchar, int delta,
 			  void (*mymap)(int, char *, int, 

### 2011-01-06 src/mapreduce.cpp

@@ -1149,6 +1149,162 @@ uint64_t MapReduce::map(char *file,
 }
 
 /* ----------------------------------------------------------------------
+   create a KV via a parallel map operation for list of files in dir
+   make one call to appmap() for each file in dir
+   if oneflag = 1, all processors process all files in single dir
+     mapstyle determines how tasks are partitioned to processors
+   if oneflag = 0, each processor reads own dir, processes those files
+------------------------------------------------------------------------- */
+
+uint64_t MapReduce::map(char *file, int oneflag,
+			void (*appmap)(int, char *, KeyValue *, void *),
+			void *appptr, int addflag)
+{
+  int n;
+  char line[MAXLINE];
+  MPI_Status status;
+
+  if (timer) start_timer();
+  if (verbosity) file_stats(0);
+
+  if (!allocated) allocate();
+  if (kmv) myfree(kmv->memtag);
+  delete kmv;
+  kmv = NULL;
+
+  if (addflag == 0) {
+    if (kv) myfree(kv->memtag);
+    delete kv;
+    kv = new KeyValue(this,kalign,valign,memory,error,comm);
+    kv->set_page();
+  } else if (kv == NULL) {
+    kv = new KeyValue(this,kalign,valign,memory,error,comm);
+    kv->set_page();
+  } else {
+    kv->append();
+  }
+
+  // open file and extract filenames
+  // bcast each filename to all procs
+  // trim whitespace from beginning and end of filename
+
+  int nmap = 0;
+  int maxfiles = 0;
+  char **files = NULL;
+  FILE *fp;
+
+  if (me == 0) {
+    fp = fopen(file,"r");
+    if (fp == NULL) error->one("Could not open file of file names");
+  }
+
+  while (1) {
+    if (me == 0) {
+      if (fgets(line,MAXLINE,fp) == NULL) n = 0;
+      else n = strlen(line) + 1;
+    }
+    MPI_Bcast(&n,1,MPI_INT,0,comm);
+    if (n == 0) {
+      if (me == 0) fclose(fp);
+      break;
+    }
+
+    MPI_Bcast(line,n,MPI_CHAR,0,comm);
+
+    char *ptr = line;
+    while (isspace(*ptr)) ptr++;
[…]

### 2011-01-06 src/mapreduce.h

@@ -77,6 +77,8 @@ class MapReduce {
 	       void *, int addflag = 0);
   uint64_t map(char *, void (*)(int, char *, class KeyValue *, void *),
 	       void *, int addflag = 0);
+  uint64_t map(char *, int, void (*)(int, char *, class KeyValue *, void *),
+	       void *, int addflag = 0);
   uint64_t map(int, int, char **, char, int, 
 	       void (*)(int, char *, int, class KeyValue *, void *),
 	       void *, int addflag = 0);

### 2011-01-06 src/mapreduce.cpp
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@702 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -19,6 +19,7 @@
 #include "stdint.h"
 #include "sys/types.h"
 #include "sys/stat.h"
+#include "dirent.h"
 #include "mapreduce.h"
 #include "mrtype.h"
 #include "keyvalue.h"
@@ -1156,7 +1157,7 @@ uint64_t MapReduce::map(char *file,
    if oneflag = 0, each processor reads own dir, processes those files
 ------------------------------------------------------------------------- */
 
-uint64_t MapReduce::map(char *file, int oneflag,
+uint64_t MapReduce::map(char *dir, int oneflag,
 			void (*appmap)(int, char *, KeyValue *, void *),
 			void *appptr, int addflag)
 {
@@ -1184,58 +1185,59 @@ uint64_t MapReduce::map(char *file, int oneflag,
     kv->append();
   }
 
-  // open file and extract filenames
-  // bcast each filename to all procs
-  // trim whitespace from beginning and end of filename
+  // open dir and read filenames
+  // if oneflag = 0, each proc reads filenames from own dir
+  // if oneflag = 1, only proc 0 reads filenames, bcast list to all procs
 
   int nmap = 0;
   int maxfiles = 0;
   char **files = NULL;
-  FILE *fp;
 
-  if (me == 0) {
-    fp = fopen(file,"r");
-    if (fp == NULL) error->one("Could not open file of file names");
+  if (oneflag == 0 || me == 0) {
+    struct dirent *ep;
+    DIR *dp = opendir(dir);
+    if (dp == NULL) error->one("Cannot open dir to search for files");
+    while (ep = readdir(dp)) {
+      if (nmap == maxfiles) {
+	maxfiles += FILECHUNK;
+	files = (char **)
+	  memory->srealloc(files,maxfiles*sizeof(char *),"MR:files");
+      }
+      n = strlen(ep->d_name) + 1;
+      files[nmap] = new char[n];
+      strcpy(files[nmap],ep->d_name);
+      nmap++;
+    }
+    closedir(dp);
   }
 
-  while (1) {
-    if (me == 0) {
-      if (fgets(line,MAXLINE,fp) == NULL) n = 0;
-      else n = strlen(line) + 1;
-    }
-    MPI_Bcast(&n,1,MPI_INT,0,comm);
-    if (n == 0) {
-      if (me == 0) fclose(fp);
[…]

### 2011-01-06 python/mrmpi.py
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@704 3fc257bb-8550-0410-81b3-a5c103da83f7

@@ -53,9 +53,9 @@ class mrmpi:
     MAP_FILE_FUNC = CFUNCTYPE(c_void_p,c_int,c_char_p,c_void_p,c_void_p)
     self.map_file_def = MAP_FILE_FUNC(self.map_file_callback)
     
-    MAP_FILE_STR_FUNC = CFUNCTYPE(c_void_p,c_int,POINTER(c_char),c_int,
-                                 c_void_p,c_void_p)
-    self.map_file_str_def = MAP_FILE_STR_FUNC(self.map_file_str_callback)
+    MAP_STR_FUNC = CFUNCTYPE(c_void_p,c_int,POINTER(c_char),c_int,
+                             c_void_p,c_void_p)
+    self.map_str_def = MAP_STR_FUNC(self.map_str_callback)
     
     MAP_MR_FUNC = CFUNCTYPE(c_void_p,c_int,POINTER(c_char),c_int,
                             POINTER(c_char),c_int,c_void_p,c_void_p)
@@ -178,7 +178,20 @@ class mrmpi:
     if self.map_argcount == 2: self.map_caller(itask,self)
     else: self.map_caller(itask,self,self.map_ptr)
     
-  def map_file_list(self,file,map,ptr=None,addflag=0):
+  def map_file(self,files,map,ptr=None,addflag=0):
+    self.map_caller = map
+    self.map_argcount = map.func_code.co_argcount
+    self.map_ptr = ptr
+    cfiles = (c_char_p*len(files))(*files)   # array of C strings from list
+    if not addflag:
+      n = self.lib.MR_map_file(self.mr,len(cfiles),cfiles,
+                               self.map_file_def,None)
+    else:
+      n = self.lib.MR_map_file_add(self.mr,len(cfiles),cfiles,
+                                   self.map_file_def,None,addflag)
+    return n
+
+  def map_file_file(self,file,map,ptr=None,addflag=0):
     self.map_caller = map
     self.map_argcount = map.func_code.co_argcount
     self.map_ptr = ptr
@@ -189,14 +202,14 @@ class mrmpi:
                                         None,addflag)
     return n
 
-  def map_dir(self,dir,oneflag,map,ptr=None,addflag=0):
+  def map_dir(self,dir,selfflag,map,ptr=None,addflag=0):
     self.map_caller = map
     self.map_argcount = map.func_code.co_argcount
     self.map_ptr = ptr
     if not addflag:
[…]

### 2011-01-06 src/cmapreduce.cpp

@@ -138,7 +138,27 @@ uint64_t MR_map_add(void *MRptr, int nmap,
   return mr->map(nmap,appmap,APPptr,addflag);
 }
 
-uint64_t MR_map_file_list(void *MRptr, char *file,
+uint64_t MR_map_file(void *MRptr, int nfile, char **files,
+			  void (*mymap)(int, char *, void *, void *),
+			  void *APPptr)
+{
+  typedef void (MapFunc)(int, char *, KeyValue *, void *);
+  MapReduce *mr = (MapReduce *) MRptr;
+  MapFunc *appmap = (MapFunc *) mymap;
+  return mr->map(nfile,files,appmap,APPptr);
+}
+
+uint64_t MR_map_file_add(void *MRptr, int nfile, char **files,
+			      void (*mymap)(int, char *, void *, void *),
+			      void *APPptr, int addflag)
+{
+  typedef void (MapFunc)(int, char *, KeyValue *, void *);
+  MapReduce *mr = (MapReduce *) MRptr;
+  MapFunc *appmap = (MapFunc *) mymap;
+  return mr->map(nfile,files,appmap,APPptr,addflag);
+}
+
+uint64_t MR_map_file_file(void *MRptr, char *file,
 			  void (*mymap)(int, char *, void *, void *),
 			  void *APPptr)
 {
@@ -148,7 +168,7 @@ uint64_t MR_map_file_list(void *MRptr, char *file,
   return mr->map(file,appmap,APPptr);
 }
 
-uint64_t MR_map_file_list_add(void *MRptr, char *file,
+uint64_t MR_map_file_file_add(void *MRptr, char *file,
 			      void (*mymap)(int, char *, void *, void *),
 			      void *APPptr, int addflag)
 {
@@ -158,9 +178,9 @@ uint64_t MR_map_file_list_add(void *MRptr, char *file,
   return mr->map(file,appmap,APPptr,addflag);
 }
 
-uint64_t MR_map_dir(void *MRptr, char *dir, int oneflag,
-		    void (*mymap)(int, char *, void *, void *),
-		    void *APPptr)
+uint64_t MR_map_file_dir(void *MRptr, char *dir, int oneflag,
+			 void (*mymap)(int, char *, void *, void *),
+			 void *APPptr)
 {
   typedef void (MapFunc)(int, char *, KeyValue *, void *);
   MapReduce *mr = (MapReduce *) MRptr;
[…]

### 2011-01-06 src/cmapreduce.h

@@ -48,22 +48,30 @@ uint64_t MR_map(void *MRptr, int nmap,
 uint64_t MR_map_add(void *MRptr, int nmap,
 		    void (*mymap)(int, void *KVptr, void *APPptr),
 		    void *APPptr, int addflag);
-uint64_t MR_map_file_list(void *MRptr, char *file,
+uint64_t MR_map_file(void *MRptr, int nfile, char *files,
+		     void (*mymap)(int, char *, 
+				   void *KVptr, void *APPptr),
+		     void *APPptr);
+uint64_t MR_map_file_add(void *MRptr, int nfile, char *files,
+			 void (*mymap)(int, char *, 
+				       void *KVptr, void *APPptr),
+			 void *APPptr, int addflag);
+uint64_t MR_map_file_file(void *MRptr, char *file,
 			  void (*mymap)(int, char *, 
 					void *KVptr, void *APPptr),
 			  void *APPptr);
-uint64_t MR_map_file_list_add(void *MRptr, char *file,
+uint64_t MR_map_file_file_add(void *MRptr, char *file,
 			      void (*mymap)(int, char *, 
 					    void *KVptr, void *APPptr),
 			      void *APPptr, int addflag);
-uint64_t MR_map_dir(void *MRptr, char *dir, int oneflag,
-		    void (*mymap)(int, char *, 
-				  void *KVptr, void *APPptr),
-		    void *APPptr);
-uint64_t MR_map_dir_add(void *MRptr, char *dir, int oneflag,
-			void (*mymap)(int, char *, 
-				      void *KVptr, void *APPptr),
-			void *APPptr, int addflag);
+uint64_t MR_map_file_dir(void *MRptr, char *dir, int oneflag,
+			 void (*mymap)(int, char *, 
+				       void *KVptr, void *APPptr),
+			 void *APPptr);
+uint64_t MR_map_file_dir_add(void *MRptr, char *dir, int oneflag,
+			     void (*mymap)(int, char *, 
+					   void *KVptr, void *APPptr),
+			     void *APPptr, int addflag);
 uint64_t MR_map_file_char(void *MRptr, int nmap, int nfiles, char **files,
 			  char sepchar, int delta,
 			  void (*mymap)(int, char *, int, 
@@ -92,6 +100,7 @@ uint64_t MR_map_mr_add(void *MRptr, void *MRptr2,
 		       void (*mymap)(uint64_t, char *, int, char *, int, 
[…]

### 2011-01-06 src/mapreduce.cpp

@@ -904,101 +904,25 @@ uint64_t MapReduce::gather(int numprocs)
 uint64_t MapReduce::map(int nmap, void (*appmap)(int, KeyValue *, void *),
 			void *appptr, int addflag)
 {
-  MPI_Status status;
-
-  if (timer) start_timer();
-  if (verbosity) file_stats(0);
-
-  if (!allocated) allocate();
-  if (kmv) myfree(kmv->memtag);
-  delete kmv;
-  kmv = NULL;
-
-  if (addflag == 0) {
-    if (kv) myfree(kv->memtag);
-    delete kv;
-    kv = new KeyValue(this,kalign,valign,memory,error,comm);
-    kv->set_page();
-  } else if (kv == NULL) {
-    kv = new KeyValue(this,kalign,valign,memory,error,comm);
-    kv->set_page();
-  } else {
-    kv->append();
-  }
-
-  // nprocs = 1 = all tasks to single processor
-  // mapstyle 0 = chunk of tasks to each proc
-  // mapstyle 1 = strided tasks to each proc
-  // mapstyle 2 = master/slave assignment of tasks
-
-  if (nprocs == 1) {
-    for (int itask = 0; itask < nmap; itask++)
-      appmap(itask,kv,appptr);
-
-  } else if (mapstyle == 0) {
-    uint64_t nmap64 = nmap;
-    int lo = me * nmap64 / nprocs;
-    int hi = (me+1) * nmap64 / nprocs;
-    for (int itask = lo; itask < hi; itask++)
-      appmap(itask,kv,appptr);
-
-  } else if (mapstyle == 1) {
-    for (int itask = me; itask < nmap; itask += nprocs)
-      appmap(itask,kv,appptr);
-
-  } else if (mapstyle == 2) {
-    if (me == 0) {
-      int doneflag = -1;
-      int ndone = 0;
-      int itask = 0;
-      for (int iproc = 1; iproc < nprocs; iproc++) {
-	if (itask < nmap) {
-	  MPI_Send(&itask,1,MPI_INT,iproc,0,comm);
-	  itask++;
-	} else {
-	  MPI_Send(&doneflag,1,MPI_INT,iproc,0,comm);
-	  ndone++;
-	}
-      }
-      while (ndone < nprocs-1) {
-	int iproc,tmp;
[…]

### 2011-01-06 src/mapreduce.h

@@ -75,6 +75,8 @@ class MapReduce {
 
   uint64_t map(int, void (*)(int, class KeyValue *, void *),
 	       void *, int addflag = 0);
+  uint64_t map(int, char **, void (*)(int, char *, class KeyValue *, void *),
+	       void *, int addflag = 0);
   uint64_t map(char *, void (*)(int, char *, class KeyValue *, void *),
 	       void *, int addflag = 0);
   uint64_t map(char *, int, void (*)(int, char *, class KeyValue *, void *),
@@ -206,9 +208,13 @@ class MapReduce {
   void copy_kv(KeyValue *);
   void copy_kmv(KeyMultiValue *);
 
+  uint64_t map_task(int, char **,
+		    void (*)(int, KeyValue *, void *),
+		    void (*)(int, char *, KeyValue *, void *),
+		    void *, int, int);
   uint64_t map_file(int, int, char **,
 		    void (*)(int, char *, int, class KeyValue *, void *),
-		    void *, int addflag);
+		    void *, int);
 
   void sort_kv(int);
   void sort_onepage(int, int, char *, char *, char *);

### 2011-02-23 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
-linear-algebra concepts implemented in a MapReduce framework, with
-data stored in four MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$
-stores the matrix; $MT$ stores the row index of ``empty'' rows of $A$; 
-and $x$ and $y$ are
-vectors of length $|V|$.  The nonzeros of matrix $A$ are initialized
-as described above; the PageRank vector $x$ is initialized uniformly.
-A MapReduce object $MT$ contains the indices of all-zero rows of $A$,
+linear-algebra concepts implemented in a MapReduce framework.  Related
+algorithms, each with their own optimizations, are discussed in
+\cite{Kang09,Lin10}.  In this version data is stored in four MapReduce
+objects: $A$, $MT$, $x$, and $y$.  $A$ stores the matrix; $MT$ stores
+the row index of ``empty'' rows of $A$; and $x$ and $y$ are vectors of
+length $|V|$.  The nonzeros of matrix $A$ are initialized as described
+above; the PageRank vector $x$ is initialized uniformly.  A MapReduce
+object $MT$ contains the indices of all-zero rows of $A$,
 corresponding to vertices with no outlinks.  The PageRank algorithm is
@@
 strategy of pre-aggregating to improve data locality is useful in many
-MapReduce algorithms, but assumes that identical keys in different
+MapReduce algorithms; it assumes that identical keys in different
 MapReduce objects can be mapped to the same processor.
@@
 
-A connected component of a graph is a set of vertices where all pairs
-of vertices in the set are connected by a path of edges.  A sparse
-graph may contain many such components.  Figure~\ref{fig:cc} outlines
-a MapReduce algorithm that labels each vertex in a graph with a
-component ID.  All vertices in the same component are labeled with
+A connected component (CC) of a graph is a set of vertices where all
+pairs of vertices in the set are connected by a path of edges.  A
+sparse graph may contain many such components.  Figure~\ref{fig:cc}
+outlines a MapReduce algorithm that labels each vertex in a graph with
+a component ID.  All vertices in the same component are labeled with
 the same ID, which is the ID of some vertex in the component.  We
-assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
-exposition follows the connected-component algorithm presented
-in~\cite{Cohen09}, with the addition of logic that load-balances data
-across processors, which is important when one or a few giant
-components exist in the graph.
+assume an input graph $E$ of undirected edges $(V_i,V_j)$.  Kang
+et.~al.~\cite{Kang09} propose a MapReduce CC algorithm based on a
+matrix-vector multiplication kernel.  This exposition follows the CC
+algorithm presented in~\cite{Cohen09}, with the addition of logic that
+load-balances data across processors, which is important when one or a
+few giant components exist in the graph.
 
@@
 and the MR object $V$ holds the zone assignment for every vertex.  The
-algorithm thus finds all connected components in the graph
-simultaneously.  The number of iterations required depends on the
-largest diameter of any component in the graph.
+algorithm thus finds all CCs in the graph simultaneously.  The number
+of iterations required depends on the largest diameter of any
+component in the graph.
 

### 2011-02-23 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 post-processing analysis of simulation data~\cite{Tu08}, graph
-algorithmics~\cite{Cohen09}, and linear algebra
-operations~\cite{Ekanayake09}.  The paper by Tu et al.~\cite{Tu08} was
-particularly insightful to us, because it described how MapReduce
-could be implemented on top of the ubiquitous distributed-memory
-message-passing interface (MPI), and how the intermediate data-shuffle
-operation is conceptually identical to the familiar MPI\_Alltoall
-operation.  Their implementation of MapReduce was within a Python
-wrapper to simplify the writing of user programs.  The paper motivated
-us to develop our own C++ library built on top of MPI for use in graph
-analytics, which we initially released as open-source software in
-mid-2009.  We have since worked to optimize several of the library's
-underlying algorithms and to enable its operation in out-of-core mode
-for larger data sets.  These algorithmic improvements are described in
-this paper and are part of the current downloadable version; see
-Section~\ref{sec:thanks} for details.
+algorithmics~\cite{Cohen09,Kang09,Lin10}, and linear algebra
+operations~\cite{Ekanayake09}.  The paper by Tu et
+al.~\cite{Tu08} was particularly insightful to us, because it
+described how MapReduce could be implemented on top of the ubiquitous
+distributed-memory message-passing interface (MPI), and how the
+intermediate data-shuffle operation is conceptually identical to the
+familiar MPI\_Alltoall operation.  Their implementation of MapReduce
+was within a Python wrapper to simplify the writing of user programs.
+The paper motivated us to develop our own C++ library built on top of
+MPI for use in graph analytics, which we initially released as
+open-source software in mid-2009.  We have since worked to optimize
+several of the library's underlying algorithms and to enable its
+operation in out-of-core mode for larger data sets.  These algorithmic
+improvements are described in this paper and are part of the current
+downloadable version; see Section~\ref{sec:thanks} for details.
 

### 2011-02-23 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 MapReduce algorithms described in this section, we list the line
-counts of the C++ codes used to perform the benchmarks.  Each code
-makes calls to the MR-MPI library and implements the algorithm in the
-indicated figure.  The line counts exclude comments and blank lines
-which typically add about 30\% to the total.  They also exclude the
-routines used to generate the R-MAT matrix input (except in the case
-of the two R-MAT algorithms themselves) or process it into
-upper-triangular form, since that is common to several of the
-benchmarks.
+counts of the C++ codes used to perform the benchmarks in
+Table~\ref{t:linecount}.  Each code makes calls to the MR-MPI library
+and implements the algorithm in the indicated figure.  The line counts
+exclude comments and blank lines which typically add about 30\% to the
+total.  They also exclude the routines used to generate the R-MAT
+matrix input (except in the case of the two R-MAT algorithms
+themselves) or process it into upper-triangular form, since that is
+common to several of the benchmarks.
 
@@
   algorithms in the associated figures.}
-\label{t:rmat}
+\label{t:linecount}
 \end{center}

### 2011-02-23 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@815 3fc257bb-8550-0410-81b3-a5c103da83f7
Many redone figures so they look better in B&W.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@817 3fc257bb-8550-0410-81b3-a5c103da83f7

--- 2010-11-29 paper/results.tex
@@
 To give a rough sense of the effort involved to implement the various
-MapReduce algorithms described in this section, we list the line
-counts of the C++ codes used to perform the benchmarks.  Each code
-makes calls to the MR-MPI library and implements the algorithm in the
-indicated figure.  The line counts exclude comments and blank lines
-which typically add about 30\% to the total.  They also exclude the
-routines used to generate the R-MAT matrix input (except in the case
-of the two R-MAT algorithms themselves) or process it into
-upper-triangular form, since that is common to several of the
-benchmarks.
+MapReduce algorithms described in this section, we list in Table~\ref{t:linecount} the line
+counts of the C++ codes used to perform the benchmarks.
+Each code makes calls to the MR-MPI library
+and implements the algorithm in the indicated figure.  The line counts
+exclude comments and blank lines which typically add about 30\% to the
+total.  They also exclude the routines used to generate the R-MAT
+matrix input (except in the case of the two R-MAT algorithms
+themselves) or process it into upper-triangular form, since that is
+common to several of the benchmarks.
 
@@
 that fit in memory, so out-of-core operations were not needed.  For
-these experiments, we used two R-MAT data sets with with $2^{25}$
-vertices and $2^{28}$ edges, with parameters given in
-Table~\ref{t:rmat}.  The ``nice'' parameters produce a less-skewed
-sparsity pattern in the matrix than the ``nasty'' parameters, as
-indicated by the maximum degree metric.  We ran both the PageRank and
-Connected Components algorithms on these matrices.
+these experiments, we used two R-MAT data sets with $2^{25}$ vertices
+and $2^{28}$ edges, with parameters given in Table~\ref{t:rmat}.  The
+``nice'' parameters produce a less-skewed sparsity pattern in the
+matrix than the ``nasty'' parameters, as indicated by the maximum
+degree metric.  We ran both the PageRank and Connected Components
+algorithms on these matrices.
 

### 2011-03-10 paper/results.tex
Updated line counts. Finished proofs for submission
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@866 3fc257bb-8550-0410-81b3-a5c103da83f7

@@
 \hline
-Lines of code & 208 & 216 & 603 & 367 & 453 & 393 & 394 & 448 \\
+Lines of code & 161 & 163 & 463 & 266 & 254 & 291 & 243 & 262 \\
 Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
@@
 (non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
-\ref{fig:rmat2}.  It required 845 lines of code, mainly because it
+\ref{fig:rmat2}.  It required 453 lines of code, mainly because it
 must implement the parallelism explicitly, whereas the codes of this
 section relegate that task to the MR-MPI library.  A similar all-MPI
-implementation of the PageRank algorithm required 718 lines.  By
+implementation of the PageRank algorithm required 506 lines.  By
 contrast an implementation of PageRank using the Trilinos parallel
 linear algebra library (discussed in Section
-\ref{subsec:results_pagerank}) required only 153 lines of code, since
+\ref{subsec:results_pagerank}) required only 76 lines of code, since
 entire dot-products and matrix-vector multiplications can be invoked by single function calls to
@@
 implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
-required 697 lines of Java.
+required 527 lines of Java.
 
